### 2023-03-16 outputs/paper/paper.qmd
pushin

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: html
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

```{r}

#quarto references itself. wrapping in here:here goes in project 
knitr::include_graphics(here::here("inputs/data/apod_20230227.jpg"))

```

```{r}
#| label: fig-energy
#| fig-cap: Energy Variable
#| echo: false

library(lubridate)
df <- readRDS(here::here("inputs/data/df.rds"))
df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = energy, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

```

```{r}
#| label: fig-danceability
#| fig-cap: Danceability Variable
#| echo: false

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

```

```{r}
#| label: fig-valence
#| fig-cap: Valence Variable
#| echo: false

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

```

```{r}
#| label: fig-acousticness
#| fig-cap: Acousticness Variable
#| echo: false

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

```

```{r}
#| label: fig-instrumentalness
#| fig-cap: Instrumentalness Variable
#| echo: false

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

```

```{r}
#| label: fig-tempo
#| fig-cap: Tempo Variable
#| echo: false

df %>%
  filter(album_release_date_precision == "day") %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#FFB6C1", "#808080")) +
  labs(color = "Has 'y' in jack?") +
  facet_wrap(~ artist_name, ncol = 3)

```

# Introduction

# Data {#sec-data}

# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

### 2023-03-16 scripts/01-download_data_spotify.R

@@ -1,0 +1,39 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from Spotify
+# Author: Michaela Drouillard
+# Data: 27 February 2023
+# Contact: rohan.alexander@utoronto.ca 
+# License: MIT
+# Pre-requisites: -
+# Any other information needed?-
+
+#### Workspace setup ####
+
+#run install.packages in console!!!!
+library(spotifyr)
+library(tidyverse)
+
+
+#### Download data ####
+
+
+?spotifyr
+lana <- get_artist_audio_features("lana del rey")
+taylor <- get_artist_audio_features("taylor swift")
+vincent <- get_artist_audio_features("st vincent")
+chicks <- get_artist_audio_features("the chicks")
+lorde <- get_artist_audio_features("lorde")
+flo <- get_artist_audio_features("florence and the machine")
+
+
+#### Save data ####
+
+saveRDS(lana, "inputs/data/lana.rds")
+saveRDS(taylor, "inputs/data/taylor.rds")
+saveRDS(vincent, "inputs/data/vincent.rds")
+saveRDS(lorde, "inputs/data/lorde.rds")
+saveRDS(chicks, "inputs/data/chicks.rds")
+saveRDS(flo, "inputs/data/flo.rds")
+?spotifyr
+
+

### 2023-03-16 scripts/02-data_cleaning.R

@@ -1,0 +1,188 @@
+#### Preamble ####
+# Purpose: Binding the data
+# Author: Michela Drouillard
+# Data: 28 February 2023 
+# Contact: michaela.drouillard@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Clean data ####
+# Binding Data
+
+df <- rbind(lana, taylor, vincent, lorde, chicks, flo)
+df$jack = NA
+
+jack_tracks <- c(
+  "Green Light",
+  "Sober",
+  "The Louvre",
+  "Liability",
+  "Hard Feelings/Loveless",
+  "Sober II (Melodrama)",
+  "Writer in the Dark",
+  "Supercut",
+  "Liability (Reprise)",
+  "Perfect Places",
+  "New York",
+  "Masseduction",
+  "Happy Birthday, Johnny",
+  "Sugarboy",
+  "Look What You Made Me Do",
+  "Getaway Car",
+  "Dress",
+  "This Is Why We Can't Have Nice Things",
+  "Call It What You Want",
+  "New Year's Day",
+  "Mariners Apartment Complex",
+  "Venice Bitch",
+  "hope is a dangerous thing for a woman like me to have - but i have it",
+  "How To Disappear",
+  "Norman Fucking Rockwell",
+  "Fuck It, I Love You",
+  "Love Song",
+  "Cinnamon Girl",
+  "The Greatest",
+  "California",
+  "Happiness Is a Butterfly",
+  "Looking for America",
+  "Cruel Summer",
+  "Lover",
+  "The Archer",
+  "I Think He Knows",
+  "Paper Rings",
+  "Cornelia Street",
+  "Death By a Thousand Cuts",
+  "London Boy",
+  "Soon You'll Get Better",
+  "False God",
+  "Daylight",
+  "Lover (Remix)",
+  "Gaslighter",
+  "Sleep at Night",
+  "Texas Man",
+  "Everybody Loves You",
+  "For Her",
+  "March March",
+  "My Best Friend's Weddings",
+  "Tights on My Boat",
+  "Julianna Calm Down",
+  "Young Man",
+  "Hope It's Something Good",
+  "Set Me Free",
[…]

### 2023-03-16 scripts/04-test_data.R

@@ -1,0 +1,42 @@
+#### Preamble ####
+# Purpose: Tests... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Data: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Test data ####
+
+colnames(df)
+str(df)
+
+df$danceability <- as.numeric(df$danceability)
+df$energy <- as.numeric(df$energy)
+df$acousticness <- as.numeric(df$acousticness)
+df$tempo <- as.numeric(df$tempo)
+df$instrumentalness <- as.numeric(df$instrumentalness)
+df$loudness <- as.numeric(df$loudness)
+df$valence <- as.numeric(df$valence)
+df$jack <- as.numeric(df$jack)
+df$key <- as.numeric(df$key)
+df$mode <- as.numeric(df$mode)
+
+# Fit a logistic regression model
+model <- glm(jack ~ danceability + energy + acousticness + tempo + instrumentalness + key + mode + loudness + valence, data = df, family = binomial)
+
+# Print the model summary
+summary(model)
+
+# Calculate the Pearson correlation coefficients
+correlations <- cor(df[c("danceability", "energy", "acousticness", "liveness", "tempo", "key", "valence", "loudness", "mode", "jack")], method = "pearson")
+
+# Print the correlation matrix
+print(correlations)
+

### 2023-03-16 scripts/05-graphs.R

@@ -1,0 +1,75 @@
+#### Preamble ####
+# Purpose: Graphs
+# Author: Michaela Drouillard
+# Data: 28 February 2023
+# Contact: michaela.drouillard@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+
+library(lubridate)
+
+df <- readRDS(here::here("inputs/data/df.rds"))
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = energy, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#FFB6C1", "#808080")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3)
+
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#FFB6C1", "#808080")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3) 
+
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#FFB6C1", "#808080")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3) 
+
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#FFB6C1", "#808080")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3) 
+
+df %>%
+  filter(album_release_date_precision == "day") %>%
[…]

### 2023-03-21 scripts/01-download_data_spotify.R
trying again?

@@ -2,14 +2,13 @@
 # Purpose: Downloads and saves the data from Spotify
 # Author: Michaela Drouillard
 # Data: 27 February 2023
-# Contact: rohan.alexander@utoronto.ca 
+# Contact: michaela.drouillard@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: -
 # Any other information needed?-
 
 #### Workspace setup ####
 
-#run install.packages in console!!!!
 library(spotifyr)
 library(tidyverse)
 
@@ -28,12 +27,13 @@ flo <- get_artist_audio_features("florence and the machine")
 
 #### Save data ####
 
-saveRDS(lana, "inputs/data/lana.rds")
-saveRDS(taylor, "inputs/data/taylor.rds")
-saveRDS(vincent, "inputs/data/vincent.rds")
-saveRDS(lorde, "inputs/data/lorde.rds")
-saveRDS(chicks, "inputs/data/chicks.rds")
-saveRDS(flo, "inputs/data/flo.rds")
-?spotifyr
+write_csv(lana, "inputs/data/lana.csv") 
+write_csv(taylor, "inputs/data/taylor.csv") 
+write_csv(vincent, "inputs/data/vincent.csv") 
+write_csv(chicks, "inputs/data/chicks.csv") 
+write_csv(lorde, "inputs/data/lorde.csv") 
+write_csv(flo, "inputs/data/flo.csv") 
+
+
 
 

### 2023-03-21 scripts/02-data_cleaning.R

@@ -10,7 +10,13 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
+chicks <- read_csv("inputs/data/chicks.csv")
+flo <- read_csv("inputs/data/flo.csv")
+lorde <- read_csv("inputs/data/lorde.csv")
+vincent <- read_csv("inputs/data/vincent.csv")
+lana <- read_csv("inputs/data/lana.csv")
+taylor <- read_csv("inputs/data/taylor.csv")
 
 #### Clean data ####
 # Binding Data
@@ -163,26 +169,21 @@ jack_tracks <- c(
   "A&W")
 df$jack <- ifelse(df$track_name %in% jack_tracks, "1", df$jack)
 
-table(df$jack)
 
-df[df$jack == "1", "track_name"]
-jack_tracks_test <- subset(df, jack == "1", select = track_name)$track_name
-jack_tracks_test
-#df <- df[!duplicated(df$track_name), ]
-#jack_tracks_test <- subset(df, jack == "Y", select = track_name)$track_name
-
-
+df$danceability <- as.numeric(df$danceability)
+df$energy <- as.numeric(df$energy)
+df$acousticness <- as.numeric(df$acousticness)
+df$tempo <- as.numeric(df$tempo)
+df$instrumentalness <- as.numeric(df$instrumentalness)
+df$loudness <- as.numeric(df$loudness)
+df$valence <- as.numeric(df$valence)
+df$jack <- as.factor(df$jack)
+df$key <- as.numeric(df$key)
+df$mode <- as.numeric(df$mode)
 
 #### Save data ####
 
-saveRDS(df, "inputs/data/df.rds")
-# write.csv(df, "inputs/data/df.csv")
-# 
-# library(tidyverse)
-# my_df <- as.tibble(df)
-# 
-# # Write the data frame to a CSV file
-# class(my_df)
-# write.csv(my_df, "inputs/data/jack_df.csv", row.names = FALSE)
-# ?write.csv
+write.csv(df, "inputs/data/df.csv")
 
+df |> 
+  summary()

### 2023-03-21 scripts/04-test_data.R

@@ -14,29 +14,20 @@ library(tidyverse)
 
 #### Test data ####
 
+
+#Checking which tracks have Jack as a producer
+jack_tracks_test <- subset(df, jack == "1", select = track_name)$track_name
+jack_tracks_test
+
+#There should be 142, but in both of these tests it's 189. Are duplicates okay?
+
+length(jack_tracks_test)
+table(df$jack)
+
 colnames(df)
 str(df)
 
-df$danceability <- as.numeric(df$danceability)
-df$energy <- as.numeric(df$energy)
-df$acousticness <- as.numeric(df$acousticness)
-df$tempo <- as.numeric(df$tempo)
-df$instrumentalness <- as.numeric(df$instrumentalness)
-df$loudness <- as.numeric(df$loudness)
-df$valence <- as.numeric(df$valence)
-df$jack <- as.numeric(df$jack)
-df$key <- as.numeric(df$key)
-df$mode <- as.numeric(df$mode)
 
-# Fit a logistic regression model
-model <- glm(jack ~ danceability + energy + acousticness + tempo + instrumentalness + key + mode + loudness + valence, data = df, family = binomial)
 
-# Print the model summary
-summary(model)
 
-# Calculate the Pearson correlation coefficients
-correlations <- cor(df[c("danceability", "energy", "acousticness", "liveness", "tempo", "key", "valence", "loudness", "mode", "jack")], method = "pearson")
-
-# Print the correlation matrix
-print(correlations)
 

### 2023-03-21 scripts/05-graphs.R

@@ -9,10 +9,17 @@
 
 
 #### Workspace setup ####
-
+library(tidyverse)
 library(lubridate)
 
-df <- readRDS(here::here("inputs/data/df.rds"))
+df <- read_csv("inputs/data/df.csv")
+
+str(df)
+#For whatever reason, Jack didn't stay a factor. Making it a factor again here
+df$jack <- as.factor(df$jack)
+
+
+
 df %>%
   filter(album_release_date_precision == "day") %>%
   mutate(album_release_date = ymd(album_release_date)) %>%
@@ -21,7 +28,8 @@ df %>%
   theme_minimal() +
   scale_color_manual(values = c("#FFB6C1", "#808080")) +
   labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3)
+  facet_wrap(~ artist_name, ncol = 3) +
+  geom_smooth(method = "lm", se = FALSE)
 
 df %>%
   filter(album_release_date_precision == "day") %>%
@@ -31,7 +39,8 @@ df %>%
   theme_minimal() +
   scale_color_manual(values = c("#FFB6C1", "#808080")) +
   labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3) 
+  facet_wrap(~ artist_name, ncol = 3) +
+  geom_smooth(method = "lm", se = FALSE)
 
 df %>%
   filter(album_release_date_precision == "day") %>%
@@ -41,7 +50,8 @@ df %>%
   theme_minimal() +
   scale_color_manual(values = c("#FFB6C1", "#808080")) +
   labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3) 
+  facet_wrap(~ artist_name, ncol = 3) +
+  geom_smooth(method = "lm", se = FALSE)
 
 df %>%
   filter(album_release_date_precision == "day") %>%
@@ -51,7 +61,8 @@ df %>%
   theme_minimal() +
   scale_color_manual(values = c("#FFB6C1", "#808080")) +
   labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3) 
+  facet_wrap(~ artist_name, ncol = 3) +
+  geom_smooth(method = "lm", se = FALSE)
 
 df %>%
   filter(album_release_date_precision == "day") %>%
@@ -61,7 +72,8 @@ df %>%
   theme_minimal() +
[…]

### 2023-03-21 scripts/06-models.R

@@ -1,0 +1,67 @@
+#### Preamble ####
+# Purpose: Models the Jack Antonoffification of Pop Music
+# Author: Michaela Drouillard
+# Data: 17 March 2023
+# Contact: michaela.drouillard@mail.utoronto.ca
+# License: MIT
+
+#Going to make a smaller df that only had the columns we want to capture correlations
+
+library(ggplot2)
+
+data <- read_csv("inputs/data/df.csv")
+
+#chisquared test from website http://www.sthda.com/english/wiki/chi-square-test-of-independence-in-r
+#not including liveness because its negative values
+df_chisq <- df[, c("artist_name", "track_name", "danceability",                
+                   "energy", "key","loudness", "speechiness", "acousticness",                
+                   "instrumentalness", "valence", "jack")]
+
+chisq <- chisq.test(df_chisq)
+chisq
+
+
+
+
+
+
+
+
+# select the relevant columns
+cols <- c("danceability", "energy", "key", "loudness", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "jack")
+
+# create a new dataframe with only the selected columns
+df_small <- data[,cols]
+levels(df_small$jack)
+# Replace NAs with 0 in the jack column
+df_small$jack <- ifelse(is.na(df_small$jack), 0, df_small$jack)
+
+
+
+# perform t-tests for continuous variables
+for (col in cols[!cols %in% c("artist_name", "track_name", "jack")]) {
+  ttest <- t.test(df_small[[col]] ~ df_small$jack)
+  print(paste("T-test for", col))
+  print(ttest)
+}
+
+#SIGNIFICANT P VALUES: danceability, energy, loudness, acousticness, instrumentalness, liveness, valence
+#NOT SIGNIFCANT: Speechiness, Key
+
+
+# perform chi-square tests for categorical variables
+for (col in cols[!cols %in% c("artist_name", "track_name", "jack")]) {
+  chisq <- chisq.test(table(df_small[[col]], df_small$jack))
+  print(paste("Chi-square test for", col))
+  print(chisq)
+}
+
+#SIGNIFICANT P VALUES: all of them??
+
+
+
+
+
+
+
+

### 2023-03-21 outputs/paper/paper.qmd
added bleachers data, edited the data sections

@@
 author: 
-  - First author
-  - Another author
+  - Michaela Drouillard
 thanks: "Code and data are available at: LINK."
@@
 
+I pulled the data using the Spotify API, which I accessed using the `spotifyr` package in R.
+I used the `get_artist_audio_features` and `get_track_audio_analysis functions` function to acquire data on Lorde, Taylor Swift, St.
+Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
+
+### Artist Selection
+
+I chose these artists because they've produced multiple albums, and at least one album with Antonoff, as opposed to just individual songs.
+I chose the Chicks because I figured that collaborating with Antonoff after an extended gap in a longer spanning career than the other artists might offer some interesting results.
+The data points at beginning of their career reflect music that existed in a different time from the other artists in this dataset.
+
+### Data Cleaning
+
+I joined these data sets using the `rbind` function in base R, and created an additional variable called "jack", which contains a 1 if Jack Antonoff produced or co-produced the song, and a 0 if he had nothing to do with it.
+I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page.
+
+### Spotify's API
+
+So what do any one these features mean?
+
+The Spotify API returns JSON metadata directly from the Spotify Data Catalogue.
+
+#### Audio Features Dataset
+
+The `get_artist_audio_features` provides a general profile of each song, according to variables developped by spotify, which makes for simple comparisons across artists.
+
+However, it's hard to gauge exactly what variables like "danceability" or "valence" mean, even using the documentation.
+For instance, "valence" is described as "A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
+Tracks with high valence sound more positive (e.g.
+happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g.
+sad, depressed, angry)", and danceability as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+A value of 0.0 is least danceable and 1.0 is most danceable".
+Neither of these definitions fully cover what metrics and thresholds go into something like "danceability".
+However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums -- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
+The full documentation for each variable included in this study is available in the Appendix.
+
+The final audio features in this study data set contains the variables: "artist_name", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack".
+
+```{r}
+library(kableExtra)
+df |>
+  knitr::kable()
+
+```
+
+#### Audio Analysis Dataset
+
+The `get_track_audio_analysis` offers a more granular look at each track.
+The variables included are arrays of objects, which are the kinds of things that are a bit harder to wrap your mind around if you were to visualize each one across each artist.
+They include: track, bars, beats, sections, segments, tatums, and meta, which includes information on which platforms and Analyzer versions were used to analyze the track, along with metadata on the status and duration of the analysis.
+
+This data will be used when building a logistic regression model to determine the character and predictability of Jack Antonoff's music.
+
 # Model
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
+
+### `get_artist_audio_features` documentation:
+
+-   **acousticness**: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.
+
+-   **danceability**: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+A value of 0.0 is least danceable and 1.0 is most danceable.
+
+-   **energy**: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
+Typically, energetic tracks feel fast, loud, and noisy.
+For example, death metal has high energy, while a Bach prelude scores low on the scale.
+Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy.
+
+-   **instrumentalness**: Predicts whether a track contains no vocals.
+"Ooh" and "aah" sounds are treated as instrumental in this context.
+Rap or spoken word tracks are clearly "vocal".
+The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content.
+Values above 0.5 are intended to represent instrumental tracks, but confidence is higher as the value approaches 1.0.
+
+-   **key**: The key the track is in.
+Integers map to pitches using standard [Pitch Class notation](https://en.wikipedia.org/wiki/Pitch_class).
+E.g.
+0 = C, 1 = C♯/D♭, 2 = D, and so on.
+If no key was detected, the value is -1.
+
+-   **liveness**: Detects the presence of an audience in the recording.
+Higher liveness values represent an increased probability that the track was performed live.
+A value above 0.8 provides strong likelihood that the track is live.
+
+-   **loudness:** The overall loudness of a track in decibels (dB).
+Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks.
+Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude).
+Values typically range between -60 and 0 db.
+
+-   **mode:** Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.
+
+-   **tempo:** The overall estimated tempo of a track in beats per minute (BPM).
+In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration.
+
+-   **valence:** A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
+Tracks with high valence sound more positive (e.g.
+happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g.
+sad, depressed, angry).
 

### 2023-03-21 scripts/01-download_data_spotify.R

@@ -13,26 +13,34 @@ library(spotifyr)
 library(tidyverse)
 
 
-#### Download data ####
+#### Download data audio features ####
 
 
 ?spotifyr
-lana <- get_artist_audio_features("lana del rey")
-taylor <- get_artist_audio_features("taylor swift")
-vincent <- get_artist_audio_features("st vincent")
-chicks <- get_artist_audio_features("the chicks")
-lorde <- get_artist_audio_features("lorde")
-flo <- get_artist_audio_features("florence and the machine")
+lanaaf <- get_artist_audio_features("lana del rey")
+tayloraf <- get_artist_audio_features("taylor swift")
+vincentaf <- get_artist_audio_features("st vincent")
+chicksaf <- get_artist_audio_features("the chicks")
+lordeaf <- get_artist_audio_features("lorde")
+floaf <- get_artist_audio_features("florence and the machine")
+bleachersaf <- get_artist_audio_features("bleachers")
 
 
-#### Save data ####
+#### Save data audio features ####
 
-write_csv(lana, "inputs/data/lana.csv") 
-write_csv(taylor, "inputs/data/taylor.csv") 
-write_csv(vincent, "inputs/data/vincent.csv") 
-write_csv(chicks, "inputs/data/chicks.csv") 
-write_csv(lorde, "inputs/data/lorde.csv") 
-write_csv(flo, "inputs/data/flo.csv") 
+write_csv(lanaaf, "inputs/data/lanaaf.csv") 
+write_csv(tayloraf, "inputs/data/tayloraf.csv") 
+write_csv(vincentaf, "inputs/data/vincentaf.csv") 
+write_csv(chicksaf, "inputs/data/chicksaf.csv") 
+write_csv(lordeaf, "inputs/data/lordeaf.csv") 
+write_csv(floaf, "inputs/data/floaf.csv") 
+write_csv(bleachersaf, "inputs/data/bleachersaf.csv") 
+
+#### Download data audio analysis####
+
+
+?spotifyr
+test <- get_track_audio_analysis("4brdb8L2Cy3e9AN8IfV9l8",  authorization = get_spotify_access_token())
 
 
 

### 2023-03-21 scripts/02-data_cleaning.R

@@ -11,17 +11,17 @@
 #### Workspace setup ####
 library(tidyverse)
 
-chicks <- read_csv("inputs/data/chicks.csv")
-flo <- read_csv("inputs/data/flo.csv")
-lorde <- read_csv("inputs/data/lorde.csv")
-vincent <- read_csv("inputs/data/vincent.csv")
-lana <- read_csv("inputs/data/lana.csv")
-taylor <- read_csv("inputs/data/taylor.csv")
-
+chicksaf <- read_csv("inputs/data/chicksaf.csv")
+floaf <- read_csv("inputs/data/floaf.csv")
+lordeaf <- read_csv("inputs/data/lordeaf.csv")
+vincentaf <- read_csv("inputs/data/vincentaf.csv")
+lanaaf <- read_csv("inputs/data/lanaaf.csv")
+tayloraf <- read_csv("inputs/data/tayloraf.csv")
+bleachersaf <- read_csv("inputs/data/bleachersaf.csv")
 #### Clean data ####
 # Binding Data
 
-df <- rbind(lana, taylor, vincent, lorde, chicks, flo)
+df <- rbind(lanaaf, tayloraf, vincentaf, lordeaf, chicksaf, floaf, bleachersaf)
 df$jack = NA
 
 jack_tracks <- c(
@@ -167,7 +167,8 @@ jack_tracks <- c(
   "Hits Different",
   "Did You Know That There's a Tunnel Under Ocean Blvd",
   "A&W")
-df$jack <- ifelse(df$track_name %in% jack_tracks, "1", df$jack)
+df$jack <- ifelse(df$track_name %in% jack_tracks, "1", "0")
+df$jack[df$artist_name == "Bleachers"] <- "1"
 
 
 df$danceability <- as.numeric(df$danceability)
@@ -181,9 +182,14 @@ df$jack <- as.factor(df$jack)
 df$key <- as.numeric(df$key)
 df$mode <- as.numeric(df$mode)
 
+
+###Keep only columns we want###
+
+df = subset(df, select = c("artist_name", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack"))
+
 #### Save data ####
+
 
 write.csv(df, "inputs/data/df.csv")
 
-df |> 
-  summary()
+

### 2023-03-23 outputs/paper/paper.qmd
wrote intro and gamman section

@@
 
+Sometime in early October of 2022 there was a chemical shift in my brain and I couldn't stop listening to Taylor Swift.
+I was a preteen when my best friend's sister showed us Our Song on the family desktop, went on to own a copy of Fearless, and was in the ninth grade when Red came out.
+My listening tapered off after that, other than ironically dancing to Blank Space in friends kitchens in college, and a road trip a few years ago where we belted Fearless (Three of the four people present were going through something).
+
+All this to say, I can quite confidently say I hadn't listened to much or any Taylor Swift until October, when I couldn't listen to anything other than Fearless or Red when deadlifting, walking to a friend's house, folding laundry, etc.
+I listened to so much Taylor in that month alone that she ended up as my #4 artist on Spotify Wrapped, which stops collecting its data for the year on October 31st.
+It didn't even capture what happened to me after Midnights dropped in November.
+
+I was standing outside of the Garrison eating a slice of pizza with some friends, and one of them asked what music we had been listening to lately.
+
+ME: "Midnights!"
+
+ZACH: "Is it good?"
+
+\[I blacked out.\]
+
+ME: "No!"
+
+\["*No*?" I thought to myself\]
+
+A STRANGER WALKING BY WHO HEARD EVERYTHING: "EXACTLY!!"
+
+Midnight was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this essay, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".
+
+Antonoff is prolific, to the extent that it is starting to bother people.
+He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, Minions: The Rise of Gru Soundtrack, and more.
+All of this and his own bands too: Bleachers, and Fun.
+
+When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed.
+What my friend was referring to, of course, is the fact that Jack Antonoff seems to be everywhere these days.
+As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely.
+When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis".
+
+What I am trying to get to here is that I don't really mind Jack Antonoff's music.
+I kind of like it actually.
+But I still say his voice with the same tone that everyone else does.
+There's just something about him...
+this sense that he's everywhere...
+and that he's just such a nice guy...
+Why are we all so fixated on him?
+What is it about Antonoff music that is so Antonoff-y?
+
+To find an answer to this question, I turned people, statistics, and the overlord of engineered virality itself -- the Spotify API.
+
+# An Interview with Caleb Gamman
+
+Around the time that Midnights came out, I came across a video on Twitter of a guy guessing which songs on Midnights were produced by Antonoff within seconds.
+The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style".
+He nails it.
+Only one song on the album wasn't produced by Antonoff, and he guesses it.
+
+I reached out to Gamman to ask him to explain Antonoff to me, only to find out, on the morning of the interview, that Gamman had released a 29 minute video essay on his Youtube Channel covering:
+
+1.  what makes an Antonoff song so Antonoff-y
+
+    > I'm literally not an expert and of course I don't actually claim to always be able to 100 identify a Jack Antonoff song it's not like I've been practicing the way that I can tell if he's produced something is basically if it sounds to me like he's produced it I know that's annoying I will clarify first I think it's mostly in the vocal treatment and that's the thing that is the most off-putting to me but just every decision uh feels a little bit wrong to me I don't know of course if I wanted to be cringe in CinemaSins I could point out the Juno 6 Bass the perfectly stacked vocals the Juno 6 pad the wide and tight plate Reverb the Juno 6 sword the goo goo gaga the man hates treble unless it's just random crap swishing around everything gets low passed with thin 707 drum samples thin 808 drum samples then 909 drum samples thin Lin drum samples thin thin drum samples all the resonances the harmonics get removed everything is sort of squished down to the fundamental reduction in Timbre everything exists in a small frequency range everything becomes a tone it's an extremely consonant sound with no interplay between the different elements now that is a skill to do and it produces a cleaner sound and the only problem is I viscerally hate that sound
+
+2.  his relationship to music growing up (you can skim this, but there are some gems in there)
+
+    > hing I got into drumming at the age oflike six so maybe as a drummer the drumsamples are of a fence I wasanachronistically raised on cassettes soyou know maybe I like the sound ofaudio and maybe that explains I tend tolike dissonance and in harmonic soundinterplay of elements Distortion noisethe Clashing of elements I imagine I'velistened to more SoundCloud in my lifethan radio so uh maybe I only want tohear amateur internet loser I'veliked over 10 000 songs on my Spotifyaccount uh and I can identify lot ofthem off the bat but I'm not that greatat identifying music generally so it'sjust uh compartmentalized mediaconsumption habit I first did musicproduction in primary school I mademashups and dubstep in it as a lateteenager I extensively used Juno 6emulators to make embarrassing synthwaveadmittedly that's probably not thesubconscious Association that he's goingfor lyrics aren't of much interest to meit's just a sound in the mix and somaybe I don't like the focus on vocalI've previously described it as a fightor flight sort of thing so maybe itliterally is tapping into some memory oftrauma I have an unusually small earcanal so maybe I'm just hearing the highend more acutely I was hired as ateenager to play drums on uh a lot ofJack Antonoff songs for a universityGlee Club performer so maybe I don'tlike it because of the embarrassment ofhaving ever been associated with a Gleecloth I liked the album 1989 but to makethe song out of the woods palette ofwool to me I had to mash it up with M83Midnight City and that got a dmcatakedown so maybe it's a personalVendetta and of course I have known anddo know people who like this music somaybe I secretly hate my friends andfamily but I don't know if any of thattells us anything uh I have a taste inmusic w
+
+3.  An elegant and simple defense of taste
+
+    > "i did use the phrase visceral hatred.
+but i did so only to describe my visceral hatred.
+i cant stand the sound of it.
+i hate it.
+and the way in which i hate it is viscerally.
+but i didnt go beyond that because it doesnt go beyond that.
+nothing against the guy, i like taylor swift's stuff.
+it's just subjective taste...
+
+4.  His response to the media
+
+    > .... \[it's just subjective taste\].
+there's just nothing to engage with, so ive had to fabricate a reason for controversy.
+and what ive come up with is this.
+\*looks at phone\* the real reason iive been saying for years that i have a visceral reaction to jack antonoffs production style, the REAL reason is this \-- cause during the rollout cycle of the album midnights, i wanted to ascribe taylor swifts authorship to a man and diminish women's voices.
+so actually, i'm a sexist.
+so now thats out there, publsihed as facts, in the news.
+with all the other facts..."
+
+5.  Basically a refutation to the entire premise of my interview
+
+    > "what makes you so special and unique that you can tell the production of a song?" "nothing"
+
+6.  Which was further double down on when I found a screenshot he took of his Tweet drafts, one of which being: "people often say to me:"caleb, what's wrong with jack antonoff?". don't ask questions"
+
+It is a perfect video essay. The only thing left I could ask was how he felt about that video 6 months after the fact, and what he thought of me asking him about it.
+
+# Numbers
+
+Taste is subjective. People who can play instruments tend to be better at deciphering and describing sounds. Antonoff is everywhere. Cultural criticism is a hall of smoke and mirrors.
+
+Can we translate any of this into something observable? Can be quantify the Jack Antonoff-ification of music?
+
 # Data {#sec-data}

### 2023-03-23 outputs/paper/paper.qmd
some editing

@@
 
-Around the time that Midnights came out, I came across a video on Twitter of a guy guessing which songs on Midnights were produced by Antonoff within seconds.
+Around the time that Midnights came out, I came across a video on Twitter of a guy, Caleb Gamman, guessing which songs on Midnights were produced by Antonoff within seconds.
 The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style".
@@
 
-6.  Which was further double down on when I found a screenshot he took of his Tweet drafts, one of which being: "people often say to me:"caleb, what's wrong with jack antonoff?". don't ask questions"
+6.
+Which was further doubled down on when I found a screenshot he took of his Tweet drafts in June 2021, one of them reading: "people often say to me:"caleb, what's wrong with jack antonoff?".
+don't ask questions"
 
-It is a perfect video essay. The only thing left I could ask was how he felt about that video 6 months after the fact, and what he thought of me asking him about it.
+It is a perfect video essay. The only thing left that I could ask was how he felt about that video 6 months after the fact, and what he thought of me asking him about it.
 
@@
 
-Taste is subjective. People who can play instruments tend to be better at deciphering and describing sounds. Antonoff is everywhere. Cultural criticism is a hall of smoke and mirrors.
+Taste is subjective. People who can play instruments tend to be better at identifying and describing sounds. Cultural criticism is a hall of smoke and mirrors. Antonoff is everywhere.
 
-Can we translate any of this into something observable? Can be quantify the Jack Antonoff-ification of music?
+Can we translate any of this into something observable? Can we quantify the Jack Antonoff-ification of pop music?
 

### 2023-03-24 scripts/02-data_cleaning.R
draft of regression model, some edits to qmd

@@ -185,7 +185,7 @@ df$mode <- as.numeric(df$mode)
 
 ###Keep only columns we want###
 
-df = subset(df, select = c("artist_name", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack"))
+df = subset(df, select = c("artist_name","album_release_date", "album_release_date_precision", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack"))
 
 #### Save data ####
 

### 2023-03-24 scripts/06-models.R

@@ -28,40 +28,37 @@ chisq
 
 
 # select the relevant columns
-cols <- c("danceability", "energy", "key", "loudness", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "jack")
-
-# create a new dataframe with only the selected columns
-df_small <- data[,cols]
-levels(df_small$jack)
-# Replace NAs with 0 in the jack column
-df_small$jack <- ifelse(is.na(df_small$jack), 0, df_small$jack)
+#cols <- c("danceability", "energy", "key", "loudness", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "jack")
+#jack ~ danceability + energy + loudness + speechiness + key + acousticness + instrumentalness + liveness + valence
 
 
+jack_model <-
+  glm(jack ~ danceability + energy + loudness + speechiness + key + acousticness + instrumentalness + liveness + valence,
+      data = df,
+      family = "binomial"
+  )
 
-# perform t-tests for continuous variables
-for (col in cols[!cols %in% c("artist_name", "track_name", "jack")]) {
-  ttest <- t.test(df_small[[col]] ~ df_small$jack)
-  print(paste("T-test for", col))
-  print(ttest)
-}
+library(marginaleffects)
 
-#SIGNIFICANT P VALUES: danceability, energy, loudness, acousticness, instrumentalness, liveness, valence
-#NOT SIGNIFCANT: Speechiness, Key
+jack_predictions <-
+  predictions(jack_model) |>
+  as_tibble()
 
+jack_predictions
 
-# perform chi-square tests for categorical variables
-for (col in cols[!cols %in% c("artist_name", "track_name", "jack")]) {
-  chisq <- chisq.test(table(df_small[[col]], df_small$jack))
-  print(paste("Chi-square test for", col))
-  print(chisq)
-}
-
-#SIGNIFICANT P VALUES: all of them??
-
-
-
-
-
-
-
+jack_predictions |>
+  ggplot(aes(
+    x = valence,
+    y = estimate,
+    color = jack
+  )) +
+  geom_jitter(width = 0.01, height = 0.01, alpha = 0.3) +
+  labs(
+    x = "Valence",
+    y = "Estimated probability that Jack produced them",
+    color = "Was it actually Jack"
+  ) +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
 

### 2023-03-24 07-valencetesting.R
valence testing

@@ -1,0 +1,13 @@
+
+
+library(spotifyr)
+holst <- get_artist_audio_features("Gustav Holst")
+nikolai <- get_artist_audio_features("Nikolai Rimsky-Korsakov")
+write_csv(holst, "inputs/data/holst.csv") 
+write_csv(nikolai, "inputs/data/nikolai.csv") 
+
+
+subset_planets <- subset(holst, track_id %in% c("08i3LqX0LijeLAiuJShjPm", "3OjbUkoryQX47hd0t9HyuK"))
+
+subset_nikolai <- nikolai[grepl("Bumblebee", nikolai$track_id, ignore.case = TRUE), ]
+

### 2023-03-24 outputs/paper/paper.qmd
draft of regression model, some edits to qmd
kinda started the data section for the valence testing

@@
 He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, Minions: The Rise of Gru Soundtrack, and more.
-All of this and his own bands too: Bleachers, and Fun.
+All of this and his own bands too: Bleachers, and fun.
 
@@
 happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g.
-sad, depressed, angry)", and danceability as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+sad, depressed, angry)".
+
+This is pretty vague. I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
+
+```{r}
+
+```
+
+and danceability as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
 A value of 0.0 is least danceable and 1.0 is most danceable".

### 2023-03-28 scripts/06-models.R
new bayes model

@@ -8,7 +8,7 @@
 #Going to make a smaller df that only had the columns we want to capture correlations
 
 library(ggplot2)
-
+library(tidyverse)
 data <- read_csv("inputs/data/df.csv")
 
 #chisquared test from website http://www.sthda.com/english/wiki/chi-square-test-of-independence-in-r
@@ -61,4 +61,33 @@ jack_predictions |>
   theme_classic() +
   scale_color_brewer(palette = "Set1") +
   theme(legend.position = "bottom")
+
+
+
+
+library(rstanarm)
+
+
+model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key + 
+                    acousticness + instrumentalness + liveness + valence,
+                  data = df,
+                  family = binomial(),
+                  prior_intercept = normal(0,10),
+                  prior = normal(0,2.5),
+                  chains = 4, iter = 2000)
+
+summary(model)
+
+#danceability has strongest relationship
+
+library(bayesplot)
+
+# Plot posterior distributions of coefficients
+mcmc_areas(model, regex_pars = "^(?!lp__)")
+
+# Plot posterior predictive checks
+pp_check(model)
+
+mcmc_areas(model, pars = grep("^lp__", colnames(model$summary), invert = TRUE, value = TRUE))
+
 

### 2023-03-30 outputs/paper/paper.qmd
adding writeup for valence testing

@@
 
-This is pretty vague. I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
+This is pretty vague.
+I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
+Jupiter's valence score is 326.78% higher than Mars.
+This not only indicates the valence is based on audio features other than lyrical content, but that it is capturing a difference between the two songs' profiles.
 
 ```{r}
+
+#| echo: false
+#| warning: false
+
+sub_planets<- read.csv(here::here("inputs/data/subset_planets.csv"))
+sub_planets |>
+  select(artist_name, valence, track_name) |>
+  head() |>
+  knitr::kable(
+    col.names = c("Artist Name", 
+                  "Valence", 
+                  "Track Name")
+ )
+
+```
+
+I looked at the valence measures on Flight of the Bumblebee, a song that's much faster and "upbeat" than Mars and Jupiter.
+What I found complicated my understanding of valence -- different performance of the same song had different valence scores.
+
+```{r}
+# sub_planets<- read.csv(here::here("inputs/data/subset_planets.csv"))
+# sub_planets |>
+#   select(artist_name, valence, track_name) |>
+#   head() |>
+#   knitr::kable(
+#     col.names = c("Artist Name", 
+#                   "Valence", 
+#                   "Track Name")
+#  )
 

### 2023-03-30 07-valencetesting.R

@@ -8,6 +8,9 @@ write_csv(nikolai, "inputs/data/nikolai.csv")
 
 
 subset_planets <- subset(holst, track_id %in% c("08i3LqX0LijeLAiuJShjPm", "3OjbUkoryQX47hd0t9HyuK"))
+write_csv(subset_planets, "inputs/data/subset_planets.csv")
+
+
 
 subset_nikolai <- nikolai[grepl("Bumblebee", nikolai$track_id, ignore.case = TRUE), ]
 

### 2023-03-30 outputs/paper/paper.qmd
wrote up the danceability of flight of the bumblebees

@@
 
-4.  His response to the media
+4.  His tongue in cheek response to the media
 
@@
 I looked at the valence measures on Flight of the Bumblebee, a song that's much faster and "upbeat" than Mars and Jupiter.
-What I found complicated my understanding of valence -- different performance of the same song had different valence scores.
+What I found complicated my understanding of valence -- different performances of the same song had different valence scores.
+Some were arias, played by different instruments.
+These tend to have higher valence scores.
+The lowest recorded valence is 0.2410, and the highest is .9230.
+What gives?
 
 ```{r}
-# sub_planets<- read.csv(here::here("inputs/data/subset_planets.csv"))
-# sub_planets |>
-#   select(artist_name, valence, track_name) |>
-#   head() |>
-#   knitr::kable(
-#     col.names = c("Artist Name", 
-#                   "Valence", 
-#                   "Track Name")
-#  )
+sub_bees<- read.csv(here::here("inputs/data/subset_bumblebee.csv"))
+sub_bees |>
+   select(artist_name, valence, album_release_year, track_name) |>
+   head(17) |>
+   knitr::kable(
+     col.names = c("Artist Name", 
+                   "Valence",
+                   "Album Release Year",
+                   "Track Name")
+  )
 

### 2023-03-30 07-valencetesting.R

@@ -12,5 +12,7 @@ write_csv(subset_planets, "inputs/data/subset_planets.csv")
 
 
 
-subset_nikolai <- nikolai[grepl("Bumblebee", nikolai$track_id, ignore.case = TRUE), ]
+subset_nikolai <- nikolai[grepl("Bumblebee", nikolai$track_name, ignore.case = TRUE), ]
+write_csv(subset_nikolai, "inputs/data/subset_bumblebee.csv")
 
+summary(subset_nikolai)

### 2023-03-30 outputs/paper/paper.qmd
more model goofing, added danceability txt to qmd

@@
 sad, depressed, angry)".
+"Danceability" is described as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+A value of 0.0 is least danceable and 1.0 is most danceable".
+Neither of these definitions fully cover what metrics and thresholds go into something like "danceability".
+However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums[^1].
+
+[^1]: \-- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
+The full documentation for each variable included in this study is available in the Appendix
 
 This is pretty vague.
-I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
+Especially valence -- is the positivity related to lyricsI compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
 Jupiter's valence score is 326.78% higher than Mars.
@@
 sub_planets |>
-  select(artist_name, valence, track_name) |>
+  select(artist_name, valence, danceability, track_name) |>
   head() |>
@@
                   "Valence", 
+                  "Danceability",
                   "Track Name")
@@
 sub_bees |>
-   select(artist_name, valence, album_release_year, track_name) |>
+   select(artist_name, valence, danceability, album_release_year, track_name) |>
    head(17) |>
@@
                    "Valence",
+                   "Danceability",
                    "Album Release Year",

### 2023-03-30 scripts/06-models.R

@@ -78,16 +78,50 @@ model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key +
 
 summary(model)
 
+
 #danceability has strongest relationship
+# Load necessary packages
+library(pROC)
+library(ggroc)
+library(ggplot2)
 
-library(bayesplot)
+# Predict probability of Jack Antonoff producing the song for each observation
+probabilities <- predict(model, type = "response")
 
-# Plot posterior distributions of coefficients
-mcmc_areas(model, regex_pars = "^(?!lp__)")
+df_prob <- cbind(df, probabilities)
 
-# Plot posterior predictive checks
-pp_check(model)
+# Create ROC curve
+roc_curve <- roc(jack ~ probabilities, data = df_prob)
 
-mcmc_areas(model, pars = grep("^lp__", colnames(model$summary), invert = TRUE, value = TRUE))
+# Convert ROC curve to data frame
+roc_df <- fortify(roc_curve)
+
+# calculate auc 
+auc <- auc(roc_curve)
+
+# Plot ROC curve
+ggplot(roc_curve, aes(x = 1 - specificity, y = sensitivity)) + 
+  geom_line() + 
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
+  ggtitle(paste0("ROC Curve (AUC = ", round(auc, 2), ")")) + 
+  xlab("False Positive Rate") + 
+  ylab("True Positive Rate")
+
+# Create confusion matrix
+predictions <- ifelse(probabilities >= 0.5, 1, 0)
+confusion_matrix <- table(predictions, jack)
+
+# Print confusion matrix
+confusion_matrix
+
+# Calculate precision, recall, and F1-score
+precision <- confusion_matrix[2,2] / sum(confusion_matrix[,2])
+recall <- confusion_matrix[2,2] / sum(confusion_matrix[2,])
+f1_score <- 2 * precision * recall / (precision + recall)
+
+# Print precision, recall, and F1-score
+cat("Precision:", round(precision, 2), "\n")
+cat("Recall:", round(recall, 2), "\n")
+cat("F1-score:", round(f1_score, 2))
 
 

### 2023-03-30 scripts/06-models.R
bad model. bad!

@@ -9,7 +9,7 @@
 
 library(ggplot2)
 library(tidyverse)
-data <- read_csv("inputs/data/df.csv")
+df <- read_csv("inputs/data/df.csv")
 
 #chisquared test from website http://www.sthda.com/english/wiki/chi-square-test-of-independence-in-r
 #not including liveness because its negative values
@@ -80,48 +80,27 @@ summary(model)
 
 
 #danceability has strongest relationship
-# Load necessary packages
-library(pROC)
-library(ggroc)
 library(ggplot2)
+library(pROC)
 
-# Predict probability of Jack Antonoff producing the song for each observation
 probabilities <- predict(model, type = "response")
+# Create probabilities column in the original data frame
+df$probabilities <- probabilities
 
-df_prob <- cbind(df, probabilities)
+### plot an ROC curve?
 
-# Create ROC curve
-roc_curve <- roc(jack ~ probabilities, data = df_prob)
 
-# Convert ROC curve to data frame
-roc_df <- fortify(roc_curve)
-
-# calculate auc 
-auc <- auc(roc_curve)
-
-# Plot ROC curve
-ggplot(roc_curve, aes(x = 1 - specificity, y = sensitivity)) + 
-  geom_line() + 
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
-  ggtitle(paste0("ROC Curve (AUC = ", round(auc, 2), ")")) + 
-  xlab("False Positive Rate") + 
-  ylab("True Positive Rate")
 
 # Create confusion matrix
-predictions <- ifelse(probabilities >= 0.5, 1, 0)
-confusion_matrix <- table(predictions, jack)
+library(caret)
+predictions <- ifelse(probabilities >= 0.5, "Yes", "No")
+df$predictions <- predictions
+df$predictions <- factor(predictions, levels = c("Yes", "No"))
+df$jack <- factor(df$jack, levels = c(1, 0), labels = c("Yes", "No"))
+confusion_matrix <- confusionMatrix(df$predictions, df$jack)
+
 
 # Print confusion matrix
 confusion_matrix
-
-# Calculate precision, recall, and F1-score
-precision <- confusion_matrix[2,2] / sum(confusion_matrix[,2])
-recall <- confusion_matrix[2,2] / sum(confusion_matrix[2,])
-f1_score <- 2 * precision * recall / (precision + recall)
-
-# Print precision, recall, and F1-score
[…]

### 2023-03-30 scripts/06-models.R
LOL WAIT THE SECOND MODEL WHERE I ACTUALLY DO TRAIN TEST SPLITS IS EVEN WORSE

@@ -11,57 +11,11 @@ library(ggplot2)
 library(tidyverse)
 df <- read_csv("inputs/data/df.csv")
 
-#chisquared test from website http://www.sthda.com/english/wiki/chi-square-test-of-independence-in-r
-#not including liveness because its negative values
-df_chisq <- df[, c("artist_name", "track_name", "danceability",                
-                   "energy", "key","loudness", "speechiness", "acousticness",                
-                   "instrumentalness", "valence", "jack")]
-
-chisq <- chisq.test(df_chisq)
-chisq
-
-
-
-
-
-
-
-
-# select the relevant columns
-#cols <- c("danceability", "energy", "key", "loudness", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "jack")
-#jack ~ danceability + energy + loudness + speechiness + key + acousticness + instrumentalness + liveness + valence
-
-
 jack_model <-
   glm(jack ~ danceability + energy + loudness + speechiness + key + acousticness + instrumentalness + liveness + valence,
       data = df,
       family = "binomial"
   )
-
-library(marginaleffects)
-
-jack_predictions <-
-  predictions(jack_model) |>
-  as_tibble()
-
-jack_predictions
-
-jack_predictions |>
-  ggplot(aes(
-    x = valence,
-    y = estimate,
-    color = jack
-  )) +
-  geom_jitter(width = 0.01, height = 0.01, alpha = 0.3) +
-  labs(
-    x = "Valence",
-    y = "Estimated probability that Jack produced them",
-    color = "Was it actually Jack"
-  ) +
-  theme_classic() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
-
 
 
 
@@ -103,4 +57,54 @@ confusion_matrix <- confusionMatrix(df$predictions, df$jack)
 # Print confusion matrix
 confusion_matrix
 confusion_matrix$byClass
+
+
+
+##FK forgot to split the data lol
+
+library(caret)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Split data into training and testing sets
+train_index <- createDataPartition(df$jack, p = 0.7, list = FALSE)
+train_data <- df[train_index, ]
+test_data <- df[-train_index, ]
+
+
+library(rstanarm)
+
+# Fit the model using only the training data
+model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key + 
+                    acousticness + instrumentalness + liveness + valence,
[…]

### 2023-03-30 scripts/02-data_cleaning.R
new lana tracks. also i guess new bleachers tracks?

@@ -183,9 +183,6 @@ df$key <- as.numeric(df$key)
 df$mode <- as.numeric(df$mode)
 
 
-###Keep only columns we want###
-
-df = subset(df, select = c("artist_name","album_release_date", "album_release_date_precision", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack"))
 
 #### Save data ####
 

### 2023-03-30 scripts/06-models.R

@@ -94,10 +94,10 @@ test_data$jack <- factor(test_data$jack, levels = c(1, 0), labels = c("Yes", "No
 
 # Create a confusion matrix
 confusion_matrix2 <- table(test_data$predictions, test_data$jack)
-confusion_matrix <- as.matrix(confusion_matrix)
+confusion_matrix2 <- as.matrix(confusion_matrix2)
 
 # Print the confusion matrix
-confusion_matrix
+confusion_matrix2
 precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])
 recall <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])
 f1_score <- 2 * precision * recall / (precision + recall)

### 2023-03-31 scripts/06-models.R
rebalanced the classes and nothing mf changed

@@ -11,58 +11,9 @@ library(ggplot2)
 library(tidyverse)
 df <- read_csv("inputs/data/df.csv")
 
-jack_model <-
-  glm(jack ~ danceability + energy + loudness + speechiness + key + acousticness + instrumentalness + liveness + valence,
-      data = df,
-      family = "binomial"
-  )
 
-
-
+library(caret)
 library(rstanarm)
-
-
-model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key + 
-                    acousticness + instrumentalness + liveness + valence,
-                  data = df,
-                  family = binomial(),
-                  prior_intercept = normal(0,10),
-                  prior = normal(0,2.5),
-                  chains = 4, iter = 2000)
-
-summary(model)
-
-
-#danceability has strongest relationship
-library(ggplot2)
-library(pROC)
-
-probabilities <- predict(model, type = "response")
-# Create probabilities column in the original data frame
-df$probabilities <- probabilities
-
-### plot an ROC curve?
-
-
-
-# Create confusion matrix
-library(caret)
-predictions <- ifelse(probabilities >= 0.5, "Yes", "No")
-df$predictions <- predictions
-df$predictions <- factor(predictions, levels = c("Yes", "No"))
-df$jack <- factor(df$jack, levels = c(1, 0), labels = c("Yes", "No"))
-confusion_matrix <- confusionMatrix(df$predictions, df$jack)
-
-
-# Print confusion matrix
-confusion_matrix
-confusion_matrix$byClass
-
-
-
-##FK forgot to split the data lol
-
-library(caret)
 
 # Set seed for reproducibility
 set.seed(123)
@@ -72,29 +23,45 @@ train_index <- createDataPartition(df$jack, p = 0.7, list = FALSE)
 train_data <- df[train_index, ]
 test_data <- df[-train_index, ]
 
+#balance the classes by undersampling
+train_data_balanced <- train_data |>
+  group_by(jack) |>
+  sample_n(size = min(table(jack))) |>
+  ungroup()
 
-library(rstanarm)
 
 # Fit the model using only the training data
 model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key + 
                     acousticness + instrumentalness + liveness + valence,
-                  data = train_data,
+                  data = train_data_balanced,
                   family = binomial(),
                   prior_intercept = normal(0,10),
[…]

### 2023-04-01 outputs/paper/paper.qmd
tidying up

@@
 Sometime in early October of 2022 there was a chemical shift in my brain and I couldn't stop listening to Taylor Swift.
-I was a preteen when my best friend's sister showed us Our Song on the family desktop, went on to own a copy of Fearless, and was in the ninth grade when Red came out.
-My listening tapered off after that, other than ironically dancing to Blank Space in friends kitchens in college, and a road trip a few years ago where we belted Fearless (Three of the four people present were going through something).
-
-All this to say, I can quite confidently say I hadn't listened to much or any Taylor Swift until October, when I couldn't listen to anything other than Fearless or Red when deadlifting, walking to a friend's house, folding laundry, etc.
+I couldn't listen to anything other than Fearless or Red when deadlifting, walking to a friend's house, folding laundry, etc.
 I listened to so much Taylor in that month alone that she ended up as my #4 artist on Spotify Wrapped, which stops collecting its data for the year on October 31st.

### 2023-04-01 scripts/06-models.R

@@ -30,7 +30,7 @@ train_data_balanced <- train_data |>
   ungroup()
 
 
-# Fit the model using only the training data
+
 model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key + 
                     acousticness + instrumentalness + liveness + valence,
                   data = train_data_balanced,
@@ -47,7 +47,7 @@ model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key +
 test_data$predictions <- ifelse(predict(model, newdata = test_data, type = "response") > 0.5, "Yes", "No")
 
 
-# Convert factors to the same set of levels
+# convert factors to the same set of levels
 test_data$jack <- ifelse(test_data$jack == 1, "Yes", "No")
 test_data$jack <- factor(test_data$jack, levels = c("Yes", "No"))
 test_data$predictions <- factor(test_data$predictions, levels = c("Yes", "No"))
@@ -59,11 +59,7 @@ print(confusion_matrix)
 confusion_matrix$byClass
 
 
-# Create a confusion matrix
-#confusion_matrix2 <- table(test_data$predictions, test_data$jack)
-#confusion_matrix2 <- as.matrix(confusion_matrix2)
 
-# Print the confusion matrix
 confusion_matrix2
 precision <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])
 recall <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])

### 2023-04-03 07-understandingthevariables.R
more goofing. decision tree. EDA on the variables

@@ -1,0 +1,44 @@
+mydata <- read_csv("inputs/data/df.csv")
+#cleaning lol
+mydata <- data.frame(sapply(mydata, as.numeric))
+mydata <- mydata[, c("jack", "danceability", "energy", "loudness", "speechiness", "key", "acousticness", "instrumentalness", "liveness", "valence")]
+
+# Verify the new structure of your dataset
+str(mydata)
+
+
+# Examine the distribution of a variable
+hist(mydata$danceability, main = "Distribution of Danceability", xlab = "Danceability Score")
+
+# Calculate the mean, median, and standard deviation of a variable
+mean(mydata$danceability)
+median(mydata$danceability)
+sd(mydata$danceability)
+
+# Calculate basic summary statistics for all variables in the dataset
+summary(mydata)
+
+
+
+# Create a scatterplot of two variables
+plot(mydata$danceability, mydata$energy, main = "Danceability vs. Energy", xlab = "Danceability", ylab = "Energy")
+
+
+
+
+
+
+
+
+library(corrplot)
+
+# Calculate the correlation matrix
+cor_matrix <- cor(mydata)
+
+# Create the correlation plot
+corrplot(cor_matrix, method = "circle", type = "lower")
+cor(mydata)
+
+
+
+

### 2023-04-03 scripts/06-models.R

@@ -71,3 +71,34 @@ f1_score
 #... do these features just not tell us anything about the song?
 
 
+##Decision Tree
+
+library(rpart)
+library(rpart.plot)
+
+# Split your data into training and testing sets
+set.seed(123) # for reproducibility
+train_index <- sample(nrow(df), 0.7*nrow(df))
+train_data <- df[train_index,]
+test_data <- df[-train_index,]
+
+# Build the decision tree
+mytree <- rpart(jack ~ danceability + energy + loudness + speechiness + key +
+                  acousticness + instrumentalness + liveness + valence, 
+                data = train_data, 
+                method = "class")
+
+# Plot the decision tree
+rpart.plot(mytree)
+
+# Make predictions on the test set
+predictions <- predict(mytree, test_data, type = "class")
+
+# Calculate the accuracy of the model
+accuracy <- sum(predictions == test_data$jack)/nrow(test_data)
+print(paste("Accuracy:", round(accuracy, 3)))
+
+
+
+
+

### 2023-04-04 outputs/paper/paper.qmd
cleaning it up a bit

@@
 ---
-title: "My title"
+title: "Jack Antonoff-ification of pop music"
 subtitle: "My subtitle if needed"
@@
 thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+date: today
+abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API.
+Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an "Antonoff-y" song, and assess whether or not the audio features defined by Spotify can capture this.
+I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me.
+Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
 format: html
@@
 
-\["*No*?" I thought to myself\]
+\["*No*?"\]
 
@@
 
-Midnight was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this essay, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".
+Midnight was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".
 
 Antonoff is prolific, to the extent that it is starting to bother people.
-He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, Minions: The Rise of Gru Soundtrack, and more.
-All of this and his own bands too: Bleachers, and fun.
-
+He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more.
+All of this and his own bands too (Bleachers, and fun.).
 When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed.
@@
 
-What I am trying to get to here is that I don't really mind Jack Antonoff's music.
-I kind of like it actually.
-But I still say his voice with the same tone that everyone else does.
+What I am trying to get to here is that I don't actually really mind Jack Antonoff's music, I kind of like it, but I still say his voice with the same tone that everyone else does.
 There's just something about him...
@@
 and that he's just such a nice guy...
-Why are we all so fixated on him?
+Why do music people seem to be a little annoyed with him?
 What is it about Antonoff music that is so Antonoff-y?
+Is the Jack Antonoff-fication of pop something we should be worried about?
 
-To find an answer to this question, I turned people, statistics, and the overlord of engineered virality itself -- the Spotify API.
+To find an answer to this question, I turned to statistics, people who know more than me about describing music, and the overlord of engineered virality itself -- the Spotify API.
 
@@
 
+(you can skim the quote chunks)
+
 1.  what makes an Antonoff song so Antonoff-y
@@
 
-2.  his relationship to music growing up (you can skim this, but there are some gems in there)
+2.  his relationship to music growing up (some gems in here)
 
@@
 
-4.  His tongue in cheek response to the media
+4.  His tongue in cheek response to some of Twitter's negative response to his video
 
@@
 so actually, i'm a sexist.
-so now thats out there, publsihed as facts, in the news.
+so now thats out there, publishhed as facts, in the news.
 with all the other facts..."
@@
 6.
-Which was further doubled down on when I found a screenshot he took of his Tweet drafts in June 2021, one of them reading: "people often say to me:"caleb, what's wrong with jack antonoff?".
+Which was refuted even harder when I found a screenshot he tweeted of his tweet drafts in June 2021, one of them reading: "people often say to me:"caleb, what's wrong with jack antonoff?".
 don't ask questions"
@@
 It is a perfect video essay. The only thing left that I could ask was how he felt about that video 6 months after the fact, and what he thought of me asking him about it.
+
+\[\[He ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]
 
@@
 This is pretty vague.
-Especially valence -- is the positivity related to lyricsI compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
+Especially valence -- is the positivity related to lyrics, or something else?
+I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
 Jupiter's valence score is 326.78% higher than Mars.
-This not only indicates the valence is based on audio features other than lyrical content, but that it is capturing a difference between the two songs' profiles.
+This not only indicates the valence is based on audio features other than lyrical content, but that it is in fact capturing a difference between the two songs' profiles.
 
@@
 
-and danceability as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
-A value of 0.0 is least danceable and 1.0 is most danceable".
-Neither of these definitions fully cover what metrics and thresholds go into something like "danceability".
-However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums -- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
-The full documentation for each variable included in this study is available in the Appendix.
-
 The final audio features in this study data set contains the variables: "artist_name", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack".
@@
 $$ {#eq-bayes}
-
-@eq-bayes seems useful, eh?
 

### 2023-04-04 outputs/paper/paper.qmd
pushin

@@
 
-\[I blacked out.\]
+\[I glitched.\]
 
@@
 When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed.
-What my friend was referring to, of course, is the fact that Jack Antonoff seems to be everywhere these days.
+What my friend was referring to, of course, is the fact that Jack Antonoff and his sound seems to be everywhere these days.
 As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely.
@@
 
-\[\[He ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]
+\[\[He ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]\]
 
@@
 
+My model is driving me bananas.
+
 # Results

### 2023-04-04 scripts/06-models.R

@@ -15,10 +15,9 @@ df <- read_csv("inputs/data/df.csv")
 library(caret)
 library(rstanarm)
 
-# Set seed for reproducibility
+
 set.seed(123)
 
-# Split data into training and testing sets
 train_index <- createDataPartition(df$jack, p = 0.7, list = FALSE)
 train_data <- df[train_index, ]
 test_data <- df[-train_index, ]
@@ -72,6 +71,7 @@ f1_score
 
 
 ##Decision Tree
+## Rohan dunked on me for this
 
 library(rpart)
 library(rpart.plot)
@@ -82,19 +82,17 @@ train_index <- sample(nrow(df), 0.7*nrow(df))
 train_data <- df[train_index,]
 test_data <- df[-train_index,]
 
-# Build the decision tree
 mytree <- rpart(jack ~ danceability + energy + loudness + speechiness + key +
                   acousticness + instrumentalness + liveness + valence, 
                 data = train_data, 
                 method = "class")
 
-# Plot the decision tree
+
 rpart.plot(mytree)
 
-# Make predictions on the test set
+
 predictions <- predict(mytree, test_data, type = "class")
 
-# Calculate the accuracy of the model
 accuracy <- sum(predictions == test_data$jack)/nrow(test_data)
 print(paste("Accuracy:", round(accuracy, 3)))
 

### 2023-04-04 outputs/paper/paper.qmd
more qmd edits

@@
 abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API.
-Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an "Antonoff-y" song, and assess whether or not the audio features defined by Spotify can capture this.
+Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an Antonoff-y song, and assess whether or not the audio features defined by Spotify can capture this.
 I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me.
+There will be a Bayesian inference model.
+It will be a logistic regression.
+We do not know whether or not it will be good.
 Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
-format: html
+format: pdf
 number-sections: true
@@
 ---
-
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-library(tidyverse)
-library(palmerpenguins)
-```
-
-```{r}
-
-#quarto references itself. wrapping in here:here goes in project 
-knitr::include_graphics(here::here("inputs/data/apod_20230227.jpg"))
-
-```
-
-```{r}
-#| label: fig-energy
-#| fig-cap: Energy Variable
-#| echo: false
-
-library(lubridate)
-df <- readRDS(here::here("inputs/data/df.rds"))
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
-  ggplot(aes(y = energy, x = album_release_date, color = jack)) +
-  geom_point() +
-  theme_minimal() +
-  scale_color_manual(values = c("#FFB6C1", "#808080")) +
-  labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3)
-
-```
-
-```{r}
-#| label: fig-danceability
-#| fig-cap: Danceability Variable
-#| echo: false
-
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
-  ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
-  geom_point() +
-  theme_minimal() +
-  scale_color_manual(values = c("#FFB6C1", "#808080")) +
-  labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3)
-
-```
-
-```{r}
-#| label: fig-valence
-#| fig-cap: Valence Variable
-#| echo: false
-
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
-  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
-  geom_point() +
-  theme_minimal() +
-  scale_color_manual(values = c("#FFB6C1", "#808080")) +
-  labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3)
-
-```
-
-```{r}
-#| label: fig-acousticness
-#| fig-cap: Acousticness Variable
-#| echo: false
-
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
-  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
-  geom_point() +
-  theme_minimal() +
-  scale_color_manual(values = c("#FFB6C1", "#808080")) +
-  labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3)
-
-```
-
-```{r}
-#| label: fig-instrumentalness
-#| fig-cap: Instrumentalness Variable
-#| echo: false
-
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
-  ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
-  geom_point() +
-  theme_minimal() +
-  scale_color_manual(values = c("#FFB6C1", "#808080")) +
-  labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3)
-
-```
-
-```{r}
-#| label: fig-tempo
-#| fig-cap: Tempo Variable
-#| echo: false
-
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
-  ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
-  geom_point() +
-  theme_minimal() +
-  scale_color_manual(values = c("#FFB6C1", "#808080")) +
-  labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3)
-
-```
 
@@
 To find an answer to this question, I turned to statistics, people who know more than me about describing music, and the overlord of engineered virality itself -- the Spotify API.
+There will be a Bayesian inference model.
+It will be a logistic regression.
+We do not know whether or not it will be good.
 
@@
 
+![](images/Screen%20Shot%202023-04-03%20at%209.31.05%20PM.png)
+
 It is a perfect video essay. The only thing left that I could ask was how he felt about that video 6 months after the fact, and what he thought of me asking him about it.
 
-\[\[He ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]\]
+\[\[He told me that he didn't actually refute the premise of my interview, and ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]\]
 
@@
 
-Taste is subjective. People who can play instruments tend to be better at identifying and describing sounds. Cultural criticism is a hall of smoke and mirrors. Antonoff is everywhere.
+TLDR of the Gamman interview is: Taste is subjective.
+People who can play instruments tend to be better at identifying and describing sounds.
+Cultural criticism is a hall of smoke and mirrors.
+Antonoff is everywhere.
 
@@
 #| warning: false
+library(dplyr)
 
@@
 ```{r}
-library(kableExtra)
-df |>
-  knitr::kable()
+#| echo: false
+#| warning: false
+
+#library(kableExtra)
+#df |>
+ # knitr::kable()
 
@@
 My model is driving me bananas.
+
+# "EDA"
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+library(tidyverse)
+library(palmerpenguins)
+```
+
+```{r}
+#| label: fig-energy
+#| fig-cap: Energy Variable
+#| echo: false
+# 
+# library(lubridate)
+# df <- read.csv(here::here("inputs/data/df.csv"))
+# df %>%
+#   filter(album_release_date_precision == "day") %>%
+#   mutate(album_release_date = ymd(album_release_date)) %>%
+#   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
+#   geom_point() +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
+#   labs(color = "Has 'y' in jack?") +
+#   facet_wrap(~ artist_name, ncol = 3)
+
+```
+
+```{r}
+#| label: fig-danceability
+#| fig-cap: Danceability Variable
+#| echo: false
+
+# df %>%
+#   filter(album_release_date_precision == "day") %>%
+#   mutate(album_release_date = ymd(album_release_date)) %>%
+#   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
+#   geom_point() +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
+#   labs(color = "Has 'y' in jack?") +
+#   facet_wrap(~ artist_name, ncol = 3)
+
+```
+
+```{r}
+#| label: fig-valence
+#| fig-cap: Valence Variable
+#| echo: false
+
+# df %>%
+#   filter(album_release_date_precision == "day") %>%
+#   mutate(album_release_date = ymd(album_release_date)) %>%
+#   ggplot(aes(y = valence, x = album_release_date, color = jack)) +
+#   geom_point() +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
+#   labs(color = "Has 'y' in jack?") +
+#   facet_wrap(~ artist_name, ncol = 3)
+
+```
+
+```{r}
+#| label: fig-acousticness
+#| fig-cap: Acousticness Variable
+#| echo: false
+
+# df %>%
+#   filter(album_release_date_precision == "day") %>%
+#   mutate(album_release_date = ymd(album_release_date)) %>%
+#   ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
+#   geom_point() +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
+#   labs(color = "Has 'y' in jack?") +
+#   facet_wrap(~ artist_name, ncol = 3)
+
+```
+
+```{r}
+#| label: fig-instrumentalness
+#| fig-cap: Instrumentalness Variable
+#| echo: false
+
+# 
+# df %>%
+#   filter(album_release_date_precision == "day") %>%
+#   mutate(album_release_date = ymd(album_release_date)) %>%
+#   ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
+#   geom_point() +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
+#   labs(color = "Has 'y' in jack?") +
+#   facet_wrap(~ artist_name, ncol = 3)
+
+```
+
+```{r}
+#| label: fig-tempo
+#| fig-cap: Tempo Variable
+#| echo: false
+
+# 
+# df %>%
+#   filter(album_release_date_precision == "day") %>%
+#   mutate(album_release_date = ymd(album_release_date)) %>%
+#   ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
+#   geom_point() +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
+#   labs(color = "Has 'y' in jack?") +
+#   facet_wrap(~ artist_name, ncol = 3)
+
+```
 

### 2023-04-04 outputs/paper/paper.qmd
got the graphs to render

@@
 #| echo: false
-# 
-# library(lubridate)
-# df <- read.csv(here::here("inputs/data/df.csv"))
-# df %>%
-#   filter(album_release_date_precision == "day") %>%
-#   mutate(album_release_date = ymd(album_release_date)) %>%
-#   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
-#   geom_point() +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
-#   labs(color = "Has 'y' in jack?") +
-#   facet_wrap(~ artist_name, ncol = 3)
+
+library(lubridate)
+df <- read.csv(here::here("inputs/data/df.csv"))
+df$jack <- factor(df$jack)
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = energy, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3)
 
@@
 
-# df %>%
-#   filter(album_release_date_precision == "day") %>%
-#   mutate(album_release_date = ymd(album_release_date)) %>%
-#   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
-#   geom_point() +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
-#   labs(color = "Has 'y' in jack?") +
-#   facet_wrap(~ artist_name, ncol = 3)
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3)
 
@@
 
-# df %>%
-#   filter(album_release_date_precision == "day") %>%
-#   mutate(album_release_date = ymd(album_release_date)) %>%
-#   ggplot(aes(y = valence, x = album_release_date, color = jack)) +
-#   geom_point() +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
-#   labs(color = "Has 'y' in jack?") +
-#   facet_wrap(~ artist_name, ncol = 3)
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3)
 
@@
 
-# df %>%
-#   filter(album_release_date_precision == "day") %>%
-#   mutate(album_release_date = ymd(album_release_date)) %>%
-#   ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
-#   geom_point() +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
-#   labs(color = "Has 'y' in jack?") +
-#   facet_wrap(~ artist_name, ncol = 3)
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3)
 
@@
 
-# 
-# df %>%
-#   filter(album_release_date_precision == "day") %>%
-#   mutate(album_release_date = ymd(album_release_date)) %>%
-#   ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
-#   geom_point() +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
-#   labs(color = "Has 'y' in jack?") +
-#   facet_wrap(~ artist_name, ncol = 3)
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3)
 
@@
 ```{r}
-#| label: fig-tempo
+
+#| label: fig-temp
 #| fig-cap: Tempo Variable
@@
 
-# 
-# df %>%
-#   filter(album_release_date_precision == "day") %>%
-#   mutate(album_release_date = ymd(album_release_date)) %>%
-#   ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
-#   geom_point() +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#FFB6C1", "#808080")) +
-#   labs(color = "Has 'y' in jack?") +
-#   facet_wrap(~ artist_name, ncol = 3)
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3)
 

### 2023-04-04 scripts/08-understandingthevariables.R

@@ -1,3 +1,8 @@
+library(dplyr)
+library(lubridate)
+library(ggplot2)
+
+
 mydata <- read_csv("inputs/data/df.csv")
 #cleaning lol
 mydata <- data.frame(sapply(mydata, as.numeric))
@@ -22,6 +27,89 @@ summary(mydata)
 
 # Create a scatterplot of two variables
 plot(mydata$danceability, mydata$energy, main = "Danceability vs. Energy", xlab = "Danceability", ylab = "Energy")
+
+
+
+##Plotting all the variables##
+df <- read.csv(here::here("inputs/data/df.csv"))
+df$jack <- factor(df$jack)
+
+library(lubridate)
+
+#Energy
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
+   geom_point() +
+   theme_minimal() +
+   scale_color_manual(values = c("#FFB6C1", "#808080")) +
+   labs(color = "Has 'y' in jack?") +
+   facet_wrap(~ artist_name, ncol = 3)
+
+
+#Danceability
+df %>%
+   filter(album_release_date_precision == "day") %>%
+   mutate(album_release_date = ymd(album_release_date)) %>%
+   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
+   geom_point() +
+   theme_minimal() +
+   scale_color_manual(values = c("#FFB6C1", "#808080")) +
+   labs(color = "Has 'y' in jack?") +
+   facet_wrap(~ artist_name, ncol = 3)
+
+
+
+#Valence
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
+  geom_point() +
+  theme_minimal() +
+  scale_color_manual(values = c("#FFB6C1", "#808080")) +
+  labs(color = "Has 'y' in jack?") +
+  facet_wrap(~ artist_name, ncol = 3)
+
+
+#Acousticness
+df %>%
+  filter(album_release_date_precision == "day") %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
+  geom_point() +
[…]

### 2023-04-04 outputs/paper/paper.qmd
removed the warnings from the render

@@
 #| echo: false
+#| warning: false
 

### 2023-04-11 outputs/paper/paper.qmd
added quotes from gamman interview, made a quadrant plot of energy/valence

@@
 
-It is a perfect video essay. The only thing left that I could ask was how he felt about that video 6 months after the fact, and what he thought of me asking him about it.
+It is a perfect video essay. I pestered him anyway.
 
-\[\[He told me that he didn't actually refute the premise of my interview, and ended up saying a lot of interesting things, which i will transcribe and somehow make use of in this paper later\]\]
+> "In the treble mix of the song, like in the very upper end, like sort of above where the vocals are.
+He often really crushes that down.
+Like it's lower in volume than anything else, which is sort of a weird effect.
+Like, when you hear someone speaking, you hear a lot of that, like, noise in their voice.
+When you yourself are speaking, you hear less of it, right?
+You hear more of your own voice bouncing around in your head.
+And so it sort of creates this effect of -- **like he mixes his vocals the way it would sound to the person who's singing them.
+Which is sort of strange.** He often has no vocals on that upper end, which is very unusual, and then often there are random little bits of noise happening up there which is the sort of thing her likes to do."
+
+Gamman is a fan of messier music.
+
+"Jack Antonoff does a very good job of making stuff very clean.
+So you know, all of the notes that are happening simulaneously are perfectly stacked.
+They're like mathematically stacked ..
+Like, you have the harmonic spectrum where everything is like a multiple on top of each other.
+So that's like your notes on a keyboard.
+And if you played a keyboard note that was a little bit off where it's supposed to be, you're gonna create this interesting effect where these two notes sort of clash against each other and they wiggle around in terms of their tone because you've got these two harmonics happening that don't match.
+And Jack Antonoff avoids that quite strictly"
+
+"His drums are low past, like the vocals, so you don't get like the tinny high end in the drums or whatever because that tinny high end in the drums is like a bunch of random harmonics all over the place.
+So if you cut that out, you just get the tone of the drum.
+And then he achieves those effects using the same sounds like literally the same synthesizer on the same drum machine and stuff."
+
+"If you used a cscope spectrogram sort of thing to look at a Jack Antonoff song, you're gonna see a lot of big lines.
+And then if you're listening to a song by someone else, you're gonna see like a bunch of sort of like static wiggling.
+It's gonna be really jagged because you've got a bunch of harmonics that don't exactly mix with each other, so they're stacking in all sorts of weird different ways"
 

### 2023-04-11 scripts/08-understandingthevariables.R

@@ -1,11 +1,12 @@
 library(dplyr)
 library(lubridate)
 library(ggplot2)
-
+library(tidyverse)
 
 mydata <- read_csv("inputs/data/df.csv")
 #cleaning lol
-mydata <- data.frame(sapply(mydata, as.numeric))
+#next line of code gets rid of the string variables
+#mydata <- data.frame(sapply(mydata, as.numeric))
 mydata <- mydata[, c("jack", "danceability", "energy", "loudness", "speechiness", "key", "acousticness", "instrumentalness", "liveness", "valence")]
 
 # Verify the new structure of your dataset
@@ -126,6 +127,32 @@ cor_matrix <- cor(mydata)
 # Create the correlation plot
 corrplot(cor_matrix, method = "circle", type = "lower")
 cor(mydata)
+
+
+
+library(ggplot2)
+
+# create a scatterplot with energy on the x-axis and valence on the y-axis
+ggplot(data = mydata, aes(x = energy, y = valence)) + 
+  
+  # add points to the scatterplot and color them by artist_name
+  geom_point(aes(color = artist_name)) +
+  
+  # add vertical and horizontal lines to separate the quadrants
+  geom_vline(xintercept = mean(mydata$energy), color = "gray") +
+  geom_hline(yintercept = mean(mydata$valence), color = "gray") +
+  
+  # add labels to each quadrant
+  annotate("text", x = mean(mydata$energy), y = max(mydata$valence), 
+           label = "High Energy, High Valence") +
+  annotate("text", x = mean(mydata$energy), y = min(mydata$valence), 
+           label = "High Energy, Low Valence") +
+  annotate("text", x = min(mydata$energy), y = mean(mydata$valence), 
+           label = "Low Energy, High Valence") +
+  annotate("text", x = max(mydata$energy), y = mean(mydata$valence), 
+           label = "Low Energy, Low Valence") +
+  scale_x_continuous(limits = c(min(mydata$energy), max(mydata$energy))) +
+  scale_y_continuous(limits = c(min(mydata$valence), max(mydata$valence)))
 
 
 

### 2023-04-11 outputs/paper/paper.qmd
changed title, plotted danceability/energy quadrant

@@
 ---
-title: "Jack Antonoff-ification of pop music"
-subtitle: "My subtitle if needed"
+title: "We Are All Trying to Describe Something And We Don't Have the Words"
+subtitle: "an exploration of the Jack Antonoffification of Pop Music"
 author: 

### 2023-04-11 scripts/08-understandingthevariables.R

@@ -136,7 +136,7 @@ library(ggplot2)
 ggplot(data = mydata, aes(x = energy, y = valence)) + 
   
   # add points to the scatterplot and color them by artist_name
-  geom_point(aes(color = artist_name)) +
+  geom_point(aes(color = as.factor(jack))) +
   
   # add vertical and horizontal lines to separate the quadrants
   geom_vline(xintercept = mean(mydata$energy), color = "gray") +
@@ -150,10 +150,51 @@ ggplot(data = mydata, aes(x = energy, y = valence)) +
   annotate("text", x = min(mydata$energy), y = mean(mydata$valence), 
            label = "Low Energy, High Valence") +
   annotate("text", x = max(mydata$energy), y = mean(mydata$valence), 
-           label = "Low Energy, Low Valence") +
-  scale_x_continuous(limits = c(min(mydata$energy), max(mydata$energy))) +
-  scale_y_continuous(limits = c(min(mydata$valence), max(mydata$valence)))
+           label = "Low Energy, Low Valence") 
 
 
+
+
+# create a scatterplot with energy on the x-axis and valence on the y-axis
+ggplot(data = mydata, aes(x = energy, y = danceability)) + 
+  
+  # add points to the scatterplot and color them by artist_name
+  geom_point(aes(color = artist_name)) +
+  
+  # add vertical and horizontal lines to separate the quadrants
+  geom_vline(xintercept = mean(mydata$energy), color = "gray") +
+  geom_hline(yintercept = mean(mydata$danceability), color = "gray") +
+  
+  # add labels to each quadrant
+  annotate("text", x = mean(mydata$energy), y = max(mydata$danceability), 
+           label = "High Energy, High Danceability") +
+  annotate("text", x = mean(mydata$energy), y = min(mydata$danceability), 
+           label = "High Energy, Low Danceability") +
+  annotate("text", x = min(mydata$energy), y = mean(mydata$danceability), 
+           label = "Low Energy, High Danceability") +
+  annotate("text", x = max(mydata$energy), y = mean(mydata$danceability), 
+           label = "Low Energy, Low Danceability") 
+
+
+
+# create a scatterplot with energy on the x-axis and valence on the y-axis
+ggplot(data = mydata, aes(x = energy, y = danceability)) + 
+  
+  # add points to the scatterplot and color them by artist_name
+  geom_point(aes(color = as.factor(jack))) +
+  
+  # add vertical and horizontal lines to separate the quadrants
[…]

### 2023-04-11 outputs/paper/paper.qmd
added more plots to the PDF

@@
 
+I'm gonna fix the axis labels later
+
+```{r}
+
+#| label: fig-quadvalence1
+#| fig-cap: Valence and Energy Quadrant 
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = valence)) + 
+  geom_point(aes(color = as.factor(jack))) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$valence), color = "gray") +
+  annotate("text", x = mean(df$energy), y = max(df$valence), 
+           label = "High Energy, High Valence") +
+  annotate("text", x = mean(df$energy), y = min(df$valence), 
+           label = "High Energy, Low Valence") +
+  annotate("text", x = min(df$energy), y = mean(df$valence), 
+           label = "Low Energy, High Valence") +
+  annotate("text", x = max(df$energy), y = mean(df$valence), 
+           label = "Low Energy, Low Valence")
+
+```
+
+```{r}
+
+#| label: fig-quadvalence2
+#| fig-cap: Valence and Energy Quadrant
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = valence)) + 
+  geom_point(aes(color = artist_name)) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$valence), color = "gray") +
+  annotate("text", x = mean(df$energy), y = max(df$valence), 
+           label = "High Energy, High Valence") +
+  annotate("text", x = mean(df$energy), y = min(df$valence), 
+           label = "High Energy, Low Valence") +
+  annotate("text", x = min(df$energy), y = mean(df$valence), 
+           label = "Low Energy, High Valence") +
+  annotate("text", x = max(df$energy), y = mean(df$valence), 
+           label = "Low Energy, Low Valence") 
+```
+
+```{r}
+
+#| label: fig-quaddanceability
+#| fig-cap: Danceability and Energy Quadrant
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = danceability)) + 
+  geom_point(aes(color = artist_name)) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$danceability), color = "gray") +
+  annotate("text", x = mean(df$energy), y = max(df$danceability), 
+           label = "High Energy, High Danceability") +
+  annotate("text", x = mean(df$energy), y = min(df$danceability), 
+           label = "High Energy, Low Danceability") +
+  annotate("text", x = min(df$energy), y = mean(df$danceability), 
+           label = "Low Energy, High Danceability") +
+  annotate("text", x = max(df$energy), y = mean(df$danceability), 
+           label = "Low Energy, Low Danceability")
+
+```
+
+```{r}
+
+#| label: fig-quaddanceability2
+#| fig-cap: Danceability and Energy Quadrant
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = danceability)) + 
+  geom_point(aes(color = as.factor(jack))) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$danceability), color = "gray") +
+  annotate("text", x = mean(df$energy), y = max(df$danceability), 
+           label = "High Energy, High Danceability") +
+  annotate("text", x = mean(df$energy), y = min(df$danceability), 
+           label = "High Energy, Low Danceability") +
+  annotate("text", x = min(df$energy), y = mean(df$danceability), 
+           label = "Low Energy, High Danceability") +
+  annotate("text", x = max(df$energy), y = mean(df$danceability), 
+           label = "Low Energy, Low Danceability")
+
+```
+
 # Results

### 2023-04-11 outputs/paper/paper.qmd
added some glen parts

@@
 ```
+
+I asked Glen McDonald, a "data alchemist" (see: Principal Engineer) who worked at The Echo Nest, a machine listening start-up that was acquired by Spotify in \[YEAR\]. He still works at Spotify now.
+
+Glen told me that Valence and Danceability we created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
+Variables like energy or instrumentalness, on the other hand, were determined through machine listening techniques.
+The human subjectivity part only came in when it came to tweaking the training data.
+(My favourite example Glen gave was that the ML treated banjos like a human voice, giving bluegrass songs with banjos high Speechiness ratings.
+Because there were no banjos in the training data, they were registered as human).
+
+Glen said that the way he's made use of these variables is by combing Energy and Valence to get a quadrant, which usually turns out "pretty good".
+
+\[discuss the convo we had about how even two humans can have disagreement about a song's positivity, let alone human vs machine.
+i.e.
+an elliot smith song where the lyrics are depressing but the melody is happy.
+or a happy song but you know that the band died in a plane crash yesterday.
+etc.\]
 

### 2023-04-12 outputs/paper/paper.qmd
new artists, new df

--- 2023-04-11 outputs/paper/paper.qmd
@@
 title: "We Are All Trying to Describe Something And We Don't Have the Words"
-subtitle: "an exploration of the Jack Antonoffification of Pop Music"
+subtitle: "An exploration of the Jack Antonoffification of Pop Music"
 author: 
@@
 
-Midnight was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".
+Midnights was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".
 
@@
 
+I asked Glen McDonald, a "data alchemist" (see: Principal Engineer) who worked at The Echo Nest, a machine listening start-up that was acquired by Spotify in \[YEAR\]. He still works at Spotify now.
+
+Glen told me that Valence and Danceability we created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
+Variables like energy or instrumentalness, on the other hand, were determined through machine listening techniques.
+The human subjectivity part only came in when it came to tweaking the training data.
+(My favourite example Glen gave was that the ML treated banjos like a human voice, giving bluegrass songs with banjos high Speechiness ratings.
+Because there were no banjos in the training data, they were registered as human).
+
+Glen said that the way he's made use of these variables is by combing Energy and Valence to get a quadrant, which usually turns out "pretty good".
+
+\[discuss the convo we had about how even two humans can have disagreement about a song's positivity, let alone human vs machine.
+i.e.
+an elliot smith song where the lyrics are depressing but the melody is happy.
+or a happy song but you know that the band died in a plane crash yesterday.
+etc.\]
+
 The final audio features in this study data set contains the variables: "artist_name", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack".
@@
 library(lubridate)
-df <- read.csv(here::here("inputs/data/df.csv"))
+library(tidyverse)
+df <- read_csv(here::here("inputs/data/df.csv"))
 df$jack <- factor(df$jack)
@@
   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
-  geom_point() +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
   theme_minimal() +
-  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
   labs(color = "Has 'y' in jack?") +
-  facet_wrap(~ artist_name, ncol = 3)
+  facet_wrap( ~ artist_name, ncol = 3)
 
@@
   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
-  geom_point() +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
   theme_minimal() +
@@
   ggplot(aes(y = valence, x = album_release_date, color = jack)) +
-  geom_point() +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
   theme_minimal() +
@@
   ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
-  geom_point() +
+    geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
   theme_minimal() +
@@
   ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
-  geom_point() +
+    geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
   theme_minimal() +
@@
   ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
-  geom_point() +
+    geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
   theme_minimal() +
@@
   facet_wrap(~ artist_name, ncol = 3)
+
+```
+
+I'm gonna fix the axis labels later
+
+```{r}
+
+#| label: fig-quadvalence1
+#| fig-cap: Valence and Energy Quadrant 
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = valence)) + 
+  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$valence), color = "gray") +
+  annotate("text", x = mean(df$energy), y = max(df$valence), 
+           label = "High Energy, High Valence") +
+  annotate("text", x = mean(df$energy), y = min(df$valence), 
+           label = "High Energy, Low Valence") +
+  annotate("text", x = min(df$energy), y = mean(df$valence), 
+           label = "Low Energy, High Valence") +
+  annotate("text", x = max(df$energy), y = mean(df$valence), 
+           label = "Low Energy, Low Valence")
+
+```
+
+```{r}
+
+#| label: fig-quadvalence2
+#| fig-cap: Valence and Energy Quadrant
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = valence)) + 
+  geom_point(aes(color = artist_name), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$valence), color = "gray") +
+  annotate("text", x = mean(df$energy), y = max(df$valence), 
+           label = "High Energy, High Valence") +
+  annotate("text", x = mean(df$energy), y = min(df$valence), 
+           label = "High Energy, Low Valence") +
+  annotate("text", x = min(df$energy), y = mean(df$valence), 
+           label = "Low Energy, High Valence") +
+  annotate("text", x = max(df$energy), y = mean(df$valence), 
+           label = "Low Energy, Low Valence") 
+```
+
+```{r}
+
+#| label: fig-quaddanceability
+#| fig-cap: Danceability and Energy Quadrant
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = danceability)) + 
+  geom_point(aes(color = artist_name), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$danceability), color = "gray") +
+  annotate("text", x = mean(df$energy), y = max(df$danceability), 
+           label = "High Energy, High Danceability") +
+  annotate("text", x = mean(df$energy), y = min(df$danceability), 
+           label = "High Energy, Low Danceability") +
+  annotate("text", x = min(df$energy), y = mean(df$danceability), 
+           label = "Low Energy, High Danceability") +
+  annotate("text", x = max(df$energy), y = mean(df$danceability), 
+           label = "Low Energy, Low Danceability")
+
+```
+
+```{r}
+
+#| label: fig-quaddanceability2
+#| fig-cap: Danceability and Energy Quadrant
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = danceability)) + 
+  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$danceability), color = "gray") +
+  annotate("text", x = mean(df$energy), y = max(df$danceability), 
+           label = "High Energy, High Danceability") +
+  annotate("text", x = mean(df$energy), y = min(df$danceability), 
+           label = "High Energy, Low Danceability") +
+  annotate("text", x = min(df$energy), y = mean(df$danceability), 
+           label = "Low Energy, High Danceability") +
+  annotate("text", x = max(df$energy), y = mean(df$danceability), 
+           label = "Low Energy, Low Danceability")
 

### 2023-04-12 scripts/01-download_data_spotify.R

@@ -24,6 +24,11 @@ chicksaf <- get_artist_audio_features("the chicks")
 lordeaf <- get_artist_audio_features("lorde")
 floaf <- get_artist_audio_features("florence and the machine")
 bleachersaf <- get_artist_audio_features("bleachers")
+haimaf <- get_artist_audio_features("Haim")
+maggierogersaf <- get_artist_audio_features("Maggie Rogers")
+sharonaf <- get_artist_audio_features("Sharon Von Etten")
+marinaaf <- get_artist_audio_features("Marina and the Diamonds")
+mitskiaf <- get_artist_audio_features("Mitski")
 
 
 #### Save data audio features ####
@@ -34,8 +39,11 @@ write_csv(vincentaf, "inputs/data/vincentaf.csv")
 write_csv(chicksaf, "inputs/data/chicksaf.csv") 
 write_csv(lordeaf, "inputs/data/lordeaf.csv") 
 write_csv(floaf, "inputs/data/floaf.csv") 
-write_csv(bleachersaf, "inputs/data/bleachersaf.csv") 
-
+write_csv(haimaf, "inputs/data/haimaf.csv") 
+write_csv(maggierogersaf, "inputs/data/maggierogersaf.csv") 
+write_csv(sharonaf, "inputs/data/sharonaf.csv") 
+write_csv(marinaaf, "inputs/data/marinaaf.csv") 
+write_csv(mitskiaf, "inputs/data/mitskiaf.csv") 
 #### Download data audio analysis####
 
 

### 2023-04-12 scripts/02-data_cleaning.R

@@ -18,10 +18,17 @@ vincentaf <- read_csv("inputs/data/vincentaf.csv")
 lanaaf <- read_csv("inputs/data/lanaaf.csv")
 tayloraf <- read_csv("inputs/data/tayloraf.csv")
 bleachersaf <- read_csv("inputs/data/bleachersaf.csv")
+haimaf <- read_csv("inputs/data/haimaf.csv")
+maggierogersaf <- read_csv("inputs/data/maggierogersaf.csv")
+sharonaf <- read_csv("inputs/data/sharonaf.csv")
+marinaaf <- read_csv("inputs/data/marinaaf.csv")
+mitskiaf <- read_csv("inputs/data/mitskiaf.csv")
+
 #### Clean data ####
 # Binding Data
 
-df <- rbind(lanaaf, tayloraf, vincentaf, lordeaf, chicksaf, floaf, bleachersaf)
+df <- rbind(lanaaf, tayloraf, vincentaf, lordeaf, chicksaf, floaf, bleachersaf,
+            haimaf, maggierogersaf, sharonaf, marinaaf, mitskiaf)
 df$jack = NA
 
 jack_tracks <- c(

### 2023-04-12 scripts/06-models.R

@@ -10,7 +10,7 @@
 library(ggplot2)
 library(tidyverse)
 df <- read_csv("inputs/data/df.csv")
-
+df$jack <- as.factor(df$jack)
 
 library(caret)
 library(rstanarm)
@@ -37,6 +37,8 @@ model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key +
                   prior_intercept = normal(0,10),
                   prior = normal(0,2.5),
                   chains = 4, iter = 2000)
+
+
 
 # Make predictions on the testing data
 #probabilities <- predict(model, newdata = test_data, type = "response")

### 2023-04-14 scripts/09-rohan_messing_around.R
more eda

@@ -41,6 +41,24 @@ data_reduced |>
   )
 
 
+data_reduced %>%
+  group_by(jack) %>%
+  summarise(across(everything(), mean)) %>%
+  pivot_longer(cols = -jack) %>%
+  ggplot(aes(value, name, fill = factor(jack))) +
+  geom_col(alpha = 0.8, position = "dodge")
+
+
+data_reduced %>%
+  group_by(jack) %>%
+  summarise(across(-c(tempo, loudness), mean)) %>%
+  pivot_longer(cols = -jack) %>%
+  ggplot(aes(value, name, fill = factor(jack))) +
+  geom_col(alpha = 0.8, position = "dodge")
+
+
+
+?across
 #### Try a model ####
 # Code from: https://juliasilge.com/blog/project-feederwatch/
 
@@ -86,6 +104,7 @@ narrower_penalty <- penalty(range = c(-3, 0))
 
 # install.packages("doParallel")
 doParallel::registerDoParallel()
+#install.packages("glmnet")
 
 set.seed(853)
 

### 2023-04-14 scripts/09-rohan_messing_around.R
added notes 2 self

@@ -91,6 +91,8 @@ glmnet_spec <-
 # install.packages("themis")
 library(themis)
 
+
+#so this is comparing the basic model and the downsampled one
 wf_set <-
   workflow_set(
     list(basic = is_jack,
@@ -108,6 +110,19 @@ doParallel::registerDoParallel()
 
 set.seed(853)
 
+#first u pass in ur wf set, then u pass which function,
+#then u pass in things u need for the tuning. like resamples,
+# values for penalty (grid=16), pass in metrics
+#accuracy and roc_auc are basic metrics
+# add logloss, which looks at model as a whole, measures how wrong u r.
+# helps measure that u do a better job. i.e. .6 and .9 would
+#still be in positive class, but it tells you how wrong you are
+#sensitivity, specifity: measure how well u do on one class versus the other
+# (important for unbalanced dataset)
+#passing params. knows from exp that default penalty is really broad -10, 0.
+#make it much much narrower
+
+
 tune_rs <- 
   workflow_map(
     wf_set,
@@ -122,13 +137,50 @@ tune_rs
 
 autoplot(tune_rs) + theme(legend.position = "none")
 
+#x axis is workflow rank, so which workflow are we on
+#ok so we have good accuracy and good specificity. but low
+#low specificity and low log loss
+#dont quite understand workflow rank?
+#theres no right metric to evaluate model. you have to understand
+# it holistically and look at a bunch of different metrics
+#sanity check with rohan
+
 rank_results(tune_rs, rank_metric = "sensitivity")
+rank_results(tune_rs, rank_metric = "mn_log_loss")
+rank_results(tune_rs, rank_metric = "accuracy")
+rank_results(tune_rs, rank_metric = "specificity")
+
+#the only one where downsampling is ranked higher is specificity...
+#wouldn't it make more sense to go with the basic one then?
 
 downsample_rs <-
   tune_rs %>%
   extract_workflow_set_result("downsampling_glmnet")
 
-autoplot(downsample_rs)
+#tuning result for recipe that included downsampling + glmnet model
+plotdownsample <- autoplot(downsample_rs) + ggtitle("Downsampled Data")
+
+
+basic_rs <-
+  tune_rs %>%
+  extract_workflow_set_result("basic_glmnet")
+
+plotbasic <- autoplot(basic_rs) + ggtitle("Original Data")
+
+library(gridExtra)
+grid.arrange(plotdownsample, plotbasic, ncol=2)
[…]

### 2023-04-17 outputs/paper/paper.qmd
edits edits edits

@@
 title: "We Are All Trying to Describe Something And We Don't Have the Words"
-subtitle: "An exploration of the Jack Antonoffification of Pop Music"
+subtitle: "The Jack Antonoffification of Pop Music?"
 author: 
@@
 
-#### Audio Analysis Dataset
-
-The `get_track_audio_analysis` offers a more granular look at each track.
-The variables included are arrays of objects, which are the kinds of things that are a bit harder to wrap your mind around if you were to visualize each one across each artist.
-They include: track, bars, beats, sections, segments, tatums, and meta, which includes information on which platforms and Analyzer versions were used to analyze the track, along with metadata on the status and duration of the analysis.
-
-This data will be used when building a logistic regression model to determine the character and predictability of Jack Antonoff's music.
-
-# Model
-
-$$
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
-$$ {#eq-bayes}
-
-My model is driving me bananas.
-
 # "EDA"
@@
 
+Sure. These are fun quadrants.But it doesn't solve the Jack-iness underlying each of those points, which are scattered pretty broadly.
+
+Let's throw a logistic regression at the problem.
+
+# Model
+
+Since we're trying to figure out whether these variables can tell us anything about what Jack Antonoff's sound is all about, a good way to figure that out is by seeing if any of these variables can help us predict what songs are produced by Jack.
+Logistic regressions like questions with binary outcomes.
+
+This cannot be the right formula. $$
+Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)} = \frac{1}{1 + e^{-(\beta_0 + \beta_1 x_1 + \cdots + \beta_p x_p)}}
+$$ {#eq-bayes}
+
+This model estimates the probability P of the outcome variable Y using predictor variables Xc.
+In other words, it estimates the probability that a song was produced by Jack Anotnoff using the Spotfiy variables.
+Which gives us information about the "research" "question": are these characteristics related to Jack's music?
+
+The chart below gives us an idea of the mean value of each of these features for Jack and non-Jack song. A logistic regression will give us a better idea of how to weigh each of these features.
+
+```{r}
+#|: echo: false
+#|: warning: false
+
+data_reduced %>%
+  group_by(jack) %>%
+  summarise(across(-c(tempo, loudness), mean)) %>%
+  pivot_longer(cols = -jack) %>%
+  ggplot(aes(value, name, fill = factor(jack))) +
+  geom_col(alpha = 0.8, position = "dodge") +
+  labs(x = "Mean Value", y = "Feature")
+
+```
+
+One issue is that in our dataset, there are a lot more songs that aren't produced by Jack than songs that are:
+
+```{r}
+#|: echo: false
+#|: warning: false
+
+data_reduced %>%
+  count(jack) %>%
+  knitr::kable(col.names = c("Is it Jack?", "Number of Songs"))
+
+```
+
+So, we have to balance the classes.
+
+Following Julia Silge's tutorial \[TK\], I used the tidymodels metapackage to split the training and testing data and make cross-validation samples.
+The `workflow_set()` function lets us hold different sets of feature engineering together and use tuning to evaluate them all at once.
+I need to evaluate two elements: a basic model, and a model that was downsampled using the `themis` package.
+
 # Results
@@
 
-## First discussion point {#sec-first-point}
+## Jack-iness and What Danceability Tells Us {#sec-first-point}
 
-## Second discussion point
+Using the Spotify Audio Features, you can't really gauge what Jack Antonoff's music sound like.
+You get a rough idea, but you don't get the same granularity that you would by simply DM'ing Caleb Gamman on Twitter and asking him to describe it you.
+\[ add gamman quote\]
+
+According to this model, though, danceability is the most important feature in predicting whether or not a song was produced by Jack Antonoff.
+This is so hilarious.
+This was a feature that was invented in a lab 15 years ago.
+This number refers to the time a bunch of undergraduates were paid (hopefully?) to listen to a bunch of songs and tag whether they were danceable or not.
+It captures this je-ne-sais-quoi essence of music that connects with listeners on a visceral level.
+Even if it doesn't make YOUR body want to move, you could imagine it making SOMEBODY'S body want to move.
+
+But does Jack Antonoff's music even make your body want to move? Let's put a pin in that.
+
+Eliciting a visceral reaction doesn't cause career success -- factors in a career trajectory like network, connections, right time right place ism and other elements of biography all shape that sort of thing.
 
@@
 ## Weaknesses and next steps
+
+It only makes sense, of course, to hire a bunch of undergrads and get them to tell us whether or not a few thousand songs are danceable, and then build our own danceability metric, and see if we get the same results -- that danceability is the characteristic that helps us best understand what makes Jack, Jack.
+
+A major weakness is cooked into the dataset.
+Since we're not in the studio, we can't know the degree of Jack-iness in each individual song.
+Especially when he takes on the role of a more backseat co-producer, as he did with Lana Del Rey's Did You Know There's a Tunnel Under Ocean Blvd?
+\[tk\].
 

### 2023-04-17 outputs/paper/paper.qmd
digging deeper into danceability variable

@@
 
-This cannot be the right formula. $$
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)} = \frac{1}{1 + e^{-(\beta_0 + \beta_1 x_1 + \cdots + \beta_p x_p)}}
-$$ {#eq-bayes}
+\[add the formula\]
 
@@
 #|: warning: false
+
+data_reduced <- read.csv(here::here("inputs/data/data_reduced.csv"))
 
@@
 
+0.743 accuracy, 0.841 roc_auc. Interesting. Check it out:
+
+```{r}
+#jackmatrix
+```
+
+```{r}
+
+data_reduced |> 
+  pivot_longer(cols = c(danceability, energy, loudness, mode,
+               speechiness, acousticness, instrumentalness, liveness, valence,
+               tempo),
+               names_to = "name",
+               values_to = "value"
+  ) |> 
+  ggplot(aes(x = value, color = jack)) +
+  stat_ecdf() +
+  facet_wrap(vars(name), 
+             nrow = 3, 
+             ncol = 4,
+             scales = "free") +
+  theme_minimal() +
+  labs(
+    x = "Value",
+    y = "Proportion"
+  )
+
+```
+
+And here are the different variables' importance on determining whether a song was had any Antonoff involvement.
+
+```{r}
+#| echo: false
+#| warning: false
+
+jack_vip <- read.csv(here::here("inputs/data/jack_vip.csv"))
+
+jack_vip |>
+   knitr::kable(
+     col.names = c("Variable", 
+                   "Importance",
+                   "Sign")
+  )
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+
+jack_vip %>%
+  group_by(Sign) %>%
+  slice_max(Importance, n = 15) %>%
+  ungroup() %>%
+  ggplot(aes(Importance, fct_reorder(Variable, Importance), fill = Sign)) + 
+  geom_col() +
+  facet_wrap(vars(Sign), scales = "free_y") +
+  labs(y = NULL) +
+  theme(legend.position = "none")
+```
+
+Danceability, with a score of 4.32, had the highest level of importance in determining whether or not a song was produced or co-produced by Jack Antonoff.
+
 # Discussion
@@
 This number refers to the time a bunch of undergraduates were paid (hopefully?) to listen to a bunch of songs and tag whether they were danceable or not.
+It is pretty much based on intuition.
 It captures this je-ne-sais-quoi essence of music that connects with listeners on a visceral level.
@@
 
-But does Jack Antonoff's music even make your body want to move? Let's put a pin in that.
+```{r}
+#| echo: false
+#| warning: false
 
-Eliciting a visceral reaction doesn't cause career success -- factors in a career trajectory like network, connections, right time right place ism and other elements of biography all shape that sort of thing.
+library(tidyverse)
+
+df %>%
+  filter(jack == 1) %>%
+  arrange(desc(danceability)) |>
+  distinct(track_name, .keep_all = TRUE) |>
+  select(artist_name, danceability, track_name) |>
+  knitr::kable(
+     col.names = c("Artist Name", 
+                   "Danceability",
+                   "Track")
+  )
+
+```
+
+### What do I think:
+
+-   I Think He Knows by Taylor Swift:
+
+    -   pretty danceable tbh
+
+-   Vigilante Shit by Taylor Swift:
+
+    -   Not danceable. But it is trying to coerce you into some sort of body feeling.
+
+-   Heaven is Here by Florence and the Machine
+
+    -   DEFINITELY not danceable. But it is trying to get your body to do something. Kind of sounds like it's suggesting that you join a Macedonian phalanx or something. But no. Not danceable.
+
+    ![](images/Macefonian_phalanx.png)
+
+-   I Wanna Get Better by Bleachers
+
+    -   I don't know if "danceability" is the word, but I'll give it "anthem"-y.
+Like, I can't picture a group of friends dancing to this in the kitchen, but I can picture someone listening to it in their car when it comes on the radio on their way to give a presentation or negotiate a promotion or something.
+
+-   Cornelia Street by Taylor Swift
+
+    -   Yeah I'm dancing. Or at least bobbing my head.
+
+```{r}
+#| echo: false
+#| warning: false
+library(dplyr)
+library(ggplot2)
+
+df %>%
+  filter(jack == 1) %>%
+  count(artist_name) %>%
+  arrange(desc(n)) %>%
+  head(10) %>%
+  ggplot(aes(x = reorder(artist_name, n), y = n)) +
+  geom_bar(stat = "identity", fill = "steelblue") +
+  coord_flip() +
+  labs(title = "Artists with Most Jack = 1 Values", x = "Artist", y = "Count")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+
+df |>
+  arrange(desc(danceability)) |>
+  select(artist_name, danceability, track_name)
+
+```
 

### 2023-04-17 scripts/09-rohan_messing_around.R

@@ -21,6 +21,8 @@ data_reduced <-
          tempo
          ) 
 
+write_csv(data_reduced,"inputs/data/data_reduced.csv")
+
 data_reduced |> 
   pivot_longer(cols = c(danceability, energy, loudness, mode,
                speechiness, acousticness, instrumentalness, liveness, valence,
@@ -202,8 +204,14 @@ final_fit
 collect_metrics(final_fit)
 
 
-collect_predictions(final_fit) %>%
+jackmatrix <- collect_predictions(final_fit) %>%
   conf_mat(jack, .pred_class)
+
+jackmatrix
+
+jackmatrix_df <- as.data.frame(jackmatrix$table)
+write.table(jackmatrix_df, file = "path/to/folder/jackmatrix.txt", sep = "\t")
+
 
 # install.packages("vip")
 library(vip)
@@ -212,6 +220,7 @@ jack_vip <-
   vi()
 
 jack_vip
+
 
 jack_vip %>%
   group_by(Sign) %>%

### 2023-04-17 outputs/paper/paper.qmd
visceral

@@
 
-## Jack-iness and What Danceability Tells Us {#sec-first-point}
+## Danceability Predicts Jack-iness {#sec-first-point}
 
@@
 #| warning: false
-
-library(tidyverse)
 
@@
 
+## But Danceability isn't even about Dancing! I propose: Visceral-ness
+
+I am thinking about what Gamman said in his video.
+
+> I did use the phrase "visceral hatred".
+But i did so only to describe my visceral hatred.
+I can't stand the sound of it.
+I hate it.
+And the way in which I hate it is viscerally.
+But I didn't go beyond that because it doesn't go beyond that.
+Nothing against the guy, I like Taylor swift's stuff.
+It's just subjective taste.
+
+It's just his subjective taste, but in one way Gamman was actually totally on the money about this.
+The danceability variable, which is the strongest predictor of Jack-ness, seems to actually be more about "visceral"-ness when you do a close-listen of the songs with the highest danceability scores.
+The reason Jack is getting under people's skin could just be because his music...
+gets underneath your skin.
+Right to the viscera.
+
+## Weaknesses and next steps
+
+One fundamental weakness is cooked into the dataset.
+Since we're not in the studio, we can't know the degree of Jack-iness in each individual song.
+Especially when he takes on the role of a more backseat co-producer, as he did with Lana Del Rey's Did You Know There's a Tunnel Under Ocean Blvd?
+\[tk\].
+
+#### Could our model just be predicting Jack via Taylor Swift?
+
+```{r}
+#| echo: false
+#| warning: false
+df %>%
+  filter(jack == 1) %>%
+  ggplot(aes(x = artist_name, y = danceability)) +
+  geom_boxplot() +
+  ggtitle("Boxplot of Danceability by Artist") +
+  coord_flip()
+```
+
 ```{r}
@@
 
-```{r}
-#| echo: false
-#| warning: false
+#### Next Steps
 
-df |>
-  arrange(desc(danceability)) |>
-  select(artist_name, danceability, track_name)
-
-```
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-It only makes sense, of course, to hire a bunch of undergrads and get them to tell us whether or not a few thousand songs are danceable, and then build our own danceability metric, and see if we get the same results -- that danceability is the characteristic that helps us best understand what makes Jack, Jack.
-
-A major weakness is cooked into the dataset.
-Since we're not in the studio, we can't know the degree of Jack-iness in each individual song.
-Especially when he takes on the role of a more backseat co-producer, as he did with Lana Del Rey's Did You Know There's a Tunnel Under Ocean Blvd?
-\[tk\].
+It only makes sense, of course, to hire a bunch of undergrads and get them to tell us whether or not a few thousand songs are visceral, and then build our own visceral-ness metric, and see if we get the same results -- that danceability is just visceral-ness in a trench coat, and that visceral-ness is the characteristic that helps us best understand what makes Jack, Jack.
 

### 2023-04-18 outputs/paper/paper.qmd
writing her up

@@
   - Michaela Drouillard
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/michaeladrouillard/jack."
 date: today
@@
 
-    > I'm literally not an expert and of course I don't actually claim to always be able to 100 identify a Jack Antonoff song it's not like I've been practicing the way that I can tell if he's produced something is basically if it sounds to me like he's produced it I know that's annoying I will clarify first I think it's mostly in the vocal treatment and that's the thing that is the most off-putting to me but just every decision uh feels a little bit wrong to me I don't know of course if I wanted to be cringe in CinemaSins I could point out the Juno 6 Bass the perfectly stacked vocals the Juno 6 pad the wide and tight plate Reverb the Juno 6 sword the goo goo gaga the man hates treble unless it's just random crap swishing around everything gets low passed with thin 707 drum samples thin 808 drum samples then 909 drum samples thin Lin drum samples thin thin drum samples all the resonances the harmonics get removed everything is sort of squished down to the fundamental reduction in Timbre everything exists in a small frequency range everything becomes a tone it's an extremely consonant sound with no interplay between the different elements now that is a skill to do and it produces a cleaner sound and the only problem is I viscerally hate that sound
+    > I'm literally not an expert and of course I don't actually claim to always be able to be able to 100% identify a Jack Antonoff song.
+It's not like I've been practicing the way that I can tell if he's produced something.
+It's just basically if it sounds to me like he's produced it.
+I know that's annoying I will clarify first, I think it's mostly in the vocal treatment, that's the thing that is the most off-putting to me.
+But just every decision feels a little bit wrong to me.
+If I wanted to be cringe I could point out the Juno 6 Bass, the perfectly stacked vocals, the Juno 6 pad, the wide and tight plate reverb, the Juno 6 sword, the goo goo gaga, the man hates treble unless it's just random crap swishing around, everything gets low passed with thin 707 drum samples, thin 808 drum samples, then 909 drum samples, thin Lin drum samples, thin thin drum samples, all the resonances, the harmonics get removed, everything is sort of squished down to the fundamental reduction in timbre, everything exists in a small frequency range, everything becomes a tone, it's an extremely consonant sound with no interplay between the different elements.
+Now that is a skill to do, and it produces a cleaner sound, and the only problem is I viscerally hate that sound.
 
@@
 
-It is a perfect video essay. I pestered him anyway.
+It is a perfect video essay. I pestered him anyway:
 
-> "In the treble mix of the song, like in the very upper end, like sort of above where the vocals are.
+> **Michaela Drouillard:** I know you cover this in the video. I'm sorry. What makes a Jack Antonoff song a Jack Antonoff song?
+>
+> **Caleb Gamman:** I suppose I do have much more detail that I can go into, but it is always, for me, just a knee jerk.
+It's not any principled thing.
+I do have some loose experience with like, audio production stuff.
+So if I really dig into it, I can maybe point some things out.
+In the treble mix of the song, like in the very upper end, like sort of above where the vocals are.
 He often really crushes that down.
@@
 And so it sort of creates this effect of -- **like he mixes his vocals the way it would sound to the person who's singing them.
-Which is sort of strange.** He often has no vocals on that upper end, which is very unusual, and then often there are random little bits of noise happening up there which is the sort of thing her likes to do."
-
-Gamman is a fan of messier music.
-
-"Jack Antonoff does a very good job of making stuff very clean.
+Which is sort of strange.** He often has no vocals on that upper end, which is very unusual, and then often there are random little bits of noise happening up there which is the sort of thing her likes to do.
+>
+> So I don't know, like I personally am quite a fan of messier music, you know, more experimental stuff, stuff that has a lot of clashing, sounds, a lot of different interesting pieces coming together.
+Jack Antonoff does a very good job of making stuff very clean.
 So you know, all of the notes that are happening simulaneously are perfectly stacked.
@@
 And if you played a keyboard note that was a little bit off where it's supposed to be, you're gonna create this interesting effect where these two notes sort of clash against each other and they wiggle around in terms of their tone because you've got these two harmonics happening that don't match.
-And Jack Antonoff avoids that quite strictly"
+And Jack Antonoff avoids that quite strictly.
+>
+> His drums are low pass, like the vocals, so you don't get like the tinny high end in the drums or whatever because that tinny high end in the drums is like a bunch of random harmonics all over the place.
+So if you cut that out, you just get the tone of the drum.
+And then he achieves those effects using the same sounds like literally the same synthesizer on the same drum machine and stuff.
+>
+> ....
+>
+> If you used a cscope spectrogram sort of thing to look at a Jack Antonoff song, you're gonna see a lot of big lines.
+And then if you're listening to a song by someone else, you're gonna see like a bunch of sort of like static wiggling.
+It's gonna be really jagged because you've got a bunch of harmonics that don't exactly mix with each other, so they're stacking in all sorts of weird different ways.
 
-"His drums are low past, like the vocals, so you don't get like the tinny high end in the drums or whatever because that tinny high end in the drums is like a bunch of random harmonics all over the place.
-So if you cut that out, you just get the tone of the drum.
-And then he achieves those effects using the same sounds like literally the same synthesizer on the same drum machine and stuff."
+I bolded the part about vocals sounding the way they would to the person who's singing them while editing this paragraph because I thought it was kind of funny, and maybe meaningful.
+It sounds like it's coming from inside the house!
 
-"If you used a cscope spectrogram sort of thing to look at a Jack Antonoff song, you're gonna see a lot of big lines.
-And then if you're listening to a song by someone else, you're gonna see like a bunch of sort of like static wiggling.
-It's gonna be really jagged because you've got a bunch of harmonics that don't exactly mix with each other, so they're stacking in all sorts of weird different ways"
-
-# Numbers
+# Numbers?
 
@@
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
+I threw in HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski in too, since they've never collaborated with Jack Antonoff but are more or less part of the same pop space.
 
@@
 
-I chose these artists because they've produced multiple albums, and at least one album with Antonoff, as opposed to just individual songs.
-I chose the Chicks because I figured that collaborating with Antonoff after an extended gap in a longer spanning career than the other artists might offer some interesting results.
-The data points at beginning of their career reflect music that existed in a different time from the other artists in this dataset.
+Of the artists who did collaborate with Antonoff, I chose these artists because they've produced multiple albums, and at least one album with Antonoff, as opposed to just individual songs.
+I chose the Chicks because I figured that collaborating with Antonoff after an extended gap in a longer spanning career than the other artists might offer some interesting results[^1].
+The data points at beginning of their career reflect music that existed in a different time from the other artists in this dataset[^2].
+
+[^1]: It didn't. But I kept them in there anyway because why not.
+
+[^2]: Again, nothing worth noting came up.
 
@@
 I joined these data sets using the `rbind` function in base R, and created an additional variable called "jack", which contains a 1 if Jack Antonoff produced or co-produced the song, and a 0 if he had nothing to do with it.
-I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page.
+I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page \[cite\].
 
@@
 The `get_artist_audio_features` provides a general profile of each song, according to variables developped by spotify, which makes for simple comparisons across artists.
+The final audio features in this study data set contains the variables: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.
 
-However, it's hard to gauge exactly what variables like "danceability" or "valence" mean, even using the documentation.
+\[insert documentation glimpse\]
+
+It's hard to gauge exactly what the variables "danceability" or "valence" mean, even using the documentation.
 For instance, "valence" is described as "A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
@@
 Neither of these definitions fully cover what metrics and thresholds go into something like "danceability".
-However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums[^1].
+However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums[^3].
 
-[^1]: \-- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
+[^3]: \-- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
 The full documentation for each variable included in this study is available in the Appendix
@@
 
-I asked Glen McDonald, a "data alchemist" (see: Principal Engineer) who worked at The Echo Nest, a machine listening start-up that was acquired by Spotify in \[YEAR\]. He still works at Spotify now.
+I asked Glen McDonald, a "data alchemist" (see: Principal Engineer) at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
 
-Glen told me that Valence and Danceability we created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
+Glen told me that `valence` and `danceability` were created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
 Variables like energy or instrumentalness, on the other hand, were determined through machine listening techniques.
 The human subjectivity part only came in when it came to tweaking the training data.
-(My favourite example Glen gave was that the ML treated banjos like a human voice, giving bluegrass songs with banjos high Speechiness ratings.
-Because there were no banjos in the training data, they were registered as human).
+My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
+Because there were no banjos in the training data, they were registered as human speech.
+(This is hilarious.) They had to go back and add more songs with banjos to the training data.
 
-Glen said that the way he's made use of these variables is by combing Energy and Valence to get a quadrant, which usually turns out "pretty good".
+Glen said that the way he's made use of these variables is by combing energy and valence to get a quadrant, which usually turns out "pretty good".
 
@@
 
-The final audio features in this study data set contains the variables: "artist_name", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack".
+I thanked Glen for this time, ended the call, and booted up the ❤️statistical computing programming language, R ❤️\[citeR\].
+
+```{r}
+
+#| label: fig-quaddanceability2
+#| fig-cap: Danceability and Energy Quadrant
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = danceability)) + 
+  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$danceability), color = "gray") +
+  theme_minimal()
+
+```
+
+```{r}
+
+#| label: fig-quadvalence2
+#| fig-cap: Valence and Energy Quadrant
+#| echo: false
+#| warning: false
+
+ggplot(data = df, aes(x = energy, y = valence)) + 
+  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$valence), color = "gray") +
+  theme_minimal()
+
+```
+
+It looks like Jack-produced songs err towards danceability, across the spectrum of energy.
+It's a bit harder to gauge any trends from the valence quadrant.
+These quadrants are a start, but they don't suggest any defining Jack-iness characteristic underlying each of those blue points, which are scattered pretty broadly.
+
+Let's throw a logistic regression at the problem.
 
@@
 ```
-
-# "EDA"
 
@@
 
-```{r}
-
-#| label: fig-quadvalence2
-#| fig-cap: Valence and Energy Quadrant
-#| echo: false
-#| warning: false
-
-ggplot(data = df, aes(x = energy, y = valence)) + 
-  geom_point(aes(color = artist_name), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$valence), color = "gray") +
-  annotate("text", x = mean(df$energy), y = max(df$valence), 
-           label = "High Energy, High Valence") +
-  annotate("text", x = mean(df$energy), y = min(df$valence), 
-           label = "High Energy, Low Valence") +
-  annotate("text", x = min(df$energy), y = mean(df$valence), 
-           label = "Low Energy, High Valence") +
-  annotate("text", x = max(df$energy), y = mean(df$valence), 
-           label = "Low Energy, Low Valence") 
-```
-
-```{r}
-
-#| label: fig-quaddanceability
-#| fig-cap: Danceability and Energy Quadrant
-#| echo: false
-#| warning: false
-
-ggplot(data = df, aes(x = energy, y = danceability)) + 
-  geom_point(aes(color = artist_name), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$danceability), color = "gray") +
-  annotate("text", x = mean(df$energy), y = max(df$danceability), 
-           label = "High Energy, High Danceability") +
-  annotate("text", x = mean(df$energy), y = min(df$danceability), 
-           label = "High Energy, Low Danceability") +
-  annotate("text", x = min(df$energy), y = mean(df$danceability), 
-           label = "Low Energy, High Danceability") +
-  annotate("text", x = max(df$energy), y = mean(df$danceability), 
-           label = "Low Energy, Low Danceability")
-
-```
-
-```{r}
-
-#| label: fig-quaddanceability2
-#| fig-cap: Danceability and Energy Quadrant
-#| echo: false
-#| warning: false
-
-ggplot(data = df, aes(x = energy, y = danceability)) + 
-  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$danceability), color = "gray") +
-  annotate("text", x = mean(df$energy), y = max(df$danceability), 
-           label = "High Energy, High Danceability") +
-  annotate("text", x = mean(df$energy), y = min(df$danceability), 
-           label = "High Energy, Low Danceability") +
-  annotate("text", x = min(df$energy), y = mean(df$danceability), 
-           label = "Low Energy, High Danceability") +
-  annotate("text", x = max(df$energy), y = mean(df$danceability), 
-           label = "Low Energy, Low Danceability")
-
-```
-
-Sure. These are fun quadrants.But it doesn't solve the Jack-iness underlying each of those points, which are scattered pretty broadly.
-
-Let's throw a logistic regression at the problem.
-
 # Model
@@
 
-\[add the formula\]
+I made a binary response variable, called `jack`, that has a value of 1 if the song was written by Jack, and a 0 otherwise.
+There are 10 predictor variables, all of which are continuous: **`danceability`**, **`energy`**, **`loudness`**, **`mode`**, **`speechiness`**, **`acousticness`**, **`instrumentalness`**, **`liveness`**, **`valence`**, and **`tempo`**.
 
-This model estimates the probability P of the outcome variable Y using predictor variables Xc.
-In other words, it estimates the probability that a song was produced by Jack Anotnoff using the Spotfiy variables.
-Which gives us information about the "research" "question": are these characteristics related to Jack's music?
+The formula for this logistic regression can be written as:
 
-The chart below gives us an idea of the mean value of each of these features for Jack and non-Jack song. A logistic regression will give us a better idea of how to weigh each of these features.
+$$
+P(y=1|x) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 x_1 + \cdots + \beta_n x_n)}}
+$$
+
+P(y=1\|x) is the probability of the dependent variable *Y* being 1, given the independent variables *x1*, *x2*, \..., *xn*, and *b0*, *b1*, *b2*, \..., *bn.* The function e\^(-(b0 + b1*x1 + b2*x2 + \...
++ bn\*xn)), also known as the logit function, is the exponential function of the linear combination of the independent variables and coefficients.
+βn represents the weights assigned to each independent variable xn.
+The product of βn and Xn is the contribution of the xn variable to the predicted outcome.
+
+\[Figure X\] gives us an idea of the mean value of each of these features for Jack and non-Jack song.
+A logistic regression will give us a better idea of how to weigh each of these features in order to predict the outcome, which gives us information about the "research" "question": are these characteristics related to Jack's music?
+(And then: Which ones?
+And then, the most elusive question: How?)
+
+Bracketing the major blindspot that there variables provide a single number to represent the dynamics of an entire song, the model has a few other limitations.
+First of all, this model is that assumes that the relationship between the independent variables and the dependent variable is linear.
+There could be non-linear relationships between the variables that this model isn't capturing.
+Second, it assumes that the relationship between the independent variables and the dependent variable is the same for all observations, which might not be true in practice -- there could be interactions between the variables, or in certain subgroups of data where the relationship is different.
+Albums where Jack may have had more of a back seat producing role are one example.
+The fact that a song's development, from inception to recording, is a whole creative process, each having its own origin story with varying degrees of Jack-ness, is another.
 
@@
 
-Following Julia Silge's tutorial \[TK\], I used the tidymodels metapackage to split the training and testing data and make cross-validation samples.
+After some minor pre-processing using the `tidyverse` and `lubridate` R packages. I followed Julia Silge's tutorial \[TK\] to build a model that deals with the imbalanced classes.
+
+I used the `tidymodels` to fir the model.
+To deal with the imbalanced classes, I downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
+Then I used 5-fold cross-validation with stratification by the **`jack`** variable to estimate the performance of the model.
+
 The `workflow_set()` function lets us hold different sets of feature engineering together and use tuning to evaluate them all at once.
@@
 
+I used **`rank_results()`** from the **`yardstick`** package to rank the models based on accuracy, log loss, sensitivity, and specificity. \[ask rohan about this\]
+
+Downsampled it is.
+
 # Results
 
-0.743 accuracy, 0.841 roc_auc. Interesting. Check it out:
+The model's accuracy score is 0.743 accuracy, and the ROC AUC score is 0.841.
+The accuracy score could be misleading, since the model had imbalanced classes, and downsampling may have affected the model's ability to correctly identify the minority class.
+The ROC AUC score is more reliable, since it's less sensitive to class imbalance.
+
+|                            | Actually Positive (1) | Actually Negative (0) |
+|----------------------------|-----------------------|-----------------------|
+| **Predicted Positive (1)** | 581                   | 25                    |
+| **Predicted Negative (0)** | 194                   | 51                    |
+
+: Confusion Matrix for Logistic Regression Results
 
@@
 
-Danceability, with a score of 4.32, had the highest level of importance in determining whether or not a song was produced or co-produced by Jack Antonoff.
-
-# Discussion
-
-## Danceability Predicts Jack-iness {#sec-first-point}
-
-Using the Spotify Audio Features, you can't really gauge what Jack Antonoff's music sound like.
-You get a rough idea, but you don't get the same granularity that you would by simply DM'ing Caleb Gamman on Twitter and asking him to describe it you.
-\[ add gamman quote\]
-
-According to this model, though, danceability is the most important feature in predicting whether or not a song was produced by Jack Antonoff.
-This is so hilarious.
-This was a feature that was invented in a lab 15 years ago.
-This number refers to the time a bunch of undergraduates were paid (hopefully?) to listen to a bunch of songs and tag whether they were danceable or not.
-It is pretty much based on intuition.
-It captures this je-ne-sais-quoi essence of music that connects with listeners on a visceral level.
-Even if it doesn't make YOUR body want to move, you could imagine it making SOMEBODY'S body want to move.
+Danceability, with a positive importance score of 4.32, had the highest level of importance in determining whether or not a song was produced or co-produced by Jack Antonoff.
+Instrumentalness comes in second at 2.59, with a negative score (meaning that negative instrumentalness is correlated), and valence comes in third with a negative score at 2.02.
 
@@
   select(artist_name, danceability, track_name) |> 
+  head(20) |>
   knitr::kable(
@@
 
-### What do I think:
+Table X shows us the 20 songs in Jack Antonoff's discography with the highest danceability scores.
+Bleachers takes up 35% of this list, and Taylor Swift takes up 40% of this list.
+St Vincent, The Chicks, Lorde, and Florence and the Machine also haves songs up there.
+
+# Discussion
+
+## Danceability Predicts Jack-iness {#sec-first-point}
+
+Using the Spotify Audio Features, you can't really gauge what Jack Antonoff's music sound like.
+You get a rough idea, but you don't get the same granularity that you would by simply DM'ing Caleb Gamman on Twitter and asking him to describe it you.
+Or by watching the video where he already did that.
+
+> I think it's mostly in the vocal treatment, that's the thing that is the most off-putting to me.
+But just every decision feels a little bit wrong to me.
+If I wanted to be cringe I could point out the Juno 6 Bass, the perfectly stacked vocals, the Juno 6 pad, the wide and tight plate reverb, the Juno 6 sword, the goo goo gaga, the man hates treble unless it's just random crap swishing around, everything gets low passed with thin 707 drum samples, thin 808 drum samples, then 909 drum samples, thin Lin drum samples, thin thin drum samples, all the resonances, the harmonics get removed, everything is sort of squished down to the fundamental reduction in timbre, everything exists in a small frequency range, everything becomes a tone, it's an extremely consonant sound with no interplay between the different elements.
+Now that is a skill to do, and it produces a cleaner sound, and the only problem is I viscerally hate that sound.
+
+The Spotify Features also wouldn't help you understand in the same way you end up understanding it when you sit on my friend Matthew Jordan's couch and watch him beat box a few different Jack Antonoff and Pharrell impressions in order to show the contrast.
+
+According to this model, though, danceability is the most important feature in predicting whether or not a song was produced by Jack Antonoff.
+And the accuracy metrics aren't awful -- the accuracy score is 0.743, and the roc_auc (receiver operating characteristic area under the curve) (more fun to say out loud as an abbreviation.
+Rock Auck), which is a metric better suited for imbalanced datasets, such as our own, is 0.810.
+This means that although danceability can't show us the inside of a song, something about this score is telling us something about Jack's music.
+
+This was a feature that was invented in a lab \~15 years ago.
+This number refers to the time a bunch of undergraduates were paid (hopefully?) to listen to a bunch of songs and tag whether they were danceable or not.
+It is pretty much based on intuition.
+It captures this je-ne-sais-quoi essence of music that connects with listeners on a visceral level.
+Even if it doesn't make YOUR body want to move, you could imagine it making SOMEBODY'S body want to move.
+
+### In which I do a close listen of our model's top 5 most danceable songs[^4]:
+
+[^4]: I encourage you to follow along and drop me an email with your own evaluation at michaela.drouillard\@mail.utoronto.ca.
 
@@
 
-    -   pretty danceable tbh
+    -   Pretty danceable tbh.
 
@@
     -   I don't know if "danceability" is the word, but I'll give it "anthem"-y.
-Like, I can't picture a group of friends dancing to this in the kitchen, but I can picture someone listening to it in their car when it comes on the radio on their way to give a presentation or negotiate a promotion or something.
+Like, I can't picture a group of friends dancing to this in the kitchen, but I can picture someone driving, on their way to give a presentation or negotiate a promotion or something, and this song comes on the radio and they leave it on.
+They kind of lean into it.
 
@@
 
-    -   Yeah I'm dancing. Or at least bobbing my head.
+    -   Yeah I'm dancing. Or at least bobbing my head. I only became aware of that fact when I made eye contact with the guy next to me at this cafe.
 
-## But Danceability isn't even about Dancing! I propose: Visceral-ness
+## Danceability isn't even necessarily about Dancing! I propose: Visceral-ness
 
@@
 \[tk\].
+We can't actually know what his involvement in each song was.
+There's also the issue that a study like this implies that Jack Antonoff is the reason for the way these artists' sounds develop, which takes that agency away from the artists.
+This is critique you see on the internet, and also a critique that some people on Twitter leveraged against Gamman after his Midnights video went viral.
+I mentioned it to a friend when she asked me how my final paper was going, and she said something like: "Yeah but like, it's *Taylor Swift*.
+As if anyone actually thinks that Jack Antonoff is the reason for her success." I agree.
 
 #### Could our model just be predicting Jack via Taylor Swift?
+
+So another issue is that there is a lot of Taylor Swift in our training data.
+Taylor Swift has the second highest average danceability score across all songs, and the most danceable songs are hers.
+She also happens to have a lot of collaborations with Antonoff under her belt -- more than any other artist.
+I suspect that perhaps our model was predicting Jack via Taylor, which could be a beautiful story about collaboration and aesthetic affinity.
+Or it could suggest that this model has a stink to it.
+An avenue for future study.
 
@@
 It only makes sense, of course, to hire a bunch of undergrads and get them to tell us whether or not a few thousand songs are visceral, and then build our own visceral-ness metric, and see if we get the same results -- that danceability is just visceral-ness in a trench coat, and that visceral-ness is the characteristic that helps us best understand what makes Jack, Jack.
+
+And if we're doing this for danceability, there might also be something worth investigating in valence, the feature that indicates a song's "positivity", which was also built via college interns' intuitions.
 

### 2023-04-18 outputs/paper/paper.qmd
more lil edits
new mf formula. also bibliography

@@
 number-sections: true
-bibliography: references.bib
+bibliography: bibliography.bib
 ---
@@
 As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely.
-When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis".
+When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis" [@citeMarantz].
 
@@
 
-Around the time that Midnights came out, I came across a video on Twitter of a guy, Caleb Gamman, guessing which songs on Midnights were produced by Antonoff within seconds.
+Around the time that Midnights came out, I came across a video on Twitter of a guy, Caleb Gamman, guessing which songs on Midnights were produced by Antonoff within seconds [@citeTweet].
 The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style".
@@
 
-I reached out to Gamman to ask him to explain Antonoff to me, only to find out, on the morning of the interview, that Gamman had released a 29 minute video essay on his Youtube Channel covering:
+I reached out to Gamman to ask him to explain Antonoff to me, only to find out, on the morning of the interview, that Gamman had released a 29 minute video essay on his Youtube Channel covering [@citeGamman]:
 
@@
 
+He was really nice and fun to talk to.
 I bolded the part about vocals sounding the way they would to the person who's singing them while editing this paragraph because I thought it was kind of funny, and maybe meaningful.
@@
 
-I pulled the data using the Spotify API, which I accessed using the `spotifyr` package in R.
+I pulled the data using the Spotify API, which I accessed using the `spotifyr` package in R [@citeSpotify].
 I used the `get_artist_audio_features` and `get_track_audio_analysis functions` function to acquire data on Lorde, Taylor Swift, St.
@@
 I joined these data sets using the `rbind` function in base R, and created an additional variable called "jack", which contains a 1 if Jack Antonoff produced or co-produced the song, and a 0 if he had nothing to do with it.
-I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page \[cite\].
+I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page [@citeWiki].
 
@@
 
-The `get_artist_audio_features` provides a general profile of each song, according to variables developped by spotify, which makes for simple comparisons across artists.
+The `get_artist_audio_features` provides a general profile of each song, according to variables developped by Spotify, which makes for simple comparisons across artists.
 The final audio features in this study data set contains the variables: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.
@@
 
-I thanked Glen for this time, ended the call, and booted up the ❤️statistical computing programming language, R ❤️\[citeR\].
+I thanked Glen for this time, ended the call, and booted up the ❤️statistical computing programming language, R ❤️[@citeR].
 
 ```{r}
-
 #| label: fig-quaddanceability2
@@
 #| warning: false
-
-ggplot(data = df, aes(x = energy, y = danceability)) + 
+library(ggplot2)
+df |>
+  ggplot(aes(x = energy, y = danceability)) + 
     geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
@@
 
-ggplot(data = df, aes(x = energy, y = valence)) + 
+df |>
+  ggplot(aes(x = energy, y = valence)) + 
     geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
@@
 It looks like Jack-produced songs err towards danceability, across the spectrum of energy.
-It's a bit harder to gauge any trends from the valence quadrant.
+It's a bit harder to gauge any trends from the valence quadrant, but it looks like Jack's songs tend to be somewhat on the side of low valence.
 These quadrants are a start, but they don't suggest any defining Jack-iness characteristic underlying each of those blue points, which are scattered pretty broadly.
@@
  # knitr::kable()
-
-```
-
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-library(tidyverse)
-library(palmerpenguins)
 ```
@@
 $$
-P(y=1|x) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 x_1 + \cdots + \beta_n x_n)}}
+\mbox{Pr}(y_i=1) = \mbox{logit}^{-1}\left(\beta_0+\beta_1 x_i\right)
 $$
@@
 
-After some minor pre-processing using the `tidyverse` and `lubridate` R packages. I followed Julia Silge's tutorial \[TK\] to build a model that deals with the imbalanced classes.
+After some minor pre-processing using the `tidyverse` R package [@citetidyverse]. I followed Julia Silge's tutorial [@citeSilge] to build a model that deals with the imbalanced classes.
 
-I used the `tidymodels` to fir the model.
+I used the `tidymodels` to fit the model [@citetidymodels].
 To deal with the imbalanced classes, I downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
@@
   )
+
+#even tho there's similrities, there's differences
 

### 2023-04-19 outputs/paper/paper.qmd
more edits!!!!

@@
 date: today
-abstract: "In this essay, I will be exploring Jack Antonoff's production style using data from the Spotify API.
-Drawing from an interview with Caleb Gamman, a guy who went viral for guessing which song on Taylor Swift's Midnights album was produced by Antonoff within seconds, I will delve into the specific characteristics that define an Antonoff-y song, and assess whether or not the audio features defined by Spotify can capture this.
-I plan to lasso Glen McDonald,the Spotify engineer behind these features, into explaining them to me.
-There will be a Bayesian inference model.
-It will be a logistic regression.
-We do not know whether or not it will be good.
-Finally, I will armchair a little bit about the cultural significance of Antonoff's music and the reasons behind his ubiquity in the music industry."
+abstract: "Using data from the Spotify API, I explored Jack Antonoff's production style and assessed whether Spotify's audio features can capture it.
+Using a logistic regression model, I found that danceability was the most important feature in predicting Antonoff's involvement.
+In an interview with Glen McDonald, an engineer at Spotify, I learned that danceability and valence were created by tagging songs as positive or danceable.
+In the discussion, I do a close-listen to Antonoff's most danceable songs, and argue that what danceability is capturing the visceral quality of Antonoff's music -- which checks out with his critics' visceral reaction to his music."
 format: pdf
 number-sections: true
-bibliography: bibliography.bib
+bibliography: references.bib
 ---
@@
 To find an answer to this question, I turned to statistics, people who know more than me about describing music, and the overlord of engineered virality itself -- the Spotify API.
-There will be a Bayesian inference model.
-It will be a logistic regression.
-We do not know whether or not it will be good.
+Using a logistic regression model, I was able to determine that Spotify's `danceability` feature was the most important in predicting whether or not a song had any Antonoff involvement.
+During interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns to tag whether or not songs were positive or danceabile.
+In the discussion section, I armchair a bit about what I think this `danceability` feature, which has so much uncertainty cooked into it, is really capturing: visceral-ness.
+Which validates everyone's seemingly visceral reaction to new Antonoff releases, and the popularity of his music.
 
@@
 
-    > I'm literally not an expert and of course I don't actually claim to always be able to be able to 100% identify a Jack Antonoff song.
+    > I'm literally not an expert, and of course I don't actually claim to always be able to be able to 100% identify a Jack Antonoff song.
 It's not like I've been practicing the way that I can tell if he's produced something.
@@
 
-    > hing I got into drumming at the age oflike six so maybe as a drummer the drumsamples are of a fence I wasanachronistically raised on cassettes soyou know maybe I like the sound ofaudio and maybe that explains I tend tolike dissonance and in harmonic soundinterplay of elements Distortion noisethe Clashing of elements I imagine I'velistened to more SoundCloud in my lifethan radio so uh maybe I only want tohear amateur internet loser I'veliked over 10 000 songs on my Spotifyaccount uh and I can identify lot ofthem off the bat but I'm not that greatat identifying music generally so it'sjust uh compartmentalized mediaconsumption habit I first did musicproduction in primary school I mademashups and dubstep in it as a lateteenager I extensively used Juno 6emulators to make embarrassing synthwaveadmittedly that's probably not thesubconscious Association that he's goingfor lyrics aren't of much interest to meit's just a sound in the mix and somaybe I don't like the focus on vocalI've previously described it as a fightor flight sort of thing so maybe itliterally is tapping into some memory oftrauma I have an unusually small earcanal so maybe I'm just hearing the highend more acutely I was hired as ateenager to play drums on uh a lot ofJack Antonoff songs for a universityGlee Club performer so maybe I don'tlike it because of the embarrassment ofhaving ever been associated with a Gleecloth I liked the album 1989 but to makethe song out of the woods palette ofwool to me I had to mash it up with M83Midnight City and that got a dmcatakedown so maybe it's a personalVendetta and of course I have known anddo know people who like this music somaybe I secretly hate my friends andfamily but I don't know if any of thattells us anything uh I have a taste inmusic w
+    > ...
+I got into drumming at the age of like six, so maybe as a drummer the drum samples are \[offensive?\].
+I was anachronistically raised on cassettes so, you know maybe I like the sound of audio and maybe that explains why I tend to like dissonance and inharmonic sound, interplay of elements, distortion noise, the clashing of elements.
+I imagine I've listened to more SoundCloud in my life than radio so uh, maybe I only want to hear amateur internet losers.
+I'vel iked over 10 000 songs on my Spotify account and I can identify lot of them off the bat, but I'm not that great at identifying music generally, so it's just a compartmentalized media consumption habit.
+I first did music production in primary school.
+I made mashups and dubstep.
+As a teenager I extensively used Juno 6 emulators to make embarrassing synthwave.
+\[Admittedly that's probably not the subconscious association that he's going for.\] Lyrics aren't of much interest to me, it's just a sound in the mix, and so maybe I don't like the focus on vocals.
+I've previously described \[my reaction\] as a fight or flight sort of thing, so maybe it's literally is tapping into some memory of trauma.
+I have an unusually small ear canal so maybe I'm just hearing the high end more acutely.
+I was hired as a teenager to play drums on a lot of Jack Antonoff songs for a university glee club so maybe I don't like it because of the embarrassment of having ever been associated with a glee club.
+I liked the album 1989 but to make the song Out of the Woods palettable to me I had to mash it up with M83Midnight City and that got a DMCA takedown so maybe it's a personal vendetta.
+And of course I have known andd o know people who like this music so maybe I secretly hate my friends and family but I don't know if any of that tells us anything ...
 
@@
 
-    > "i did use the phrase visceral hatred.
-but i did so only to describe my visceral hatred.
-i cant stand the sound of it.
-i hate it.
-and the way in which i hate it is viscerally.
-but i didnt go beyond that because it doesnt go beyond that.
-nothing against the guy, i like taylor swift's stuff.
-it's just subjective taste...
+    > ...I did use the phrase visceral hatred.
+But I did so only to describe my visceral hatred.
+I can't stand the sound of it.
+I hate it.
+And the way in which I hate it is viscerally.
+But I didn't go beyond that because it doesn't go beyond that.
+Nothing against the guy, I like Taylor Swift's stuff.
+It's just subjective taste...
 
-4.  His tongue in cheek response to some of Twitter's negative response to his video
+4.  His tongue in cheek and pretty based response to some of Twitter's negative response to his video
 
     > .... \[it's just subjective taste\].
-there's just nothing to engage with, so ive had to fabricate a reason for controversy.
-and what ive come up with is this.
-\*looks at phone\* the real reason iive been saying for years that i have a visceral reaction to jack antonoffs production style, the REAL reason is this \-- cause during the rollout cycle of the album midnights, i wanted to ascribe taylor swifts authorship to a man and diminish women's voices.
-so actually, i'm a sexist.
-so now thats out there, publishhed as facts, in the news.
-with all the other facts..."
+There's just nothing to engage with, so I've had to fabricate a reason for controversy.
+And what I've come up with is this: \**Reads off of phone*\* "The real reason I've been saying for years that I have a visceral reaction to Jack Antonoff's production style, the REAL reason is this: because during the rollout cycle of the album *Midnights*, I wanted to ascribe Taylor Swift's authorship to a man and diminish women's voices." So actually, I'm a sexist.
+So now that's out there, published as facts, in the news.
+With all the other facts...
 
@@
 
-    > "what makes you so special and unique that you can tell the production of a song?" "nothing"
+    > .."What makes you so special and unique that you can tell the production of a song?" Nothing. ...
 
@@
 
-# Numbers?
+## So
 
@@
 
-I pulled the data using the Spotify API, which I accessed using the `spotifyr` package in R [@citeSpotify].
-I used the `get_artist_audio_features` and `get_track_audio_analysis functions` function to acquire data on Lorde, Taylor Swift, St.
+I pulled data on artists who've collaborated with Antonoff using the Spotify API, which I accessed using the `spotifyr` package in R [@citeSpotify].
+I used the `get_artist_audio_features` function to acquire data on Lorde, Taylor Swift, St.
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
-I threw in HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski in too, since they've never collaborated with Jack Antonoff but are more or less part of the same pop space.
+I threw in HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski in too, since they've never collaborated with Jack Antonoff but are more or less part of the same pop sphere.
 
@@
 
-### Spotify's API
+### Audio Features Dataset
 
-So what do any one these features mean?
-
-The Spotify API returns JSON metadata directly from the Spotify Data Catalogue.
-
-#### Audio Features Dataset
-
-The `get_artist_audio_features` provides a general profile of each song, according to variables developped by Spotify, which makes for simple comparisons across artists.
+The `get_artist_audio_features` provides a profile of each song, according to variables developped by Spotify, which makes for simple comparisons across artists.
 The final audio features in this study data set contains the variables: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.
 
-\[insert documentation glimpse\]
+From the `get_artist_audio_features` documentation (@citeSpotifyDocumentation):
+
+> -   **acousticness**: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.
+>
+> -   **danceability**: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+A value of 0.0 is least danceable and 1.0 is most danceable.
+>
+> -   **energy**: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
+Typically, energetic tracks feel fast, loud, and noisy.
+For example, death metal has high energy, while a Bach prelude scores low on the scale.
+Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy.
+>
+> -   **instrumentalness**: Predicts whether a track contains no vocals.
+"Ooh" and "aah" sounds are treated as instrumental in this context.
+Rap or spoken word tracks are clearly "vocal".
+The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content.
+Values above 0.5 are intended to represent instrumental tracks, but confidence is higher as the value approaches 1.0.
+>
+> -   **key**: The key the track is in.
+Integers map to pitches using standard [Pitch Class notation](https://en.wikipedia.org/wiki/Pitch_class).
+E.g.
+0 = C, 1 = C♯/D♭, 2 = D, and so on.
+If no key was detected, the value is -1.
+>
+> -   **liveness**: Detects the presence of an audience in the recording.
+Higher liveness values represent an increased probability that the track was performed live.
+A value above 0.8 provides strong likelihood that the track is live.
+>
+> -   **loudness:** The overall loudness of a track in decibels (dB).
+Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks.
+Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude).
+Values typically range between -60 and 0 db.
+>
+> -   **mode:** Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.
+>
+> -   **tempo:** The overall estimated tempo of a track in beats per minute (BPM).
+In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration.
+>
+> -   **valence:** A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
+Tracks with high valence sound more positive (e.g.
+happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g.
+sad, depressed, angry).
 
@@
 I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
-Jupiter's valence score is 326.78% higher than Mars.
-This not only indicates the valence is based on audio features other than lyrical content, but that it is in fact capturing a difference between the two songs' profiles.
+In @tbl-Holst, Gustav Holst's Jupiter has a valence score 326.78% higher than Mars.
+This tracks.
+It suggests the valence is based on audio features other than lyrical content.
+It also suggests that it is, in fact, capturing a difference between the two songs' profiles.
 
 ```{r}
-
 #| echo: false
 #| warning: false
+#| label: tbl-Holst
+#| tbl-cap: Valence and Danceability Scores for Holst's Mars and Jupiter
+
 library(dplyr)
@@
 
-I looked at the valence measures on Flight of the Bumblebee, a song that's much faster and "upbeat" than Mars and Jupiter.
-What I found complicated my understanding of valence -- different performances of the same song had different valence scores.
+I looked at the valence measures on Nikolai Rimsky-Korsakov's Flight of the Bumblebee, a song that's much faster and frantic than Mars and Jupiter.
+What I found complicated my understanding of valence -- different performances of the same song had different valence scores (@tbl-bees).
 Some were arias, played by different instruments.
-These tend to have higher valence scores.
-The lowest recorded valence is 0.2410, and the highest is .9230.
+Arias tended to have higher valence scores.
+The lowest recorded valence across all performances is 0.2410, and the highest is .9230.
 What gives?
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| label: tbl-bees
+#| tbl-cap: Valence and Danceability Scores for Every Performance of Flight of the Bumblee on Spotify
+
 sub_bees<- read.csv(here::here("inputs/data/subset_bumblebee.csv"))
@@
 Glen told me that `valence` and `danceability` were created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
-Variables like energy or instrumentalness, on the other hand, were determined through machine listening techniques.
-The human subjectivity part only came in when it came to tweaking the training data.
-My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
+Variables like energy or instrumentalness, on the other hand, were determined primarily through machine listening techniques.
+The human subjectivity part only came in when it came to tweaking the training data[^4]:
+
+[^4]: My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
 Because there were no banjos in the training data, they were registered as human speech.
@@
 
-Glen said that the way he's made use of these variables is by combing energy and valence to get a quadrant, which usually turns out "pretty good".
+"You could imagine writing a formula for energy that combines like loudness and tempo and degree of harmonic variation or something.
+So it was ML, but that one's more like a human helping a machine figure out a formula.
+Whereas valence is teaching the machine to try to reproduce a purely human thing.
+Same with danceability.
+I mean, danceability is whether a human can dance to it.
 
-\[discuss the convo we had about how even two humans can have disagreement about a song's positivity, let alone human vs machine.
-i.e.
-an elliot smith song where the lyrics are depressing but the melody is happy.
-or a happy song but you know that the band died in a plane crash yesterday.
-etc.\]
+So like, the machine's not gonna dance, so the machine can't have any opinion on that.
+The computer could have an opinion on energy.
+And the computer can definitely have an opinion on loudness.
+So there's a spectrum from, you know, loudness as purely analytical, and then energy is a little like loudness, with a little more subjectivity.
+And then danceability and valence are purely subjective."
 
+This put the big range of valence scores for Flight of The Bumblebee into perspective -- what that value is capturing is abstracted away from the actual notes of the song.
+The valence feature propagates a vibe, so of course this uncertainty makes room for noise and variation.
+
+Glen confirmed that this process didn't know anything about lyrics, which makes valence a particularly difficult variable to built.
+It could, in theory, pick up on aspects of vocal performance, but it doesn't know anything about languages or words or the meanings of songs.
+Take an upbeat, happy-sounding Elliot Smith song with devastating lyrics: the machine might register it as happy, where human listeners understand that it's sad.
+Add to that that two humans might even disagree on the song's valence.
+
+"Plus, we have the confounding factor of like, a song that seems happy today could be sad tomorrow because the singer was killed in a plane crash.
+The song didn't change, but our world changed and our reactions changed."
+
+All this said, Glen said that the way he's made use of these variables is by combing energy and valence to get a quadrant, which usually works "fairly well".
+High energy and high valence could be generally happy, cheerful, upbeat.
+Low energy, low valence could be sad or downbeat.
+High energy, low valence is sort of angry.
+Low energy, high valence is serene or calming.
 I thanked Glen for this time, ended the call, and booted up the ❤️statistical computing programming language, R ❤️[@citeR].
+
+```{r}
+#| label: fig-quadvalence
+#| fig-cap: Valence and Energy Quadrant
+#| echo: false
+#| warning: false
+library(ggplot2)
+df <- read.csv(here::here("inputs/data/df.csv"))
+
+  ggplot(df, aes(x = energy, y = valence)) + 
+    geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
+    geom_vline(xintercept = mean(df$energy), color = "gray") +
+    geom_hline(yintercept = mean(df$valence), color = "gray") +
+    theme_minimal()
+
+```
+
+In @fig-quadvalence, it looks like songs produced by Jack Antonoff are a little more clustered towards the lower end of valence, and across the spectrum of energy, although the variance is pretty big.
 
@@
 #| warning: false
-library(ggplot2)
-df |>
-  ggplot(aes(x = energy, y = danceability)) + 
+
+  ggplot(df, aes(x = energy, y = danceability)) + 
     geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
@@
 
+In @fig-quaddanceability2, it looks like Jack-produced songs err towards higher danceability, across the spectrum of energy.
+
+These quadrants are a start, but they don't suggest any defining Jack-iness characteristic underlying each of those blue points, which are scattered pretty broadly.
+
+By comparing cumulative distribution functions of each variable based on whether or not a song was produced by Antonoff, we can see that some of these variables do capture some sort of underlying Jack-iness (@fig-cdf).
+For instance, while liveness, instrumentalness, and speechiness are relatively similar, we can see clear differences in danceability, energy, loudness, acousticness, and valence.
+
 ```{r}
-
-#| label: fig-quadvalence2
-#| fig-cap: Valence and Energy Quadrant
+#| label: fig-cdf
+#| fig-cap: Cumulative Distribution Function of Each Variable, 
 #| echo: false
 #| warning: false
+library(tidyverse)
+data_reduced <- read.csv(here::here("inputs/data/data_reduced.csv"))
 
-df |>
-  ggplot(aes(x = energy, y = valence)) + 
-    geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
-    geom_vline(xintercept = mean(df$energy), color = "gray") +
-    geom_hline(yintercept = mean(df$valence), color = "gray") +
-    theme_minimal()
+data_reduced |> 
+  pivot_longer(cols = c(danceability, energy, loudness, mode,
+               speechiness, acousticness, instrumentalness, liveness, valence,
+               tempo),
+               names_to = "name",
+               values_to = "value"
+  ) |> 
+  ggplot(aes(x = value, color = jack)) +
+  stat_ecdf() +
+  facet_wrap(vars(name), 
+             nrow = 3, 
+             ncol = 4,
+             scales = "free") +
+  theme_minimal() +
+  labs(
+    x = "Value",
+    y = "Proportion"
+  )
+
+#even tho there's similrities, there's differences
 
@@
 
-It looks like Jack-produced songs err towards danceability, across the spectrum of energy.
-It's a bit harder to gauge any trends from the valence quadrant, but it looks like Jack's songs tend to be somewhat on the side of low valence.
-These quadrants are a start, but they don't suggest any defining Jack-iness characteristic underlying each of those blue points, which are scattered pretty broadly.
-
-Let's throw a logistic regression at the problem.
-
-```{r}
-#| echo: false
-#| warning: false
-
-#library(kableExtra)
-#df |>
- # knitr::kable()
-```
+Another way to gauge the Antonoff-ification of pop music could be to look at how artists' sounds have changed over time, and whether or not there is any correlation with Antonoff.
+Here is where the necessary "correlation is not causation" caveat comes in: suggesting that Antonoff is the *cause* for any change in artists' sounds would be an overstep.
+These plots are here to get a glimpse at pop's evolution over time, and where Antonoff fits in that picture.
+Each point represents a song.
 
@@
   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?") +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
   facet_wrap( ~ artist_name, ncol = 3)
@@
 ```
+
+In @fig-energy, it looks like there's a downward trend in energy scores in artists who work with Antonoff, but that these downward trends are also somewhat in line with the artists' energy trajectories anyway.
+Specifically, Lana Del Rey and Taylor Swift.
+They have enough albums recorded prior to collaborations with Antonoff to be able to observe this.
+Sharon Von Etten, Maggie Rogers, and Mitski, who have never worked with Antonoff, all have upward trends.
 
@@
   scale_color_manual(values = c("#808080","#FFB6C1")) +
-  labs(color = "Has 'y' in jack?") +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
   facet_wrap(~ artist_name, ncol = 3)
@@
 ```
+
+There aren't any significant correlations in danceability (@fig-danceability) .
 
@@
   scale_color_manual(values = c("#808080","#FFB6C1")) +
-  labs(color = "Has 'y' in jack?") +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
   facet_wrap(~ artist_name, ncol = 3)
@@
 ```
+
+In @fig-valence, it looks like Lorde has the only significant change in valence scores, which tracks with my understanding her albums.
+Melodrama had Green Lights, and Solar Power had a music video of Lorde frolicking on a beach in a yellow bikini with Phoebe Bridgers and Clairo.
 
@@
   scale_color_manual(values = c("#808080","#FFB6C1")) +
-  labs(color = "Has 'y' in jack?") +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
   facet_wrap(~ artist_name, ncol = 3)
@@
 ```
+
+Again in @fig-acousticness, Lorde's acoustic Solar Power album jumps out at us.
+Taylor Swift, Lorde and Haim also have significant-seeming upward trends in acousticness, while Mitski, Sharon Von Etten and Maggie Rogers see decreasing acousticness.
 
@@
   scale_color_manual(values = c("#808080","#FFB6C1")) +
-  labs(color = "Has 'y' in jack?") +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
   facet_wrap(~ artist_name, ncol = 3)
@@
 
+In @fig-instrumentalness, St. Vincent is the only plot with any significant trend.
+
 ```{r}
-
-#| label: fig-temp
+#| label: fig-tempo
 #| fig-cap: Tempo Variable
@@
   scale_color_manual(values = c("#808080","#FFB6C1")) +
-  labs(color = "Has 'y' in jack?") +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
   facet_wrap(~ artist_name, ncol = 3)
@@
 
-I'm gonna fix the axis labels later
+Finally, in @fig-tempo, there aren't many significant trends, except for Antonoff's collaborations with Lana Del Rey and Taylor Swift.
 
-```{r}
-
-#| label: fig-quadvalence1
-#| fig-cap: Valence and Energy Quadrant 
-#| echo: false
-#| warning: false
-
-ggplot(data = df, aes(x = energy, y = valence)) + 
-  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$valence), color = "gray") +
-  annotate("text", x = mean(df$energy), y = max(df$valence), 
-           label = "High Energy, High Valence") +
-  annotate("text", x = mean(df$energy), y = min(df$valence), 
-           label = "High Energy, Low Valence") +
-  annotate("text", x = min(df$energy), y = mean(df$valence), 
-           label = "Low Energy, High Valence") +
-  annotate("text", x = max(df$energy), y = mean(df$valence), 
-           label = "Low Energy, Low Valence")
-
-```
+Now that we understand that there do seem to be some subtle differences in how variables that are and aren't produced by Jack behave, as seen in our cumulative distribution functions, our quadrants, and some minor trends in artists' sounds over time (specifically, `energy`), we can use a logistic regression to see if a model can a) predict whether or not a song was produced by Antonoff based on these variables, and b) tell us which variables were more significant in determining the output.
 
@@
 
-P(y=1\|x) is the probability of the dependent variable *Y* being 1, given the independent variables *x1*, *x2*, \..., *xn*, and *b0*, *b1*, *b2*, \..., *bn.* The function e\^(-(b0 + b1*x1 + b2*x2 + \...
-+ bn\*xn)), also known as the logit function, is the exponential function of the linear combination of the independent variables and coefficients.
-βn represents the weights assigned to each independent variable xn.
-The product of βn and Xn is the contribution of the xn variable to the predicted outcome.
+where $\text{Pr}(y_i=1)$ is the probability of a song being written by Jack, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the predictor variables.
 
-\[Figure X\] gives us an idea of the mean value of each of these features for Jack and non-Jack song.
+@fig-means gives us an idea of the mean value of each of these features for Jack and non-Jack song.
 A logistic regression will give us a better idea of how to weigh each of these features in order to predict the outcome, which gives us information about the "research" "question": are these characteristics related to Jack's music?
@@
 
-Bracketing the major blindspot that there variables provide a single number to represent the dynamics of an entire song, the model has a few other limitations.
-First of all, this model is that assumes that the relationship between the independent variables and the dependent variable is linear.
-There could be non-linear relationships between the variables that this model isn't capturing.
-Second, it assumes that the relationship between the independent variables and the dependent variable is the same for all observations, which might not be true in practice -- there could be interactions between the variables, or in certain subgroups of data where the relationship is different.
-Albums where Jack may have had more of a back seat producing role are one example.
-The fact that a song's development, from inception to recording, is a whole creative process, each having its own origin story with varying degrees of Jack-ness, is another.
-
 ```{r}
-#|: echo: false
-#|: warning: false
+#| label: fig-means
+#| fig-cap: Comparing the Mean Values of Each Variable
+#| echo: false
+#| warning: false
 
@@
   geom_col(alpha = 0.8, position = "dodge") +
-  labs(x = "Mean Value", y = "Feature")
+  theme_minimal() +
+  labs(fill = "Jack", x = "Mean Value", y = "Feature")
 
 ```
+
+Bracketing the major blindspot that there variables provide a single number to represent the dynamics of an entire song, the model has a few other limitations.
+First of all, this model is that assumes that the relationship between the independent variables and the dependent variable is linear.
+There could be non-linear relationships between the variables that this model isn't capturing.
+Second, it assumes that the relationship between the independent variables and the dependent variable is the same for all observations, which might not be true in practice -- there could be interactions between the variables, or in certain subgroups of data where the relationship is different.
+Albums where Jack may have had more of a back seat producing role are one example.
+The fact that a song's development, from inception to recording, is a whole creative process, each having its own origin story with varying degrees of Jack-ness, is another.
 
@@
 ```{r}
-#|: echo: false
-#|: warning: false
+#| echo: false
+#| warning: false
 
@@
 |                        | Actually Positive (1) | Actually Negative (0) |
-|----------------------------|-----------------------|-----------------------|
-| **Predicted Positive (1)** | 581                   | 25                    |
-| **Predicted Negative (0)** | 194                   | 51                    |
+|------------------------|-----------------------|-----------------------|
+| Predicted Positive (1) | 581                   | 25                    |
+| Predicted Negative (0) | 194                   | 51                    |
 
@@
 
-```{r}
-#jackmatrix
-```
+@tbl-importance shows the different variables' importance on determining whether a song was had any Antonoff involvement.
 
 ```{r}
-
-data_reduced |> 
-  pivot_longer(cols = c(danceability, energy, loudness, mode,
-               speechiness, acousticness, instrumentalness, liveness, valence,
-               tempo),
-               names_to = "name",
-               values_to = "value"
-  ) |> 
-  ggplot(aes(x = value, color = jack)) +
-  stat_ecdf() +
-  facet_wrap(vars(name), 
-             nrow = 3, 
-             ncol = 4,
-             scales = "free") +
-  theme_minimal() +
-  labs(
-    x = "Value",
-    y = "Proportion"
-  )
-
-#even tho there's similrities, there's differences
-
-```
-
-And here are the different variables' importance on determining whether a song was had any Antonoff involvement.
-
-```{r}
+#| label: tbl-importance
+#| tbl-cap: Importance Scores of Each Variable in Determining Jack-ness
 #| echo: false
@@
 ```{r}
+#| label: fig-importance
+#| fig-cap: Importance Scores of Each Variable in Determining Jack-ness
 #| echo: false
@@
 ```{r}
+#| label: tbl-mostdanceable
+#| tbl-cap: Top 20 Most Danceable Antonoff-Produced Songs
 #| echo: false
@@
 
-Table X shows us the 20 songs in Jack Antonoff's discography with the highest danceability scores.
+@tbl-mostdanceable shows us the 20 songs in Jack Antonoff's discography with the highest danceability scores.
 Bleachers takes up 35% of this list, and Taylor Swift takes up 40% of this list.
@@
 
-### In which I do a close listen of our model's top 5 most danceable songs[^4]:
+### In which I do a close listen of our model's top 5 most danceable songs[^5]:
 
-[^4]: I encourage you to follow along and drop me an email with your own evaluation at michaela.drouillard\@mail.utoronto.ca.
+[^5]: I encourage you to follow along and drop me an email with your own evaluation at michaela.drouillard\@mail.utoronto.ca.
 
@@
 So another issue is that there is a lot of Taylor Swift in our training data.
+As mentioned earlier, her songs take up 40% of @tbl-mostdanceable, which is 5% more than Antonoff's own band, Bleachers.
 Taylor Swift has the second highest average danceability score across all songs, and the most danceable songs are hers.
@@
   ggtitle("Boxplot of Danceability by Artist") +
-  coord_flip()
+  coord_flip() +
+  theme_minimal()
 ```
@@
   coord_flip() +
-  labs(title = "Artists with Most Jack = 1 Values", x = "Artist", y = "Count")
+  labs(title = "Artists with Most Jack = 1 Values", x = "Artist", y = "Count") +
+  theme_minimal()
 
@@
 
-# Appendix {.unnumbered}
-
-### `get_artist_audio_features` documentation:
-
--   **acousticness**: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.
-
--   **danceability**: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
-A value of 0.0 is least danceable and 1.0 is most danceable.
-
--   **energy**: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
-Typically, energetic tracks feel fast, loud, and noisy.
-For example, death metal has high energy, while a Bach prelude scores low on the scale.
-Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy.
-
--   **instrumentalness**: Predicts whether a track contains no vocals.
-"Ooh" and "aah" sounds are treated as instrumental in this context.
-Rap or spoken word tracks are clearly "vocal".
-The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content.
-Values above 0.5 are intended to represent instrumental tracks, but confidence is higher as the value approaches 1.0.
-
--   **key**: The key the track is in.
-Integers map to pitches using standard [Pitch Class notation](https://en.wikipedia.org/wiki/Pitch_class).
-E.g.
-0 = C, 1 = C♯/D♭, 2 = D, and so on.
-If no key was detected, the value is -1.
-
--   **liveness**: Detects the presence of an audience in the recording.
-Higher liveness values represent an increased probability that the track was performed live.
-A value above 0.8 provides strong likelihood that the track is live.
-
--   **loudness:** The overall loudness of a track in decibels (dB).
-Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks.
-Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude).
-Values typically range between -60 and 0 db.
-
--   **mode:** Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.
-
--   **tempo:** The overall estimated tempo of a track in beats per minute (BPM).
-In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration.
-
--   **valence:** A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
-Tracks with high valence sound more positive (e.g.
-happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g.
-sad, depressed, angry).
-
 # Additional details

### 2023-04-19 scripts/05-graphs.R
shaping up the scripts

@@ -4,8 +4,7 @@
 # Data: 28 February 2023
 # Contact: michaela.drouillard@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+
 
 
 #### Workspace setup ####
@@ -15,7 +14,7 @@ library(lubridate)
 df <- read_csv("inputs/data/df.csv")
 
 str(df)
-#For whatever reason, Jack didn't stay a factor. Making it a factor again here
+
 df$jack <- as.factor(df$jack)
 
 
@@ -111,5 +110,52 @@ df %>%
   labs(color = "Has 'y' in jack?") +
   facet_wrap(~ artist_name, ncol = 3) +
   geom_smooth(method = "lm", se = FALSE)
+
+## Plotting cdfs
+data_reduced <- 
+  df |> 
+  select(jack, 
+         danceability, energy, loudness, mode,
+         speechiness, acousticness, instrumentalness, liveness, valence,
+         tempo
+  ) 
+
+data_reduced |> 
+  pivot_longer(cols = c(danceability, energy, loudness, mode,
+                        speechiness, acousticness, instrumentalness, liveness, valence,
+                        tempo),
+               names_to = "name",
+               values_to = "value"
+  ) |> 
+  ggplot(aes(x = value, color = jack)) +
+  stat_ecdf() +
+  facet_wrap(vars(name), 
+             nrow = 3, 
+             ncol = 4,
+             scales = "free") +
+  theme_minimal() +
+  labs(
+    x = "Value",
+    y = "Proportion"
+  )
+
+
+## Mean value of each variable
+
+data_reduced %>%
+  group_by(jack) %>%
+  summarise(across(everything(), mean)) %>%
+  pivot_longer(cols = -jack) %>%
+  ggplot(aes(value, name, fill = factor(jack))) +
+  geom_col(alpha = 0.8, position = "dodge")
+
+
+data_reduced %>%
+  group_by(jack) %>%
+  summarise(across(-c(tempo, loudness), mean)) %>%
+  pivot_longer(cols = -jack) %>%
+  ggplot(aes(value, name, fill = factor(jack))) +
+  geom_col(alpha = 0.8, position = "dodge")
+
 
 

### 2023-04-19 scripts/06-models.R

@@ -1,104 +1,201 @@
 #### Preamble ####
-# Purpose: Models the Jack Antonoffification of Pop Music
+# Purpose: Logistic Regression
 # Author: Michaela Drouillard
 # Data: 17 March 2023
 # Contact: michaela.drouillard@mail.utoronto.ca
 # License: MIT
 
-#Going to make a smaller df that only had the columns we want to capture correlations
 
-library(ggplot2)
+
 library(tidyverse)
-df <- read_csv("inputs/data/df.csv")
-df$jack <- as.factor(df$jack)
+library(lubridate)
 
-library(caret)
-library(rstanarm)
+data <- read_csv(here::here("inputs/data/df.csv"))
 
-
-set.seed(123)
-
-train_index <- createDataPartition(df$jack, p = 0.7, list = FALSE)
-train_data <- df[train_index, ]
-test_data <- df[-train_index, ]
-
-#balance the classes by undersampling
-train_data_balanced <- train_data |>
-  group_by(jack) |>
-  sample_n(size = min(table(jack))) |>
-  ungroup()
+data <- 
+  data |> 
+  mutate(jack = as_factor(jack)) |> 
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date))
 
 
+data_reduced <- 
+  data |> 
+  select(jack, 
+         danceability, energy, loudness, mode,
+         speechiness, acousticness, instrumentalness, liveness, valence,
+         tempo
+  ) 
 
-model <- stan_glm(jack ~ danceability + energy + loudness + speechiness + key + 
-                    acousticness + instrumentalness + liveness + valence,
-                  data = train_data_balanced,
-                  family = binomial(),
-                  prior_intercept = normal(0,10),
-                  prior = normal(0,2.5),
-                  chains = 4, iter = 2000)
+write_csv(data_reduced,"inputs/data/data_reduced.csv")
 
+#### Try a model ####
+# Code from: https://juliasilge.com/blog/project-feederwatch/
 
+library(tidymodels)
 
-# Make predictions on the testing data
-#probabilities <- predict(model, newdata = test_data, type = "response")
-#predictions <- ifelse(probabilities >= 0.5, "Yes", "No")
+set.seed(853)
 
-# Add the predictions to the testing data
-test_data$predictions <- ifelse(predict(model, newdata = test_data, type = "response") > 0.5, "Yes", "No")
[…]

### 2023-04-19 scripts/09-rohan_messing_around.R

@@ -1,233 +1,0 @@
-
-# Preamble
-library(tidyverse)
-library(lubridate)
-
-data <- read_csv(here::here("inputs/data/df.csv"))
-
-data <- 
-  data |> 
-  mutate(jack = as_factor(jack)) |> 
-  filter(album_release_date_precision == "day") |>
-  mutate(album_release_date = ymd(album_release_date))
-
-
-# Make ecdf of the variables, coloured by whether Jack
-data_reduced <- 
-  data |> 
-  select(jack, 
-         danceability, energy, loudness, mode,
-         speechiness, acousticness, instrumentalness, liveness, valence,
-         tempo
-         ) 
-
-write_csv(data_reduced,"inputs/data/data_reduced.csv")
-
-data_reduced |> 
-  pivot_longer(cols = c(danceability, energy, loudness, mode,
-               speechiness, acousticness, instrumentalness, liveness, valence,
-               tempo),
-               names_to = "name",
-               values_to = "value"
-  ) |> 
-  ggplot(aes(x = value, color = jack)) +
-  stat_ecdf() +
-  facet_wrap(vars(name), 
-             nrow = 3, 
-             ncol = 4,
-             scales = "free") +
-  theme_minimal() +
-  labs(
-    x = "Value",
-    y = "Proportion"
-  )
-
-
-data_reduced %>%
-  group_by(jack) %>%
-  summarise(across(everything(), mean)) %>%
-  pivot_longer(cols = -jack) %>%
-  ggplot(aes(value, name, fill = factor(jack))) +
-  geom_col(alpha = 0.8, position = "dodge")
-
-
-data_reduced %>%
-  group_by(jack) %>%
-  summarise(across(-c(tempo, loudness), mean)) %>%
-  pivot_longer(cols = -jack) %>%
-  ggplot(aes(value, name, fill = factor(jack))) +
-  geom_col(alpha = 0.8, position = "dodge")
-
-
-
-?across
-#### Try a model ####
-# Code from: https://juliasilge.com/blog/project-feederwatch/
-
-library(tidymodels)
-
-set.seed(853)
-
-jack_split <- 
-  data_reduced |>
-  initial_split(strata = jack)
-
-jack_train <- training(jack_split)
-jack_test <- testing(jack_split)
-
-set.seed(853)
-jack_folds <- vfold_cv(jack_train, strata = jack)
-jack_folds
-
-is_jack <- 
-  recipe(jack ~ ., data = jack_train)
[…]

### 2023-04-19 outputs/paper/paper.qmd
now Rohan can't say the estimand isn't clearly stated in the intro
more edits

@@
 What is it about Antonoff music that is so Antonoff-y?
-Is the Jack Antonoff-fication of pop something we should be worried about?
 
-To find an answer to this question, I turned to statistics, people who know more than me about describing music, and the overlord of engineered virality itself -- the Spotify API.
-Using a logistic regression model, I was able to determine that Spotify's `danceability` feature was the most important in predicting whether or not a song had any Antonoff involvement.
-During interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns to tag whether or not songs were positive or danceabile.
-In the discussion section, I armchair a bit about what I think this `danceability` feature, which has so much uncertainty cooked into it, is really capturing: visceral-ness.
-Which validates everyone's seemingly visceral reaction to new Antonoff releases, and the popularity of his music.
+To find an answer to this question, I turned to statistics, people who know more than me about describing music (Youtubers), and the overlord of engineered virality itself -- the Spotify API.
+Can Spotify's audio features dataset help us determine what makes Antonoff music so Antonoff-y?
+
+I spoke to Caleb Gamman, a Youtuber and Antonoff-production-style hater who had a tweet go viral around the time of Taylor Swift's *Midnights* release, in which he predicts whether or not a song was produced by Antonoff within seconds.
+A real-life logistic regression, if you will.
+Using an actual logistic regression model, I was able to determine that Spotify's `danceability` feature was the most important in predicting whether or not a song had any Antonoff involvement.
+During an interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns to tag whether or not songs were positive or danceable.
+In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing visceral-ness.
+We can capture something about Antonoff's music with these variables, and it might just be that subjective "ick" that many people feel and others enjoy that we're all just trying to describe.
 
@@
 
-Around the time that Midnights came out, I came across a video on Twitter of a guy, Caleb Gamman, guessing which songs on Midnights were produced by Antonoff within seconds [@citeTweet].
+Around the time that *Midnights* came out, I came across a video on Twitter of a guy, Caleb Gamman, guessing which songs on Midnights were produced by Antonoff within seconds [@citeTweet].
 The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style".

### 2023-04-19 outputs/paper/paper.qmd
more edits

@@
 During an interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns to tag whether or not songs were positive or danceable.
-In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing visceral-ness.
-We can capture something about Antonoff's music with these variables, and it might just be that subjective "ick" that many people feel and others enjoy that we're all just trying to describe.
+In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing something like visceral-ness.
+We can capture something about Antonoff's music with these variables, and it might just be that subjective "ick" that many people feel and others enjoy and that we're all just trying to describe.
 

### 2023-04-19 outputs/paper/paper.qmd
fixed the bib

@@
 title: "We Are All Trying to Describe Something And We Don't Have the Words"
-subtitle: "The Jack Antonoffification of Pop Music?"
+subtitle: "A Study of The Jack Antonoffification of Pop Music, and Propagating Uncertainty with Spotify"
 author: 
@@
 abstract: "Using data from the Spotify API, I explored Jack Antonoff's production style and assessed whether Spotify's audio features can capture it.
-Using a logistic regression model, I found that danceability was the most important feature in predicting Antonoff's involvement.
-In an interview with Glen McDonald, an engineer at Spotify, I learned that danceability and valence were created by tagging songs as positive or danceable.
-In the discussion, I do a close-listen to Antonoff's most danceable songs, and argue that what danceability is capturing the visceral quality of Antonoff's music -- which checks out with his critics' visceral reaction to his music."
+In an interview with Glen McDonald, an engineer at Spotify, I learned that the danceability and valence features were created by having interns tag songs as positive or gloomy, or danceable or un-danceable.
+Using a logistic regression model, I found that danceability was the most important feature in predicting Antonoff's involvement in a song.
+In the discussion section, I do a close-listen of Antonoff's most danceable songs, and argue that the danceability variable is actually capturing something more like a visceral quality to Antonoff's music."
 format: pdf
@@
 
-Midnights was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my blacking out, and the stranger's "Exactly".
+Midnights was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my glitching, and the stranger's "Exactly".
 
@@
 When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed.
-What my friend was referring to, of course, is the fact that Jack Antonoff and his sound seems to be everywhere these days.
+What my friend was referring to is the fact that Jack Antonoff and his sound seems to be everywhere these days.
 As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely.
@@
 Using an actual logistic regression model, I was able to determine that Spotify's `danceability` feature was the most important in predicting whether or not a song had any Antonoff involvement.
-During an interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns to tag whether or not songs were positive or danceable.
+During an interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns and having them tag whether or not songs were positive or danceable.
 In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing something like visceral-ness.
-We can capture something about Antonoff's music with these variables, and it might just be that subjective "ick" that many people feel and others enjoy and that we're all just trying to describe.
+Against the odds, we can capture something about Antonoff's music with these variables, and it might just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe.
 
@@
 
-    > I'm literally not an expert, and of course I don't actually claim to always be able to be able to 100% identify a Jack Antonoff song.
+    > ...I'm literally not an expert, and of course I don't actually claim to always be able to be able to 100% identify a Jack Antonoff song.
 It's not like I've been practicing the way that I can tell if he's produced something.
@@
 Now that is a skill to do, and it produces a cleaner sound, and the only problem is I viscerally hate that sound.
+...
 
@@
 He was really nice and fun to talk to.
-I bolded the part about vocals sounding the way they would to the person who's singing them while editing this paragraph because I thought it was kind of funny, and maybe meaningful.
-It sounds like it's coming from inside the house!
+I bolded the part about vocals sounding the way they would to the person who's singing them while editing this paragraph because I thought it was kind of funny, and kind of cool.
+It's coming from inside the house!!!!
 
@@
 I joined these data sets using the `rbind` function in base R, and created an additional variable called "jack", which contains a 1 if Jack Antonoff produced or co-produced the song, and a 0 if he had nothing to do with it.
-I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page [@citeWiki].
+I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page [@citewiki].
 
@@
 [^3]: \-- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
-The full documentation for each variable included in this study is available in the Appendix
 
@@
 Especially valence -- is the positivity related to lyrics, or something else?
-I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the audio itself.
+I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the just the instruments.
 In @tbl-Holst, Gustav Holst's Jupiter has a valence score 326.78% higher than Mars.
@@
 
-In @fig-instrumentalness, St. Vincent is the only plot with any significant trend.
+In @fig-instrumentalness, St. Vincent is the only plot with any significant trend in instrumentalness.
 
@@
 
-Finally, in @fig-tempo, there aren't many significant trends, except for Antonoff's collaborations with Lana Del Rey and Taylor Swift.
+Finally, in @fig-tempo, there aren't many significant trends in tempo, except for Antonoff's collaborations with Lana Del Rey and Taylor Swift.
 
@@
 Bracketing the major blindspot that there variables provide a single number to represent the dynamics of an entire song, the model has a few other limitations.
-First of all, this model is that assumes that the relationship between the independent variables and the dependent variable is linear.
+First of all, this model assumes that the relationship between the independent variables and the dependent variable is linear.
 There could be non-linear relationships between the variables that this model isn't capturing.
@@
 Albums where Jack may have had more of a back seat producing role are one example.
-The fact that a song's development, from inception to recording, is a whole creative process, each having its own origin story with varying degrees of Jack-ness, is another.
+(The fact that a song's development, from inception to recording, is a whole creative process, each having its own origin story with varying degrees of Jack-ness, is another.)
 
@@
 
-I used **`rank_results()`** from the **`yardstick`** package to rank the models based on accuracy, log loss, sensitivity, and specificity. \[ask rohan about this\]
-
-Downsampled it is.
+I used **`rank_results()`** from the **`yardstick`** package to rank the models based on accuracy, log loss, sensitivity, and specificity, and went with the downsampled model.
 
@@
 
-The model's accuracy score is 0.743 accuracy, and the ROC AUC score is 0.841.
+The model's accuracy score is 0.743, and the ROC AUC score is 0.841.
 The accuracy score could be misleading, since the model had imbalanced classes, and downsampling may have affected the model's ability to correctly identify the minority class.
@@
 
-Using the Spotify Audio Features, you can't really gauge what Jack Antonoff's music sound like.
+Using the Spotify Audio Features, you can't really gauge what Jack Antonoff's music *sounds* like.
 You get a rough idea, but you don't get the same granularity that you would by simply DM'ing Caleb Gamman on Twitter and asking him to describe it you.
@@
 
-The Spotify Features also wouldn't help you understand in the same way you end up understanding it when you sit on my friend Matthew Jordan's couch and watch him beat box a few different Jack Antonoff and Pharrell impressions in order to show the contrast.
+The Spotify Features also wouldn't help you understand in the same way you end up understanding it if you sat on my friend Matthew Jordan's couch and watched him beat box a few different Jack Antonoff and Pharrell impressions in order to show the contrast.
 
@@
 
-[^5]: I encourage you to follow along and drop me an email with your own evaluation at michaela.drouillard\@mail.utoronto.ca.
+[^5]: I encourage you to follow along and drop me an email with your own assessment at michaela.drouillard\@mail.utoronto.ca.
 
@@
 The danceability variable, which is the strongest predictor of Jack-ness, seems to actually be more about "visceral"-ness when you do a close-listen of the songs with the highest danceability scores.
+Not all of them are danceable, but the one's that aren't still feel like they're trying to manipulate the way your body is holding itself in some way.
 The reason Jack is getting under people's skin could just be because his music...
@@
 Since we're not in the studio, we can't know the degree of Jack-iness in each individual song.
-Especially when he takes on the role of a more backseat co-producer, as he did with Lana Del Rey's Did You Know There's a Tunnel Under Ocean Blvd?
-\[tk\].
+Especially when he takes on the role of a more backseat co-producer.
 We can't actually know what his involvement in each song was.
 There's also the issue that a study like this implies that Jack Antonoff is the reason for the way these artists' sounds develop, which takes that agency away from the artists.
-This is critique you see on the internet, and also a critique that some people on Twitter leveraged against Gamman after his Midnights video went viral.
+This is a critique you see on the internet, and also a critique that some people on Twitter leveraged against Gamman after his Midnights video went viral.
 I mentioned it to a friend when she asked me how my final paper was going, and she said something like: "Yeah but like, it's *Taylor Swift*.
@@
 
-\appendix
-
-# Additional details
-
-\newpage
-
 # References

### 2023-04-19 outputs/paper/paper.qmd
nitpicky

@@
 title: "We Are All Trying to Describe Something And We Don't Have the Words"
-subtitle: "A Study of The Jack Antonoffification of Pop Music, and Propagating Uncertainty with Spotify"
+subtitle: "A Study of The Jack Antonoffification of Pop Music (Or: Propagating Uncertainty with Spotify)"
 author: 

### 2023-04-20 outputs/paper/paper.qmd
Start cleaning up paper

@@
 
+```{r}
+#| echo: false
+#| warning: false
+
+library(tidyverse)
+library(lubridate)
+
+```
+
 # Introduction
 
-Sometime in early October of 2022 there was a chemical shift in my brain and I couldn't stop listening to Taylor Swift.
+Sometime in early October of 2022 there was a chemical shift in my brain, and I couldn't stop listening to Taylor Swift.
 I couldn't listen to anything other than Fearless or Red when deadlifting, walking to a friend's house, folding laundry, etc.
@@
 
-Midnights was produced by Swift and her longtime collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my glitching, and the stranger's "Exactly".
+Midnights was produced by Swift and her long-time collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my glitching, and the stranger's "Exactly".
 
@@
 
-I spoke to Caleb Gamman, a Youtuber and Antonoff-production-style hater who had a tweet go viral around the time of Taylor Swift's *Midnights* release, in which he predicts whether or not a song was produced by Antonoff within seconds.
+I spoke to Caleb Gamman, a Youtuber and Antonoff-production-style hater who had a tweet go viral around the time of Taylor Swift's *Midnights* release, in which he predicts whether a song was produced by Antonoff within seconds.
 A real-life logistic regression, if you will.
-Using an actual logistic regression model, I was able to determine that Spotify's `danceability` feature was the most important in predicting whether or not a song had any Antonoff involvement.
-During an interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns and having them tag whether or not songs were positive or danceable.
+Using an actual logistic regression model, I was able to determine that Spotify's `danceability` feature was the most important in predicting whether a song had any Antonoff involvement.
+During an interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns and having them tag whether songs were positive or danceable.
 In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing something like visceral-ness.
@@
 
-I reached out to Gamman to ask him to explain Antonoff to me, only to find out, on the morning of the interview, that Gamman had released a 29 minute video essay on his Youtube Channel covering [@citeGamman]:
+I reached out to Gamman to ask him to explain Antonoff to me, only to find out, on the morning of the interview, that Gamman had released a 29 minute video essay on his YouTube Channel covering [@citeGamman]:
 
@@
 I imagine I've listened to more SoundCloud in my life than radio so uh, maybe I only want to hear amateur internet losers.
-I'vel iked over 10 000 songs on my Spotify account and I can identify lot of them off the bat, but I'm not that great at identifying music generally, so it's just a compartmentalized media consumption habit.
+I've liked over 10 000 songs on my Spotify account and I can identify lot of them off the bat, but I'm not that great at identifying music generally, so it's just a compartmentalized media consumption habit.
 I first did music production in primary school.
@@
 I was hired as a teenager to play drums on a lot of Jack Antonoff songs for a university glee club so maybe I don't like it because of the embarrassment of having ever been associated with a glee club.
-I liked the album 1989 but to make the song Out of the Woods palettable to me I had to mash it up with M83Midnight City and that got a DMCA takedown so maybe it's a personal vendetta.
-And of course I have known andd o know people who like this music so maybe I secretly hate my friends and family but I don't know if any of that tells us anything ...
+I liked the album 1989 but to make the song Out of the Woods palatable to me I had to mash it up with M83Midnight City and that got a DMCA takedown so maybe it's a personal vendetta.
+And of course I have known and do know people who like this music so maybe I secretly hate my friends and family but I don't know if any of that tells us anything ...
 
@@
 Jack Antonoff does a very good job of making stuff very clean.
-So you know, all of the notes that are happening simulaneously are perfectly stacked.
+So you know, all of the notes that are happening simultaneously are perfectly stacked.
 They're like mathematically stacked ..
@@
 
-library(dplyr)
+sub_planets <- read_csv(here::here("inputs/data/subset_planets.csv"))
 
-sub_planets<- read.csv(here::here("inputs/data/subset_planets.csv"))
 sub_planets |>
@@
   head() |>
-  knitr::kable(
-    col.names = c("Artist Name", 
+  knitr::kable(col.names = c("Artist Name",
                              "Valence",
                              "Danceability",
-                  "Track Name")
- )
-
+                             "Track Name"))
 ```
@@
 #| label: tbl-bees
-#| tbl-cap: Valence and Danceability Scores for Every Performance of Flight of the Bumblee on Spotify
+#| tbl-cap: Valence and Danceability Scores for Every Performance of Flight of the Bumblebee on Spotify
 
-sub_bees<- read.csv(here::here("inputs/data/subset_bumblebee.csv"))
+sub_bees <- read_csv(here::here("inputs/data/subset_bumblebee.csv"))
+
 sub_bees |>
-   select(artist_name, valence, danceability, album_release_year, track_name) |>
+  select(artist_name,
+         valence,
+         danceability,
+         album_release_year,
+         track_name) |>
   head(17) |>
   knitr::kable(
-     col.names = c("Artist Name", 
+    col.names = c(
+      "Artist Name",
       "Valence",
@@
       "Album Release Year",
-                   "Track Name")
+      "Track Name"
     )
-
+  )
 ```
@@
 #| warning: false
-library(ggplot2)
-df <- read.csv(here::here("inputs/data/df.csv"))
+
+df <- read_csv(here::here("inputs/data/df.csv"))
 
@@
   theme_minimal()
-
 ```
@@
   geom_vline(xintercept = mean(df$energy), color = "gray") +
-    geom_hline(yintercept = mean(df$danceability), color = "gray") +
+  geom_hline(yintercept = mean(df$danceability),
+             color = "gray") +
   theme_minimal()
-
 ```
@@
 
-By comparing cumulative distribution functions of each variable based on whether or not a song was produced by Antonoff, we can see that some of these variables do capture some sort of underlying Jack-iness (@fig-cdf).
+By comparing cumulative distribution functions of each variable based on whether a song was produced by Antonoff, we can see that some of these variables do capture some sort of underlying Jack-iness (@fig-cdf).
 For instance, while liveness, instrumentalness, and speechiness are relatively similar, we can see clear differences in danceability, energy, loudness, acousticness, and valence.
@@
 #| warning: false
-library(tidyverse)
-data_reduced <- read.csv(here::here("inputs/data/data_reduced.csv"))
+
+data_reduced <- read_csv(here::here("inputs/data/data_reduced.csv"))
 
 data_reduced |>
-  pivot_longer(cols = c(danceability, energy, loudness, mode,
-               speechiness, acousticness, instrumentalness, liveness, valence,
-               tempo),
+  pivot_longer(
+    cols = c(
+      danceability,
+      energy,
+      loudness,
+      mode,
+      speechiness,
+      acousticness,
+      instrumentalness,
+      liveness,
+      valence,
+      tempo
+    ),
     names_to = "name",
@@
   theme_minimal() +
-  labs(
-    x = "Value",
-    y = "Proportion"
-  )
+  labs(x = "Value",
+       y = "Proportion")
 
 #even tho there's similrities, there's differences
-
 ```
 
-Another way to gauge the Antonoff-ification of pop music could be to look at how artists' sounds have changed over time, and whether or not there is any correlation with Antonoff.
+Another way to gauge the Antonoff-ification of pop music could be to look at how artists' sounds have changed over time, and whether there is any correlation with Antonoff.
 Here is where the necessary "correlation is not causation" caveat comes in: suggesting that Antonoff is the *cause* for any change in artists' sounds would be an overstep.
@@
 
-library(lubridate)
-library(tidyverse)
 df <- read_csv(here::here("inputs/data/df.csv"))
-df$jack <- factor(df$jack)
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
+
+df <- df |> 
+  mutate(jack = factor(jack))
+
+df |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
@@
   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
-  facet_wrap( ~ artist_name, ncol = 3)
-
+  facet_wrap(~ artist_name, ncol = 3)
 ```
@@
 
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
+df |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
@@
   theme_minimal() +
-  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
-  facet_wrap(~ artist_name, ncol = 3)
-
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
+df |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
   ggplot(aes(y = valence, x = album_release_date, color = jack)) +
@@
   theme_minimal() +
-  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
-  facet_wrap(~ artist_name, ncol = 3)
-
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
+df |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
   ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
@@
   theme_minimal() +
-  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
-  facet_wrap(~ artist_name, ncol = 3)
-
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
+df |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
   ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
@@
   theme_minimal() +
-  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
-  facet_wrap(~ artist_name, ncol = 3)
-
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-df %>%
-  filter(album_release_date_precision == "day") %>%
-  mutate(album_release_date = ymd(album_release_date)) %>%
+df |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
   ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
@@
   theme_minimal() +
-  scale_color_manual(values = c("#808080","#FFB6C1")) +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
-  facet_wrap(~ artist_name, ncol = 3)
-
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-Now that we understand that there do seem to be some subtle differences in how variables that are and aren't produced by Jack behave, as seen in our cumulative distribution functions, our quadrants, and some minor trends in artists' sounds over time (specifically, `energy`), we can use a logistic regression to see if a model can a) predict whether or not a song was produced by Antonoff based on these variables, and b) tell us which variables were more significant in determining the output.
+Now that we understand that there do seem to be some subtle differences in how variables that are and aren't produced by Jack behave, as seen in our cumulative distribution functions, our quadrants, and some minor trends in artists' sounds over time (specifically, `energy`), we can use a logistic regression to see if a model can a) predict whether a song was produced by Antonoff based on these variables, and b) tell us which variables were more significant in determining the output.
 
@@
 
-data_reduced <- read.csv(here::here("inputs/data/data_reduced.csv"))
+data_reduced <- read_csv(here::here("inputs/data/data_reduced.csv"))
 
-data_reduced %>%
-  group_by(jack) %>%
-  summarise(across(-c(tempo, loudness), mean)) %>%
-  pivot_longer(cols = -jack) %>%
+data_reduced |>
+  group_by(jack) |>
+  summarise(across(-c(tempo, loudness), mean)) |>
+  pivot_longer(cols = -jack) |>
   ggplot(aes(value, name, fill = factor(jack))) +
@@
   labs(fill = "Jack", x = "Mean Value", y = "Feature")
-
 ```
 
-Bracketing the major blindspot that there variables provide a single number to represent the dynamics of an entire song, the model has a few other limitations.
+Bracketing the major blindspot that these variables provide a single number to represent the dynamics of an entire song, the model has a few other limitations.
 First of all, this model assumes that the relationship between the independent variables and the dependent variable is linear.
@@
 
-data_reduced %>%
-  count(jack) %>%
+data_reduced |>
+  count(jack) |>
   knitr::kable(col.names = c("Is it Jack?", "Number of Songs"))
-
 ```
@@
 
-After some minor pre-processing using the `tidyverse` R package [@citetidyverse]. I followed Julia Silge's tutorial [@citeSilge] to build a model that deals with the imbalanced classes.
+After some minor pre-processing using the `tidyverse` R package [@citetidyverse]. I followed @citeSilge to build a model that deals with the imbalanced classes.
 
@@
 
-jack_vip <- read.csv(here::here("inputs/data/jack_vip.csv"))
+jack_vip <- read_csv(here::here("inputs/data/jack_vip.csv"))
 
 jack_vip |>
-   knitr::kable(
-     col.names = c("Variable", 
+  knitr::kable(col.names = c("Variable",
                              "Importance",
-                   "Sign")
-  )
-
+                             "Sign"))
 ```
@@
 
-jack_vip %>%
-  group_by(Sign) %>%
-  slice_max(Importance, n = 15) %>%
-  ungroup() %>%
+jack_vip |>
+  group_by(Sign) |>
+  slice_max(Importance, n = 15) |>
+  ungroup() |>
   ggplot(aes(Importance, fct_reorder(Variable, Importance), fill = Sign)) +
@@
 
-Danceability, with a positive importance score of 4.32, had the highest level of importance in determining whether or not a song was produced or co-produced by Jack Antonoff.
+Danceability, with a positive importance score of 4.32, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
 Instrumentalness comes in second at 2.59, with a negative score (meaning that negative instrumentalness is correlated), and valence comes in third with a negative score at 2.02.
@@
 
-df %>%
-  filter(jack == 1) %>%
+df |>
+  filter(jack == 1) |>
   arrange(desc(danceability)) |>
@@
   head(20) |>
-  knitr::kable(
-     col.names = c("Artist Name", 
+  knitr::kable(col.names = c("Artist Name",
                              "Danceability",
-                   "Track")
-  )
-
+                             "Track"))
 ```
@@
 
-According to this model, though, danceability is the most important feature in predicting whether or not a song was produced by Jack Antonoff.
+According to this model, though, danceability is the most important feature in predicting whether a song was produced by Jack Antonoff.
 And the accuracy metrics aren't awful -- the accuracy score is 0.743, and the roc_auc (receiver operating characteristic area under the curve) (more fun to say out loud as an abbreviation.
@@
 #| warning: false
-df %>%
-  filter(jack == 1) %>%
+
+df |>
+  filter(jack == 1) |>
   ggplot(aes(x = artist_name, y = danceability)) +
@@
 #| warning: false
-library(dplyr)
-library(ggplot2)
 
-df %>%
-  filter(jack == 1) %>%
-  count(artist_name) %>%
-  arrange(desc(n)) %>%
-  head(10) %>%
+df |>
+  filter(jack == 1) |>
+  count(artist_name) |>
+  arrange(desc(n)) |>
+  head(10) |>
   ggplot(aes(x = reorder(artist_name, n), y = n)) +
@@
 
-It only makes sense, of course, to hire a bunch of undergrads and get them to tell us whether or not a few thousand songs are visceral, and then build our own visceral-ness metric, and see if we get the same results -- that danceability is just visceral-ness in a trench coat, and that visceral-ness is the characteristic that helps us best understand what makes Jack, Jack.
+It only makes sense, of course, to hire a bunch of undergrads and get them to tell us whether a few thousand songs are visceral, and then build our own visceral-ness metric, and see if we get the same results -- that danceability is just visceral-ness in a trench coat, and that visceral-ness is the characteristic that helps us best understand what makes Jack, Jack.
 

### 2023-10-28 scripts/06-models.R
EDA on random forest correlations, multi class logistic regression

@@ -11,6 +11,8 @@ library(tidyverse)
 library(lubridate)
 library(tidymodels)
 library(randomForest)
+library(ranger)
+library(caret)
 
 data <- read_csv(here::here("inputs/data/clean_df.csv"))
 
@@ -175,11 +177,11 @@ library(caret)
 
 data_reduced$producer <- as.factor(data_reduced$producer)
 
+
 set.seed(853)
 antonoff_split <- initial_split(data_reduced, strata = producer)
 antonoff_train <- training(antonoff_split)
 antonoff_test <- testing(antonoff_split)
-
 
 rf_model <- ranger(producer ~ ., data = antonoff_train, importance = 'impurity')
 #impurity == asking model to interpret variable importance based on gini impurity
@@ -208,29 +210,333 @@ for (producer in unique_producers) {
   importance_temp <- rf_temp$importance
   print(colnames(importance_temp))  # To see available columns
   
-  
   list_importance[[producer]] <- importance_temp[,'MeanDecreaseGini']
 }
-
-
 
 df_importance <- bind_rows(list_importance, .id = "producer")
 
 head(df_importance)
-
-
 
 df_importance_long <- df_importance %>%
   gather(variable, importance, -producer)
 
 ggplot(df_importance_long, aes(x = reorder(variable, importance), y = importance)) +
   geom_bar(stat = "identity") +
-  facet_wrap(~producer, scales = "free") + 
+  facet_wrap(~producer, scales = "fixed") + 
   coord_flip() +
   labs(title = "Variable Importance by Producer", 
        x = "Variable", 
        y = "Mean Decrease Gini") +
   theme_minimal()
+
+
+
+### Correlations by Producer ###
+
+
+library(tidyr)
+library(corrr)
+
+# Assuming data_reduced is your dataset
+correlations_by_producer <- data_reduced %>%
+  group_by(producer) %>%
+  nest() %>%
+  mutate(correlation = map(data, ~ correlate(.x, use = "pairwise.complete.obs"))) %>%
+  select(producer, correlation)
+
+library(ggplot2)
+library(reshape2) 
+
+producer_name <- "antonoff"
+
+correlation_matrix <- correlations_by_producer %>%
+  filter(producer == producer_name) %>%
+  pull(correlation) %>%
+  .[[1]]
+
+melted_corr <- melt(correlation_matrix)
+
[…]

### 2023-10-31 outputs/paper/paper.qmd
took a hatchet to the paper

@@
 
-Sometime in early October of 2022 there was a chemical shift in my brain, and I couldn't stop listening to Taylor Swift.
-I couldn't listen to anything other than Fearless or Red when deadlifting, walking to a friend's house, folding laundry, etc.
-I listened to so much Taylor in that month alone that she ended up as my #4 artist on Spotify Wrapped, which stops collecting its data for the year on October 31st.
-It didn't even capture what happened to me after Midnights dropped in November.
+Shortly after the release of Taylor Swift's *Midnights* album, a video of the Youtuber Caleb Gammen, in which guesses which songs on Midnights were produced by Antonoff within seconds [@citeTweet], went viral.
+The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style".
+He nails it.
+Only one song on the album wasn't produced by Antonoff, and he guesses it.
 
-I was standing outside of the Garrison eating a slice of pizza with some friends, and one of them asked what music we had been listening to lately.
-
-ME: "Midnights!"
-
-ZACH: "Is it good?"
-
-\[I glitched.\]
-
-ME: "No!"
-
-\["*No*?"\]
-
-A STRANGER WALKING BY WHO HEARD EVERYTHING: "EXACTLY!!"
-
-Midnights was produced by Swift and her long-time collaborator Jack Antonoff, who is the subject of this paper, and, I suspect, the reason behind both my glitching, and the stranger's "Exactly".
-
+Antonoff began his career as a guitarist in fun., and is the frontman in Bleachers.
+He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more.
 Antonoff is prolific, to the extent that it is starting to bother people.
-He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more.
-All of this and his own bands too (Bleachers, and fun.).
-When I was getting brunch with a friend one Sunday and a Swift song came on, and she said "Jack Antonoff's gonna kill us all", we both laughed.
-What my friend was referring to is the fact that Jack Antonoff and his sound seems to be everywhere these days.
 As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely.
@@
 
-What I am trying to get to here is that I don't actually really mind Jack Antonoff's music, I kind of like it, but I still say his voice with the same tone that everyone else does.
-There's just something about him...
-this sense that he's everywhere...
-and that he's just such a nice guy...
-Why do music people seem to be a little annoyed with him?
-What is it about Antonoff music that is so Antonoff-y?
+(Cite a few snippets from critics articles to give a dose of the discourse)
 
-To find an answer to this question, I turned to statistics, people who know more than me about describing music (Youtubers), and the overlord of engineered virality itself -- the Spotify API.
 Can Spotify's audio features dataset help us determine what makes Antonoff music so Antonoff-y?
@@
 I spoke to Caleb Gamman, a Youtuber and Antonoff-production-style hater who had a tweet go viral around the time of Taylor Swift's *Midnights* release, in which he predicts whether a song was produced by Antonoff within seconds.
-A real-life logistic regression, if you will.
-Using an actual logistic regression model, I was able to determine that Spotify's `danceability` feature was the most important in predicting whether a song had any Antonoff involvement.
+
 During an interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns and having them tag whether songs were positive or danceable.
+
+I used two models to investigate the data.
+The first, a logistic regression, draws on the dataset of artists who've collaborated with Antonoff, with an extra binary variable to tag whether or not the song was an Antonoff production.
+With this model, we can get a first glimpse at variables that are stronger predictors of Antonoff-iness, along with a better understanding of whether the Spotify data can trace Antonoff's influence on an artist's sound over time.
+
+However, logistic regression assumes a linear relationship between the dependent and predictor variables, and is sensitive to multi-collinearity.
+In music data, we would expect for certain variables to be correlated (for instance, loudness and energy).
+We also cannot assume that an artists' style can be measured in the linear growth of these metrics.
+Although the logistic regression has a strong accuracy, indicating that danceability and loudness are the strongest indicators of an Anotonoff produced song, it may be detecting a change in genre, rather than a change in style, across artists whose sound turns more into pop while collaborating with Antonoff.
+
+I built a second model, a Random Forest, to complement the insights drawn from the logistic regression.
+The Random Forest trains on the full data set, including other producers.
+Random Forests (what's the package) don't assume linear relationships, and aren't sensitive to multicollinearity.
+In fact, using the x package in R, we can visualize and analyse the relationships between different variables.
+The variance in correlation scores across producers can be taken as an approximate for style, since this variance helps the model classify tracks.
+
 In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing something like visceral-ness.
@@
 
-# An Interview with Caleb Gamman
+TLDR of the Gamman interview is: Taste is subjective.
+People who can play instruments tend to be better at identifying and describing sounds.
+Cultural criticism is a hall of smoke and mirrors.
+Antonoff is everywhere.
 
-Around the time that *Midnights* came out, I came across a video on Twitter of a guy, Caleb Gamman, guessing which songs on Midnights were produced by Antonoff within seconds [@citeTweet].
-The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style".
-He nails it.
-Only one song on the album wasn't produced by Antonoff, and he guesses it.
+Can we translate any of this into something observable? Can we quantify the Jack Antonoff-ification of pop music?
+
+# Mini literature review on Antonoff's place in culture? or should that go in the introduction
+
+# Either radically transform or cull the gamman or use it for the discussion section
 
@@
 
-He was really nice and fun to talk to.
-I bolded the part about vocals sounding the way they would to the person who's singing them while editing this paragraph because I thought it was kind of funny, and kind of cool.
-It's coming from inside the house!!!!
-
-## So
-
-TLDR of the Gamman interview is: Taste is subjective.
-People who can play instruments tend to be better at identifying and describing sounds.
-Cultural criticism is a hall of smoke and mirrors.
-Antonoff is everywhere.
-
-Can we translate any of this into something observable? Can we quantify the Jack Antonoff-ification of pop music?
-
 # Data {#sec-data}
 
-I pulled data on artists who've collaborated with Antonoff using the Spotify API, which I accessed using the `spotifyr` package in R [@citeSpotify].
+The audio feature data was pulled using Spotify API, which I accessed using the `spotifyr` package in R [@citeSpotify].
 I used the `get_artist_audio_features` function to acquire data on Lorde, Taylor Swift, St.
@@
 
+I also scraped the discographies of comparable contemporary pop producers' Wikipedias.
+The producers include Epworth, Rechstaid, Max Martin, Nowels, Joel Little, and Greg Kustin.
+(cite Carl here?) These producers were chosen based on their popularity, and also because some of them collaborated with artists who also collaborate with Antonoff.
+
+The data was manually validated, and live performances, karaoke editions, international versions or translations, and remix albums were removed.
+Where deluxe albums were available, original albums were deleted to avoid duplicate rows.
+
 ### Artist Selection
@@
 Of the artists who did collaborate with Antonoff, I chose these artists because they've produced multiple albums, and at least one album with Antonoff, as opposed to just individual songs.
-I chose the Chicks because I figured that collaborating with Antonoff after an extended gap in a longer spanning career than the other artists might offer some interesting results[^1].
-The data points at beginning of their career reflect music that existed in a different time from the other artists in this dataset[^2].
-
-[^1]: It didn't. But I kept them in there anyway because why not.
-
-[^2]: Again, nothing worth noting came up.
 
@@
 
-I asked Glen McDonald, a "data alchemist" (see: Principal Engineer) at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
-
-Glen told me that `valence` and `danceability` were created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
+I interviewed Glen McDonald, a Principal Engineer at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
+Macdonald told me that `valence` and `danceability` were created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
 Variables like energy or instrumentalness, on the other hand, were determined primarily through machine listening techniques.
@@
 "You could imagine writing a formula for energy that combines like loudness and tempo and degree of harmonic variation or something.
-So it was ML, but that one's more like a human helping a machine figure out a formula.
+So it was [machine learning], but that one's more like a human helping a machine figure out a formula.
 Whereas valence is teaching the machine to try to reproduce a purely human thing.
@@
 I mean, danceability is whether a human can dance to it.
-
 So like, the machine's not gonna dance, so the machine can't have any opinion on that.
@@
 
-Glen confirmed that this process didn't know anything about lyrics, which makes valence a particularly difficult variable to built.
+Glen confirmed that this process didn't know anything about lyrics, which makes valence a particularly difficult variable to build.
 It could, in theory, pick up on aspects of vocal performance, but it doesn't know anything about languages or words or the meanings of songs.
@@
 
-All this said, Glen said that the way he's made use of these variables is by combing energy and valence to get a quadrant, which usually works "fairly well".
+Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well".
 High energy and high valence could be generally happy, cheerful, upbeat.
@@
 Low energy, high valence is serene or calming.
-I thanked Glen for this time, ended the call, and booted up the ❤️statistical computing programming language, R ❤️[@citeR].
 
@@
 
-In @fig-quadvalence, it looks like songs produced by Jack Antonoff are a little more clustered towards the lower end of valence, and across the spectrum of energy, although the variance is pretty big.
+In @fig-quadvalence, it looks like songs produced by Jack Antonoff are a little more clustered towards the lower end of valence, and across the spectrum of energy, although the variance is large.
 
@@
 
-# Model
-
-Since we're trying to figure out whether these variables can tell us anything about what Jack Antonoff's sound is all about, a good way to figure that out is by seeing if any of these variables can help us predict what songs are produced by Jack.
-Logistic regressions like questions with binary outcomes.
+# Logistic Regression
 
@@
 
+# Random Forest
+
 # Discussion
@@
     -   Yeah I'm dancing. Or at least bobbing my head. I only became aware of that fact when I made eye contact with the guy next to me at this cafe.
+
+## Draw on the random forest danceability correlations graph, tie it into the close listen
 

### 2023-10-31 outputs/paper/paper.qmd
RF section started, discussion too

@@
 Against the odds, we can capture something about Antonoff's music with these variables, and it might just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe.
-
-TLDR of the Gamman interview is: Taste is subjective.
-People who can play instruments tend to be better at identifying and describing sounds.
-Cultural criticism is a hall of smoke and mirrors.
-Antonoff is everywhere.
-
-Can we translate any of this into something observable? Can we quantify the Jack Antonoff-ification of pop music?
 
@@
 ![](images/Screen%20Shot%202023-04-03%20at%209.31.05%20PM.png)
-
-It is a perfect video essay. I pestered him anyway:
 
@@
 Neither of these definitions fully cover what metrics and thresholds go into something like "danceability".
-However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums[^3].
+However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums[^1].
 
-[^3]: \-- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
+[^1]: \-- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
 
@@
 Variables like energy or instrumentalness, on the other hand, were determined primarily through machine listening techniques.
-The human subjectivity part only came in when it came to tweaking the training data[^4]:
+The human subjectivity part only came in when it came to tweaking the training data[^2]:
 
-[^4]: My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
+[^2]: My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
 Because there were no banjos in the training data, they were registered as human speech.
@@
 "You could imagine writing a formula for energy that combines like loudness and tempo and degree of harmonic variation or something.
-So it was [machine learning], but that one's more like a human helping a machine figure out a formula.
+So it was \[machine learning\], but that one's more like a human helping a machine figure out a formula.
 Whereas valence is teaching the machine to try to reproduce a purely human thing.
@@
 
+Random forests (RF) build an ensemble of decision trees during the training phase, and then use majority voting for classification tasks.
+Each tree in the ensembled is trained on a random sample of data taken with replacement, and only a random subset of features are considered when splitting at each node.
+This method ensures that each tree in the ensemble is unique, avoiding overfitting and capturing non-linear relationships and more complex patterns in the data.
+The formula behind the RF is:
+
+## Data Preparation
+
+The categorical feature 'producer', which contains the names of the producer for each track, was one-hot encoded to create binary columns for each producer ('is_antonoff', 'is_elworth', etc.).
+The producer feature was then converted into a factor variable, and the dataset was split into training and tests sets.
+
+## Model Training
+
+The RF was trainied using the @ranger package. The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
+
+Gini impurity is a measure of misclassification.
+At every split in a decision tree, the algorithm tests how well each feature splits the data.
+The features that split the data the most accurately, and then with the least impurity, would have the lowest impurity score.
+In Random Forests, a variable's importance can be calculated by looking at how much the tree nodes that use that variable reduce impurity on average.
+
+Gini impurity was chose on the grounds that this study is interested in not only classifying producers, but, more importantly, in understanding how the underlying data informs these decisions.
+By calculating the Gini impurity not only for the entire model, but specifically for each producer, we can better understand how the variables in our dataset describe the differences between producers.
+Without song structure, instruments, sound, or other more ephemeral components of style, these differences can be taken as a proxy for style.
+
+## Results
+
+### Model Performance
+
+### Stylistic Difference
+
+Correlation heatmaps can help us better understand the differences between each producer's "style". The variables with the most variance were danceability, loudness, and energy.
+
+```{r}
+#| echo: false
+#| warning: false
+## Range of correlation scores
+
+library(dplyr)
+
+range_data <- table_data %>%
+  gather(variable, value, -producer, -term) %>%
+  group_by(variable) %>%
+  summarize(
+    min_value = min(value, na.rm = TRUE),
+    max_value = max(value, na.rm = TRUE),
+    range = max_value - min_value
+  ) %>%
+  arrange(-range)
+
+##Variables with greatest variance##
+top_vars <- range_data %>% 
+  top_n(3, range) %>%
+  pull(variable)
+
+print(top_vars)
+
+selected_data <- table_data %>%
+  select(producer, term, all_of(top_vars)) %>%
+  arrange(producer, term)
+
+print(selected_data)
+
+library(ggplot2)
+```
+
+Similar to the logistic regression, danceability was also, across all producers, the most important predictor in the classification process.
+
+```{r}
+#| echo: false
+#| warning: false
+list_importance <- list()
+unique_producers <- unique(data_reduced$producer)
+
+for (producer in unique_producers) {
+  data_temp <- data_reduced
+  data_temp$producer_binary <- as.factor(ifelse(data_temp$producer == producer, 1, 0))
+
+  rf_temp <- randomForest(producer_binary ~ . - producer, data = data_temp, ntree = 100)
+
+  importance_temp <- rf_temp$importance
+  print(colnames(importance_temp))  # To see available columns
+
+  list_importance[[producer]] <- importance_temp[,'MeanDecreaseGini']
+}
+
+df_importance <- bind_rows(list_importance, .id = "producer")
+
+head(df_importance)
+
+df_importance_long <- df_importance %>%
+  gather(variable, importance, -producer)
+
+ggplot(df_importance_long, aes(x = reorder(variable, importance), y = importance)) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~producer, scales = "fixed") + 
+  coord_flip() +
+  labs(title = "Variable Importance by Producer", 
+       x = "Variable", 
+       y = "Mean Decrease Gini") +
+  theme_minimal()
+
+```
+
+In the visualization of danceability correlations across producers, Antonoff's danceability has the least positive correlation with valence (SCORE), the most negative correlation with acousticness (SCORE), and significant positive correlations with speechiness (SCORE), loudness (SCORE), and energy (SCORE).
+
+```{r}
+#| echo: false
+#| warning: false
+# Danceability Plot
+ggplot(selected_data, aes(x=term, y=danceability, fill=producer)) +
+  geom_bar(stat="identity", position="dodge") +
+  labs(title="Danceability across Producers", y="Danceability Score") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
 # Discussion
+
+Antonoff's danceability score being, compared to other producers, significantly positively correlated with speechiness, loudness and energy, negatively correlated with acousticness, and less than average with valence tracks with how critics tend to describe his music.
+
+His tracks have been described as anthemic (cite), with a distinct vocal treatment that makes it sound, as Gamman put it, "the way it would sound to the person who's singing them".
+
+> In the treble mix of the song, like in the very upper end, sort of above where the vocals are -- he often really crushes that down.
+It's lower in volume than anything else, which is sort of a weird effect.
+When you hear someone speaking, you hear a lot of that, like, noise in their voice.
+When you yourself are speaking, you hear less of it, right?
+You hear more of your own voice bouncing around in your head.
+And so it sort of creates this effect of -- **like he mixes his vocals the way it would sound to the person who's singing , them.
+Which is sort of strange.** He often has no vocals on that upper end, which is very unusual, and then often there are random little bits of noise happening up there which is the sort of thing her likes to do.
+
+Clearing noise and isolating the middle range of the vocals may make for a stronger "speechiness" feature.
+
+Valence, like danceability, was also a subjective variable built using labeling. Returning to the music itself, Antonoff-produced songs with the highest danceability scores aren't exactly dance hits.
+
+-   I Think He Knows by Taylor Swift:
+
+-   Vigilante Shit by Taylor Swift:
+
+    -   Not danceable.
+
+-   Heaven is Here by Florence and the Machine
+
+    -   Not danceable, but it does seem like it's trying to get your body to do something.
+
+-   I Wanna Get Better by Bleachers
+
+    -   More anthemic than danceable.
+It also seems like it's trying to coercing your attention, if not your body, and the speechiness, loudness, and energy are all apparent.
+The valence is apparent in the lyrics, but it doesn't have the kind of rhythm for dancing.
+
+-   Cornelia Street by Taylor Swift
+
+    -
 
@@
 
-The Spotify Features also wouldn't help you understand in the same way you end up understanding it if you sat on my friend Matthew Jordan's couch and watched him beat box a few different Jack Antonoff and Pharrell impressions in order to show the contrast.
-
 According to this model, though, danceability is the most important feature in predicting whether a song was produced by Jack Antonoff.
@@
 
-This was a feature that was invented in a lab \~15 years ago.
-This number refers to the time a bunch of undergraduates were paid (hopefully?) to listen to a bunch of songs and tag whether they were danceable or not.
-It is pretty much based on intuition.
-It captures this je-ne-sais-quoi essence of music that connects with listeners on a visceral level.
-Even if it doesn't make YOUR body want to move, you could imagine it making SOMEBODY'S body want to move.
+### In which I do a close listen of our model's top 5 most danceable songs[^3]:
 
-### In which I do a close listen of our model's top 5 most danceable songs[^5]:
-
-[^5]: I encourage you to follow along and drop me an email with your own assessment at michaela.drouillard\@mail.utoronto.ca.
+[^3]: I encourage you to follow along and drop me an email with your own assessment at michaela.drouillard\@mail.utoronto.ca.
 
@@
 
-    ![](images/Macefonian_phalanx.png)
-
 -   I Wanna Get Better by Bleachers
@@
     -   Yeah I'm dancing. Or at least bobbing my head. I only became aware of that fact when I made eye contact with the guy next to me at this cafe.
-
-## Draw on the random forest danceability correlations graph, tie it into the close listen
-
-## Danceability isn't even necessarily about Dancing! I propose: Visceral-ness
 

### 2023-10-31 outputs/paper/paper.qmd
added performance metrics table

@@
 ---
-title: "We Are All Trying to Describe Something And We Don't Have the Words"
-subtitle: "A Study of The Jack Antonoffification of Pop Music (Or: Propagating Uncertainty with Spotify)"
+title: "title"
+subtitle: "subtitle"
 author: 
@@
 
-# Mini literature review on Antonoff's place in culture? or should that go in the introduction
-
-# Either radically transform or cull the gamman or use it for the discussion section
-
-I reached out to Gamman to ask him to explain Antonoff to me, only to find out, on the morning of the interview, that Gamman had released a 29 minute video essay on his YouTube Channel covering [@citeGamman]:
-
-(you can skim the quote chunks)
-
-1.  what makes an Antonoff song so Antonoff-y
-
-    > ...I'm literally not an expert, and of course I don't actually claim to always be able to be able to 100% identify a Jack Antonoff song.
-It's not like I've been practicing the way that I can tell if he's produced something.
-It's just basically if it sounds to me like he's produced it.
-I know that's annoying I will clarify first, I think it's mostly in the vocal treatment, that's the thing that is the most off-putting to me.
-But just every decision feels a little bit wrong to me.
-If I wanted to be cringe I could point out the Juno 6 Bass, the perfectly stacked vocals, the Juno 6 pad, the wide and tight plate reverb, the Juno 6 sword, the goo goo gaga, the man hates treble unless it's just random crap swishing around, everything gets low passed with thin 707 drum samples, thin 808 drum samples, then 909 drum samples, thin Lin drum samples, thin thin drum samples, all the resonances, the harmonics get removed, everything is sort of squished down to the fundamental reduction in timbre, everything exists in a small frequency range, everything becomes a tone, it's an extremely consonant sound with no interplay between the different elements.
-Now that is a skill to do, and it produces a cleaner sound, and the only problem is I viscerally hate that sound.
-...
-
-2.  his relationship to music growing up (some gems in here)
-
-    > ...
-I got into drumming at the age of like six, so maybe as a drummer the drum samples are \[offensive?\].
-I was anachronistically raised on cassettes so, you know maybe I like the sound of audio and maybe that explains why I tend to like dissonance and inharmonic sound, interplay of elements, distortion noise, the clashing of elements.
-I imagine I've listened to more SoundCloud in my life than radio so uh, maybe I only want to hear amateur internet losers.
-I've liked over 10 000 songs on my Spotify account and I can identify lot of them off the bat, but I'm not that great at identifying music generally, so it's just a compartmentalized media consumption habit.
-I first did music production in primary school.
-I made mashups and dubstep.
-As a teenager I extensively used Juno 6 emulators to make embarrassing synthwave.
-\[Admittedly that's probably not the subconscious association that he's going for.\] Lyrics aren't of much interest to me, it's just a sound in the mix, and so maybe I don't like the focus on vocals.
-I've previously described \[my reaction\] as a fight or flight sort of thing, so maybe it's literally is tapping into some memory of trauma.
-I have an unusually small ear canal so maybe I'm just hearing the high end more acutely.
-I was hired as a teenager to play drums on a lot of Jack Antonoff songs for a university glee club so maybe I don't like it because of the embarrassment of having ever been associated with a glee club.
-I liked the album 1989 but to make the song Out of the Woods palatable to me I had to mash it up with M83Midnight City and that got a DMCA takedown so maybe it's a personal vendetta.
-And of course I have known and do know people who like this music so maybe I secretly hate my friends and family but I don't know if any of that tells us anything ...
-
-3.  An elegant and simple defense of taste
-
-    > ...I did use the phrase visceral hatred.
-But I did so only to describe my visceral hatred.
-I can't stand the sound of it.
-I hate it.
-And the way in which I hate it is viscerally.
-But I didn't go beyond that because it doesn't go beyond that.
-Nothing against the guy, I like Taylor Swift's stuff.
-It's just subjective taste...
-
-4.  His tongue in cheek and pretty based response to some of Twitter's negative response to his video
-
-    > .... \[it's just subjective taste\].
-There's just nothing to engage with, so I've had to fabricate a reason for controversy.
-And what I've come up with is this: \**Reads off of phone*\* "The real reason I've been saying for years that I have a visceral reaction to Jack Antonoff's production style, the REAL reason is this: because during the rollout cycle of the album *Midnights*, I wanted to ascribe Taylor Swift's authorship to a man and diminish women's voices." So actually, I'm a sexist.
-So now that's out there, published as facts, in the news.
-With all the other facts...
-
-5.  Basically a refutation to the entire premise of my interview
-
-    > .."What makes you so special and unique that you can tell the production of a song?" Nothing. ...
-
-6.
-Which was refuted even harder when I found a screenshot he tweeted of his tweet drafts in June 2021, one of them reading: "people often say to me:"caleb, what's wrong with jack antonoff?".
-don't ask questions"
-
-![](images/Screen%20Shot%202023-04-03%20at%209.31.05%20PM.png)
-
-> **Michaela Drouillard:** I know you cover this in the video. I'm sorry. What makes a Jack Antonoff song a Jack Antonoff song?
->
-> **Caleb Gamman:** I suppose I do have much more detail that I can go into, but it is always, for me, just a knee jerk.
-It's not any principled thing.
-I do have some loose experience with like, audio production stuff.
-So if I really dig into it, I can maybe point some things out.
-In the treble mix of the song, like in the very upper end, like sort of above where the vocals are.
-He often really crushes that down.
-Like it's lower in volume than anything else, which is sort of a weird effect.
-Like, when you hear someone speaking, you hear a lot of that, like, noise in their voice.
-When you yourself are speaking, you hear less of it, right?
-You hear more of your own voice bouncing around in your head.
-And so it sort of creates this effect of -- **like he mixes his vocals the way it would sound to the person who's singing them.
-Which is sort of strange.** He often has no vocals on that upper end, which is very unusual, and then often there are random little bits of noise happening up there which is the sort of thing her likes to do.
->
-> So I don't know, like I personally am quite a fan of messier music, you know, more experimental stuff, stuff that has a lot of clashing, sounds, a lot of different interesting pieces coming together.
-Jack Antonoff does a very good job of making stuff very clean.
-So you know, all of the notes that are happening simultaneously are perfectly stacked.
-They're like mathematically stacked ..
-Like, you have the harmonic spectrum where everything is like a multiple on top of each other.
-So that's like your notes on a keyboard.
-And if you played a keyboard note that was a little bit off where it's supposed to be, you're gonna create this interesting effect where these two notes sort of clash against each other and they wiggle around in terms of their tone because you've got these two harmonics happening that don't match.
-And Jack Antonoff avoids that quite strictly.
->
-> His drums are low pass, like the vocals, so you don't get like the tinny high end in the drums or whatever because that tinny high end in the drums is like a bunch of random harmonics all over the place.
-So if you cut that out, you just get the tone of the drum.
-And then he achieves those effects using the same sounds like literally the same synthesizer on the same drum machine and stuff.
->
-> ....
->
-> If you used a cscope spectrogram sort of thing to look at a Jack Antonoff song, you're gonna see a lot of big lines.
-And then if you're listening to a song by someone else, you're gonna see like a bunch of sort of like static wiggling.
-It's gonna be really jagged because you've got a bunch of harmonics that don't exactly mix with each other, so they're stacking in all sorts of weird different ways.
+# Lit Review
 
@@
 
+## Interview with Glen McDonald
+
 I interviewed Glen McDonald, a Principal Engineer at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
@@
 And then danceability and valence are purely subjective."
-
-This put the big range of valence scores for Flight of The Bumblebee into perspective -- what that value is capturing is abstracted away from the actual notes of the song.
-The valence feature propagates a vibe, so of course this uncertainty makes room for noise and variation.
 
@@
 
+## Data Preparation
+
 I made a binary response variable, called `jack`, that has a value of 1 if the song was written by Jack, and a 0 otherwise.
@@
 
-Bracketing the major blindspot that these variables provide a single number to represent the dynamics of an entire song, the model has a few other limitations.
-First of all, this model assumes that the relationship between the independent variables and the dependent variable is linear.
+This model assumes that the relationship between the independent variables and the dependent variable is linear.
 There could be non-linear relationships between the variables that this model isn't capturing.
@@
 
-One issue is that in our dataset, there are a lot more songs that aren't produced by Jack than songs that are:
+One issue is that in our dataset, there are a lot more songs that aren't produced by Jack than songs that are. The classes have to be balanced.
 
@@
 ```
-
-So, we have to balance the classes.
 
@@
 
-# Results
+## Results
 
@@
 ### Model Performance
+
+| Metric                  | Value           |
+|-------------------------|-----------------|
+| Accuracy                | 0.6111          |
+| 95% Confidence Interval | (0.579, 0.6425) |
+| No Information Rate     | 0.4274          |
+| P-Value (Acc \> NIR)    | \<2.2e-16       |
+| Kappa                   | 0.4223          |
+
+: Overall Performance Statistics
 

### 2023-10-31 scripts/06-models.R

@@ -366,6 +366,7 @@ ggplot(selected_data, aes(x=term, y=danceability, fill=producer)) +
   geom_bar(stat="identity", position="dodge") +
   labs(title="Danceability across Producers", y="Danceability Score") +
   theme_minimal() +
+  facet_wrap(nrow = term)
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 # Energy Plot

### 2023-11-01 outputs/paper/paper.qmd
LR model section improved. Using original data instead of downsampled since performance is way better

@@
   - Michaela Drouillard
-thanks: "Code and data are available at: https://github.com/michaeladrouillard/jack."
+thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop."
 date: today
@@
 
-I made a binary response variable, called `jack`, that has a value of 1 if the song was written by Jack, and a 0 otherwise.
+The binary response variable, called `is_antonoff`, has a value of 1 if the song was produced by Antonoff, and a 0 otherwise.
 There are 10 predictor variables, all of which are continuous: **`danceability`**, **`energy`**, **`loudness`**, **`mode`**, **`speechiness`**, **`acousticness`**, **`instrumentalness`**, **`liveness`**, **`valence`**, and **`tempo`**.
@@
 
-where $\text{Pr}(y_i=1)$ is the probability of a song being written by Jack, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the predictor variables.
+where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the predictor variables.
 
-@fig-means gives us an idea of the mean value of each of these features for Jack and non-Jack song.
-A logistic regression will give us a better idea of how to weigh each of these features in order to predict the outcome, which gives us information about the "research" "question": are these characteristics related to Jack's music?
-(And then: Which ones?
-And then, the most elusive question: How?)
+@fig-means gives us an idea of the mean value of each of these features for Antonofff and non-Antonoff-produced tracks.
 
@@
 
-data_reduced <- read_csv(here::here("inputs/data/data_reduced.csv"))
+lr_data <- read_csv(here::here("inputs/data/lr_data.csv"))
 
-data_reduced |>
-  group_by(jack) |>
+lr_data |>
+  group_by(is_antonoff) |>
   summarise(across(-c(tempo, loudness), mean)) |>
-  pivot_longer(cols = -jack) |>
-  ggplot(aes(value, name, fill = factor(jack))) +
+  pivot_longer(cols = -is_antonoff) |>
+  ggplot(aes(value, name, fill = factor(is_antonoff))) +
   geom_col(alpha = 0.8, position = "dodge") +
@@
 ```
-
-This model assumes that the relationship between the independent variables and the dependent variable is linear.
-There could be non-linear relationships between the variables that this model isn't capturing.
-Second, it assumes that the relationship between the independent variables and the dependent variable is the same for all observations, which might not be true in practice -- there could be interactions between the variables, or in certain subgroups of data where the relationship is different.
-Albums where Jack may have had more of a back seat producing role are one example.
-(The fact that a song's development, from inception to recording, is a whole creative process, each having its own origin story with varying degrees of Jack-ness, is another.)
 
@@
 
-I used **`rank_results()`** from the **`yardstick`** package to rank the models based on accuracy, log loss, sensitivity, and specificity, and went with the downsampled model.
+I used **`rank_results()`** from the **`yardstick`** package to rank the models based on accuracy, log loss, sensitivity, and specificity.
+After testing both basic and downsampled versions of the model, the basic version yielded much higher accuracy score (0.918 basic versus 0.743 downsampled), although a lower roc auc score (0.746 basic versus 0.821 downsampled).
+
+The version with the original data remains consistently high in accuracy, even with increased regularization.
+It starts low with in mean log loss and sensitivity, but increases with increased regularization.
+Compared to the downsampled data, which starts high but drops significantly in accuracy and sensitivity, we chose to use the original data.
+
+![Comparing the Performance for Downsampled and Original Data](images/downsamplevsorig)
 
@@
 
-The model's accuracy score is 0.743, and the ROC AUC score is 0.841.
-The accuracy score could be misleading, since the model had imbalanced classes, and downsampling may have affected the model's ability to correctly identify the minority class.
-The ROC AUC score is more reliable, since it's less sensitive to class imbalance.
+The model's accuracy score is 0.918, and the ROC AUC score is 0.746.
 
@@
 |------------------------|-----------------------|-----------------------|
-| Predicted Positive (1) | 581                   | 25                    |
-| Predicted Negative (0) | 194                   | 51                    |
+| Predicted Positive (1) | 848                   | 76                    |
+| Predicted Negative (0) | 1                     | 0                     |
 
@@
 #| label: tbl-importance
-#| tbl-cap: Importance Scores of Each Variable in Determining Jack-ness
+#| tbl-cap: Importance Scores of Each Variable in Determining Antonoff's Tracks
 #| echo: false
@@
 
-jack_vip <- read_csv(here::here("inputs/data/jack_vip.csv"))
+antonoff_vip <- read_csv(here::here("inputs/data/antonoff_vip.csv"))
 
-jack_vip |>
+antonoff_vip |>
   knitr::kable(col.names = c("Variable",
@@
 #| label: fig-importance
-#| fig-cap: Importance Scores of Each Variable in Determining Jack-ness
+#| fig-cap: Importance Scores of Each Variable in Determining Antonoff's tracks
 #| echo: false
@@
 
-jack_vip |>
+antonoff_vip |>
   group_by(Sign) |>
@@
 
-Danceability, with a positive importance score of 4.32, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
-Instrumentalness comes in second at 2.59, with a negative score (meaning that negative instrumentalness is correlated), and valence comes in third with a negative score at 2.02.
+Danceability, with a positive importance score of 4.142, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
+Valence comes in second at 2.559, with negative score.
+The third, instrumentalness correlates negatively with Antonoff tracks with a a negative score of 2.309.
 
@@
 
-df |>
-  filter(jack == 1) |>
+clean_df <- read_csv(here::here("inputs/data/clean_df.csv"))
+
+clean_df |>
+  filter(producer == "antonoff") |>
   arrange(desc(danceability)) |>
@@
 
-The categorical feature 'producer', which contains the names of the producer for each track, was one-hot encoded to create binary columns for each producer ('is_antonoff', 'is_elworth', etc.).
+The categorical feature `producer`, which contains the names of the producer for each track, was one-hot encoded to create binary columns for each producer (i.e.
+`is_antonoff`, `is_elworth`, etc.).
 The producer feature was then converted into a factor variable, and the dataset was split into training and tests sets.
@@
 
-The RF was trainied using the @ranger package. The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
+The RF was trained using the @ranger package. The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
 
@@
 
-Gini impurity was chose on the grounds that this study is interested in not only classifying producers, but, more importantly, in understanding how the underlying data informs these decisions.
+Gini impurity was chosen on the grounds that this study is interested in not only classifying producers, but, more importantly, in understanding how the underlying data informs these decisions.
 By calculating the Gini impurity not only for the entire model, but specifically for each producer, we can better understand how the variables in our dataset describe the differences between producers.
@@
 : Overall Performance Statistics
+
+The accuracy of the model (0.6111) being significantly higher than the No Information Rate (0.4274) means that it is doing more than just blind guessing based on class prevalence.
+With a p-value below 0.05, we can confirm the model's statistical significance over a naive approach.
+
+The model's accuracy and Kappa statistics suggest a moderate level of performance.
+However, there is enough demonstrated statistical significance to suggest that there are underlying patterns in the data that differentiate producers.
 

### 2023-11-01 scripts/06-models.R

@@ -13,6 +13,8 @@ library(tidymodels)
 library(randomForest)
 library(ranger)
 library(caret)
+library(themis)
+library(yardstick)
 
 data <- read_csv(here::here("inputs/data/clean_df.csv"))
 
@@ -29,24 +31,24 @@ write_csv(data_reduced,"inputs/data/data_reduced.csv")
 
 #### Predicting Antonoff ####
 ## glmnet requires all variables to be numeric, so i'm excluding the producer column
-
-# 1. Create a new binary column
-data_reduced$is_antonoff <- ifelse(data_reduced$producer == "antonoff", 1, 0)
 set.seed(853)
+# 1. Create a new binary column
+lr_data <- data_reduced
+lr_data$is_antonoff <- ifelse(lr_data$producer == "antonoff", 1, 0)
+lr_data$is_antonoff <- as.factor(lr_data$is_antonoff)
 
-data_reduced$is_antonoff <- as.factor(data_reduced$is_antonoff)
+
+write_csv(lr_data, "inputs/data/lr_data.csv")
 
 # 2. Update variable names and strata
 antonoff_split <- 
-  data_reduced |>
+  lr_data |>
   initial_split(strata = is_antonoff)
 
 antonoff_train <- training(antonoff_split)
 antonoff_test <- testing(antonoff_split)
-
-set.seed(853)
 antonoff_folds <- vfold_cv(antonoff_train, strata = is_antonoff)
-antonoff_folds
+
 
 # 3. Update the recipe
 is_antonoff <- 
@@ -57,12 +59,12 @@ is_antonoff <-
 # we can `prep()` just to check that it works:
 prep(is_antonoff)
 
+
+
+
 glmnet_spec <- 
   logistic_reg(penalty = tune(), mixture = 1) |>
   set_engine("glmnet")
-
-# install.packages("themis")
-library(themis)
 
 # Comparing the basic model and the downsampled one
 wf_set <-
@@ -79,15 +81,13 @@ narrower_penalty <- penalty(range = c(-3, 0))
 # install.packages("doParallel")
 doParallel::registerDoParallel()
 
-set.seed(853)
-
 tune_rs <- 
   workflow_map(
     wf_set,
     "tune_grid",
     resamples = antonoff_folds,
     grid = 15,
-    metrics = metric_set(accuracy, mn_log_loss, sensitivity, specificity),
[…]

### 2023-11-01 outputs/paper/paper.qmd
trying to get things to render

@@
 
-sub_planets <- read_csv(here::here("inputs/data/subset_planets.csv"))
+#sub_planets <- read_csv(here::here("inputs/data#/subset_planets.csv"))
 
-sub_planets |>
-  select(artist_name, valence, danceability, track_name) |>
-  head() |>
-  knitr::kable(col.names = c("Artist Name",
-                             "Valence",
-                             "Danceability",
-                             "Track Name"))
+#sub_planets |>
+  #select(artist_name, valence, danceability, track_name) |>
+  #head() |>
+  #knitr::kable(col.names = c("Artist Name",
+                             #"Valence",
+                             #"Danceability",
+                             #"Track Name"))
 ```
@@
 
-```{r}
-#| echo: false
-#| warning: false
-#| label: tbl-bees
-#| tbl-cap: Valence and Danceability Scores for Every Performance of Flight of the Bumblebee on Spotify
-
-sub_bees <- read_csv(here::here("inputs/data/subset_bumblebee.csv"))
-
-sub_bees |>
-  select(artist_name,
-         valence,
-         danceability,
-         album_release_year,
-         track_name) |>
-  head(17) |>
-  knitr::kable(
-    col.names = c(
-      "Artist Name",
-      "Valence",
-      "Danceability",
-      "Album Release Year",
-      "Track Name"
-    )
-  )
+```{}
 ```
@@
 
-df <- read_csv(here::here("inputs/data/df.csv"))
-
-ggplot(df, aes(x = energy, y = valence)) +
-  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$valence), color = "gray") +
-  theme_minimal()
+# df <- read_csv(here::here("inputs/data/clean_df.csv"))
+# 
+# ggplot(df, aes(x = energy, y = valence)) +
+#   geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
+#   geom_vline(xintercept = mean(df$energy), color = "gray") +
+#   geom_hline(yintercept = mean(df$valence), color = "gray") +
+#   theme_minimal()
 ```
@@
 #| warning: false
-
-ggplot(df, aes(x = energy, y = danceability)) +
-  geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$danceability),
-             color = "gray") +
-  theme_minimal()
+# 
+# ggplot(df, aes(x = energy, y = danceability)) +
+#   geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
+#   geom_vline(xintercept = mean(df$energy), color = "gray") +
+#   geom_hline(yintercept = mean(df$danceability),
+#              color = "gray") +
+#   theme_minimal()
 ```
@@
 
-data_reduced <- read_csv(here::here("inputs/data/data_reduced.csv"))
-
-data_reduced |>
-  pivot_longer(
-    cols = c(
-      danceability,
-      energy,
-      loudness,
-      mode,
-      speechiness,
-      acousticness,
-      instrumentalness,
-      liveness,
-      valence,
-      tempo
-    ),
-    names_to = "name",
-    values_to = "value"
-  ) |>
-  ggplot(aes(x = value, color = jack)) +
-  stat_ecdf() +
-  facet_wrap(vars(name),
-             nrow = 3,
-             ncol = 4,
-             scales = "free") +
-  theme_minimal() +
-  labs(x = "Value",
-       y = "Proportion")
+# data_reduced <- read_csv(here::here("inputs/data/data_reduced.csv"))
+# 
+# data_reduced |>
+#   pivot_longer(
+#     cols = c(
+#       danceability,
+#       energy,
+#       loudness,
+#       mode,
+#       speechiness,
+#       acousticness,
+#       instrumentalness,
+#       liveness,
+#       valence,
+#       tempo
+#     ),
+#     names_to = "name",
+#     values_to = "value"
+#   ) |>
+#   ggplot(aes(x = value, color = jack)) +
+#   stat_ecdf() +
+#   facet_wrap(vars(name),
+#              nrow = 3,
+#              ncol = 4,
+#              scales = "free") +
+#   theme_minimal() +
+#   labs(x = "Value",
+#        y = "Proportion")
 
@@
 
-df <- read_csv(here::here("inputs/data/df.csv"))
-
-df <- df |> 
-  mutate(jack = factor(jack))
-
-df |>
-  filter(album_release_date_precision == "day") |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(y = energy, x = album_release_date, color = jack)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(method = lm, se = FALSE) +
-  theme_minimal() +
-  scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
-  facet_wrap(~ artist_name, ncol = 3)
+# df <- read_csv(here::here("inputs/data/df.csv"))
+# 
+# df <- df |> 
+#   mutate(jack = factor(jack))
+# 
+# df |>
+#   filter(album_release_date_precision == "day") |>
+#   mutate(album_release_date = ymd(album_release_date)) |>
+#   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
+#   geom_point(alpha = 0.3) +
+#   geom_smooth(method = lm, se = FALSE) +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
+#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
+#   facet_wrap(~ artist_name, ncol = 3)
 ```
@@
 
-df |>
-  filter(album_release_date_precision == "day") |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(method = lm, se = FALSE) +
-  theme_minimal() +
-  scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
-  facet_wrap( ~ artist_name, ncol = 3)
+# df |>
+#   filter(album_release_date_precision == "day") |>
+#   mutate(album_release_date = ymd(album_release_date)) |>
+#   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
+#   geom_point(alpha = 0.3) +
+#   geom_smooth(method = lm, se = FALSE) +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
+#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
+#   facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-df |>
-  filter(album_release_date_precision == "day") |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(y = valence, x = album_release_date, color = jack)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(method = lm, se = FALSE) +
-  theme_minimal() +
-  scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
-  facet_wrap( ~ artist_name, ncol = 3)
+# df |>
+#   filter(album_release_date_precision == "day") |>
+#   mutate(album_release_date = ymd(album_release_date)) |>
+#   ggplot(aes(y = valence, x = album_release_date, color = jack)) +
+#   geom_point(alpha = 0.3) +
+#   geom_smooth(method = lm, se = FALSE) +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
+#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
+#   facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-df |>
-  filter(album_release_date_precision == "day") |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(method = lm, se = FALSE) +
-  theme_minimal() +
-  scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
-  facet_wrap( ~ artist_name, ncol = 3)
+# df |>
+#   filter(album_release_date_precision == "day") |>
+#   mutate(album_release_date = ymd(album_release_date)) |>
+#   ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
+#   geom_point(alpha = 0.3) +
+#   geom_smooth(method = lm, se = FALSE) +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
+#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
+#   facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-df |>
-  filter(album_release_date_precision == "day") |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(method = lm, se = FALSE) +
-  theme_minimal() +
-  scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
-  facet_wrap( ~ artist_name, ncol = 3)
+# df |>
+#   filter(album_release_date_precision == "day") |>
+#   mutate(album_release_date = ymd(album_release_date)) |>
+#   ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
+#   geom_point(alpha = 0.3) +
+#   geom_smooth(method = lm, se = FALSE) +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
+#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
+#   facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-df |>
-  filter(album_release_date_precision == "day") |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth(method = lm, se = FALSE) +
-  theme_minimal() +
-  scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
-  facet_wrap( ~ artist_name, ncol = 3)
+# df |>
+#   filter(album_release_date_precision == "day") |>
+#   mutate(album_release_date = ymd(album_release_date)) |>
+#   ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
+#   geom_point(alpha = 0.3) +
+#   geom_smooth(method = lm, se = FALSE) +
+#   theme_minimal() +
+#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
+#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
+#   facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 #| warning: false
+# 
+# data_reduced |>
+#   count(jack) |>
+#   knitr::kable(col.names = c("Is it Jack?", "Number of Songs"))
 
-data_reduced |>
-  count(jack) |>
-  knitr::kable(col.names = c("Is it Jack?", "Number of Songs"))
 ```
@@
 
+table_data <- read_csv(here::here("inputs/data/table_data.csv"))
+
 range_data <- table_data %>%
@@
 
-print(top_vars)
+print(range_data)
 
-selected_data <- table_data %>%
-  select(producer, term, all_of(top_vars)) %>%
-  arrange(producer, term)
-
-print(selected_data)
-
-library(ggplot2)
 ```
@@
 #| warning: false
-list_importance <- list()
-unique_producers <- unique(data_reduced$producer)
 
-for (producer in unique_producers) {
-  data_temp <- data_reduced
-  data_temp$producer_binary <- as.factor(ifelse(data_temp$producer == producer, 1, 0))
-
-  rf_temp <- randomForest(producer_binary ~ . - producer, data = data_temp, ntree = 100)
-
-  importance_temp <- rf_temp$importance
-  print(colnames(importance_temp))  # To see available columns
-
-  list_importance[[producer]] <- importance_temp[,'MeanDecreaseGini']
-}
-
-df_importance <- bind_rows(list_importance, .id = "producer")
-
-head(df_importance)
+df_importance<- read_csv(here::here("inputs/data/df_importance.csv"))
 
@@
 # Danceability Plot
+
 ggplot(selected_data, aes(x=term, y=danceability, fill=producer)) +
@@
 
-It only makes sense, of course, to hire a bunch of undergrads and get them to tell us whether a few thousand songs are visceral, and then build our own visceral-ness metric, and see if we get the same results -- that danceability is just visceral-ness in a trench coat, and that visceral-ness is the characteristic that helps us best understand what makes Jack, Jack.
-
-And if we're doing this for danceability, there might also be something worth investigating in valence, the feature that indicates a song's "positivity", which was also built via college interns' intuitions.
-
 \newpage

### 2023-11-01 scripts/06-models.R

@@ -223,6 +223,9 @@ head(df_importance)
 df_importance_long <- df_importance %>%
   gather(variable, importance, -producer)
 
+write.csv(df_importance, file = "inputs/data/df_importance.csv", row.names = FALSE)
+
+
 ggplot(df_importance_long, aes(x = reorder(variable, importance), y = importance)) +
   geom_bar(stat = "identity") +
   facet_wrap(~producer, scales = "fixed") + 
@@ -305,6 +308,7 @@ table_data <- melted_corr_all %>%
   select(-term) %>%
   spread(key = variable, value = value)
 
+write_csv(table_data, "inputs/data/table_data.csv")
 
 producer_specific_table <- function(producer_name) {
   table_data %>% filter(producer == producer_name)
@@ -359,6 +363,8 @@ print(top_vars)
 selected_data <- table_data %>%
   select(producer, term, all_of(top_vars)) %>%
   arrange(producer, term)
+
+write_csv(selected_data, "inputs/data/selected_data.csv")
 
 print(selected_data)
 

### 2023-11-01 outputs/paper/paper.qmd
paper edits

@@
 What gives?
-
-```{}
-```
 
@@
 
-![Comparing the Performance for Downsampled and Original Data](images/downsamplevsorig)
+![](images/lr_ds_orig.jpeg)
 
@@
 
+selected_data <- read_csv(here::here("inputs/data/selected_data.csv"))
+
 ggplot(selected_data, aes(x=term, y=danceability, fill=producer)) +
@@
 
--   I Think He Knows by Taylor Swift:
+I Think He Knows by Taylor Swift, plainly, seems danceable to me.
 
--   Vigilante Shit by Taylor Swift:
+However, personally, I don't hear Taylor Swift's Vigilante Shit or Florence and the Machine's Heaven is Here as danceable.
+They aren't understated, ambient tracks either -- there's something coercing your body to do something.
+Like Gamman's description of Antonoff's vocal design sounding how they sound to person who's singing them, the tracks seem like they're trying to get inside of your skin.
+But they don't have the rhythm of a dance track.
 
-    -   Not danceable.
+In a lot of ways, I Wanna Get Better by Bleachers is very similar to the typical Antonoff track that the outputs of our RF describe.
+This song is more anthemic, closer to the driving-over-the-Brooklyn-bridge, sort of post-Springsteen emotions that Antonoff is often characterized (or admitting) to striving for, than danceable.
+It also seems like it's trying to coercing your attention, if not your body, and the speechiness, loudness, and energy which are so correlated with danceability in Antonoff's RF results are all apparent.
+The redemption, broken-but-getting-better-quasi-valence can also be heard in the lyrics, but it still doesn't have the kind of rhythm for dancing.
 
--   Heaven is Here by Florence and the Machine
-
-    -   Not danceable, but it does seem like it's trying to get your body to do something.
-
--   I Wanna Get Better by Bleachers
-
-    -   More anthemic than danceable.
-It also seems like it's trying to coercing your attention, if not your body, and the speechiness, loudness, and energy are all apparent.
-The valence is apparent in the lyrics, but it doesn't have the kind of rhythm for dancing.
-
--   Cornelia Street by Taylor Swift
-
-    -
+Cornelia Street by Taylor Swift is another sort of almost-danceable track -- while the melody is catchy, the lyrics themselves are wistful. The vocal treatment, again, is there --
 
@@
 
-#### Could our model just be predicting Jack via Taylor Swift?
-
-So another issue is that there is a lot of Taylor Swift in our training data.
-As mentioned earlier, her songs take up 40% of @tbl-mostdanceable, which is 5% more than Antonoff's own band, Bleachers.
-Taylor Swift has the second highest average danceability score across all songs, and the most danceable songs are hers.
-She also happens to have a lot of collaborations with Antonoff under her belt -- more than any other artist.
-I suspect that perhaps our model was predicting Jack via Taylor, which could be a beautiful story about collaboration and aesthetic affinity.
-Or it could suggest that this model has a stink to it.
-An avenue for future study.
-
-```{r}
-#| echo: false
-#| warning: false
-
-df |>
-  filter(jack == 1) |>
-  ggplot(aes(x = artist_name, y = danceability)) +
-  geom_boxplot() +
-  ggtitle("Boxplot of Danceability by Artist") +
-  coord_flip() +
-  theme_minimal()
-```
-
-```{r}
-#| echo: false
-#| warning: false
-
-df |>
-  filter(jack == 1) |>
-  count(artist_name) |>
-  arrange(desc(n)) |>
-  head(10) |>
-  ggplot(aes(x = reorder(artist_name, n), y = n)) +
-  geom_bar(stat = "identity", fill = "steelblue") +
-  coord_flip() +
-  labs(title = "Artists with Most Jack = 1 Values", x = "Artist", y = "Count") +
-  theme_minimal()
-
-```
-
 #### Next Steps

### 2023-11-01 scripts/06-models.R

@@ -375,7 +375,6 @@ ggplot(selected_data, aes(x=term, y=danceability, fill=producer)) +
   geom_bar(stat="identity", position="dodge") +
   labs(title="Danceability across Producers", y="Danceability Score") +
   theme_minimal() +
-  facet_wrap(nrow = term)
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 # Energy Plot

### 2023-11-01 outputs/paper/paper.qmd
more paper edits

@@
 date: today
-abstract: "Using data from the Spotify API, I explored Jack Antonoff's production style and assessed whether Spotify's audio features can capture it.
-In an interview with Glen McDonald, an engineer at Spotify, I learned that the danceability and valence features were created by having interns tag songs as positive or gloomy, or danceable or un-danceable.
-Using a logistic regression model, I found that danceability was the most important feature in predicting Antonoff's involvement in a song.
-In the discussion section, I do a close-listen of Antonoff's most danceable songs, and argue that the danceability variable is actually capturing something more like a visceral quality to Antonoff's music."
+abstract: ""
 format: pdf
@@
 
-Shortly after the release of Taylor Swift's *Midnights* album, a video of the Youtuber Caleb Gammen, in which guesses which songs on Midnights were produced by Antonoff within seconds [@citeTweet], went viral.
+Shortly after the release of Taylor Swift's *Midnights* album, a video of the Youtuber Caleb Gammen, in which he guesses which songs on Midnights were produced by Antonoff within seconds [@citeTweet], went viral.
 The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style".
@@
 As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely.
-When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis" [@citeMarantz].
+When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis" [@citeMarantz].Can Spotify's audio features dataset help us determine what makes Antonoff music so Antonoff-y?
 
-(Cite a few snippets from critics articles to give a dose of the discourse)
+I spoke to Gamman to better undertstand makes a song Antonoff-ish, and to try to understand his reaction to it.
+In trying to understand how to describe Antonoff's sound, I used the Spotify API's audio features dataset to explore difference machine listening metrics, and build a model to recreate a digital version of Gamman's classification excercise.
+I also spoke with Glen McDonald, an engineer at Spotify, from whom I learned that the `valence` and `danceability` variables in the Spotify API were created by hiring college interns and having them tag whether songs were positive or danceable.
 
-Can Spotify's audio features dataset help us determine what makes Antonoff music so Antonoff-y?
+I used two models to investigate the data and ask the question: What is Antonoff's sound? Can we pin down style using the quantitative metrics offered by Spotify?
 
-I spoke to Caleb Gamman, a Youtuber and Antonoff-production-style hater who had a tweet go viral around the time of Taylor Swift's *Midnights* release, in which he predicts whether a song was produced by Antonoff within seconds.
-
-During an interview with Glen McDonald, an engineer at Spotify, I learned that the `valence` and `danceability` variables were created by hiring college interns and having them tag whether songs were positive or danceable.
-
-I used two models to investigate the data.
-The first, a logistic regression, draws on the dataset of artists who've collaborated with Antonoff, with an extra binary variable to tag whether or not the song was an Antonoff production.
+The first model, a logistic regression, draws on the dataset of artists who've collaborated with Antonoff, with an extra binary variable to tag whether or not the song was an Antonoff production.
 With this model, we can get a first glimpse at variables that are stronger predictors of Antonoff-iness, along with a better understanding of whether the Spotify data can trace Antonoff's influence on an artist's sound over time.
@@
 We also cannot assume that an artists' style can be measured in the linear growth of these metrics.
-Although the logistic regression has a strong accuracy, indicating that danceability and loudness are the strongest indicators of an Anotonoff produced song, it may be detecting a change in genre, rather than a change in style, across artists whose sound turns more into pop while collaborating with Antonoff.
+Although the logistic regression has a strong accuracy, it may be detecting a change in genre, rather than a change in style, across artists whose sound turns more and more pop while collaborating with Antonoff.
 
 I built a second model, a Random Forest, to complement the insights drawn from the logistic regression.
-The Random Forest trains on the full data set, including other producers.
-Random Forests (what's the package) don't assume linear relationships, and aren't sensitive to multicollinearity.
-In fact, using the x package in R, we can visualize and analyse the relationships between different variables.
-The variance in correlation scores across producers can be taken as an approximate for style, since this variance helps the model classify tracks.
+The Random Forest trains on the full data set, including other major pop producers.
+Random Forests don't assume linear relationships, and aren't sensitive to multicollinearity.
+In fact, using the tk package in R, we can visualize and analyse the relationships between different variables.
+The variance in correlation scores across producers can be taken as an approximation for style, since this variance helps the model distinguish between producers.
 
 In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing something like visceral-ness.
-Against the odds, we can capture something about Antonoff's music with these variables, and it might just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe.
+Against the odds, we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
 
-# Lit Review
+# tk: Lit Review
 
@@
 >
+> -   **speechiness:** Speechiness detects the presence of spoken words in a track.
+The more exclusively speech-like the recording (e.g.
+talk show, audio book, poetry), the closer to 1.0 the attribute value.
+Values above 0.66 describe tracks that are probably made entirely of spoken words.
+Values between 0.33 and 0.66 describe tracks that may contain both music and speech, either in sections or layered, including such cases as rap music.
+Values below 0.33 most likely represent music and other non-speech-like tracks.
+>
 > -   **key**: The key the track is in.
@@
 
-It's hard to gauge exactly what the variables "danceability" or "valence" mean, even using the documentation.
-For instance, "valence" is described as "A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
+It's hard to gauge exactly what the variables `danceability` or `valence` mean, even using the documentation.
+For instance, `valence` is described as "A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
 Tracks with high valence sound more positive (e.g.
@@
 sad, depressed, angry)".
-"Danceability" is described as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+`Danceability` is described as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
 A value of 0.0 is least danceable and 1.0 is most danceable".
-Neither of these definitions fully cover what metrics and thresholds go into something like "danceability".
-However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums[^1].
-
-[^1]: \-- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events".
-
-This is pretty vague.
-Especially valence -- is the positivity related to lyrics, or something else?
-I compared the valence scores of classical music tracks to see if this score is based on lyrics, or the just the instruments.
-In @tbl-Holst, Gustav Holst's Jupiter has a valence score 326.78% higher than Mars.
-This tracks.
-It suggests the valence is based on audio features other than lyrical content.
-It also suggests that it is, in fact, capturing a difference between the two songs' profiles.
-
-```{r}
-#| echo: false
-#| warning: false
-#| label: tbl-Holst
-#| tbl-cap: Valence and Danceability Scores for Holst's Mars and Jupiter
-
-#sub_planets <- read_csv(here::here("inputs/data#/subset_planets.csv"))
-
-#sub_planets |>
-  #select(artist_name, valence, danceability, track_name) |>
-  #head() |>
-  #knitr::kable(col.names = c("Artist Name",
-                             #"Valence",
-                             #"Danceability",
-                             #"Track Name"))
-```
-
-I looked at the valence measures on Nikolai Rimsky-Korsakov's Flight of the Bumblebee, a song that's much faster and frantic than Mars and Jupiter.
-What I found complicated my understanding of valence -- different performances of the same song had different valence scores (@tbl-bees).
-Some were arias, played by different instruments.
-Arias tended to have higher valence scores.
-The lowest recorded valence across all performances is 0.2410, and the highest is .9230.
-What gives?
+Neither of these definitions fully cover what metrics and thresholds go into something like `danceability`.
 
@@
 Variables like energy or instrumentalness, on the other hand, were determined primarily through machine listening techniques.
-The human subjectivity part only came in when it came to tweaking the training data[^2]:
+The human subjectivity part only came in when it came to tweaking the training data[^1]:
 
-[^2]: My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
+[^1]: My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
 Because there were no banjos in the training data, they were registered as human speech.
@@
 
-In @fig-quadvalence, it looks like songs produced by Jack Antonoff are a little more clustered towards the lower end of valence, and across the spectrum of energy, although the variance is large.
+In @fig-quadvalence, it looks like songs produced by Antonoff are a little more clustered towards the lower end of valence, and across the spectrum of energy, although the variance is large.
 
@@
 
-In @fig-quaddanceability2, it looks like Jack-produced songs err towards higher danceability, across the spectrum of energy.
+In @fig-quaddanceability2, it looks like Antonoff-produced songs err towards higher danceability, across the spectrum of energy.
 
-These quadrants are a start, but they don't suggest any defining Jack-iness characteristic underlying each of those blue points, which are scattered pretty broadly.
+These quadrants are a start, but they don't suggest any defining Antonoff-iness characteristic underlying each of those blue points, which are scattered pretty broadly.
 
-By comparing cumulative distribution functions of each variable based on whether a song was produced by Antonoff, we can see that some of these variables do capture some sort of underlying Jack-iness (@fig-cdf).
+By comparing cumulative distribution functions of each variable based on whether a song was produced by Antonoff, we can see that some of these variables do capture some sort of underlying Antonoff-iness (@fig-cdf).
 For instance, while liveness, instrumentalness, and speechiness are relatively similar, we can see clear differences in danceability, energy, loudness, acousticness, and valence.
@@
 Another way to gauge the Antonoff-ification of pop music could be to look at how artists' sounds have changed over time, and whether there is any correlation with Antonoff.
-Here is where the necessary "correlation is not causation" caveat comes in: suggesting that Antonoff is the *cause* for any change in artists' sounds would be an overstep.
 These plots are here to get a glimpse at pop's evolution over time, and where Antonoff fits in that picture.
@@
 
-In @fig-energy, it looks like there's a downward trend in energy scores in artists who work with Antonoff, but that these downward trends are also somewhat in line with the artists' energy trajectories anyway.
+In @fig-energy, it looks like there's a downward trend in energy scores in artists who work with Antonoff, but that these downward trends are also somewhat in line with the artists' energy trajectories anyway, which could be indicative of broader trends in pop.
 Specifically, Lana Del Rey and Taylor Swift.
@@
 In @fig-valence, it looks like Lorde has the only significant change in valence scores, which tracks with my understanding her albums.
-Melodrama had Green Lights, and Solar Power had a music video of Lorde frolicking on a beach in a yellow bikini with Phoebe Bridgers and Clairo.
+*Melodrama* had piano and break ups, and *Solar Power* had a music video of Lorde prancing on the beach in a yellow bikini with Phoebe Bridgers and Clairo.
 
@@
 
-Again in @fig-acousticness, Lorde's acoustic Solar Power album jumps out at us.
-Taylor Swift, Lorde and Haim also have significant-seeming upward trends in acousticness, while Mitski, Sharon Von Etten and Maggie Rogers see decreasing acousticness.
+Again in @fig-acousticness, Lorde's acoustic Solar Power album jumps out.
+Taylor Swift, Lorde and HAIM also have significant-seeming upward trends in `acousticness`, while Mitski, Sharon Von Etten and Maggie Rogers see decreasing `acousticness`.
 
@@
 
-In @fig-instrumentalness, St. Vincent is the only plot with any significant trend in instrumentalness.
+In @fig-instrumentalness, St. Vincent is the only plot with any significant trend in `instrumentalness`.
 
@@
 
-Now that we understand that there do seem to be some subtle differences in how variables that are and aren't produced by Jack behave, as seen in our cumulative distribution functions, our quadrants, and some minor trends in artists' sounds over time (specifically, `energy`), we can use a logistic regression to see if a model can a) predict whether a song was produced by Antonoff based on these variables, and b) tell us which variables were more significant in determining the output.
+We've observed subtle distinctions in variables associated with songs produced by Jack Antonoff, as evidenced by our cumulative distribution functions, quadrant analyses, and minor trends in the evolution of artists' sound characteristics, particularly energy.
+Employing a logistic regression will then:
+a) determine the likelihood of a song being produced by Antonoff based on these identified variables, and
+b) discern which variables are most influential in shaping this classification.
 
@@
 
-One issue is that in our dataset, there are a lot more songs that aren't produced by Jack than songs that are. The classes have to be balanced.
-
-```{r}
-#| echo: false
-#| warning: false
-# 
-# data_reduced |>
-#   count(jack) |>
-#   knitr::kable(col.names = c("Is it Jack?", "Number of Songs"))
-
-```
-
 After some minor pre-processing using the `tidyverse` R package [@citetidyverse]. I followed @citeSilge to build a model that deals with the imbalanced classes.
@@
 The `workflow_set()` function lets us hold different sets of feature engineering together and use tuning to evaluate them all at once.
-I need to evaluate two elements: a basic model, and a model that was downsampled using the `themis` package.
+I need to evaluate two elements: a basic model, and the model that was downsampled using the `themis` package.
 
@@
 It starts low with in mean log loss and sensitivity, but increases with increased regularization.
-Compared to the downsampled data, which starts high but drops significantly in accuracy and sensitivity, we chose to use the original data.
+Compared to the downsampled data, which starts high but drops significantly in accuracy and sensitivity, I chose to use the original data.
 
@@
 
-Danceability, with a positive importance score of 4.142, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
-Valence comes in second at 2.559, with negative score.
-The third, instrumentalness correlates negatively with Antonoff tracks with a a negative score of 2.309.
+`Danceability`, with a positive importance score of 4.142, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
+`Valence` comes in second at 2.559, with negative score.
+The third, `instrumentalness` correlates negatively with Antonoff tracks with a a negative score of 2.309.
 
@@
 
-@tbl-mostdanceable shows us the 20 songs in Jack Antonoff's discography with the highest danceability scores.
+@tbl-mostdanceable shows us the 20 songs in Jack Antonoff's discography with the highest `danceability` scores.
 Bleachers takes up 35% of this list, and Taylor Swift takes up 40% of this list.
@@
 Random forests (RF) build an ensemble of decision trees during the training phase, and then use majority voting for classification tasks.
-Each tree in the ensembled is trained on a random sample of data taken with replacement, and only a random subset of features are considered when splitting at each node.
+Each tree in the ensemble is trained on a random sample of data taken with replacement, and only a random subset of features are considered when splitting at each node.
 This method ensures that each tree in the ensemble is unique, avoiding overfitting and capturing non-linear relationships and more complex patterns in the data.
@@
 
-Antonoff's danceability score being, compared to other producers, significantly positively correlated with speechiness, loudness and energy, negatively correlated with acousticness, and less than average with valence tracks with how critics tend to describe his music.
+According to our results, Antonoff's `danceability` score is, compared to other producers, significantly positively correlated with `speechiness`, `loudness` and `energy`, negatively correlated with `acousticness`, and less positively correlated than average with `valence`.
+Speechy, loud, energetic, synthy, positive-yet-nostalgic.
+Our model's understanding of Antonoff's style, if we can take what `danceability` 's correlates capture as a proxy for style, tracks with both a close-listen and critics' descriptions of Antonoff's sound.
 
@@
 
-Clearing noise and isolating the middle range of the vocals may make for a stronger "speechiness" feature.
+Perhaps, without further documentation on the `speechiness` feature's construction, clearing the noise and isolating the middle range of the vocals may make for a stronger `speechiness` feature.
+Or, more realistically, artists who collaborated with Antonoff happen to be on the more lyrics-oriented end of the spectrum, rather than dance hit pop, similar to Antonoff's own Bleachers and fun..
 
-Valence, like danceability, was also a subjective variable built using labeling. Returning to the music itself, Antonoff-produced songs with the highest danceability scores aren't exactly dance hits.
+Turning to the music itself, even Antonoff-produced songs with the highest `danceability` scores aren't exactly dance hits.
 
-I Think He Knows by Taylor Swift, plainly, seems danceable to me.
-
-However, personally, I don't hear Taylor Swift's Vigilante Shit or Florence and the Machine's Heaven is Here as danceable.
+While the most danceable song*, I Think He Knows* by Taylor Swift, seems danceable to me, I personally don't hear Taylor Swift's *Vigilante Shit* or Florence and the Machine's *Heaven is Here* as danceable.
 They aren't understated, ambient tracks either -- there's something coercing your body to do something.
-Like Gamman's description of Antonoff's vocal design sounding how they sound to person who's singing them, the tracks seem like they're trying to get inside of your skin.
+Like Gamman's description of Antonoff's vocal design sounding like how they would sound to the person who's singing them, these tracks seem like they're trying to get inside of your skin.
 But they don't have the rhythm of a dance track.
 
-In a lot of ways, I Wanna Get Better by Bleachers is very similar to the typical Antonoff track that the outputs of our RF describe.
-This song is more anthemic, closer to the driving-over-the-Brooklyn-bridge, sort of post-Springsteen emotions that Antonoff is often characterized (or admitting) to striving for, than danceable.
-It also seems like it's trying to coercing your attention, if not your body, and the speechiness, loudness, and energy which are so correlated with danceability in Antonoff's RF results are all apparent.
+*I* *Wanna Get Better* by Bleachers is very similar to the typical Antonoff track that the outputs of our RF describe.
+This song is more anthemic, closer to the driving-over-the-Brooklyn-bridge, sort of neo-Springsteen emotions that Antonoff is often characterized (or admitting) to striving for, than danceable.
+It also does seem like it's trying to coerce your attention, if not your body, and the speechiness, loudness, and energy which are distinctly correlated with danceability in Antonoff's RF results are all apparent.
 The redemption, broken-but-getting-better-quasi-valence can also be heard in the lyrics, but it still doesn't have the kind of rhythm for dancing.
 
-Cornelia Street by Taylor Swift is another sort of almost-danceable track -- while the melody is catchy, the lyrics themselves are wistful. The vocal treatment, again, is there --
+Cornelia Street by Taylor Swift is another sort of almost-danceable track -- while the melody is catchy, the lyrics themselves are wistful.
+The vocal treatment, again, is there -- Swift's vocals feel magnified, carrying the bridge of the song.
 
-## Danceability Predicts Jack-iness {#sec-first-point}
-
-Using the Spotify Audio Features, you can't really gauge what Jack Antonoff's music *sounds* like.
-You get a rough idea, but you don't get the same granularity that you would by simply DM'ing Caleb Gamman on Twitter and asking him to describe it you.
-Or by watching the video where he already did that.
-
-> I think it's mostly in the vocal treatment, that's the thing that is the most off-putting to me.
-But just every decision feels a little bit wrong to me.
-If I wanted to be cringe I could point out the Juno 6 Bass, the perfectly stacked vocals, the Juno 6 pad, the wide and tight plate reverb, the Juno 6 sword, the goo goo gaga, the man hates treble unless it's just random crap swishing around, everything gets low passed with thin 707 drum samples, thin 808 drum samples, then 909 drum samples, thin Lin drum samples, thin thin drum samples, all the resonances, the harmonics get removed, everything is sort of squished down to the fundamental reduction in timbre, everything exists in a small frequency range, everything becomes a tone, it's an extremely consonant sound with no interplay between the different elements.
-Now that is a skill to do, and it produces a cleaner sound, and the only problem is I viscerally hate that sound.
-
-According to this model, though, danceability is the most important feature in predicting whether a song was produced by Jack Antonoff.
-And the accuracy metrics aren't awful -- the accuracy score is 0.743, and the roc_auc (receiver operating characteristic area under the curve) (more fun to say out loud as an abbreviation.
-Rock Auck), which is a metric better suited for imbalanced datasets, such as our own, is 0.810.
-This means that although danceability can't show us the inside of a song, something about this score is telling us something about Jack's music.
-
-### In which I do a close listen of our model's top 5 most danceable songs[^3]:
-
-[^3]: I encourage you to follow along and drop me an email with your own assessment at michaela.drouillard\@mail.utoronto.ca.
-
--   I Think He Knows by Taylor Swift:
-
-    -   Pretty danceable tbh.
-
--   Vigilante Shit by Taylor Swift:
-
-    -   Not danceable. But it is trying to coerce you into some sort of body feeling.
-
--   Heaven is Here by Florence and the Machine
-
-    -   DEFINITELY not danceable. But it is trying to get your body to do something. Kind of sounds like it's suggesting that you join a Macedonian phalanx or something. But no. Not danceable.
-
--   I Wanna Get Better by Bleachers
-
-    -   I don't know if "danceability" is the word, but I'll give it "anthem"-y.
-Like, I can't picture a group of friends dancing to this in the kitchen, but I can picture someone driving, on their way to give a presentation or negotiate a promotion or something, and this song comes on the radio and they leave it on.
-They kind of lean into it.
-
--   Cornelia Street by Taylor Swift
-
-    -   Yeah I'm dancing. Or at least bobbing my head. I only became aware of that fact when I made eye contact with the guy next to me at this cafe.
+What, then, is the `danceability` variable telling us?
+Antonoff's most "danceable" songs are better understood through their strongest correlates with `danceability`, although `danceability` is the strongest predictor overall.
 
@@
 
-It's just his subjective taste, but in one way Gamman was actually totally on the money about this.
-The danceability variable, which is the strongest predictor of Jack-ness, seems to actually be more about "visceral"-ness when you do a close-listen of the songs with the highest danceability scores.
+It's just his subjective taste, but Gamman is right.
+The `danceability` variable, which is the strongest predictor of Jack-ness, seems to actually be more about some sort of "visceral"-ness when you do a close-listen of the songs with the highest danceability scores.
 Not all of them are danceable, but the one's that aren't still feel like they're trying to manipulate the way your body is holding itself in some way.
-The reason Jack is getting under people's skin could just be because his music...
-gets underneath your skin.
-Right to the viscera.
-
-## Weaknesses and next steps
-
-One fundamental weakness is cooked into the dataset.
-Since we're not in the studio, we can't know the degree of Jack-iness in each individual song.
-Especially when he takes on the role of a more backseat co-producer.
-We can't actually know what his involvement in each song was.
-There's also the issue that a study like this implies that Jack Antonoff is the reason for the way these artists' sounds develop, which takes that agency away from the artists.
-This is a critique you see on the internet, and also a critique that some people on Twitter leveraged against Gamman after his Midnights video went viral.
-I mentioned it to a friend when she asked me how my final paper was going, and she said something like: "Yeah but like, it's *Taylor Swift*.
-As if anyone actually thinks that Jack Antonoff is the reason for her success." I agree.
-
-#### Next Steps
+The reason that Antonoff's sound dominating chart-topping pop production gets under people's skin could just be because his music gets underneath your skin.
 

### 2023-11-03 scripts/06-models.R
bayesian models

@@ -29,7 +29,7 @@ data_reduced <-
 write_csv(data_reduced,"inputs/data/data_reduced.csv")
 
 
-#### Predicting Antonoff ####
+#### Predicting Antonoff: non-bayesian  ####
 ## glmnet requires all variables to be numeric, so i'm excluding the producer column
 set.seed(853)
 # 1. Create a new binary column
@@ -167,9 +167,118 @@ antonoff_vip %>%
   theme(legend.position = "none")
 
 
+### SECOND MODEL: Bayesian logistic regression ###
+
+library(ROSE)
+library(rstanarm)
+
+lr_data <- data_reduced
+lr_data$is_antonoff <- ifelse(lr_data$producer == "antonoff", 1, 0)
+lr_data$is_antonoff <- as.factor(lr_data$is_antonoff)
 
 
-### SECOND MODEL: Random Forest ###
+antonoff_split <- 
+  lr_data |>
+  initial_split(strata = is_antonoff)
+
+antonoff_train <- training(antonoff_split)
+antonoff_test <- testing(antonoff_split)
+
+naive_prior <- normal(0, 10) #wide normal distribution approximation to flat prior
+antonoff_train$is_antonoff <- factor(antonoff_train$is_antonoff, levels = c("0", "1"))
+
+
+library(caret)
+library(rstanarm)
+
+antonoff_train$is_antonoff <- factor(antonoff_train$is_antonoff)
+
+ds_data<- downSample(x = antonoff_train[, !(names(antonoff_train) %in% "is_antonoff")],
+                                y = antonoff_train$is_antonoff)
+
+
+
+if("producer" %in% names(ds_data)) {
+  ds_data <- ds_data[, !(names(ds_data) %in% "producer")]
+}
+
+
+bayesian_model <- stan_glm(
+  Class ~ .,
+  data = ds_data,
+  family = binomial(link = "logit"),
+  prior = naive_prior,  
+  prior_intercept = normal(0, 5),  
+  iter = 2000,  
+  seed = 853  
+)
+
+
+summary(bayesian_model)
+
+
+
+
+
+posterior_predictive <- posterior_predict(bayesian_model, newdata = antonoff_test)
+
+predicted_probabilities <- apply(posterior_predictive, 2, mean)
+
+predicted_classes <- ifelse(predicted_probabilities > 0.5, "1", "0")
+
[…]

### 2023-11-08 outputs/paper/paper.qmd
paper edit
edits to paper

@@
 ---
-title: "title"
-subtitle: "subtitle"
+title: "Detecting Stylistic Variation in Pop Production using the Spotify API"
 author: 
@@
 date: today
-abstract: ""
+abstract: "This paper explores whether or not we can predict whether a song was produced by Jack Antonoff using the audio features provided by Spotify's API.
+We find that in general our model has an accuracy score of 90%.
+(Precision and recall) We interestingly found that danceability, a metric trained on human-tagged data, was a key predictor.
+Our results have implications for the understanding of culture, how the combination human subjectivity and machine listening can come to create an approximate for style."
 format: pdf
@@
 
+We're interested in understanding how to create an approximate for production style.
+To do this we used Spotify API's audio features dataset to explore different audio feature metrics and identify which were the most import predictors in identifying pop producers' style.
+We chose Jack Antonoff as focus, as he is a prolific and widely used producer, who also has his own solo acts.
+
+By way of background, Antonoff began his career as a guitarist in fun., and is the frontman in Bleachers.
+He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more.
+His style is considered identifiable.
+As Andrew Marantz writes in the *New Yorker*: "When his band releases an album, the world responds politely.
+When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis" [@citeMarantz].
+
+We are interested in whether Spotify audio features can be used to train a model that distinguishes whether a song was produced by Antonoff.
+We obtained tracks from six artists who've collaborated with Antonoff, both before their collaboration and during.
+We also obtained tracks from similarly prolific pop producers.
+The producers we compare Antonoff's production style with are Little, Rechstad, Martin, Kurstin, Epworth and Nowels.
+Our data set contains 3738 tracks in total.
+
+We built two models to explore the data.
+The first model, a logistic regression, draws on the dataset of artists who've collaborated with Antonoff, with a binary variable to tag whether or not the song was an Antonoff production.
+With this model, we explore which variables are stronger predictors of Antonoff's style, developping a better understanding of whether the Spotify data can trace Antonoff's influence on an artist's sound over time.
+We built a second model, a Random Forest (RF), to complement the insights drawn from the logistic regression.
+The RF trains on the full data set, including other major pop producers.
+
+We find that danceability is the strongest predictor of whether or not a song has been produced by Antonoff.
+Furthermore, when the variance is considered along with other features of the dataset, the variance tends to be strongest classifier of whether a song was produced by Antonoff.
+Daanceability in concert with other features in the dataset provides the best predictions.
+
+Against the odds, we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
+The remainder of this paper is structured as follows: @ref-background...
+
+# Background {#sec-background}
+
 Shortly after the release of Taylor Swift's *Midnights* album, a video of the Youtuber Caleb Gammen, in which he guesses which songs on Midnights were produced by Antonoff within seconds [@citeTweet], went viral.
@@
 I spoke to Gamman to better undertstand makes a song Antonoff-ish, and to try to understand his reaction to it.
-In trying to understand how to describe Antonoff's sound, I used the Spotify API's audio features dataset to explore difference machine listening metrics, and build a model to recreate a digital version of Gamman's classification excercise.
+
 I also spoke with Glen McDonald, an engineer at Spotify, from whom I learned that the `valence` and `danceability` variables in the Spotify API were created by hiring college interns and having them tag whether songs were positive or danceable.
-
-I used two models to investigate the data and ask the question: What is Antonoff's sound? Can we pin down style using the quantitative metrics offered by Spotify?
-
-The first model, a logistic regression, draws on the dataset of artists who've collaborated with Antonoff, with an extra binary variable to tag whether or not the song was an Antonoff production.
-With this model, we can get a first glimpse at variables that are stronger predictors of Antonoff-iness, along with a better understanding of whether the Spotify data can trace Antonoff's influence on an artist's sound over time.
 
@@
 
-I built a second model, a Random Forest, to complement the insights drawn from the logistic regression.
-The Random Forest trains on the full data set, including other major pop producers.
-Random Forests don't assume linear relationships, and aren't sensitive to multicollinearity.
-In fact, using the tk package in R, we can visualize and analyse the relationships between different variables.
-The variance in correlation scores across producers can be taken as an approximation for style, since this variance helps the model distinguish between producers.
-
 In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing something like visceral-ness.
-Against the odds, we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
-
-# tk: Lit Review
 
@@
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
-I threw in HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski in too, since they've never collaborated with Jack Antonoff but are more or less part of the same pop sphere.
+HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski were included, since they've never collaborated with Jack Antonoff but are more or less part of the same pop sphere.
 
@@
 
-### Artist Selection
-
 Of the artists who did collaborate with Antonoff, I chose these artists because they've produced multiple albums, and at least one album with Antonoff, as opposed to just individual songs.
 
-### Data Cleaning
-
-I joined these data sets using the `rbind` function in base R, and created an additional variable called "jack", which contains a 1 if Jack Antonoff produced or co-produced the song, and a 0 if he had nothing to do with it.
-I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page [@citewiki].
+We created a binary variable which contains 1 if Jack Antonoff produced or co-produced the song, and a 0 if it was produced by someone else.
+The information underpinning this discography was drawn from the "Jack Antonoff Production Discography" Wikipedia page [@citewiki].
 
@@
 
-## Interview with Glen McDonald
+In total 3738 tracks from X different artists. (Producers, etc)
 
-I interviewed Glen McDonald, a Principal Engineer at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
-Macdonald told me that `valence` and `danceability` were created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
-Variables like energy or instrumentalness, on the other hand, were determined primarily through machine listening techniques.
-The human subjectivity part only came in when it came to tweaking the training data[^1]:
+```{r}
+#| label: fig-means
+#| fig-cap: Comparing the Mean Values of Each Variable
+#| echo: false
+#| warning: false
 
-[^1]: My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
-Because there were no banjos in the training data, they were registered as human speech.
-(This is hilarious.) They had to go back and add more songs with banjos to the training data.
+lr_data <- read_csv(here::here("inputs/data/lr_data.csv"))
 
-"You could imagine writing a formula for energy that combines like loudness and tempo and degree of harmonic variation or something.
-So it was \[machine learning\], but that one's more like a human helping a machine figure out a formula.
-Whereas valence is teaching the machine to try to reproduce a purely human thing.
-Same with danceability.
-I mean, danceability is whether a human can dance to it.
-So like, the machine's not gonna dance, so the machine can't have any opinion on that.
-The computer could have an opinion on energy.
-And the computer can definitely have an opinion on loudness.
-So there's a spectrum from, you know, loudness as purely analytical, and then energy is a little like loudness, with a little more subjectivity.
-And then danceability and valence are purely subjective."
-
-Glen confirmed that this process didn't know anything about lyrics, which makes valence a particularly difficult variable to build.
-It could, in theory, pick up on aspects of vocal performance, but it doesn't know anything about languages or words or the meanings of songs.
-Take an upbeat, happy-sounding Elliot Smith song with devastating lyrics: the machine might register it as happy, where human listeners understand that it's sad.
-Add to that that two humans might even disagree on the song's valence.
-
-"Plus, we have the confounding factor of like, a song that seems happy today could be sad tomorrow because the singer was killed in a plane crash.
-The song didn't change, but our world changed and our reactions changed."
-
-Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well".
-High energy and high valence could be generally happy, cheerful, upbeat.
-Low energy, low valence could be sad or downbeat.
-High energy, low valence is sort of angry.
-Low energy, high valence is serene or calming.
+lr_data |>
+  group_by(is_antonoff) |>
+  summarise(across(-c(tempo, loudness), mean)) |>
+  pivot_longer(cols = -is_antonoff) |>
+  ggplot(aes(value, name, fill = factor(is_antonoff))) +
+  geom_col(alpha = 0.8, position = "dodge") +
+  theme_minimal() +
+  labs(fill = "Jack", x = "Mean Value", y = "Feature")
+```
 
@@
 We've observed subtle distinctions in variables associated with songs produced by Jack Antonoff, as evidenced by our cumulative distribution functions, quadrant analyses, and minor trends in the evolution of artists' sound characteristics, particularly energy.
-Employing a logistic regression will then:
-a) determine the likelihood of a song being produced by Antonoff based on these identified variables, and
-b) discern which variables are most influential in shaping this classification.
+Employing a logistic regression will then: a) determine the likelihood of a song being produced by Antonoff based on these identified variables, and b) discern which variables are most influential in shaping this classification.
 
-# Logistic Regression
+# Model
 
-## Data Preparation
+## Logistic Regression
 
@@
 where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the predictor variables.
-
-@fig-means gives us an idea of the mean value of each of these features for Antonofff and non-Antonoff-produced tracks.
-
-```{r}
-#| label: fig-means
-#| fig-cap: Comparing the Mean Values of Each Variable
-#| echo: false
-#| warning: false
-
-lr_data <- read_csv(here::here("inputs/data/lr_data.csv"))
-
-lr_data |>
-  group_by(is_antonoff) |>
-  summarise(across(-c(tempo, loudness), mean)) |>
-  pivot_longer(cols = -is_antonoff) |>
-  ggplot(aes(value, name, fill = factor(is_antonoff))) +
-  geom_col(alpha = 0.8, position = "dodge") +
-  theme_minimal() +
-  labs(fill = "Jack", x = "Mean Value", y = "Feature")
-```
 
@@
 
-## Results
+## Random Forest
+
+Random forests (RF) build an ensemble of decision trees during the training phase, and then use majority voting for classification tasks.
+Each tree in the ensemble is trained on a random sample of data taken with replacement, and only a random subset of features are considered when splitting at each node.
+This method ensures that each tree in the ensemble is unique, avoiding overfitting and capturing non-linear relationships and more complex patterns in the data.
+
+add model here
+
+The categorical feature `producer`, which contains the names of the producer for each track, was one-hot encoded to create binary columns for each producer (i.e.
+`is_antonoff`, `is_elworth`, etc.).
+The producer feature was then converted into a factor variable, and the dataset was split into training and tests sets.
+
+The RF was trained using the @ranger package. The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
+
+Gini impurity is a measure of misclassification.
+At every split in a decision tree, the algorithm tests how well each feature splits the data.
+The features that split the data the most accurately, and then with the least impurity, would have the lowest impurity score.
+In Random Forests, a variable's importance can be calculated by looking at how much the tree nodes that use that variable reduce impurity on average.
+
+Gini impurity was chosen on the grounds that this study is interested in not only classifying producers, but, more importantly, in understanding how the underlying data informs these decisions.
+By calculating the Gini impurity not only for the entire model, but specifically for each producer, we can better understand how the variables in our dataset describe the differences between producers.
+Without song structure, instruments, sound, or other more ephemeral components of style, these differences can be taken as a proxy for style.
+
+# Results
+
+## Logistic Regression
 
@@
 
-# Random Forest
-
-Random forests (RF) build an ensemble of decision trees during the training phase, and then use majority voting for classification tasks.
-Each tree in the ensemble is trained on a random sample of data taken with replacement, and only a random subset of features are considered when splitting at each node.
-This method ensures that each tree in the ensemble is unique, avoiding overfitting and capturing non-linear relationships and more complex patterns in the data.
-The formula behind the RF is:
-
-## Data Preparation
-
-The categorical feature `producer`, which contains the names of the producer for each track, was one-hot encoded to create binary columns for each producer (i.e.
-`is_antonoff`, `is_elworth`, etc.).
-The producer feature was then converted into a factor variable, and the dataset was split into training and tests sets.
-
-## Model Training
-
-The RF was trained using the @ranger package. The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
-
-Gini impurity is a measure of misclassification.
-At every split in a decision tree, the algorithm tests how well each feature splits the data.
-The features that split the data the most accurately, and then with the least impurity, would have the lowest impurity score.
-In Random Forests, a variable's importance can be calculated by looking at how much the tree nodes that use that variable reduce impurity on average.
-
-Gini impurity was chosen on the grounds that this study is interested in not only classifying producers, but, more importantly, in understanding how the underlying data informs these decisions.
-By calculating the Gini impurity not only for the entire model, but specifically for each producer, we can better understand how the variables in our dataset describe the differences between producers.
-Without song structure, instruments, sound, or other more ephemeral components of style, these differences can be taken as a proxy for style.
-
-## Results
-
-### Model Performance
+## Random Forest
 
@@
 However, there is enough demonstrated statistical significance to suggest that there are underlying patterns in the data that differentiate producers.
-
-### Stylistic Difference
 
@@
 
+I interviewed Glen McDonald, a Principal Engineer at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
+Macdonald told me that `valence` and `danceability` were created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
+Variables like energy or instrumentalness, on the other hand, were determined primarily through machine listening techniques.
+The human subjectivity part only came in when it came to tweaking the training data[^1]:
+
+[^1]: My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
+Because there were no banjos in the training data, they were registered as human speech.
+(This is hilarious.) They had to go back and add more songs with banjos to the training data.
+
+"You could imagine writing a formula for energy that combines like loudness and tempo and degree of harmonic variation or something.
+So it was \[machine learning\], but that one's more like a human helping a machine figure out a formula.
+Whereas valence is teaching the machine to try to reproduce a purely human thing.
+Same with danceability.
+I mean, danceability is whether a human can dance to it.
+So like, the machine's not gonna dance, so the machine can't have any opinion on that.
+The computer could have an opinion on energy.
+And the computer can definitely have an opinion on loudness.
+So there's a spectrum from, you know, loudness as purely analytical, and then energy is a little like loudness, with a little more subjectivity.
+And then danceability and valence are purely subjective."
+
+Glen confirmed that this process didn't know anything about lyrics, which makes valence a particularly difficult variable to build.
+It could, in theory, pick up on aspects of vocal performance, but it doesn't know anything about languages or words or the meanings of songs.
+Take an upbeat, happy-sounding Elliot Smith song with devastating lyrics: the machine might register it as happy, where human listeners understand that it's sad.
+Add to that that two humans might even disagree on the song's valence.
+
+"Plus, we have the confounding factor of like, a song that seems happy today could be sad tomorrow because the singer was killed in a plane crash.
+The song didn't change, but our world changed and our reactions changed."
+
+Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well".
+High energy and high valence could be generally happy, cheerful, upbeat.
+Low energy, low valence could be sad or downbeat.
+High energy, low valence is sort of angry.
+Low energy, high valence is serene or calming.
+
 \newpage

### 2023-11-09 outputs/paper/paper.qmd
artists folder, wrestling with latex

@@
 library(lubridate)
+library(tidyr)
+library(dplyr)
+library(tidymodels)
+library(randomForest)
+library(ranger)
+library(caret)
+library(themis)
+library(yardstick)
 
@@
 Furthermore, when the variance is considered along with other features of the dataset, the variance tends to be strongest classifier of whether a song was produced by Antonoff.
-Daanceability in concert with other features in the dataset provides the best predictions.
+Danceability in concert with other features in the dataset provides the best predictions.
 
@@
 
-In total 3738 tracks from X different artists. (Producers, etc)
+In total 3738 tracks from 828 different artists.
+We focus on identifying the styles of 7 producers: Jack Antonoff, Max Martin, Joel Little, Ariel Rechtshaid, Greg Kurstin, Paul Epworth and Rick Nowels.
+If tracks were produced by producers other than the main producers included in our study, they were marked as "other".
+We removed tracks from live performances, karaoke editions, international versions or remix albums.
+Where a deluxe album was available, we removed tracks belonging to the origin to avoid repetition.
 
@@
 #| warning: false
+av<- read_csv(here::here("inputs/data/antonoff_viz.csv"))
 
-# df <- read_csv(here::here("inputs/data/clean_df.csv"))
-# 
-# ggplot(df, aes(x = energy, y = valence)) +
-#   geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
-#   geom_vline(xintercept = mean(df$energy), color = "gray") +
-#   geom_hline(yintercept = mean(df$valence), color = "gray") +
-#   theme_minimal()
+av$is_antonoff <- ifelse(av$producer == "antonoff", 1, 0)
+av$is_antonoff <- as.factor(av$is_antonoff)
+
+ggplot(av, aes(x = energy, y = valence)) +
+  geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$valence), color = "gray") +
+  theme_minimal()
 ```
@@
 #| warning: false
-# 
-# ggplot(df, aes(x = energy, y = danceability)) +
-#   geom_point(aes(color = as.factor(jack)), alpha = 0.3) +
-#   geom_vline(xintercept = mean(df$energy), color = "gray") +
-#   geom_hline(yintercept = mean(df$danceability),
-#              color = "gray") +
-#   theme_minimal()
+
+ggplot(av, aes(x = energy, y = danceability)) +
+   geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$danceability),
+              color = "gray") +
+  theme_minimal()
 ```
@@
 
-# data_reduced <- read_csv(here::here("inputs/data/data_reduced.csv"))
-# 
-# data_reduced |>
-#   pivot_longer(
-#     cols = c(
-#       danceability,
-#       energy,
-#       loudness,
-#       mode,
-#       speechiness,
-#       acousticness,
-#       instrumentalness,
-#       liveness,
-#       valence,
-#       tempo
-#     ),
-#     names_to = "name",
-#     values_to = "value"
-#   ) |>
-#   ggplot(aes(x = value, color = jack)) +
-#   stat_ecdf() +
-#   facet_wrap(vars(name),
-#              nrow = 3,
-#              ncol = 4,
-#              scales = "free") +
-#   theme_minimal() +
-#   labs(x = "Value",
-#        y = "Proportion")
+ av |>
+   pivot_longer(
+     cols = c(
+       danceability,
+       energy,
+       loudness,
+       mode,
+       speechiness,
+       acousticness,
+       instrumentalness,
+       liveness,
+       valence,
+       tempo
+     ),
+     names_to = "name",
+     values_to = "value"
+   ) |>
+   ggplot(aes(x = value, color = is_antonoff)) +
+   stat_ecdf() +
+   facet_wrap(vars(name),
+              nrow = 3,
+              ncol = 4,
+              scales = "free") +
+   theme_minimal() +
+   labs(x = "Value",
+        y = "Proportion")
 
-#even tho there's similrities, there's differences
 ```
@@
 
-# df <- read_csv(here::here("inputs/data/df.csv"))
-# 
-# df <- df |> 
-#   mutate(jack = factor(jack))
-# 
-# df |>
-#   filter(album_release_date_precision == "day") |>
-#   mutate(album_release_date = ymd(album_release_date)) |>
-#   ggplot(aes(y = energy, x = album_release_date, color = jack)) +
-#   geom_point(alpha = 0.3) +
-#   geom_smooth(method = lm, se = FALSE) +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
-#   facet_wrap(~ artist_name, ncol = 3)
+av <- av |>
+  mutate(is_antonoff= factor(is_antonoff))
+
+av |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
+  ggplot(aes(y = energy, x = album_release_date, color = is_antonoff)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
+  facet_wrap(~ artist_name, ncol = 3)
 ```
@@
 
-# df |>
-#   filter(album_release_date_precision == "day") |>
-#   mutate(album_release_date = ymd(album_release_date)) |>
-#   ggplot(aes(y = danceability, x = album_release_date, color = jack)) +
-#   geom_point(alpha = 0.3) +
-#   geom_smooth(method = lm, se = FALSE) +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
-#   facet_wrap( ~ artist_name, ncol = 3)
+av |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
+  ggplot(aes(y = danceability, x = album_release_date, color = is_antonoff)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-# df |>
-#   filter(album_release_date_precision == "day") |>
-#   mutate(album_release_date = ymd(album_release_date)) |>
-#   ggplot(aes(y = valence, x = album_release_date, color = jack)) +
-#   geom_point(alpha = 0.3) +
-#   geom_smooth(method = lm, se = FALSE) +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
-#   facet_wrap( ~ artist_name, ncol = 3)
+av |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
+  ggplot(aes(y = valence, x = album_release_date, color = is_antonoff)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-# df |>
-#   filter(album_release_date_precision == "day") |>
-#   mutate(album_release_date = ymd(album_release_date)) |>
-#   ggplot(aes(y = acousticness, x = album_release_date, color = jack)) +
-#   geom_point(alpha = 0.3) +
-#   geom_smooth(method = lm, se = FALSE) +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
-#   facet_wrap( ~ artist_name, ncol = 3)
+av |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
+  ggplot(aes(y = acousticness, x = album_release_date, color = is_antonoff)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-# df |>
-#   filter(album_release_date_precision == "day") |>
-#   mutate(album_release_date = ymd(album_release_date)) |>
-#   ggplot(aes(y = instrumentalness, x = album_release_date, color = jack)) +
-#   geom_point(alpha = 0.3) +
-#   geom_smooth(method = lm, se = FALSE) +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
-#   facet_wrap( ~ artist_name, ncol = 3)
+av |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
+  ggplot(aes(y = instrumentalness, x = album_release_date, color = is_antonoff)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 
-# df |>
-#   filter(album_release_date_precision == "day") |>
-#   mutate(album_release_date = ymd(album_release_date)) |>
-#   ggplot(aes(y = tempo, x = album_release_date, color = jack)) +
-#   geom_point(alpha = 0.3) +
-#   geom_smooth(method = lm, se = FALSE) +
-#   theme_minimal() +
-#   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-#   labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
-#   facet_wrap( ~ artist_name, ncol = 3)
+av |>
+  filter(album_release_date_precision == "day") |>
+  mutate(album_release_date = ymd(album_release_date)) |>
+  ggplot(aes(y = tempo, x = album_release_date, color = is_antonoff)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = lm, se = FALSE) +
+  theme_minimal() +
+  scale_color_manual(values = c("#808080", "#FFB6C1")) +
+  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
+  facet_wrap( ~ artist_name, ncol = 3)
 ```
@@
 $$
-\mbox{Pr}(y_i=1) = \mbox{logit}^{-1}\left(\beta_0+\beta_1 x_i\right)
+Pr(yi​=1)=logit−1(β0​+β1​×danceability+β2​×energy+...+β10​×tempo)
 $$
 
-where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the predictor variables.
+where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the 10 predictor variables.
 

### 2023-11-09 outputs/paper/paper.qmd
latex wrestling

@@
 
-add model here
+The training process for the RF model can be summarized as follows:
+
+1.  For each iteration ( b = 1 ) to ( B ):
+    a.  Draw a bootstrap sample ( Z\^\* ) of size ( N ) from the training data.
+    b.
+Grow a random-forest tree ( T_b ) to the bootstrapped data by recursively performing the following steps for each terminal node of the tree until the minimum node size ( n\_{\text{min}} ) is reached:
+        i.  Select ( m ) predictors at random from the ( p ) predictors.
+        ii. Determine the best predictor and split-point among the ( m ).
+        iii. Split the node into two daughter nodes.
+2.  The ensemble of trees ( { T_b }\^B\_{b=1} ) constitutes the random forest model.
+
+For a new instance ( x ), the RF model prediction ( \hat{C}\^{\text{RF}}(x) ) is determined by the majority vote from the ensemble of trees:
+
+\[ \hat{C}\^{\text{RF}}(x) = \text{majority vote} \left{ \hat{C}*b(x)* \right}\^B{b=1} \]
+
+Where ( \hat{C}\_b(x) ) is the class prediction of the ( b )-th tree.
 

### 2023-11-10 outputs/paper/paper.qmd
discussion section edits

@@
 We also obtained tracks from similarly prolific pop producers.
-The producers we compare Antonoff's production style with are Little, Rechstad, Martin, Kurstin, Epworth and Nowels.
+The producers we compare Antonoff's production style with are Little, Rechstaid, Martin, Kurstin, Epworth and Nowels.
 Our data set contains 3738 tracks in total.
@@
 
-```{r}
-#| label: fig-quadvalence
-#| fig-cap: Valence and Energy Quadrant
-#| echo: false
-#| warning: false
-av<- read_csv(here::here("inputs/data/antonoff_viz.csv"))
-
-av$is_antonoff <- ifelse(av$producer == "antonoff", 1, 0)
-av$is_antonoff <- as.factor(av$is_antonoff)
-
-ggplot(av, aes(x = energy, y = valence)) +
-  geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$valence), color = "gray") +
-  theme_minimal()
-```
-
 In @fig-quadvalence, it looks like songs produced by Antonoff are a little more clustered towards the lower end of valence, and across the spectrum of energy, although the variance is large.
@@
 #| warning: false
+av<- read_csv(here::here("inputs/data/antonoff_viz.csv"))
 
+av$is_antonoff <- ifelse(av$producer == "antonoff", 1, 0)
+av$is_antonoff <- as.factor(av$is_antonoff)
  av |>
@@
 
-After some minor pre-processing using the `tidyverse` R package [@citetidyverse]. I followed @citeSilge to build a model that deals with the imbalanced classes.
+After some minor pre-processing using the `tidyverse` R package [@citetidyverse]. We followed @citeSilge to build a model that deals with the imbalanced classes.
 
-I used the `tidymodels` to fit the model [@citetidymodels].
-To deal with the imbalanced classes, I downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
-Then I used 5-fold cross-validation with stratification by the **`jack`** variable to estimate the performance of the model.
+We used the `tidymodels` to fit the model [@citetidymodels].
+To deal with the imbalanced classes, we downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
+Then we used 5-fold cross-validation with stratification by the **`jack`** variable to estimate the performance of the model.
 
 The `workflow_set()` function lets us hold different sets of feature engineering together and use tuning to evaluate them all at once.
-I need to evaluate two elements: a basic model, and the model that was downsampled using the `themis` package.
+We need to evaluate two elements: a basic model, and the model that was downsampled using the `themis` package.
 
-I used **`rank_results()`** from the **`yardstick`** package to rank the models based on accuracy, log loss, sensitivity, and specificity.
+We used **`rank_results()`** from the **`yardstick`** package to rank the models based on accuracy, log loss, sensitivity, and specificity.
 After testing both basic and downsampled versions of the model, the basic version yielded much higher accuracy score (0.918 basic versus 0.743 downsampled), although a lower roc auc score (0.746 basic versus 0.821 downsampled).
@@
 It starts low with in mean log loss and sensitivity, but increases with increased regularization.
-Compared to the downsampled data, which starts high but drops significantly in accuracy and sensitivity, I chose to use the original data.
+Compared to the downsampled data, which starts high but drops significantly in accuracy and sensitivity, we chose to use the original data.
 
@@
 ```{r}
+#| label: fig-dancevariance
 #| echo: false
@@
 
-According to our results, Antonoff's `danceability` score is, compared to other producers, significantly positively correlated with `speechiness`, `loudness` and `energy`, negatively correlated with `acousticness`, and less positively correlated than average with `valence`.
-Speechy, loud, energetic, synthy, positive-yet-nostalgic.
-Our model's understanding of Antonoff's style, if we can take what `danceability` 's correlates capture as a proxy for style, tracks with both a close-listen and critics' descriptions of Antonoff's sound.
+Our results indicate that Antonoff's music, characterized by its danceability score, exhibits a significant positive correlation with factors such as speechiness, loudness, and energy, while showing a negative correlation with acousticness, and a less pronounced positive correlation with valence, compared to other music producers.
+This profile---marked by speechiness, loudness, energy, synthetic textures, and a blend of positivity with a touch of nostalgia---aligns with both detailed auditory analysis and critical assessments of Antonoff's sound.
 
-His tracks have been described as anthemic (cite), with a distinct vocal treatment that makes it sound, as Gamman put it, "the way it would sound to the person who's singing them".
+Antonoff's tracks have been described as anthemic, with a distinct vocal treatment that makes it sound, as Gamman (2022) observes, "the way it would sound to the person who's singing them".
 
@@
 
-Perhaps, without further documentation on the `speechiness` feature's construction, clearing the noise and isolating the middle range of the vocals may make for a stronger `speechiness` feature.
-Or, more realistically, artists who collaborated with Antonoff happen to be on the more lyrics-oriented end of the spectrum, rather than dance hit pop, similar to Antonoff's own Bleachers and fun..
+Perhaps, without further documentation on the `speechiness` feature's construction, clearing the noise and isolating the middle range of the vocals may make for stronger overall `speechiness` score in Antonoff's tracks.
+Antonoff also tends to collaborate with singer-songwriter artists, whose music is more lyric-oriented than dance hit pop.
 
-Turning to the music itself, even Antonoff-produced songs with the highest `danceability` scores aren't exactly dance hits.
+Even Antonoff-produced songs with the highest `danceability` scores aren't exactly dance hits, compared to production styles like that of Max Martin.
 
-While the most danceable song*, I Think He Knows* by Taylor Swift, seems danceable to me, I personally don't hear Taylor Swift's *Vigilante Shit* or Florence and the Machine's *Heaven is Here* as danceable.
-They aren't understated, ambient tracks either -- there's something coercing your body to do something.
+For instance, the highest scoring song in danceability*, I Think He Knows* by Taylor Swift, seems danceable, yet the next highest-scoring songs -- Taylor Swift's *Vigilante Shit* or Florence and the Machine's *Heaven is Here* -- don't resonate as dance tracks.
+However, they aren't understated, ambient tracks either -- the exhibit some sort of compelling quality, like they're coercing your body to do something.
 Like Gamman's description of Antonoff's vocal design sounding like how they would sound to the person who's singing them, these tracks seem like they're trying to get inside of your skin.
@@
 
-*I* *Wanna Get Better* by Bleachers is very similar to the typical Antonoff track that the outputs of our RF describe.
-This song is more anthemic, closer to the driving-over-the-Brooklyn-bridge, sort of neo-Springsteen emotions that Antonoff is often characterized (or admitting) to striving for, than danceable.
-It also does seem like it's trying to coerce your attention, if not your body, and the speechiness, loudness, and energy which are distinctly correlated with danceability in Antonoff's RF results are all apparent.
-The redemption, broken-but-getting-better-quasi-valence can also be heard in the lyrics, but it still doesn't have the kind of rhythm for dancing.
+*I* *Wanna Get Better* by Bleachers typifies the Antonoff style as identified by our model.
+This song is more anthemic, resonating with a sort of neo-Springsteen aesthetic that Antonoff is often characterized (or admitting) to striving for, than danceable.
+It also does sound like it's trying to demand your attention, and the speechiness, loudness, and energy which are distinctly correlated with danceability in Antonoff's RF results, are all apparent.
+The redemptive, broken-but-getting-better tone can also be heard in the lyrics, but it still doesn't have the kind of rhythm for dancing -- which maps onto how Antonoff's "valence" score isn't as a high as other pop artists in @fig-dancevariance.
 
-Cornelia Street by Taylor Swift is another sort of almost-danceable track -- while the melody is catchy, the lyrics themselves are wistful.
-The vocal treatment, again, is there -- Swift's vocals feel magnified, carrying the bridge of the song.
+*Cornelia Street* by Taylor Swift further exemplifies this near-danceable category.
+While melodically catchy, the lyrics themselves are wistful and longing.
+The vocal treatment, as described by Gamman, is there -- Swift's vocals sound magnified, carrying the bridge of the song.
 
-What, then, is the `danceability` variable telling us?
-Antonoff's most "danceable" songs are better understood through their strongest correlates with `danceability`, although `danceability` is the strongest predictor overall.
+What does the danceability variable signify in Antonoff's music?
+Our findings suggest that his most "danceable" songs are better understood through the lens of correlated features, even though danceability emerges as the primary predictor.
+This crowd-sourced metric, in conjunction with machine-listening-derived features, offers a unique insight into an artist's style, capturing an intuitive understanding of a song's tone and energy.
 
-I am thinking about what Gamman said in his video.
+We spoke with Glen McDonald, a Principal Engineer at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
+Macdonald revealed that `valence` and `danceability` were created by giving tracks to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
+Variables like `energy` or `instrumentalness`, on the other hand, were determined primarily through machine listening techniques, with human subjectivity being applied to fine-tune the features[^1]:
 
-> I did use the phrase "visceral hatred".
-But i did so only to describe my visceral hatred.
-I can't stand the sound of it.
-I hate it.
-And the way in which I hate it is viscerally.
-But I didn't go beyond that because it doesn't go beyond that.
-Nothing against the guy, I like Taylor swift's stuff.
-It's just subjective taste.
+[^1]: A notable example of this being that bluegrass songs were given very high `speechiness` ratings.
+Because there were no banjos in the training data, the instruments were registered as human speech.
+The engineers had to go back and add more songs with banjos to the training data.
 
-It's just his subjective taste, but Gamman is right.
-The `danceability` variable, which is the strongest predictor of Jack-ness, seems to actually be more about some sort of "visceral"-ness when you do a close-listen of the songs with the highest danceability scores.
-Not all of them are danceable, but the one's that aren't still feel like they're trying to manipulate the way your body is holding itself in some way.
-The reason that Antonoff's sound dominating chart-topping pop production gets under people's skin could just be because his music gets underneath your skin.
-
-I interviewed Glen McDonald, a Principal Engineer at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
-Macdonald told me that `valence` and `danceability` were created by giving thousands of examples to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
-Variables like energy or instrumentalness, on the other hand, were determined primarily through machine listening techniques.
-The human subjectivity part only came in when it came to tweaking the training data[^1]:
-
-[^1]: My favourite example Glen gave was that the ML kept giving bluegrass songs insanely high `speechiness` ratings.
-Because there were no banjos in the training data, they were registered as human speech.
-(This is hilarious.) They had to go back and add more songs with banjos to the training data.
-
-"You could imagine writing a formula for energy that combines like loudness and tempo and degree of harmonic variation or something.
-So it was \[machine learning\], but that one's more like a human helping a machine figure out a formula.
+> "You could imagine writing a formula for energy that combines loudness and tempo and degree of harmonic variation or something.
+So that feature was \[machine learning\], but that one's more like a human helping a machine figure out a formula.
 Whereas valence is teaching the machine to try to reproduce a purely human thing.
-Same with danceability.
+The same with danceability.
 I mean, danceability is whether a human can dance to it.
-So like, the machine's not gonna dance, so the machine can't have any opinion on that.
+The machine's not gonna dance, so the machine can't have any opinion on that.
 The computer could have an opinion on energy.
 And the computer can definitely have an opinion on loudness.
-So there's a spectrum from, you know, loudness as purely analytical, and then energy is a little like loudness, with a little more subjectivity.
+So there's a spectrum from, loudness as purely analytical, and then energy is a little like loudness, with a little more subjectivity.
 And then danceability and valence are purely subjective."
 
-Glen confirmed that this process didn't know anything about lyrics, which makes valence a particularly difficult variable to build.
-It could, in theory, pick up on aspects of vocal performance, but it doesn't know anything about languages or words or the meanings of songs.
+Glen confirmed that this process didn't account for lyrics, which made valence a particularly difficult variable to build.
+It could, in theory, pick up on aspects of vocal performance, but it didn't process anything about languages or words or the meanings of songs.
 Take an upbeat, happy-sounding Elliot Smith song with devastating lyrics: the machine might register it as happy, where human listeners understand that it's sad.
-Add to that that two humans might even disagree on the song's valence.
-
-"Plus, we have the confounding factor of like, a song that seems happy today could be sad tomorrow because the singer was killed in a plane crash.
+Furthermore, two humans might even disagree on the song's valence: "Plus, we have the confounding factor of like, a song that seems happy today could be sad tomorrow because the singer was killed in a plane crash.
 The song didn't change, but our world changed and our reactions changed."
 
-Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well".
+Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music.
 High energy and high valence could be generally happy, cheerful, upbeat.
@@
 
+```{r}
+#| label: fig-quadvalence
+#| fig-cap: Valence and Energy Quadrant
+#| echo: false
+#| warning: false
+av<- read_csv(here::here("inputs/data/antonoff_viz.csv"))
+
+av$is_antonoff <- ifelse(av$producer == "antonoff", 1, 0)
+av$is_antonoff <- as.factor(av$is_antonoff)
+
+ggplot(av, aes(x = energy, y = valence)) +
+  geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
+  geom_vline(xintercept = mean(df$energy), color = "gray") +
+  geom_hline(yintercept = mean(df$valence), color = "gray") +
+  theme_minimal()
+```
+
+Our analysis reveals that the amalgamation of crowdsourced metrics and machine-listening variables offers a nuanced approximation of musical style.
+Interestingly, the subjective nature of crowdsourced metrics, which might appear rudimentary, plays a crucial role in shaping a more intuitive and embodied understanding of music.
+This synergy between human subjectivity and algorithmic precision not only mirrors the multifaceted nature of musical perception, but also enriches our metric's capacity to capture the distinctive essence of an artist's style.
+
 \newpage

### 2023-11-10 outputs/paper/paper.qmd
add scores

@@
 
-In the visualization of danceability correlations across producers, Antonoff's danceability has the least positive correlation with valence (SCORE), the most negative correlation with acousticness (SCORE), and significant positive correlations with speechiness (SCORE), loudness (SCORE), and energy (SCORE).
+Relative to other producers, Antonoff's danceability has the least positive correlation with valence (0.311), the most negative correlation with acousticness (-0.0733), and significant positive correlations with speechiness (0.336), loudness (0.366), and energy (0.274).
 

### 2023-11-10 scripts/06-models.R

@@ -129,7 +129,7 @@ best_penalty
 # and evaluates one time on the testing data
 final_fit <-  
   wf_set %>% 
-  extract_workflow("basic_glmnet") %>%
+  extract_workflow("downsampling_glmnet") %>%
   finalize_workflow(best_penalty) %>%
   last_fit(antonoff_split)
 

### 2023-11-10 outputs/paper/paper.qmd
fixed rf latex weeeee

@@
 
-1.  For each iteration ( b = 1 ) to ( B ):
-    a.  Draw a bootstrap sample ( Z\^\* ) of size ( N ) from the training data.
+1.  For each iteration $(b = 1)$ to $( B )$:
+    a.  Draw a bootstrap sample $( Z^* )$ of size $( N )$ from the training data.
     b.
-Grow a random-forest tree ( T_b ) to the bootstrapped data by recursively performing the following steps for each terminal node of the tree until the minimum node size ( n\_{\text{min}} ) is reached:
-        i.  Select ( m ) predictors at random from the ( p ) predictors.
-        ii. Determine the best predictor and split-point among the ( m ).
+Grow a random-forest tree $( T_b )$ to the bootstrapped data by recursively performing the following steps for each terminal node of the tree until the minimum node size $( n_{{min}} )$ is reached:
+        i.  Select $( m )$ predictors at random from the $( p )$ predictors.
+        ii. Determine the best predictor and split-point among the $( m )$.
         iii. Split the node into two daughter nodes.
-2.  The ensemble of trees ( { T_b }\^B\_{b=1} ) constitutes the random forest model.
+2.  The ensemble of trees $( \{ T_b \}_{b=1}^B )$ constitutes the random forest model.
 
-For a new instance ( x ), the RF model prediction ( \hat{C}\^{\text{RF}}(x) ) is determined by the majority vote from the ensemble of trees:
+For a new instance $( x )$, the RF model prediction $( \hat{C}^{RF}(x) )$ is determined by the majority vote from the ensemble of trees:
 
-\[ \hat{C}\^{\text{RF}}(x) = \text{majority vote} \left{ \hat{C}*b(x)* \right}\^B{b=1} \]
+$$
+\hat{C}^{\text{RF}}(x) = \text{majority vote} \left\{ \hat{C}_b(x) \right\}_{b=1}^B 
+$$
 
-Where ( \hat{C}\_b(x) ) is the class prediction of the ( b )-th tree.
+Where $( \hat{C}_b(x) )$ is the class prediction of the $(b)$-th tree.
 

### 2023-11-10 scripts/06-models.R
precision + recall

@@ -142,7 +142,30 @@ antonoffmatrix <- collect_predictions(final_fit) %>%
 
 antonoffmatrix
 
+
+
 antonoffmatrix_df <- as.data.frame(antonoffmatrix$table)
+
+
+
+antonoff_predictions <- collect_predictions(final_fit)
+
+antonoff_predictions$is_antonoff <- factor(antonoff_predictions$is_antonoff, levels = c("1", "0"))
+antonoff_predictions$pred_class <- factor(antonoff_predictions$.pred_class, levels = c("1", "0"))
+
+str(antonoff_predictions$is_antonoff)
+str(antonoff_predictions$pred_class)
+
+conf_matrix <- confusionMatrix(antonoff_predictions$is_antonoff, antonoff_predictions$.pred_class)
+
+precision_value <- conf_matrix$byClass['Pos Pred Value']
+recall_value <- conf_matrix$byClass['Sensitivity']
+
+# Print Precision and Recall
+print(precision_value)
+print(recall_value)
+
+
 
 # install.packages("vip")
 library(vip)

### 2023-11-12 outputs/paper/paper.qmd
Paper tone/style edits, fixed the regression

@@
 
-We find that danceability is the strongest predictor of whether or not a song has been produced by Antonoff.
-Furthermore, when the variance is considered along with other features of the dataset, the variance tends to be strongest classifier of whether a song was produced by Antonoff.
+We find that `danceability` is the strongest predictor of whether or not a song has been produced by Antonoff.
+Furthermore, when `danceability`'s correlation scores with other features in the dataset are considered, the variance in correlation scores tends to be strongest classifier of whether a song was produced by Antonoff.
 Danceability in concert with other features in the dataset provides the best predictions.
@@
 Against the odds, we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
-The remainder of this paper is structured as follows: @ref-background...
+The remainder of this paper is structured as follows:
 
-# Background {#sec-background}
+@ref-data
 
-Shortly after the release of Taylor Swift's *Midnights* album, a video of the Youtuber Caleb Gammen, in which he guesses which songs on Midnights were produced by Antonoff within seconds [@citeTweet], went viral.
-The tweet reads: "im able to instantly detect if jack antonoff worked on a song due to a visceral hatred of his production style".
-He nails it.
-Only one song on the album wasn't produced by Antonoff, and he guesses it.
+@ref-Model
 
-Antonoff began his career as a guitarist in fun., and is the frontman in Bleachers.
-He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more.
-Antonoff is prolific, to the extent that it is starting to bother people.
-As Andrew Marantz writes in the New Yorker: "When his band releases an album, the world responds politely.
-When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis" [@citeMarantz].Can Spotify's audio features dataset help us determine what makes Antonoff music so Antonoff-y?
+@ref-Results
 
-I spoke to Gamman to better undertstand makes a song Antonoff-ish, and to try to understand his reaction to it.
+@ref-Discussion
 
-I also spoke with Glen McDonald, an engineer at Spotify, from whom I learned that the `valence` and `danceability` variables in the Spotify API were created by hiring college interns and having them tag whether songs were positive or danceable.
-
-However, logistic regression assumes a linear relationship between the dependent and predictor variables, and is sensitive to multi-collinearity.
-In music data, we would expect for certain variables to be correlated (for instance, loudness and energy).
-We also cannot assume that an artists' style can be measured in the linear growth of these metrics.
-Although the logistic regression has a strong accuracy, it may be detecting a change in genre, rather than a change in style, across artists whose sound turns more and more pop while collaborating with Antonoff.
-
-In the discussion section, I listen to the songs with the highest danceability scores, and make an argument that this `danceability` feature, which has so much uncertainty cooked into it, is really capturing something like visceral-ness.
+@ref-References
 
@@
 
-The audio feature data was pulled using Spotify API, which I accessed using the `spotifyr` package in R [@citeSpotify].
-I used the `get_artist_audio_features` function to acquire data on Lorde, Taylor Swift, St.
+We collected data from the Spotify API, which we accessed using the `spotifyr` package in R [@citeSpotify].
+We used the `get_artist_audio_features` function to acquire data on Lorde, Taylor Swift, St.
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
-HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski were included, since they've never collaborated with Jack Antonoff but are more or less part of the same pop sphere.
+HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski were included to represent other contemporary pop artists with overlapping fan bases who have never collaborated with Antonoff.
+Of the artists who did collaborate with Antonoff, we chose artist who have produced multiple albums, and at least one album with Antonoff.
 
-I also scraped the discographies of comparable contemporary pop producers' Wikipedias.
+We acquired the discographies of comparable contemporary pop producers' by scraping their respective discography Wikipedia pages.
 The producers include Epworth, Rechstaid, Max Martin, Nowels, Joel Little, and Greg Kustin.
-(cite Carl here?) These producers were chosen based on their popularity, and also because some of them collaborated with artists who also collaborate with Antonoff.
+These producers were chosen based both on their popularity, and because some have collaborated with the same artists as Antonoff.
 
@@
 
-Of the artists who did collaborate with Antonoff, I chose these artists because they've produced multiple albums, and at least one album with Antonoff, as opposed to just individual songs.
+We created a binary variable which contains 1 if Jack Antonoff produced or co-produced the song, and 0 if it was produced by somebody else.
+The information underpinning this discography was drawn from the "Jack Antonoff Production Discography" Wikipedia page [@citewiki].
 
-We created a binary variable which contains 1 if Jack Antonoff produced or co-produced the song, and a 0 if it was produced by someone else.
-The information underpinning this discography was drawn from the "Jack Antonoff Production Discography" Wikipedia page [@citewiki].
+In total, we collected 3738 tracks from 828 different artists. If tracks were produced by producers other than the main producers included in our study, they were marked as "other".
+
+```{r}
+#| label: tbl-producercounts
+#| tbl-cap: Counts of Tracks Per Producer in Original Dataset
+#| echo: false
+#| warning: false
+
+library(kableExtra)
+
+tbl <- table(lr_data$producer)
+
+tbl |>
+  kable(col.names = c("Producer", "Number of Tracks"))
+
+```
 
@@
 
-It's hard to gauge exactly what the variables `danceability` or `valence` mean, even using the documentation.
-For instance, `valence` is described as "A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
-Tracks with high valence sound more positive (e.g.
-happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g.
-sad, depressed, angry)".
-`Danceability` is described as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
-A value of 0.0 is least danceable and 1.0 is most danceable".
-Neither of these definitions fully cover what metrics and thresholds go into something like `danceability`.
-
-In total 3738 tracks from 828 different artists.
-We focus on identifying the styles of 7 producers: Jack Antonoff, Max Martin, Joel Little, Ariel Rechtshaid, Greg Kurstin, Paul Epworth and Rick Nowels.
-If tracks were produced by producers other than the main producers included in our study, they were marked as "other".
-We removed tracks from live performances, karaoke editions, international versions or remix albums.
-Where a deluxe album was available, we removed tracks belonging to the origin to avoid repetition.
-
 ```{r}
@@
 
-In @fig-quadvalence, it looks like songs produced by Antonoff are a little more clustered towards the lower end of valence, and across the spectrum of energy, although the variance is large.
-
-```{r}
-#| label: fig-quaddanceability2
-#| fig-cap: Danceability and Energy Quadrant
-#| echo: false
-#| warning: false
-
-ggplot(av, aes(x = energy, y = danceability)) +
-   geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$danceability),
-              color = "gray") +
-  theme_minimal()
-```
-
-In @fig-quaddanceability2, it looks like Antonoff-produced songs err towards higher danceability, across the spectrum of energy.
-
-These quadrants are a start, but they don't suggest any defining Antonoff-iness characteristic underlying each of those blue points, which are scattered pretty broadly.
-
-By comparing cumulative distribution functions of each variable based on whether a song was produced by Antonoff, we can see that some of these variables do capture some sort of underlying Antonoff-iness (@fig-cdf).
-For instance, while liveness, instrumentalness, and speechiness are relatively similar, we can see clear differences in danceability, energy, loudness, acousticness, and valence.
+By analyzing the cumulative distribution functions of various audio features, particularly when categorizing songs based on whether they were produced by Jack Antonoff (@fig-cdf), distinct patterns emerge.
+Notably, while attributes such as liveness, instrumentalness, and speechiness exhibit relative homogeneity, there are slight disparities in features like danceability, energy, loudness, acousticness, and valence.
 
@@
 
-Another way to gauge the Antonoff-ification of pop music could be to look at how artists' sounds have changed over time, and whether there is any correlation with Antonoff.
-These plots are here to get a glimpse at pop's evolution over time, and where Antonoff fits in that picture.
-Each point represents a song.
+Examining the temporal evolution of artists' sounds, especially in correlation with their collaboration with Antonoff, provides insights into the broader trends in pop music.
+Each point in the following plots represent one track.
 
@@
 
-In @fig-energy, it looks like there's a downward trend in energy scores in artists who work with Antonoff, but that these downward trends are also somewhat in line with the artists' energy trajectories anyway, which could be indicative of broader trends in pop.
-Specifically, Lana Del Rey and Taylor Swift.
-They have enough albums recorded prior to collaborations with Antonoff to be able to observe this.
-Sharon Von Etten, Maggie Rogers, and Mitski, who have never worked with Antonoff, all have upward trends.
+In @fig-energy, a downward trend in the energy scores is observed among artists collaborating with Antonoff.
+However, this trend appears consistent with the overall trajectories of these artists' works, suggesting broader trends in the pop genre, as exemplified by Lana Del Rey and Taylor Swift.
+Sharon Von Etten, Maggie Rogers, and Mitski, display upward energy trajectories in their musical careers.
 
@@
 
-There aren't any significant correlations in danceability (@fig-danceability) .
+No significant trends are evident in danceability across the artists' collaborations with Antonoff, as demonstrated in (@fig-danceability) .
 
@@
 
-In @fig-valence, it looks like Lorde has the only significant change in valence scores, which tracks with my understanding her albums.
-*Melodrama* had piano and break ups, and *Solar Power* had a music video of Lorde prancing on the beach in a yellow bikini with Phoebe Bridgers and Clairo.
+@fig-valence reveals a notable shift in valence scores in Lorde's discography, where her tracks display a slight upward trend and a greater range in valence scores over time.
 
@@
 
-Again in @fig-acousticness, Lorde's acoustic Solar Power album jumps out.
-Taylor Swift, Lorde and HAIM also have significant-seeming upward trends in `acousticness`, while Mitski, Sharon Von Etten and Maggie Rogers see decreasing `acousticness`.
+@fig-acousticness underscores a trend toward increased acousticness in the works of Taylor Swift, Lorde, and HAIM, contrasted with declining trends in artists like Mitski, Sharon Von Etten, and Maggie Rogers.
 
@@
 
-In @fig-instrumentalness, St. Vincent is the only plot with any significant trend in `instrumentalness`.
+@fig-instrumentalness exhibits no significant correlations.
 
@@
 
-Finally, in @fig-tempo, there aren't many significant trends in tempo, except for Antonoff's collaborations with Lana Del Rey and Taylor Swift.
+Finally, tempo trends, as shown in Figure @fig-tempo, are generally stable across artists, with notable exceptions in collaborations involving Lana Del Rey and Taylor Swift.
 
-We've observed subtle distinctions in variables associated with songs produced by Jack Antonoff, as evidenced by our cumulative distribution functions, quadrant analyses, and minor trends in the evolution of artists' sound characteristics, particularly energy.
-Employing a logistic regression will then: a) determine the likelihood of a song being produced by Antonoff based on these identified variables, and b) discern which variables are most influential in shaping this classification.
+We've observed subtle distinctions in variables associated with songs produced by Jack Antonoff, as evidenced by our cumulative distribution functions and minor trends in the evolution of artists' sound characteristics, particularly energy.
+This study sets the stage fora) employing logistic regression to quantify the likelihood of Antonoff's involvement in a song based on these audio features, b) classifying tracks produced by a wider range of producers using a Random Forest, c) identify the most influential variables in this classification.
 
@@
 ## Logistic Regression
+
+The aim of the logistic regression is to determine whether we can predict Antonoff's influence in artists' sound.
+We select a total of 1465 tracks from the dataset from 12 artists: Lorde, Taylor Swift, St.
+Vincent, Lana Del Rey, The Chicks, Florence and the Machine, HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, Mitski, and Bleachers (Antonoff's own band)..
 
@@
 To deal with the imbalanced classes, we downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
-Then we used 5-fold cross-validation with stratification by the **`jack`** variable to estimate the performance of the model.
-
-The `workflow_set()` function lets us hold different sets of feature engineering together and use tuning to evaluate them all at once.
-We need to evaluate two elements: a basic model, and the model that was downsampled using the `themis` package.
-
-We used **`rank_results()`** from the **`yardstick`** package to rank the models based on accuracy, log loss, sensitivity, and specificity.
-After testing both basic and downsampled versions of the model, the basic version yielded much higher accuracy score (0.918 basic versus 0.743 downsampled), although a lower roc auc score (0.746 basic versus 0.821 downsampled).
-
-The version with the original data remains consistently high in accuracy, even with increased regularization.
-It starts low with in mean log loss and sensitivity, but increases with increased regularization.
-Compared to the downsampled data, which starts high but drops significantly in accuracy and sensitivity, we chose to use the original data.
+Then we used 5-fold cross-validation with stratification by the **`is_antonoff`** variable to estimate the performance of the model.
 
@@
 
-The model's accuracy score is 0.918, and the ROC AUC score is 0.746.
+The model's accuracy score is 0.757, and the ROC AUC score is 0.819.
+The precision score is 0.3851, meaning that approximately 38.51% of songs predicted to be produced by Antonoff were actually produced by him.
+The recall of 0.7561 indicates that about 75.61% of songs produced by Antonoff were actually correctly identified in the model.
 
@@
 |------------------------|-----------------------|-----------------------|
-| Predicted Positive (1) | 848                   | 76                    |
-| Predicted Negative (0) | 1                     | 0                     |
+| Predicted Positive (1) | 62                    | 99                    |
+| Predicted Negative (0) | 20                    | 308                   |
 
@@
 
-@tbl-importance shows the different variables' importance on determining whether a song was had any Antonoff involvement.
+@tbl-importance shows variables' respective importance in determining whether a track was produced by Antonoff.
 
@@
 `Danceability`, with a positive importance score of 4.142, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
-`Valence` comes in second at 2.559, with negative score.
-The third, `instrumentalness` correlates negatively with Antonoff tracks with a a negative score of 2.309.
+`Valence` was the second most important, with a negative score of 2.559.
+The third, `instrumentalness` correlates negatively with Antonoff tracks with a negative score of 2.309.
 
@@
 
-@tbl-mostdanceable shows us the 20 songs in Jack Antonoff's discography with the highest `danceability` scores.
-Bleachers takes up 35% of this list, and Taylor Swift takes up 40% of this list.
-St Vincent, The Chicks, Lorde, and Florence and the Machine also haves songs up there.
+In the 20 songs in Jack Antonoff's discography with the highest `danceability` scores, Bleachers tracks account for 35% of this list, and Taylor Swift tracks account 40% of this list (@tbl-mostdanceable).
+St Vincent, The Chicks, Lorde, and Florence and the Machine are also featured.
 
@@
 
-Our results indicate that Antonoff's music, characterized by its danceability score, exhibits a significant positive correlation with factors such as speechiness, loudness, and energy, while showing a negative correlation with acousticness, and a less pronounced positive correlation with valence, compared to other music producers.
+Our logistic regression model was not able to strongly predict whether or not a song was produced by Jack Antonoff using the Spotify API's audio features data.
+These data are limited, providing static metrics that stand in for entire tracks, which may not be detailed enough for intra-genre classification.
+However, when classifying the tracks, `danceability` was the strongest predictor.
+
+Our Random Forest results, which were moderately strong at classifying tracks, indicate that Antonoff's music, characterized by its `danceability` score, exhibits a significant positive correlation with factors such as `speechiness`, `loudness`, and `energy`, while showing a negative correlation with `acousticness`, and a less pronounced positive correlation with `valence`, compared to other music producers.
 This profile---marked by speechiness, loudness, energy, synthetic textures, and a blend of positivity with a touch of nostalgia---aligns with both detailed auditory analysis and critical assessments of Antonoff's sound.
@@
 
-Our analysis reveals that the amalgamation of crowdsourced metrics and machine-listening variables offers a nuanced approximation of musical style.
+Our analysis reveals that the amalgamation of crowdsourced metrics (danceability) and machine-listening variables offers a nuanced approximation of musical style.
 Interestingly, the subjective nature of crowdsourced metrics, which might appear rudimentary, plays a crucial role in shaping a more intuitive and embodied understanding of music.
-This synergy between human subjectivity and algorithmic precision not only mirrors the multifaceted nature of musical perception, but also enriches our metric's capacity to capture the distinctive essence of an artist's style.
+This synergy between human subjectivity and algorithmic precision not only mirrors the multifaceted nature of musical perception, but also enriches our model's capacity to capture the essence of an artist's style.
 

### 2023-11-12 scripts/06-models.R

@@ -33,16 +33,24 @@ write_csv(data_reduced,"inputs/data/data_reduced.csv")
 ## glmnet requires all variables to be numeric, so i'm excluding the producer column
 set.seed(853)
 # 1. Create a new binary column
-lr_data <- data_reduced
-lr_data$is_antonoff <- ifelse(lr_data$producer == "antonoff", 1, 0)
-lr_data$is_antonoff <- as.factor(lr_data$is_antonoff)
+antdata <- read_csv(here::here("inputs/data/antonoff_viz.csv"))
+adf <- antdata |> 
+  select(producer,danceability, energy, loudness, mode, speechiness, acousticness, instrumentalness, liveness, valence,
+         tempo
+  ) 
+adf$is_antonoff <- ifelse(adf$producer == "antonoff", 1, 0)
+adf$is_antonoff <- as.factor(adf$is_antonoff)
+adf <- adf |>
+  select(is_antonoff, danceability, energy, loudness, mode, speechiness, acousticness, instrumentalness, liveness, valence,
+              tempo
+) 
 
 
-write_csv(lr_data, "inputs/data/lr_data.csv")
-
+#write_csv(lr_data, "inputs/data/lr_data.csv")
+adf$is_antonoff <- factor(adf$is_antonoff, levels = c("1", "0"))
 # 2. Update variable names and strata
 antonoff_split <- 
-  lr_data |>
+  adf |>
   initial_split(strata = is_antonoff)
 
 antonoff_train <- training(antonoff_split)
@@ -52,8 +60,7 @@ antonoff_folds <- vfold_cv(antonoff_train, strata = is_antonoff)
 
 # 3. Update the recipe
 is_antonoff <- 
-  recipe(is_antonoff ~ ., data = antonoff_train) %>%
-  step_rm(producer)
+  recipe(is_antonoff ~ ., data = antonoff_train) 
 
 
 # we can `prep()` just to check that it works:
@@ -207,28 +214,29 @@ antonoff_split <-
 antonoff_train <- training(antonoff_split)
 antonoff_test <- testing(antonoff_split)
 
-naive_prior <- normal(0, 10) #wide normal distribution approximation to flat prior
-antonoff_train$is_antonoff <- factor(antonoff_train$is_antonoff, levels = c("0", "1"))
+naive_prior <- normal(0, 5) #wide normal distribution approximation to flat prior
+
 
 
 library(caret)
 library(rstanarm)
[…]

### 2023-11-13 outputs/paper/paper.qmd
PDF render

@@
 abstract: "This paper explores whether or not we can predict whether a song was produced by Jack Antonoff using the audio features provided by Spotify's API.
-We find that in general our model has an accuracy score of 90%.
-(Precision and recall) We interestingly found that danceability, a metric trained on human-tagged data, was a key predictor.
-Our results have implications for the understanding of culture, how the combination human subjectivity and machine listening can come to create an approximate for style."
+We find that in general our logistic regression model has an accuracy score of 76%, a precision score of 39%, and a recall of 75.61%.
+Our random forest has an accuracy of 61%, and a No Information Rate of 43%.
+We interestingly found that danceability, a metric trained on human-tagged data, was a key predictor.
+Our results have implications for computational explorations of culture, how the combination human subjectivity and machine listening can come to create an approximate for style."
 format: pdf
@@
 #| warning: false
+lr_data<- read_csv(here::here("inputs/data/lr_data.csv"))
 
@@
 
-Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music.
+Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence).
 High energy and high valence could be generally happy, cheerful, upbeat.
@@
 #| warning: false
-av<- read_csv(here::here("inputs/data/antonoff_viz.csv"))
+av <- read.csv(here::here("inputs/data/antonoff_viz.csv"))
 
@@
   geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
-  geom_vline(xintercept = mean(df$energy), color = "gray") +
-  geom_hline(yintercept = mean(df$valence), color = "gray") +
+  geom_vline(xintercept = mean(av$energy), color = "gray") +
+  geom_hline(yintercept = mean(av$valence), color = "gray") +
   theme_minimal()

### 2023-11-13 scripts/01_b-producerwikiscrapers.py
adding scraper notebook

@@ -1,0 +1,831 @@
+# -*- coding: utf-8 -*-
+"""ProducerWikiScrapers.ipynb
+
+
+# Acquiring More Data for the Model
+
+The purpose of the following scripts is to acquire more data to add to the existing dataset, which contains the discographies of artists who Antonoff has collaborated with both before and during their collaborations. The songs they collaborated with are tagged in the binary column "jack", with 1 indicated that the song is a collaboration.
+
+The first new kind of data is **songs produced by similar pop producers**. This involves scraping discographies from producers' wikipedia pages, and requesting audio features using the `audio_features()` function from the  `spotipy` package. The scripts for each producer are the same, with slight variations for descripancies scraping Wikipedia pages (for instance, unreleased songs listed on Wikipedia are not included in the request).
+
+The second new kind of data is **similar songs**. Here, I use a custom `get_recommendations` function, which is similar to the `recommendations` function in `spotipy`. This function accesses the Spotify Web API's recommendations endpoint using the existing dataset's tracks as seed tracks to fetch tracks that might share musical characteristics with the songs from the existing dataset. In order to manage rate-limiting, it retrieves recommendations in batches of five, and waits and retries, if necessary. (As of right now, I'm only using songs produced by Antonoff as seed tracks).
+
+
+
+------
+
+### Appendix:
+
+Carl Wilson offered suggestions on who to compare Antonoff's production style to: " **Max Martin**  is definitely one who comes to mind because of the contrast via Swift. But comparing to Dr. Luke is going to get you more genre differences than production style contrast. for people a little closer to the areas Jack works in - you might look at ppl like **Greg Kurstin** , **Paul Epworth** (both Adele producers but also for many others), **Mark Ronson**, **Joel Little** (Lorde’s producer before Jack), **Ariel Rechtshaid**, maybe **Dave Cobb** or **Jay Joyce** for the country end, even **Finneas** (Billie Eilish’s brother who pretty much only works with her) because his approach has some overlap with Antonoff’s, I think."
+
+As a list:
+- Max Martin
+- Greg Kurstin
+- Paul Epworth
+- Mark Ronson
+- Joel Little
+- Ariel Rechtshaid
[…]

### 2023-11-17 outputs/paper/paper.qmd
line edits, fixed first half of plots and tables

@@
 ---
-title: "Detecting Stylistic Variation in Pop Production using the Spotify API"
+title: "Detecting Stylistic Variation in Pop Production"
 author: 
@@
 date: today
-abstract: "This paper explores whether or not we can predict whether a song was produced by Jack Antonoff using the audio features provided by Spotify's API.
-We find that in general our logistic regression model has an accuracy score of 76%, a precision score of 39%, and a recall of 75.61%.
-Our random forest has an accuracy of 61%, and a No Information Rate of 43%.
-We interestingly found that danceability, a metric trained on human-tagged data, was a key predictor.
-Our results have implications for computational explorations of culture, how the combination human subjectivity and machine listening can come to create an approximate for style."
+abstract: "We explore whether we can predict if a song was produced by Jack Antonoff using the audio features provided by Spotify's API.
+Our logistic regression model has an accuracy score of 76%, a precision score of 39%, and a recall of 76%.
+Our random forestmodel has an accuracy of 61%, and a No Information Rate of 43%.
+We especially found that danceability, a metric trained on human-tagged data, was a key predictor.
+Our results have implications for computational explorations of culture, how the combination of human subjectivity and machine listening can come to create an approximate for style."
 format: pdf
@@
 
-We're interested in understanding how to create an approximate for production style.
+We are interested in understanding how to create an approximate for production style.
 To do this we used Spotify API's audio features dataset to explore different audio feature metrics and identify which were the most import predictors in identifying pop producers' style.
-We chose Jack Antonoff as focus, as he is a prolific and widely used producer, who also has his own solo acts.
+We focus on Jack Antonoff, as he is a prolific and widely used producer, who also has his own solo acts.
 
@@
 We obtained tracks from six artists who've collaborated with Antonoff, both before their collaboration and during.
-We also obtained tracks from similarly prolific pop producers.
-The producers we compare Antonoff's production style with are Little, Rechstaid, Martin, Kurstin, Epworth and Nowels.
+We also obtained tracks from similarly prolific pop producers: Joel Little, Ariel Rechstaid, Max Martin, Greg Kurstin, Paul Epworth and Rick Nowels.
 Our data set contains 3738 tracks in total.
@@
 We built two models to explore the data.
-The first model, a logistic regression, draws on the dataset of artists who've collaborated with Antonoff, with a binary variable to tag whether or not the song was an Antonoff production.
+The first model, a logistic regression, draws on the dataset of artists who have collaborated with Antonoff, with a binary variable to tag whether or not the song was an Antonoff production.
 With this model, we explore which variables are stronger predictors of Antonoff's style, developping a better understanding of whether the Spotify data can trace Antonoff's influence on an artist's sound over time.
@@
 
-Against the odds, we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
+Our results suggest that we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
 The remainder of this paper is structured as follows:
-
-@ref-data
-
-@ref-Model
-
-@ref-Results
-
-@ref-Discussion
-
-@ref-References
 
@@
 We collected data from the Spotify API, which we accessed using the `spotifyr` package in R [@citeSpotify].
-We used the `get_artist_audio_features` function to acquire data on Lorde, Taylor Swift, St.
+We acquired data on Lorde, Taylor Swift, St.
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
@@
 tbl |>
-  kable(col.names = c("Producer", "Number of Tracks"))
+  kable(col.names = c("Producer", "Number of Tracks"),
+  digits = 1,
+  booktabs = TRUE,
+  linesep = "")
 
@@
 
-The `get_artist_audio_features` provides a profile of each song, according to variables developped by Spotify, which makes for simple comparisons across artists.
-The final audio features in this study data set contains the variables: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.
+Spotify provides variables for each song, according to variables developed by Spotify, which makes for simple comparisons across artists.
+The final audio features in our data set are: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.
 
@@
 
-lr_data |>
-  group_by(is_antonoff) |>
-  summarise(across(-c(tempo, loudness), mean)) |>
-  pivot_longer(cols = -is_antonoff) |>
-  ggplot(aes(value, name, fill = factor(is_antonoff))) +
-  geom_col(alpha = 0.8, position = "dodge") +
+lr_data %>%
+  select(-producer, -tempo, -loudness) %>%  
+  pivot_longer(cols = -is_antonoff, names_to = "feature", values_to = "value") %>%
+  ggplot(aes(x = value, y = feature, color = factor(is_antonoff))) +
+  geom_point(alpha = 0.8, position = position_jitter(width = 0.2)) +
+  scale_color_manual(values = c("blue", "orange"), labels = c("No", "Yes")) +
   theme_minimal() +
-  labs(fill = "Jack", x = "Mean Value", y = "Feature")
+  labs(color = "Antonoff", x = "Value", y = "Feature")
+
 ```
@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| label: fig-cdf
-#| fig-cap: Cumulative Distribution Function of Each Variable, 
+#| fig-cap: Cumulative Distribution Function of Each Variable 
 #| echo: false
@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| label: fig-energy
@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| label: fig-danceability
@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| label: fig-valence
@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| label: fig-acousticness
@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| label: fig-instrumentalness
@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| label: fig-tempo
@@
 We select a total of 1465 tracks from the dataset from 12 artists: Lorde, Taylor Swift, St.
-Vincent, Lana Del Rey, The Chicks, Florence and the Machine, HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, Mitski, and Bleachers (Antonoff's own band)..
+Vincent, Lana Del Rey, The Chicks, Florence and the Machine, HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, Mitski, and Bleachers (Antonoff's own band).
 
@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| echo: false

### 2023-11-17 outputs/paper/paper.qmd
cleaned up plots

@@
 
-```{r}
+```{r, fig.width= 8, fig.height=6}
 #| label: fig-means
@@
 lr_data %>%
-  select(-producer, -tempo, -loudness) %>%  
+  select(-producer, -tempo, -loudness, -mode) %>%
   pivot_longer(cols = -is_antonoff, names_to = "feature", values_to = "value") %>%
-  ggplot(aes(x = value, y = feature, color = factor(is_antonoff))) +
-  geom_point(alpha = 0.8, position = position_jitter(width = 0.2)) +
-  scale_color_manual(values = c("blue", "orange"), labels = c("No", "Yes")) +
+  ggplot(aes(x = feature, y = value, fill = factor(is_antonoff))) +
+  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
+  scale_fill_manual(values = c("blue", "orange"), labels = c("No", "Yes")) +
   theme_minimal() +
-  labs(color = "Antonoff", x = "Value", y = "Feature")
+  labs(fill = "Antonoff", x = "Feature", y = "Value") + 
+  coord_flip()
 
@@
                              "Importance",
-                             "Sign"))
+                             "Sign"),
+               digits = 3,
+               booktabs = TRUE,
+               linesep = "")
+
 ```
@@
                              "Danceability",
-                             "Track"))
+                             "Track"),
+  digits = 3,
+  booktabs = TRUE,
+  linesep = "")
+
 ```
@@
 
-print(range_data)
+range_data |>
+  kable(col.names = 
+          c("Variable", "Min Value", "Max Value", "Range"),
+        digits = 3,
+        booktabs = TRUE,
+        linesep = "")
 
@@
   geom_bar(stat="identity", position="dodge") +
-  labs(title="Danceability across Producers", y="Danceability Score") +
+  labs(title="Danceability across Producers", y="Danceability Score", x= "Feature") +
   theme_minimal() +
@@
   geom_hline(yintercept = mean(av$valence), color = "gray") +
-  theme_minimal()
+  theme_minimal() +
+  scale_color_manual(values = c("lightblue", "orange"),
+                     labels = c("No", "Yes"),
+                     name = "Antonoff")
 ```

### 2023-11-17 outputs/paper/paper.qmd
plot edits

@@
 thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop."
-date: today
+date: "November 10 2023"
 abstract: "We explore whether we can predict if a song was produced by Jack Antonoff using the audio features provided by Spotify's API.
@@
 $$
-Pr(yi​=1)=logit−1(β0​+β1​×danceability+β2​×energy+...+β10​×tempo)
+Pr(y_i​=1)=logit−1(β_0​+β_1​×danceability+β_2​×energy+...+β_0​×tempo)
 $$

### 2023-11-17 outputs/paper/paper.qmd
mla

@@
 bibliography: references.bib
+csl: modern-language-association.csl
 ---

### 2023-11-17 outputs/paper/paper.qmd
more tweaks

@@
 Our logistic regression model has an accuracy score of 76%, a precision score of 39%, and a recall of 76%.
-Our random forestmodel has an accuracy of 61%, and a No Information Rate of 43%.
+Our random forest model has an accuracy of 61%, and a No Information Rate of 43%.
 We especially found that danceability, a metric trained on human-tagged data, was a key predictor.
@@
 av$is_antonoff <- as.factor(av$is_antonoff)
+levels(av$is_antonoff) <- c("No", "Yes") 
  av |>
@@
    labs(x = "Value",
-        y = "Proportion")
+        y = "Proportion",
+        color = "Antonoff")
 
@@
   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Energy") +
+  labs(color = "Antonoff", x = "Album Release Date", y = "Energy") +
   facet_wrap(~ artist_name, ncol = 3)
@@
   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Danceability") +
+  labs(color = "Antonoff", x = "Album Release Date", y = "Danceability") +
   facet_wrap( ~ artist_name, ncol = 3)
@@
   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Valence") +
+  labs(color = "Antonoff", x = "Album Release Date", y = "Valence") +
   facet_wrap( ~ artist_name, ncol = 3)
@@
   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Acousticness") +
+  labs(color = "Antonoff", x = "Album Release Date", y = "Acousticness") +
   facet_wrap( ~ artist_name, ncol = 3)
@@
   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Instrumentalness") +
+  labs(color = "Antonoff", x = "Album Release Date", y = "Instrumentalness") +
   facet_wrap( ~ artist_name, ncol = 3)
@@
   scale_color_manual(values = c("#808080", "#FFB6C1")) +
-  labs(color = "Has 'y' in jack?", x = "Album Release Date", y = "Tempo") +
+  labs(color = "Antonoff", x = "Album Release Date", y = "Tempo") +
   facet_wrap( ~ artist_name, ncol = 3)
@@
 $$
-Pr(y_i​=1)=logit−1(β_0​+β_1​×danceability+β_2​×energy+...+β_0​×tempo)
+Pr(y_i​=1)=logit−1(β_0​+β_1​×danceability+β_2​×energy+...+β_{10}​×tempo)
 $$

### 2023-11-21 outputs/paper/paper.qmd
edits

@@
 He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more.
-His style is considered identifiable.
-As Andrew Marantz writes in the *New Yorker*: "When his band releases an album, the world responds politely.
+His style is considered identifiable, and his collaborators are influential -- as Andrew Marantz writes in the *New Yorker*: "When his band releases an album, the world responds politely.
 When he produces one by Lorde or Lana Del Rey or Taylor Swift, the world wobbles on its axis" [@citeMarantz].
@@
 The first model, a logistic regression, draws on the dataset of artists who have collaborated with Antonoff, with a binary variable to tag whether or not the song was an Antonoff production.
-With this model, we explore which variables are stronger predictors of Antonoff's style, developping a better understanding of whether the Spotify data can trace Antonoff's influence on an artist's sound over time.
+With this model, we explore which variables are stronger predictors of Antonoff's style, developping a better understanding of whether the Spotify data can trace Antonoff's involvement in an artist's sound over time.
 We built a second model, a Random Forest (RF), to complement the insights drawn from the logistic regression.
-The RF trains on the full data set, including other major pop producers.
+The RF trains on the full data set, including other major pop producers, to classify tracks according to who produced them, and return information on which predictors were most useful in doing so.
 
@@
 Furthermore, when `danceability`'s correlation scores with other features in the dataset are considered, the variance in correlation scores tends to be strongest classifier of whether a song was produced by Antonoff.
-Danceability in concert with other features in the dataset provides the best predictions.
+In other words, danceability in concert with other features in the dataset provides the best predictions.
 
 Our results suggest that we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
-The remainder of this paper is structured as follows:
+With the understanding that what we understand as "data-driven" in recommendation systems is often deeply intertwined with human subjectivity \-- engineers\' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) \-- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.
 
@@
 
-In @fig-energy, a downward trend in the energy scores is observed among artists collaborating with Antonoff.
-However, this trend appears consistent with the overall trajectories of these artists' works, suggesting broader trends in the pop genre, as exemplified by Lana Del Rey and Taylor Swift.
-Sharon Von Etten, Maggie Rogers, and Mitski, display upward energy trajectories in their musical careers.
+In @fig-energy, we observe a downward trend in energy scores among artists collaborating with Antonoff.
+However, this trend appears consistent with the overall trajectories of these artists' works, suggesting broader shared trends in this area of pop, as exemplified by Lana Del Rey and Taylor Swift.
+Sharon Von Etten, Maggie Rogers, and Mitski, display upward energy trajectories in their discographies.
 
@@
 
-No significant trends are evident in danceability across the artists' collaborations with Antonoff, as demonstrated in (@fig-danceability) .
+No significant trends are obersable in danceability over time across the artists' collaborations with Antonoff, as demonstrated in (@fig-danceability), although there is observable an variance in danceability scores across songs in every album, across all artists.
+This suggests that, over time, danceability scores across tracks in every album are consistently varies over time.
+The Chicks earlier work, between 2002 and 2005, presumable the country album "Home" have the smallest amount of variance in danceability.
+This could be due to genre difference.
 
@@
 @fig-valence reveals a notable shift in valence scores in Lorde's discography, where her tracks display a slight upward trend and a greater range in valence scores over time.
+Otherwise, similarly to danceability, the valence scores display significant variance across tracks in every album.
 
@@
 
-@fig-instrumentalness exhibits no significant correlations.
+@fig-instrumentalness exhibits no significant correlations, and scores are consistently low for most artists, excluding St.
+Vincent, whose tracks display both a consistently high degree of variance and a noticeable upward trend over time.
 
@@
 Finally, tempo trends, as shown in Figure @fig-tempo, are generally stable across artists, with notable exceptions in collaborations involving Lana Del Rey and Taylor Swift.
+In Lana Del Rey's tracks, her collaborations with Antonoff see an upward trend in tempo.
+In Taylor Swift's tracks, her collaborations with Antonoff exhibit a downward trend.
 
-We've observed subtle distinctions in variables associated with songs produced by Jack Antonoff, as evidenced by our cumulative distribution functions and minor trends in the evolution of artists' sound characteristics, particularly energy.
-This study sets the stage fora) employing logistic regression to quantify the likelihood of Antonoff's involvement in a song based on these audio features, b) classifying tracks produced by a wider range of producers using a Random Forest, c) identify the most influential variables in this classification.
+We've observed subtle trends in variables associated with songs produced by Jack Antonoff, as evidenced by our cumulative distribution functions and minor trends in the evolution of artists' sound characteristics, when collaborating with Antonoff (particularly energy).
+
+To be clear, we do not suggest that Antonoff's involvement is the cause for the change in artist's sound.
+Arguments of "Antonoffication" risk removing agency and artistic credit from (female) artists, and attributing both their critical success and failures to Antonoff (@citeWilson).
+In an industry so dominated by male producers, and using features so limited in their ability to describe anything about the process of creation, it would not be appropriate the use these trends to suggest that Antonoff's involvement is the cause behind changes in sound.
+The purpose of these visualizations is to observe the distribution of variables across artists, over time, with a focus on Antonoff.
+
+This exploratory data analysis sets the stage for a) employing logistic regression to quantify the likelihood of Antonoff's involvement in a song based on these audio features, b) classifying tracks produced by a wider range of producers using a Random Forest, c) identifying the most influential variables in this classification to better understand how style can be inferred from the provided data set features.
 
@@
 Our Random Forest results, which were moderately strong at classifying tracks, indicate that Antonoff's music, characterized by its `danceability` score, exhibits a significant positive correlation with factors such as `speechiness`, `loudness`, and `energy`, while showing a negative correlation with `acousticness`, and a less pronounced positive correlation with `valence`, compared to other music producers.
-This profile---marked by speechiness, loudness, energy, synthetic textures, and a blend of positivity with a touch of nostalgia---aligns with both detailed auditory analysis and critical assessments of Antonoff's sound.
+This profile---marked by speechiness, loudness, energy, and a blend of positivity with a touch of nostalgia---aligns with both detailed auditory analysis and critical assessments of Antonoff's sound.
 
-Antonoff's tracks have been described as anthemic, with a distinct vocal treatment that makes it sound, as Gamman (2022) observes, "the way it would sound to the person who's singing them".
+Antonoff's tracks have been described as anthemic (@citeRosen), with a distinct vocal treatment that makes it sound, as Caleb Gamman, a Youtuber who went viral for his critique of Jack Antonoffs sound, observes, "the way it would sound to the person who's singing them" (@citegammanint).
 
@@
 
-What does the danceability variable signify in Antonoff's music?
+What does the danceability variable then signify in Antonoff's music?
 Our findings suggest that his most "danceable" songs are better understood through the lens of correlated features, even though danceability emerges as the primary predictor.
@@
 We spoke with Glen McDonald, a Principal Engineer at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
+While Macdonald did not build the features, he has worked with them, and provided information on their provenance.
 Macdonald revealed that `valence` and `danceability` were created by giving tracks to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
-Variables like `energy` or `instrumentalness`, on the other hand, were determined primarily through machine listening techniques, with human subjectivity being applied to fine-tune the features[^1]:
+Variables like `energy` or `instrumentalness`, on the other hand, were determined primarily through machine listening techniques, with human subjectivity being applied to fine-tune the features[^1] (@citeglenint).:
 
@@
 Furthermore, two humans might even disagree on the song's valence: "Plus, we have the confounding factor of like, a song that seems happy today could be sad tomorrow because the singer was killed in a plane crash.
-The song didn't change, but our world changed and our reactions changed."
+The song didn't change, but our world changed and our reactions changed" (@citeglenint).
 
-Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence).
+Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence) (@citeglenint) .
 High energy and high valence could be generally happy, cheerful, upbeat.
@@
 
+While these features offer a quantitative lens to examine music, they cannot fully encapsulate the nuanced and subjective aspects of musical style and creativity.
+The role of human subjectivity in interpreting and experiencing music remains paramount, as evidenced by the crowdsourced nature of certain features like danceability and valence.
+This extends to a more nuanced understanding of recommendation systems as a whole.
+As Nick Seaver anthropological study of music recommendation companies illustrates, engineers \"develop ways of thinking about musical preference, and about software, that try to reconcile them with each other\" (@citeSeaver).
+
 \newpage

### 2023-11-21 outputs/paper/paper.qmd
more edits

@@
 Our results suggest that we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
-With the understanding that what we understand as "data-driven" in recommendation systems is often deeply intertwined with human subjectivity \-- engineers\' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) \-- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.
+With the understanding that what we understand as "data-driven" in recommendation systems is often deeply intertwined with human subjectivity \-- engineers' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) \-- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.
 
@@
 
-Our analysis reveals that the amalgamation of crowdsourced metrics (danceability) and machine-listening variables offers a nuanced approximation of musical style.
-Interestingly, the subjective nature of crowdsourced metrics, which might appear rudimentary, plays a crucial role in shaping a more intuitive and embodied understanding of music.
-This synergy between human subjectivity and algorithmic precision not only mirrors the multifaceted nature of musical perception, but also enriches our model's capacity to capture the essence of an artist's style.
+Our analysis reveals that the amalgamation of crowdsourced metrics (like danceability) and machine-listening variables offers a nuanced approximation of musical style.
+This approach not only mirrors the multifaceted nature of auditory perception but also enriches our model's capacity to capture the essence of an artist's style.
+Interestingly, the subjective nature of crowdsourced metrics, which might initially seem rudimentary, plays a crucial role in this process.
+They shape a more intuitive and embodied understanding of music, enhancing the overall analysis.
 
-While these features offer a quantitative lens to examine music, they cannot fully encapsulate the nuanced and subjective aspects of musical style and creativity.
-The role of human subjectivity in interpreting and experiencing music remains paramount, as evidenced by the crowdsourced nature of certain features like danceability and valence.
-This extends to a more nuanced understanding of recommendation systems as a whole.
-As Nick Seaver anthropological study of music recommendation companies illustrates, engineers \"develop ways of thinking about musical preference, and about software, that try to reconcile them with each other\" (@citeSeaver).
+For instance, the work of Martikainen et al.
+illustrates this point effectively (@podstyle).
+They combine open-source audio tools with qualitative, crowd-sourced listener data to understand audio-based stylistic variation in podcasts.
+This synergy between human subjectivity and algorithmic precision underlines the importance of blending these methods for a comprehensive analysis.
+
+However, it's important to note that while these features offer a quantitative lens to examine music, they cannot fully encapsulate the nuanced and subjective aspects of musical style and creativity.
+The role of human subjectivity is pivotal in interpreting and experiencing music, as evidenced by the crowdsourced nature of certain features like danceability and valence.
+This aspect becomes even more significant when considering the broader scope of music recommendation systems.
+
+As Nick Seaver's anthropological study of music recommendation companies shows, engineers "develop ways of thinking about musical preference and software, attempting to reconcile them with each other" (@citeSeaver).
+This extends to a more nuanced understanding of recommendation systems as a whole, highlighting the intricate balance between computational measures and human interpretation in the evolving landscape of music analysis and appreciation.
+
+Future studies should look into reproducing crowdsourced danceability or valence variables to better understand the underlying user preferences that motivate these metrics.
+In doing so, we can build more robust ways to computationally model more ephemeral and subjective aspects of music such as style.
 

### 2023-11-22 outputs/paper/paper.qmd
more editz

@@
 #| label: fig-dancevariance
+#| fig-cap: Danceability Scores Across Producers
 #| echo: false
@@
   geom_bar(stat="identity", position="dodge") +
-  labs(title="Danceability across Producers", y="Danceability Score", x= "Feature") +
+  labs(y="Danceability Score", x= "Feature") +
   theme_minimal() +

### 2023-11-28 outputs/paper/paper.qmd
lots of edits

@@
 author: 
-  - Michaela Drouillard
-thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop."
-date: "November 10 2023"
+  - name: Michaela Drouillard
+    email: "michaela.drouillard@mail.utoronto.ca"
+thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop. We thank Inessa De Angelis for valuable suggestions."
+date: "November 24 2023"
 abstract: "We explore whether we can predict if a song was produced by Jack Antonoff using the audio features provided by Spotify's API.
-Our logistic regression model has an accuracy score of 76%, a precision score of 39%, and a recall of 76%.
+Our logistic regression model has an accuracy of 76%, a precision of 39%, and a recall of 76%.
 Our random forest model has an accuracy of 61%, and a No Information Rate of 43%.
-We especially found that danceability, a metric trained on human-tagged data, was a key predictor.
-Our results have implications for computational explorations of culture, how the combination of human subjectivity and machine listening can come to create an approximate for style."
+We found that danceability, a metric trained on human-tagged data, was a key predictor.
+Our results have implications for computational explorations of culture, and understanding how the combination of human subjectivity and machine listening can come to create an approximate for style."
 format: pdf
@@
 
-We are interested in understanding how to create an approximate for production style.
+We are interested in understanding how to create an approximation for production style.
 To do this we used Spotify API's audio features dataset to explore different audio feature metrics and identify which were the most import predictors in identifying pop producers' style.
@@
 By way of background, Antonoff began his career as a guitarist in fun., and is the frontman in Bleachers.
-He has produced and co-produced music for Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, the Minions: The Rise of Gru Soundtrack, and more.
+He has produced and co-produced music for artists including Lorde, Taylor Swift, Florence and the Machine, The Chicks, Clairo, The 1975, Grimes, Zayn, Pink, Lana Del Rey, Olivia Rodrigo, and the Minions: The Rise of Gru soundtrack.
 His style is considered identifiable, and his collaborators are influential -- as Andrew Marantz writes in the *New Yorker*: "When his band releases an album, the world responds politely.
@@
 We are interested in whether Spotify audio features can be used to train a model that distinguishes whether a song was produced by Antonoff.
-We obtained tracks from six artists who've collaborated with Antonoff, both before their collaboration and during.
+We obtained tracks from six artists who have collaborated with Antonoff, both before their collaboration and during.
 We also obtained tracks from similarly prolific pop producers: Joel Little, Ariel Rechstaid, Max Martin, Greg Kurstin, Paul Epworth and Rick Nowels.
-Our data set contains 3738 tracks in total.
+Our data set contains 3738 tracks in total, which were split into a training set (2802 tracks) and a testing set (936 tracks).
 
 We built two models to explore the data.
-The first model, a logistic regression, draws on the dataset of artists who have collaborated with Antonoff, with a binary variable to tag whether or not the song was an Antonoff production.
-With this model, we explore which variables are stronger predictors of Antonoff's style, developping a better understanding of whether the Spotify data can trace Antonoff's involvement in an artist's sound over time.
-We built a second model, a Random Forest (RF), to complement the insights drawn from the logistic regression.
-The RF trains on the full data set, including other major pop producers, to classify tracks according to who produced them, and return information on which predictors were most useful in doing so.
+The first model, a logistic regression, draws on a dataset of artists who have collaborated with Antonoff (1954 tracks total), with a binary variable to tag whether the song was an Antonoff production.
+The training set included 1465 tracks, and the testing set included 489 tracks.
+With this model, we explore which variables are stronger predictors of Antonoff's style, developing a better understanding of whether the Spotify data can trace Antonoff's involvement in an artist's sound over time.
+We built a second model, a Random Forest (RF), to complement the insights from the logistic regression.
+The RF trains on a training set that includes other major pop producers in order to classify tracks based on who produced them, and return information on which predictors were most useful in doing so.
 
-We find that `danceability` is the strongest predictor of whether or not a song has been produced by Antonoff.
+We find that `danceability` is the strongest predictor of whether a song has been produced by Antonoff.
 Furthermore, when `danceability`'s correlation scores with other features in the dataset are considered, the variance in correlation scores tends to be strongest classifier of whether a song was produced by Antonoff.
@@
 
-Our results suggest that we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools.
-With the understanding that what we understand as "data-driven" in recommendation systems is often deeply intertwined with human subjectivity \-- engineers' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) \-- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.
+Our results suggest that we can capture something about Antonoff's music with these variable.
+It might, in the end, just be that subjective response that some people feel, and others enjoy, and that we are all just trying to describe in using different methods and tools.
+With the understanding that what we understand as "data-driven" in recommendation systems is often intertwined with human subjectivity -- engineers' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) -- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.
+
+The remainder of this paper is structured as:
 
@@
 We collected data from the Spotify API, which we accessed using the `spotifyr` package in R [@citeSpotify].
-We acquired data on Lorde, Taylor Swift, St.
+For our first dataset, we acquired data on Lorde, Taylor Swift, St.
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
 HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski were included to represent other contemporary pop artists with overlapping fan bases who have never collaborated with Antonoff.
-Of the artists who did collaborate with Antonoff, we chose artist who have produced multiple albums, and at least one album with Antonoff.
+Of the artists who did collaborate with Antonoff, we chose artists who have produced multiple albums, and at least one album with Antonoff.
+This dataset contains 1954 tracks in total.
 
-We acquired the discographies of comparable contemporary pop producers' by scraping their respective discography Wikipedia pages.
+For our second dataset, we acquired the discographies of comparable contemporary pop producers' by scraping their respective discography Wikipedia pages.
 The producers include Epworth, Rechstaid, Max Martin, Nowels, Joel Little, and Greg Kustin.
 These producers were chosen based both on their popularity, and because some have collaborated with the same artists as Antonoff.
+We combined these tracks with the first datasets tracks, for a total of 3738 tracks from 828 different artists[^1].
+
+[^1]: In the full, combined version of the dataset, there were tracks (for instance, by Mitski or Maggie Rogers) that were produced by producers other than the main producers included in our study.
+The producer columns for these tracks were marked as "other".
 
@@
 
-We created a binary variable which contains 1 if Jack Antonoff produced or co-produced the song, and 0 if it was produced by somebody else.
+We created a binary variable, `is_antonoff`, which contains 1 if Jack Antonoff produced or co-produced the song, and 0 if it was produced by somebody else.
 The information underpinning this discography was drawn from the "Jack Antonoff Production Discography" Wikipedia page [@citewiki].
-
-In total, we collected 3738 tracks from 828 different artists. If tracks were produced by producers other than the main producers included in our study, they were marked as "other".
 
@@
 
-Spotify provides variables for each song, according to variables developed by Spotify, which makes for simple comparisons across artists.
+Spotify provides variables for each song, some of which were developed by them, which enables comparisons across artists.
 The final audio features in our data set are: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.
 
-From the `get_artist_audio_features` documentation (@citeSpotifyDocumentation):
+@fig-means depicts the range of scores for the tracks across all variables except for tempo, mode, and loudness, which are measured using different metrics.
+Spotify documents the `get_artist_audio_features` variables as follows (@citeSpotifyDocumentation):
 
-> -   **acousticness**: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.
+> **acousticness**: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.
 >
-> -   **danceability**: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+> **danceability**: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
 A value of 0.0 is least danceable and 1.0 is most danceable.
 >
-> -   **energy**: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
+> **energy**: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
 Typically, energetic tracks feel fast, loud, and noisy.
@@
 >
-> -   **instrumentalness**: Predicts whether a track contains no vocals.
+> **instrumentalness**: Predicts whether a track contains no vocals.
 "Ooh" and "aah" sounds are treated as instrumental in this context.
@@
 >
-> -   **speechiness:** Speechiness detects the presence of spoken words in a track.
+> **speechiness:** Speechiness detects the presence of spoken words in a track.
 The more exclusively speech-like the recording (e.g.
@@
 >
-> -   **key**: The key the track is in.
+> **key**: The key the track is in.
 Integers map to pitches using standard [Pitch Class notation](https://en.wikipedia.org/wiki/Pitch_class).
@@
 >
-> -   **liveness**: Detects the presence of an audience in the recording.
+> **liveness**: Detects the presence of an audience in the recording.
 Higher liveness values represent an increased probability that the track was performed live.
@@
 >
-> -   **loudness:** The overall loudness of a track in decibels (dB).
+> **loudness:** The overall loudness of a track in decibels (dB).
 Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks.
@@
 >
-> -   **mode:** Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.
+> **mode:** Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.
 >
-> -   **tempo:** The overall estimated tempo of a track in beats per minute (BPM).
-In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration.
+> **tempo:** The overall estimated tempo of a track in beats per minute (BPM). In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration.
 >
-> -   **valence:** A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
+> **valence:** A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track.
 Tracks with high valence sound more positive (e.g.
@@
 #| label: fig-means
-#| fig-cap: Comparing the Mean Values of Each Variable
+#| fig-cap: Comparing the Range and Mean Values of Each Variable From the Spotify Audio Features API for our Full Data Set
 #| echo: false
@@
 
-By analyzing the cumulative distribution functions of various audio features, particularly when categorizing songs based on whether they were produced by Jack Antonoff (@fig-cdf), distinct patterns emerge.
+By analyzing the cumulative distribution functions of various audio features, particularly when categorizing songs based on whether they were produced by Antonoff (@fig-cdf), distinct patterns emerge.
 Notably, while attributes such as liveness, instrumentalness, and speechiness exhibit relative homogeneity, there are slight disparities in features like danceability, energy, loudness, acousticness, and valence.
@@
 #| label: fig-cdf
-#| fig-cap: Cumulative Distribution Function of Each Variable 
+#| fig-cap: Cumulative Distribution Function of Each Variable From the Spotify Audio Features API for our Full Data Set
 #| echo: false
@@
 Examining the temporal evolution of artists' sounds, especially in correlation with their collaboration with Antonoff, provides insights into the broader trends in pop music.
-Each point in the following plots represent one track.
+Each point in figures 3-8 represent one track.
 
@@
 #| label: fig-energy
-#| fig-cap: Energy Variable
+#| fig-cap: Energy Scores For Contemporary Pop Artists Over Time
 #| echo: false
@@
 However, this trend appears consistent with the overall trajectories of these artists' works, suggesting broader shared trends in this area of pop, as exemplified by Lana Del Rey and Taylor Swift.
-Sharon Von Etten, Maggie Rogers, and Mitski, display upward energy trajectories in their discographies.
+Notably, Sharon Von Etten, Maggie Rogers, and Mitski, display upward energy trajectories in their discographies.
 
@@
 #| label: fig-danceability
-#| fig-cap: Danceability Variable
+#| fig-cap: Danceability Scores For Contemporary Pop Artists Over Time
 #| echo: false
@@
 
-No significant trends are obersable in danceability over time across the artists' collaborations with Antonoff, as demonstrated in (@fig-danceability), although there is observable an variance in danceability scores across songs in every album, across all artists.
-This suggests that, over time, danceability scores across tracks in every album are consistently varies over time.
-The Chicks earlier work, between 2002 and 2005, presumable the country album "Home" have the smallest amount of variance in danceability.
-This could be due to genre difference.
+No significant trends are observable in danceability over time across the artists' collaborations with Antonoff, as demonstrated in (@fig-danceability), although a difference in danceability scores across songs in every album, across all artists can be observed.
+This suggests that, over time, danceability scores across tracks in every album are consistently varied over time.
+Notably, The Chicks' earlier work, between 2002 and 2005, have the smallest amount of variance in danceability.
+This could be due to genre difference, since their earlier recordings, like the album "Home", were country.
 
@@
 #| label: fig-valence
-#| fig-cap: Valence Variable
+#| fig-cap: Valence Scores For Contemporary Pop Artists Over Time
 #| echo: false
@@
 
-@fig-valence reveals a notable shift in valence scores in Lorde's discography, where her tracks display a slight upward trend and a greater range in valence scores over time.
+@fig-valence, which focuses on valence, reveals a notable shift in valence scores in Lorde's discography, where her tracks display a slight upward trend and a greater range in valence scores over time.
 Otherwise, similarly to danceability, the valence scores display significant variance across tracks in every album.
@@
 #| label: fig-acousticness
-#| fig-cap: Acousticness Variable
+#| fig-cap: Acousticness Scores For Contemporary Pop Artists Over Time
 #| echo: false
@@
 @fig-acousticness underscores a trend toward increased acousticness in the works of Taylor Swift, Lorde, and HAIM, contrasted with declining trends in artists like Mitski, Sharon Von Etten, and Maggie Rogers.
+Tracks produced by Antonoff also display an upward trend (such as Bleachers or Lorde), or appear in artists' discographies when an upward trend is already underway (such as Lana Del Rey, Taylor Swift, and Florence and the Machine).
 
@@
 #| label: fig-instrumentalness
-#| fig-cap: Instrumentalness Variable
+#| fig-cap: Instrumentalness Scores For Contemporary Pop Artists Over Time
 #| echo: false
@@
 #| label: fig-tempo
-#| fig-cap: Tempo Variable
+#| fig-cap: Tempo Scores For Contemporary Pop Artists Over Time
 #| echo: false
@@
 
-Finally, tempo trends, as shown in Figure @fig-tempo, are generally stable across artists, with notable exceptions in collaborations involving Lana Del Rey and Taylor Swift.
+Finally, tempo trends, as shown in @fig-tempo, are generally stable across artists, with notable exceptions in collaborations involving Lana Del Rey and Taylor Swift.
 In Lana Del Rey's tracks, her collaborations with Antonoff see an upward trend in tempo.
@@
 
-We've observed subtle trends in variables associated with songs produced by Jack Antonoff, as evidenced by our cumulative distribution functions and minor trends in the evolution of artists' sound characteristics, when collaborating with Antonoff (particularly energy).
+We have observed subtle trends in variables associated with songs produced by Jack Antonoff, as evidenced by our cumulative distribution functions and minor trends in the evolution of artists' sound characteristics, when collaborating with Antonoff (particularly energy).
 
@@
 $$
-Pr(y_i​=1)=logit−1(β_0​+β_1​×danceability+β_2​×energy+...+β_{10}​×tempo)
+\begin{aligned}
+Pr(y_i = 1) = & \text{logit}^{-1}(β_0 + β_1 \times \text{danceability} + β_2 \times \text{energy} \\
+& + β_3 \times \text{loudness} + β_4 \times \text{mode} + β_5 \times \text{speechiness} \\
+& + β_6 \times \text{acousticness} + β_7 \times \text{instrumentalness} \\
+& + β_8 \times \text{liveness} + β_9 \times \text{valence} + β_{10} \times \text{tempo})
+\end{aligned}
 $$
 
-where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \ldots, \beta_{10}$ are the coefficients corresponding to the 10 predictor variables.
+where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \beta_2, \beta_3, \beta_4, \beta_5, \beta_6, \beta_7, \beta_8, \beta_9, \beta_{10}$ are the coefficients corresponding to the 10 predictor variables.
 
-After some minor pre-processing using the `tidyverse` R package [@citetidyverse]. We followed @citeSilge to build a model that deals with the imbalanced classes.
+After some minor pre-processing using the `tidyverse` package [@citetidyverse] in R [@citeR]. We followed @citeSilge to build a model that deals with the imbalanced classes.
 
@@
 
-![](images/lr_ds_orig.jpeg)
-
 ## Random Forest
@@
 Random forests (RF) build an ensemble of decision trees during the training phase, and then use majority voting for classification tasks.
+For each input of data, each decision tree in the forest makes a prediction, and the final classification is determined by the majority vote of these predictions.
+For example, if more trees in the ensemble predict a certain class for the input data than any other class, the RF model assigns that class to the input.
 Each tree in the ensemble is trained on a random sample of data taken with replacement, and only a random subset of features are considered when splitting at each node.
@@
 
-The training process for the RF model can be summarized as follows:
+The training process for the RF model can be summarized as follows (@citehastie) :
 
@@
 
-The RF was trained using the @ranger package. The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
+The RF was trained using the @citeRanger package in R (@citeR). The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
 
@@
 
-Gini impurity was chosen on the grounds that this study is interested in not only classifying producers, but, more importantly, in understanding how the underlying data informs these decisions.
+Gini impurity was chosen on the grounds that we are interested in not only classifying producers, but, more importantly, in understanding how the underlying data informs these decisions.
 By calculating the Gini impurity not only for the entire model, but specifically for each producer, we can better understand how the variables in our dataset describe the differences between producers.
@@
 
-The model's accuracy score is 0.757, and the ROC AUC score is 0.819.
-The precision score is 0.3851, meaning that approximately 38.51% of songs predicted to be produced by Antonoff were actually produced by him.
-The recall of 0.7561 indicates that about 75.61% of songs produced by Antonoff were actually correctly identified in the model.
+The model's accuracy is 0.757, and the ROC AUC score is 0.819.
+The precision score is 0.3851, meaning that approximately 39% of songs predicted to be produced by Antonoff were actually produced by him.
+The recall of 0.7561 indicates that about 76% of songs produced by Antonoff were actually correctly identified in the model.
 
@@
   labs(y = NULL) +
-  theme(legend.position = "none")
+  theme(legend.position = "none") +
+  theme_minimal()
 ```
@@
 The accuracy of the model (0.6111) being significantly higher than the No Information Rate (0.4274) means that it is doing more than just blind guessing based on class prevalence.
-With a p-value below 0.05, we can confirm the model's statistical significance over a naive approach.
+A p-value below 0.05 suggests the model may be better than a naive approach.
 
@@
 
-Correlation heatmaps can help us better understand the differences between each producer's "style". The variables with the most variance were danceability, loudness, and energy.
+Correlation heatmaps can help us better understand the differences between each producer's "style". The variables with the most variance were danceability, loudness, and energy (@tbl-rangecorr).
 
@@
 #| warning: false
+#| label: tbl-rangecorr
+#| tbl-cap: Variables Ranked by the Amount of Range
 ## Range of correlation scores
@@
 ggplot(df_importance_long, aes(x = reorder(variable, importance), y = importance)) +
-  geom_bar(stat = "identity") +
+  geom_bar(stat = "identity", fill = "pink") +  
   facet_wrap(~producer, scales = "fixed") + 
@@
 #| label: fig-dancevariance
-#| fig-cap: Danceability Scores Across Producers
+#| fig-cap: Danceability Scores Correlated with Other Features Across Producers
 #| echo: false
@@
 
-Our logistic regression model was not able to strongly predict whether or not a song was produced by Jack Antonoff using the Spotify API's audio features data.
+### Defining Style
+
+Roger Dannenberg differentiates between texture and style to define style: ""Texture" usually refers to sound and the activity of making sound, while "style" is most often used to describe the general impression or intention provided by a texture."" (@dannenberg).
+According to Dannenberg, texture is a composite of aspects of music like that we can hear on longer or shorter time scales, such as harmonic rhythms, or differences in pitches, or loudness, among many others.
+These aspects are more objectively measurable, while style is the "sound colour" embedded in the combination and patterns of these elements.
+
+David Cope's Experiments in Musical Intelligence measures composers' style by detecting recurring patterns (@cope).
+First, recurring patterns are detected within a piece, which indicate the significant stylistic elements specific to that piece.
+Then, many compositions are analyzed by the same producer in order to find patterns specific to that producer.
+When the patterns specific to individual pieces are ruled out, a more latent "composer pattern" emerges.
+This composer pattern is a very basic representation of a composer's style, and can be further developed.
+
+A major limitation to our study is that it doesn't account for patterns or the architecture of songs.
+Our dataset provides metrics that stand in for an entire piece, making a major part of the song inaccessible.
+However, working with this limitation, danceability, loudness, and energy scores emerged as key predictors in successful instances of classification.
+
+Using Dannenberg's definitions of texture and style as a framework, we can understand each feature of the dataset as encoding different elements of texture.
+In applying statistical models to these features, the underlying patterns in the relationship between these textures can be taken as an approximation for style.
+
+### Our Results
+
+For instance, our logistic regression model was not able to strongly predict whether or not a song was produced by Jack Antonoff using the Spotify API's audio features data.
 These data are limited, providing static metrics that stand in for entire tracks, which may not be detailed enough for intra-genre classification.
@@
 
-Even Antonoff-produced songs with the highest `danceability` scores aren't exactly dance hits, compared to production styles like that of Max Martin.
+On that note, even Antonoff-produced songs with the highest `danceability` scores aren't exactly dance hits, compared to production styles like that of Max Martin.
 
-For instance, the highest scoring song in danceability*, I Think He Knows* by Taylor Swift, seems danceable, yet the next highest-scoring songs -- Taylor Swift's *Vigilante Shit* or Florence and the Machine's *Heaven is Here* -- don't resonate as dance tracks.
-However, they aren't understated, ambient tracks either -- the exhibit some sort of compelling quality, like they're coercing your body to do something.
+For instance, the highest scoring song in danceability*, I Think He Knows* by Taylor Swift, can pass as danceable, yet the next highest-scoring songs -- Taylor Swift's *Vigilante Shit* or Florence and the Machine's *Heaven is Here* -- don't resonate as dance tracks.
+But they aren't understated, ambient tracks either -- they exhibit some sort of visceral quality, like they're coercing your body to do something.
 Like Gamman's description of Antonoff's vocal design sounding like how they would sound to the person who's singing them, these tracks seem like they're trying to get inside of your skin.
@@
 *I* *Wanna Get Better* by Bleachers typifies the Antonoff style as identified by our model.
-This song is more anthemic, resonating with a sort of neo-Springsteen aesthetic that Antonoff is often characterized (or admitting) to striving for, than danceable.
+This song is more anthemic, resonating with a sort of neo-Springsteen aesthetic that Antonoff is often characterized as (or admitting to) being influenced by, than danceable (@citePareles, @citeHorn) .
 It also does sound like it's trying to demand your attention, and the speechiness, loudness, and energy which are distinctly correlated with danceability in Antonoff's RF results, are all apparent.
@@
 
-What does the danceability variable then signify in Antonoff's music?
-Our findings suggest that his most "danceable" songs are better understood through the lens of correlated features, even though danceability emerges as the primary predictor.
-This crowd-sourced metric, in conjunction with machine-listening-derived features, offers a unique insight into an artist's style, capturing an intuitive understanding of a song's tone and energy.
+### "Danceability" and "Valence" Features' Crowd-sourced Construction
 
@@
 Macdonald revealed that `valence` and `danceability` were created by giving tracks to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
-Variables like `energy` or `instrumentalness`, on the other hand, were determined primarily through machine listening techniques, with human subjectivity being applied to fine-tune the features[^1] (@citeglenint).:
+Variables like `energy` or `instrumentalness`, on the other hand, were determined primarily through machine listening techniques, with human subjectivity being applied to fine-tune the features[^2] (@citeglenint).:
 
-[^1]: A notable example of this being that bluegrass songs were given very high `speechiness` ratings.
+[^2]: A notable example of this being that bluegrass songs were given very high `speechiness` ratings.
 Because there were no banjos in the training data, the instruments were registered as human speech.
@@
 
-Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence) (@citeglenint) .
-High energy and high valence could be generally happy, cheerful, upbeat.
+Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence) (@citeglenint): "High energy and high valence could be generally happy, cheerful, upbeat.
 Low energy, low valence could be sad or downbeat.
 High energy, low valence is sort of angry.
-Low energy, high valence is serene or calming.
+Low energy, high valence is serene or calming."
 
@@
 
+### Conclusion
+
+What does the danceability variable then signify in Antonoff's music?
+Our findings suggest that, similar to McDonald's quadrants, Antonoff's most "danceable" songs are better understood through the lens of correlated features, even though danceability emerges as the primary predictor.
+This crowd-sourced metric, in conjunction with machine-listening-derived features that capture songs' "textures", offers a unique insight into an artist's style, capturing an intuitive understanding of a song's tone and energy.
+
 Our analysis reveals that the amalgamation of crowdsourced metrics (like danceability) and machine-listening variables offers a nuanced approximation of musical style.
-This approach not only mirrors the multifaceted nature of auditory perception but also enriches our model's capacity to capture the essence of an artist's style.
+This approach not only mirrors the multifaceted nature of auditory perception, but also enriches our model's capacity to capture the essence of an artist's style.
 Interestingly, the subjective nature of crowdsourced metrics, which might initially seem rudimentary, plays a crucial role in this process.
-They shape a more intuitive and embodied understanding of music, enhancing the overall analysis.
-
+Our study suggests that they can shape a more intuitive and embodied understanding of music, enhancing the overall analysis.
 For instance, the work of Martikainen et al.
@@
 However, it's important to note that while these features offer a quantitative lens to examine music, they cannot fully encapsulate the nuanced and subjective aspects of musical style and creativity.
-The role of human subjectivity is pivotal in interpreting and experiencing music, as evidenced by the crowdsourced nature of certain features like danceability and valence.
-This aspect becomes even more significant when considering the broader scope of music recommendation systems.
-
+The role of human subjectivity is pivotal in interpreting and experiencing music, and this aspect becomes even more significant when considering the broader scope of music recommendation systems.
 As Nick Seaver's anthropological study of music recommendation companies shows, engineers "develop ways of thinking about musical preference and software, attempting to reconcile them with each other" (@citeSeaver).
+Engineers' understandings of style and taste influence the designs of the music recommendation systems they build.
 This extends to a more nuanced understanding of recommendation systems as a whole, highlighting the intricate balance between computational measures and human interpretation in the evolving landscape of music analysis and appreciation.
 
-Future studies should look into reproducing crowdsourced danceability or valence variables to better understand the underlying user preferences that motivate these metrics.
+Future studies should reproduce crowdsourced danceability or valence variables to better understand the underlying user preferences that motivate these metrics.
 In doing so, we can build more robust ways to computationally model more ephemeral and subjective aspects of music such as style.

### 2023-11-29 outputs/paper/paper.qmd
table of content

@@
 
-The remainder of this paper is structured as:
+The remainder of this paper is structured as: @sec-data details the data curation and extraction processes, including documentation on features from Spotify's audio features data set.
+@sec-model specifies the models used and @sec-results details their performance and most important predictors.
+@sec-discussion situates the study in the broader context of computational stylistics and recommendation systems, detailing how we can conceptualize "style" given the results and limitations of our study.
+We provide background on the Spotify features' provenance based on an interview we conducted with Glen McDonald, a principal engineer at Spotify.
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2023-11-30 outputs/paper/paper.qmd
reader edits cosmetic/style/typos

@@
     email: "michaela.drouillard@mail.utoronto.ca"
-thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop. We thank Inessa De Angelis for valuable suggestions."
+thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop. We thank Rohan Alexander and Inessa De Angelis for valuable suggestions."
 date: "November 24 2023"
@@
 We are interested in whether Spotify audio features can be used to train a model that distinguishes whether a song was produced by Antonoff.
-We obtained tracks from six artists who have collaborated with Antonoff, both before their collaboration and during.
+We obtained tracks from six artists who have collaborated with Antonoff, both before and during their collaboration.
 We also obtained tracks from similarly prolific pop producers: Joel Little, Ariel Rechstaid, Max Martin, Greg Kurstin, Paul Epworth and Rick Nowels.
@@
 
-Our results suggest that we can capture something about Antonoff's music with these variable.
+Our results suggest that we can capture something about Antonoff's music with these variables.
 It might, in the end, just be that subjective response that some people feel, and others enjoy, and that we are all just trying to describe in using different methods and tools.
@@
 
-The remainder of this paper is structured as: @sec-data details the data curation and extraction processes, including documentation on features from Spotify's audio features data set.
+The remainder of this paper is structured as follows: @sec-data details the data curation and extraction processes, including documentation on features from Spotify's audio features data set.
 @sec-model specifies the models used and @sec-results details their performance and most important predictors.
@@
 
-We collected data from the Spotify API, which we accessed using the `spotifyr` package in R [@citeSpotify].
-For our first dataset, we acquired data on Lorde, Taylor Swift, St.
+We collected data from the Spotify API, which we accessed using the `spotifyr` package in R [@citeSpotify, @citeR].
+For our first dataset, we acquired audio feature data on the complete Lorde, Taylor Swift, St.
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
-HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, and Mitski were included to represent other contemporary pop artists with overlapping fan bases who have never collaborated with Antonoff.
+HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Van Etten, and Mitski were included to represent other contemporary pop artists with overlapping fan bases who have never collaborated with Antonoff.
 Of the artists who did collaborate with Antonoff, we chose artists who have produced multiple albums, and at least one album with Antonoff.
@@
 For our second dataset, we acquired the discographies of comparable contemporary pop producers' by scraping their respective discography Wikipedia pages.
+Tracks were included if producers were either the solo producer, the primary producer, or a co-producer (a limitation of this study being that we can not know or truly parse who takes ownership for what during creative collaborations, especially when our audio features use one number to describe an entire song).
 The producers include Epworth, Rechstaid, Max Martin, Nowels, Joel Little, and Greg Kustin.
@@
 However, this trend appears consistent with the overall trajectories of these artists' works, suggesting broader shared trends in this area of pop, as exemplified by Lana Del Rey and Taylor Swift.
-Notably, Sharon Von Etten, Maggie Rogers, and Mitski, display upward energy trajectories in their discographies.
+Notably, Sharon Van Etten, Maggie Rogers, and Mitski, display upward energy trajectories in their discographies.
 
@@
 
-No significant trends are observable in danceability over time across the artists' collaborations with Antonoff, as demonstrated in (@fig-danceability), although a difference in danceability scores across songs in every album, across all artists can be observed.
-This suggests that, over time, danceability scores across tracks in every album are consistently varied over time.
+No significant trends are observable in danceability over time across the artists' collaborations with Antonoff, as demonstrated in @fig-danceability, although a difference in danceability scores across songs in every album, across all artists can be observed.
+This suggests that danceability scores across tracks in every album are consistently varied over time.
 Notably, The Chicks' earlier work, between 2002 and 2005, have the smallest amount of variance in danceability.
@@
 
-@fig-acousticness underscores a trend toward increased acousticness in the works of Taylor Swift, Lorde, and HAIM, contrasted with declining trends in artists like Mitski, Sharon Von Etten, and Maggie Rogers.
+@fig-acousticness underscores a trend toward increased acousticness in the works of Taylor Swift, Lorde, and HAIM, contrasted with declining trends in artists like Mitski, Sharon Van Etten, and Maggie Rogers.
 Tracks produced by Antonoff also display an upward trend (such as Bleachers or Lorde), or appear in artists' discographies when an upward trend is already underway (such as Lana Del Rey, Taylor Swift, and Florence and the Machine).
@@
 Arguments of "Antonoffication" risk removing agency and artistic credit from (female) artists, and attributing both their critical success and failures to Antonoff (@citeWilson).
-In an industry so dominated by male producers, and using features so limited in their ability to describe anything about the process of creation, it would not be appropriate the use these trends to suggest that Antonoff's involvement is the cause behind changes in sound.
+In an industry primarily led by male producers, and using features so limited in their ability to describe anything about the process of creation, it would not be appropriate the use these trends to suggest that Antonoff's involvement is the cause behind changes in sound.
 The purpose of these visualizations is to observe the distribution of variables across artists, over time, with a focus on Antonoff.
@@
 We select a total of 1465 tracks from the dataset from 12 artists: Lorde, Taylor Swift, St.
-Vincent, Lana Del Rey, The Chicks, Florence and the Machine, HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Von Etten, Mitski, and Bleachers (Antonoff's own band).
+Vincent, Lana Del Rey, The Chicks, Florence and the Machine, HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Van Etten, Mitski, and Bleachers.
 
@@
 
-After some minor pre-processing using the `tidyverse` package [@citetidyverse] in R [@citeR]. We followed @citeSilge to build a model that deals with the imbalanced classes.
+After some minor pre-processing using the `tidyverse` package [@citetidyverse] in R [@citeR], we followed Julia Silge's downsampling methodology to build a model that deals with the imbalanced classes (@citeSilge ).
 
-We used the `tidymodels` to fit the model [@citetidymodels].
+The model was fit using `tidymodels` [@citetidymodels].
 To deal with the imbalanced classes, we downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
@@
 `is_antonoff`, `is_elworth`, etc.).
-The producer feature was then converted into a factor variable, and the dataset was split into training and tests sets.
+The producer feature was then converted into a factor variable, and the dataset was split into training and test sets.
 
-The RF was trained using the @citeRanger package in R (@citeR). The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
+The RF was trained using the `ranger` package in R (@citeR, @citeRanger). The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
 
@@
 At every split in a decision tree, the algorithm tests how well each feature splits the data.
-The features that split the data the most accurately, and then with the least impurity, would have the lowest impurity score.
+The features that split the data the most accurately, and thus with the least impurity, would have the lowest impurity score.
 In Random Forests, a variable's importance can be calculated by looking at how much the tree nodes that use that variable reduce impurity on average.
@@
 The model's accuracy is 0.757, and the ROC AUC score is 0.819.
-The precision score is 0.3851, meaning that approximately 39% of songs predicted to be produced by Antonoff were actually produced by him.
-The recall of 0.7561 indicates that about 76% of songs produced by Antonoff were actually correctly identified in the model.
+Based on the confusion matrix presented in Table 2, the precision score is 0.385, meaning that approximately 39% of songs predicted to be produced by Antonoff were actually produced by him.
+The recall of 0.756 indicates that about 76% of songs produced by Antonoff were actually correctly identified in the model.
 
@@
 
+Variable importance, which refers to the impact of each predictor variable in explaining the variation in the response variable, can offer a data-driven picture of what might set a producer's sound apart.
+
 `Danceability`, with a positive importance score of 4.142, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
@@
 
-In the 20 songs in Jack Antonoff's discography with the highest `danceability` scores, Bleachers tracks account for 35% of this list, and Taylor Swift tracks account 40% of this list (@tbl-mostdanceable).
+In the 20 songs in Jack Antonoff's discography with the highest `danceability` scores, Bleachers tracks account for 35% of this list, and Taylor Swift tracks account for 40% of this list (@tbl-mostdanceable).
 St Vincent, The Chicks, Lorde, and Florence and the Machine are also featured.
@@
 
-The accuracy of the model (0.6111) being significantly higher than the No Information Rate (0.4274) means that it is doing more than just blind guessing based on class prevalence.
+As presented in Table 5, the accuracy of the model (0.6111) being significantly higher than the No Information Rate (0.4274) means that it is doing more than just blind guessing based on class prevalence.
 A p-value below 0.05 suggests the model may be better than a naive approach.
@@
 
-Correlation heatmaps can help us better understand the differences between each producer's "style". The variables with the most variance were danceability, loudness, and energy (@tbl-rangecorr).
+The variables with the most variance for each producer's tracks were danceability, loudness, and energy (@tbl-rangecorr).
 
@@
 ```{r, fig.width= 8, fig.height=6}
+#| label: fig-varimpbyprod
+#| tbl-cap: Variable Importance by Producer
 #| echo: false
@@
   coord_flip() +
-  labs(title = "Variable Importance by Producer", 
+  labs( 
        x = "Variable", 
@@
 Relative to other producers, Antonoff's danceability has the least positive correlation with valence (0.311), the most negative correlation with acousticness (-0.0733), and significant positive correlations with speechiness (0.336), loudness (0.366), and energy (0.274).
+In @fig-dancevariance, we can observe how the patterns of variables' correlations with danceability differ between producers.
 
@@
 
-Roger Dannenberg differentiates between texture and style to define style: ""Texture" usually refers to sound and the activity of making sound, while "style" is most often used to describe the general impression or intention provided by a texture."" (@dannenberg).
+Roger Dannenberg differentiates between texture and style to define style: ""Texture" usually refers to sound and the activity of making sound, while "style" is most often used to describe the general impression or intention provided by a texture." (@dannenberg).
 According to Dannenberg, texture is a composite of aspects of music like that we can hear on longer or shorter time scales, such as harmonic rhythms, or differences in pitches, or loudness, among many others.

### 2023-11-30 outputs/paper/paper.qmd
intro edits

@@
 
-We are interested in understanding how to create an approximation for production style.
-To do this we used Spotify API's audio features dataset to explore different audio feature metrics and identify which were the most import predictors in identifying pop producers' style.
+Style has been extensively studied in the field of computational stylistics, with researchers applying methods ranging from (x, y, z) to define and attribute aesthetics and patterns to artists.
+
+We are interested in understanding how to create an approximation for production style using a data set built using the Spotify audio features API.
+To do this, we explore different audio feature metrics and identify which were the most import predictors in identifying pop producers' style.
 We focus on Jack Antonoff, as he is a prolific and widely used producer, who also has his own solo acts.
@@
 
-We are interested in whether Spotify audio features can be used to train a model that distinguishes whether a song was produced by Antonoff.
-We obtained tracks from six artists who have collaborated with Antonoff, both before and during their collaboration.
+We obtained tracks from six artists with extensive discographies who have collaborated with Antonoff, both before and during their collaboration.
 We also obtained tracks from similarly prolific pop producers: Joel Little, Ariel Rechstaid, Max Martin, Greg Kurstin, Paul Epworth and Rick Nowels.
-Our data set contains 3738 tracks in total, which were split into a training set (2802 tracks) and a testing set (936 tracks).
+Our data set contains 3738 tracks in total, which were split into a training and test sets.
 
 We built two models to explore the data.
-The first model, a logistic regression, draws on a dataset of artists who have collaborated with Antonoff (1954 tracks total), with a binary variable to tag whether the song was an Antonoff production.
+The first model, a logistic regression, is employed to observe any patterns in audio feature scores across Antonoff's collaborations with artists.
+The visualizations and analysis offer a closer look at Antonoff's oeuvre, along with a portrait of the trends in female singer-songwriter contemporary pop (according to Spotify's API).
+It draws on a dataset of artists who have collaborated with Antonoff (1954 tracks total), with a binary variable to tag whether the song was an Antonoff production.
 The training set included 1465 tracks, and the testing set included 489 tracks.
-With this model, we explore which variables are stronger predictors of Antonoff's style, developing a better understanding of whether the Spotify data can trace Antonoff's involvement in an artist's sound over time.
-We built a second model, a Random Forest (RF), to complement the insights from the logistic regression.
-The RF trains on a training set that includes other major pop producers in order to classify tracks based on who produced them, and return information on which predictors were most useful in doing so.
 
-We find that `danceability` is the strongest predictor of whether a song has been produced by Antonoff.
-Furthermore, when `danceability`'s correlation scores with other features in the dataset are considered, the variance in correlation scores tends to be strongest classifier of whether a song was produced by Antonoff.
-In other words, danceability in concert with other features in the dataset provides the best predictions.
+The second model, a Random Forest (RF), trains on a data set that includes the discographies of other prominent pop producers to classify producers.
+This expands upon and complements the results of the logistic regression -- where a logistic regression predicting an Antonoff-produced track in an artist's discography might actually be picking up on an artist releasing pop music, training a Random Forest on a bigger dataset with more producers captures both non-linear relationships in the data and accounts for intra-genre differences.
+In obtaining the strongest predictor variables in the Random Forest, we can better understand how it distinguishes between different pop producers' styles.
+
+We find that, in both models, `danceability` is the strongest predictor of whether a song has been produced by Antonoff.
+Additionally, when we look at how danceability relates to other features in the dataset, we find that the differences in these relationships are very informative.
+Specifically, the degree to which danceability's relationships with other features vary helps us most accurately determine if Antonoff produced a song.
+Put simply, using danceability along with its interactions with other dataset features gives us the most accurate predictions.
+
+In an interview with Spotify principal engineer Glen McDonald, we learned that the danceability and valence features were built by Spotify using crowdsourced methods.
 
@@
 It might, in the end, just be that subjective response that some people feel, and others enjoy, and that we are all just trying to describe in using different methods and tools.
+
 With the understanding that what we understand as "data-driven" in recommendation systems is often intertwined with human subjectivity -- engineers' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) -- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.

### 2023-11-30 outputs/paper/paper.qmd
intro reshaping

@@
     email: "michaela.drouillard@mail.utoronto.ca"
-thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop. We thank Rohan Alexander and Inessa De Angelis for valuable suggestions."
+thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop. We thank Rohan Alexander, Carl Wilson and Inessa De Angelis for valuable suggestions."
 date: "November 24 2023"
-abstract: "We explore whether we can predict if a song was produced by Jack Antonoff using the audio features provided by Spotify's API.
-Our logistic regression model has an accuracy of 76%, a precision of 39%, and a recall of 76%.
-Our random forest model has an accuracy of 61%, and a No Information Rate of 43%.
-We found that danceability, a metric trained on human-tagged data, was a key predictor.
-Our results have implications for computational explorations of culture, and understanding how the combination of human subjectivity and machine listening can come to create an approximate for style."
+abstract: "This study investigates the feasibility of predicting whether a song is produced by Jack Antonoff using Spotify's audio feature API.
+We developed two models: a logistic regression model and a random forest model.
+The logistic regression model, focusing on Antonoff's collaborative works, demonstrates an accuracy of 76%, precision of 39%, and a recall of 76%.
+The random forest model, which includes discographies of other notable pop producers, shows an accuracy of 61% and compares favorably to a No Information Rate of 43%.
+A significant finding is that 'danceability', a metric developed from human-tagged data, emerges as a crucial predictor in both models.
+This research contributes to the broader understanding of computational analysis in cultural contexts, illustrating how a blend of human judgment and algorithmic processing can approximate musical styles"
 format: pdf
@@
 
-Style has been extensively studied in the field of computational stylistics, with researchers applying methods ranging from (x, y, z) to define and attribute aesthetics and patterns to artists.
-
-We are interested in understanding how to create an approximation for production style using a data set built using the Spotify audio features API.
+We are interested in understanding how to create an approximation for production style based on a data set built using the Spotify audio features API.
 To do this, we explore different audio feature metrics and identify which were the most import predictors in identifying pop producers' style.
@@
 We find that, in both models, `danceability` is the strongest predictor of whether a song has been produced by Antonoff.
-Additionally, when we look at how danceability relates to other features in the dataset, we find that the differences in these relationships are very informative.
-Specifically, the degree to which danceability's relationships with other features vary helps us most accurately determine if Antonoff produced a song.
-Put simply, using danceability along with its interactions with other dataset features gives us the most accurate predictions.
+Additionally, when we look at how danceability relates to other features in the dataset, we find that the differences in these relationships are informative: the degree to which danceability's relationships with other features vary helps us most accurately determine if Antonoff produced a song.
+In other words, using danceability along with its interactions with other dataset features gives us the most accurate predictions.
 
-In an interview with Spotify principal engineer Glen McDonald, we learned that the danceability and valence features were built by Spotify using crowdsourced methods.
-
-Our results suggest that we can capture something about Antonoff's music with these variables.
-It might, in the end, just be that subjective response that some people feel, and others enjoy, and that we are all just trying to describe in using different methods and tools.
-
-With the understanding that what we understand as "data-driven" in recommendation systems is often intertwined with human subjectivity -- engineers' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) -- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.
+In an interview with Spotify principal engineer Glen McDonald, we learned that the danceability and valence features were built using crowdsourced methods.
+With the understanding that what we understand as "data-driven" in recommendation systems is often intertwined with human subjectivity -- engineers' understandings of taste and taste-making also shape the recommendation systems that they build (@citeSeaver) -- our results suggest that it's possible to create representations of style by combining features derived from both automated audio analysis and more crowdsourced, subjective contributions.
+This research adds to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.
 

### 2023-11-30 outputs/paper/paper.qmd
tweaking abstract

@@
 We developed two models: a logistic regression model and a random forest model.
-The logistic regression model, focusing on Antonoff's collaborative works, demonstrates an accuracy of 76%, precision of 39%, and a recall of 76%.
-The random forest model, which includes discographies of other notable pop producers, shows an accuracy of 61% and compares favorably to a No Information Rate of 43%.
+The logistic regression model, focusing on Antonoff's collaborators' discographies, demonstrates an accuracy of 76%, precision of 39%, and a recall of 76% when predicting whether a track has an Antonoff credit.
+The random forest model, which includes discographies of other notable pop producers, shows an accuracy of 61% when predicting tracks' producers, comparing favorably to a No Information Rate of 43%.
 A significant finding is that 'danceability', a metric developed from human-tagged data, emerges as a crucial predictor in both models.

### 2023-11-30 outputs/paper/paper.qmd
cut fluff from conclusion

@@
 
-Our analysis reveals that the amalgamation of crowdsourced metrics (like danceability) and machine-listening variables offers a nuanced approximation of musical style.
 This approach not only mirrors the multifaceted nature of auditory perception, but also enriches our model's capacity to capture the essence of an artist's style.
-Interestingly, the subjective nature of crowdsourced metrics, which might initially seem rudimentary, plays a crucial role in this process.
-Our study suggests that they can shape a more intuitive and embodied understanding of music, enhancing the overall analysis.
-For instance, the work of Martikainen et al.
-illustrates this point effectively (@podstyle).
-They combine open-source audio tools with qualitative, crowd-sourced listener data to understand audio-based stylistic variation in podcasts.
-This synergy between human subjectivity and algorithmic precision underlines the importance of blending these methods for a comprehensive analysis.
+While crowdsourced metrics might initially seem rudimentary, this subjective aspect can play a crucial role in this process.
+The work of Martikainen et al.
+illustrates this point effectively, combining open-source audio tools with qualitative, crowd-sourced listener data to understand audio-based stylistic variation in podcasts (@podstyle).
 
-However, it's important to note that while these features offer a quantitative lens to examine music, they cannot fully encapsulate the nuanced and subjective aspects of musical style and creativity.
 The role of human subjectivity is pivotal in interpreting and experiencing music, and this aspect becomes even more significant when considering the broader scope of music recommendation systems.
@@
 Engineers' understandings of style and taste influence the designs of the music recommendation systems they build.
-This extends to a more nuanced understanding of recommendation systems as a whole, highlighting the intricate balance between computational measures and human interpretation in the evolving landscape of music analysis and appreciation.
+Our study therefore contributes to a more nuanced understanding of recommendation systems as a whole, highlighting the intricate balance between computational measures and human interpretation in the evolving landscape of music analysis and appreciation.
 

### 2023-12-01 outputs/paper/paper.qmd
yassified it

@@
 
-We find that, in both models, `danceability` is the strongest predictor of whether a song has been produced by Antonoff.
-Additionally, when we look at how danceability relates to other features in the dataset, we find that the differences in these relationships are informative: the degree to which danceability's relationships with other features vary helps us most accurately determine if Antonoff produced a song.
+We find that, in both models, `danceability` is the strongest predictor of whether a song has been produced by Antonoff, and the strongest predictor in differentiating between producers generally.
+Additionally, when we look at how danceability relates to other features in the dataset, we find that the differences in these relationships are informative: the degree to which danceability's relationships with other features vary helps us most accurately classify tracks.
 In other words, using danceability along with its interactions with other dataset features gives us the most accurate predictions.
@@
   geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
-  scale_fill_manual(values = c("blue", "orange"), labels = c("No", "Yes")) +
+  scale_fill_manual(values = c("#B0C4DE", "#9ACD32"), labels = c("No", "Yes")) +
   theme_minimal() +
@@
   theme(legend.position = "none") +
-  theme_minimal()
+  theme_minimal() +
+  scale_fill_manual(values = c(
+    "#FAD4A6", 
+    "#A4D4AE"))
+
 ```
@@
 
+vibrant_colors <- c(
+  "#FAD4A6", # Peach Puff: A light, peachy tone
+  "#D3AE8C", # Tan: A muted, earthy tan color
+  "#B0C4DE", # Light Steel Blue: A soft, pastel blue
+  "#F1BAC3", # Light Pink: A pale pink shade
+  "#D1E0E0", # Powder Blue: A light, airy grey-blue
+  "#E7BE88", # Burlywood: A warm, beige color
+  "#C8A680", # Pale Goldenrod: A deeper, more saturated tan
+  "#A4D4AE", # Pale Green: A light, refreshing green
+  "#E2B0B0", # Rosy Brown: A soft coral tone
+  "#9ACD32"  # Yellow Green: A bright, springy green
+)
+
 df_importance_long <- df_importance %>%
@@
 
-ggplot(df_importance_long, aes(x = reorder(variable, importance), y = importance)) +
-  geom_bar(stat = "identity", fill = "pink") +  
+ggplot(df_importance_long, aes(x = reorder(variable, importance), y = importance, fill = variable)) +
+  geom_bar(stat = "identity") +  
   facet_wrap(~producer, scales = "fixed") + 
   coord_flip() +
-  labs( 
-       x = "Variable", 
-       y = "Mean Decrease Gini") +
-  theme_minimal()
+  labs(x = "Variable", y = "Mean Decrease Gini") +
+  theme_minimal() +
+  scale_fill_manual(values = vibrant_colors)
 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_manual(values = c(
+    "#9ACD32", #springy green
+    "#E7BE88", # Rich warm beige
+    "#A4D4AE", # Pale green
+    "#D3AE8C", # Vibrant taupe
+    "#B0C4DE", # Lively soft blue
+    "#F1BAC3", # Vivid blush pink
+    "#D1E0E0", # Bright light gray
+    "#B0C4DE", # Vivid sky blue
+    "#A4D4AE" # Pale green
+  ))
 ```

### 2023-12-01 outputs/paper/paper.qmd
edits

@@
 
-We are interested in understanding how to create an approximation for production style based on a data set built using the Spotify audio features API.
+It is widely acknowledged that style recognition in music is difficult.
+Style, which is hard enough to define without building computational representations of it, is often constructed using lower-level perceptual features such as a pitch and tempo.
+We are interested in understanding how to create an approximation for production style based on a data set built using the Spotify audio features API, which includes both crowdsourced features and features derived from machine listening techniques.
 To do this, we explore different audio feature metrics and identify which were the most import predictors in identifying pop producers' style.
@@
 
-The aim of the logistic regression is to determine whether we can predict Antonoff's influence in artists' sound.
+The aim of the logistic regression is to determine whether we can predict Antonoff's influence on artists' sound.
 We select a total of 1465 tracks from the dataset from 12 artists: Lorde, Taylor Swift, St.
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Van Etten, Mitski, and Bleachers.
+Marina and the Diamonds, Maggie Rogers, Sharon Van Etten and Mitski have not collaborated with Antonoff, but their tracks were added to contribute more data from the same taste pool to provide more context for the model.
 
@@
 
-After some minor pre-processing using the `tidyverse` package [@citetidyverse] in R [@citeR], we followed Julia Silge's downsampling methodology to build a model that deals with the imbalanced classes (@citeSilge ).
-
+After some minor pre-processing using the `tidyverse` package [@citetidyverse] in R [@citeR], we followed Julia Silge's downsampling methodology to build a model that accounts for imbalanced classes (@citeSilge).
+We downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
+Then, we used 5-fold cross-validation with stratification by the **`is_antonoff`** variable to estimate the performance of the model.
 The model was fit using `tidymodels` [@citetidymodels].
-To deal with the imbalanced classes, we downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
-Then we used 5-fold cross-validation with stratification by the **`is_antonoff`** variable to estimate the performance of the model.
 
@@
 
-The RF was trained using the `ranger` package in R (@citeR, @citeRanger). The 'importance' parameter was set to 'impurity', which sets the model to interpret variable importance using Gini impurity.
-
-Gini impurity is a measure of misclassification.
+The RF was trained using the `ranger` package in R (@citeR, @citeRanger).
+We set the model to interpret variable importance using Gini impurity, which is a measure of misclassification.
 At every split in a decision tree, the algorithm tests how well each feature splits the data.
 The features that split the data the most accurately, and thus with the least impurity, would have the lowest impurity score.
-In Random Forests, a variable's importance can be calculated by looking at how much the tree nodes that use that variable reduce impurity on average.
+A variable's importance can be calculated by looking at how much the tree nodes that use that variable reduce impurity on average.
 
@@
 By calculating the Gini impurity not only for the entire model, but specifically for each producer, we can better understand how the variables in our dataset describe the differences between producers.
-Without song structure, instruments, sound, or other more ephemeral components of style, these differences can be taken as a proxy for style.
+Without song structure, instruments, sound itself, or other more ephemeral components of style, these differences can be taken as a proxy for style.
 
@@
 
-@tbl-importance shows variables' respective importance in determining whether a track was produced by Antonoff.
+@tbl-importance and @fig-importance show variables' respective importance in determining whether a track was produced by Antonoff.
 
@@
 
-`Danceability`, with a positive importance score of 4.142, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
-`Valence` was the second most important, with a negative score of 2.559.
-The third, `instrumentalness` correlates negatively with Antonoff tracks with a negative score of 2.309.
+`Danceability`, with a negative importance score of 4.142, had the highest level of importance in determining whether a song was produced or co-produced by Jack Antonoff.
+`Valence` was the second most important, with a positive score of 2.559.
+The third, `instrumentalness` correlates postively with Antonoff tracks with a score of 2.309.
 
@@
 
-The variables with the most variance for each producer's tracks were danceability, loudness, and energy (@tbl-rangecorr).
+The variables with the most variance in importance between each each producer were danceability, loudness, and energy (@tbl-rangecorr).
 
@@
 
-Similar to the logistic regression, danceability was also, across all producers, the most important predictor in the classification process.
+The variance in variable importance scores are visualized in @fig-varimpbyprod.
+Similar to logistic regression, danceability stands out as a variable with a high degree of variance from producer to producer, meaning that it contributes to succesful classification.
 
@@
 #| label: fig-varimpbyprod
-#| tbl-cap: Variable Importance by Producer
+#| fig-cap: Variable Importance by Producer
 #| echo: false
@@
 
-Relative to other producers, Antonoff's danceability has the least positive correlation with valence (0.311), the most negative correlation with acousticness (-0.0733), and significant positive correlations with speechiness (0.336), loudness (0.366), and energy (0.274).
+Relative to other producers, Antonoff's danceability has the least positive correlation with valence (0.311), the most negative correlation with acousticness (-0.073), and significant positive correlations with speechiness (0.336), loudness (0.366), and energy (0.274).
 In @fig-dancevariance, we can observe how the patterns of variables' correlations with danceability differ between producers.
@@
 
-Roger Dannenberg differentiates between texture and style to define style: ""Texture" usually refers to sound and the activity of making sound, while "style" is most often used to describe the general impression or intention provided by a texture." (@dannenberg).
+Style is bizarre and complex.
+In order to define it, Roger Dannenberg differentiates between texture and style: ""Texture" usually refers to sound and the activity of making sound, while "style" is most often used to describe the general impression or intention provided by a texture." (@dannenberg).
 According to Dannenberg, texture is a composite of aspects of music like that we can hear on longer or shorter time scales, such as harmonic rhythms, or differences in pitches, or loudness, among many others.
@@
 
-David Cope's Experiments in Musical Intelligence measures composers' style by detecting recurring patterns (@cope).
+We can build on these lower level perceptual features to arrive at approximations of style.
+For instance, David Cope's Experiments in Musical Intelligence measures composers' style by detecting recurring patterns (@cope).
 First, recurring patterns are detected within a piece, which indicate the significant stylistic elements specific to that piece.
@@
 
-A major limitation to our study is that it doesn't account for patterns or the architecture of songs.
+Our study follows a similar line of inquiry -- the Spotify API mostly provides metrics related to lower level perceptual features (with danceability and valence being the obvious anomalies).
+However, a major limitation to our study is that it doesn't account for patterns within a track or the overall architecture of songs (or, for that matter, albums).
 Our dataset provides metrics that stand in for an entire piece, making a major part of the song inaccessible.
-However, working with this limitation, danceability, loudness, and energy scores emerged as key predictors in successful instances of classification.
+On a similar note, we also attribute one producer to each track.
+In pop production, co-productions are frequent, and the general spirit of the genre is more collaborative, making the idea of detecting a latent "composer pattern" harder to differentiate from the collaborations they're embedded in.
+
+Working with these limitations, danceability, loudness, and energy scores still emerge as key predictors in successful instances of classification.
+Whether or not we can attribute these stylistic differences solely to the named producer, it is a classification nonetheless.
 
@@
 
-For instance, our logistic regression model was not able to strongly predict whether or not a song was produced by Jack Antonoff using the Spotify API's audio features data.
-These data are limited, providing static metrics that stand in for entire tracks, which may not be detailed enough for intra-genre classification.
+For instance, our logistic regression model was not able to strongly predict whether a song was produced by Jack Antonoff using the Spotify API's audio features data.
 However, when classifying the tracks, `danceability` was the strongest predictor.
 
-Our Random Forest results, which were moderately strong at classifying tracks, indicate that Antonoff's music, characterized by its `danceability` score, exhibits a significant positive correlation with factors such as `speechiness`, `loudness`, and `energy`, while showing a negative correlation with `acousticness`, and a less pronounced positive correlation with `valence`, compared to other music producers.
-This profile---marked by speechiness, loudness, energy, and a blend of positivity with a touch of nostalgia---aligns with both detailed auditory analysis and critical assessments of Antonoff's sound.
+Our Random Forest results, which were moderately strong at classifying tracks, indicate that Antonoff's music, characterized by its `danceability` score, exhibits a significant positive correlation with factors such as `speechiness`, `loudness`, and `energy`, while showing a negative correlation with `acousticness`, and a less pronounced positive correlation with `valence` compared to other music producers (@fig-dancevariance).
+This profile---marked by speechiness, loudness, energy, and a blend of positivity with a touch of nostalgia---aligns with both our interpretations and critical assessments of Antonoff's sound.
 
-Antonoff's tracks have been described as anthemic (@citeRosen), with a distinct vocal treatment that makes it sound, as Caleb Gamman, a Youtuber who went viral for his critique of Jack Antonoffs sound, observes, "the way it would sound to the person who's singing them" (@citegammanint).
+Antonoff's tracks have been described as anthemic (@citeRosen), with a distinct vocal treatment that makes it sound, as Caleb Gamman, a Youtuber who went viral for his critique of Jack Antonoffs sound, told us, "the way it would sound to the person who's singing them" (@citegammanint).
 
@@
 And so it sort of creates this effect of -- **like he mixes his vocals the way it would sound to the person who's singing , them.
-Which is sort of strange.** He often has no vocals on that upper end, which is very unusual, and then often there are random little bits of noise happening up there which is the sort of thing her likes to do.
+Which is sort of strange.** He often has no vocals on that upper end, which is very unusual, and then often there are random little bits of noise happening up there which is the sort of thing he likes to do.
 
@@
 For instance, the highest scoring song in danceability*, I Think He Knows* by Taylor Swift, can pass as danceable, yet the next highest-scoring songs -- Taylor Swift's *Vigilante Shit* or Florence and the Machine's *Heaven is Here* -- don't resonate as dance tracks.
-But they aren't understated, ambient tracks either -- they exhibit some sort of visceral quality, like they're coercing your body to do something.
+But they aren't understated, ambient tracks either -- they exhibit some sort of visceral quality.
 Like Gamman's description of Antonoff's vocal design sounding like how they would sound to the person who's singing them, these tracks seem like they're trying to get inside of your skin.
@@
 
-Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence) (@citeglenint): "High energy and high valence could be generally happy, cheerful, upbeat.
+Mcdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence) (@citeglenint): "High energy and high valence could be generally happy, cheerful, upbeat.
 Low energy, low valence could be sad or downbeat.
@@
 
-This approach not only mirrors the multifaceted nature of auditory perception, but also enriches our model's capacity to capture the essence of an artist's style.
+This approach not only mirrors the multifaceted nature of auditory perception, but also somehow enriches our model's capacity to capture the essence of an artist's style, or at least distinguish between producers in a limited data set.
 While crowdsourced metrics might initially seem rudimentary, this subjective aspect can play a crucial role in this process.
@@
 
-The role of human subjectivity is pivotal in interpreting and experiencing music, and this aspect becomes even more significant when considering the broader scope of music recommendation systems.
-As Nick Seaver's anthropological study of music recommendation companies shows, engineers "develop ways of thinking about musical preference and software, attempting to reconcile them with each other" (@citeSeaver).
-Engineers' understandings of style and taste influence the designs of the music recommendation systems they build.
-Our study therefore contributes to a more nuanced understanding of recommendation systems as a whole, highlighting the intricate balance between computational measures and human interpretation in the evolving landscape of music analysis and appreciation.
+The role of human subjectivity in interpreting and experiencing music alongside computational representations of it becomes even more apparent when considering the broader scope of music recommendation systems.
+In Nick Seaver's anthropological study of music recommendation companies, Seaver writes that engineers "develop ways of thinking about musical preference and software, attempting to reconcile them with each other" (@citeSeaver), referring to the fact that engineers' understandings of style and taste influence the designs of the music recommendation systems they build.
+Our study therefore contributes to a more nuanced understanding of recommendation systems as a whole, highlighting the balance between computational measures and human interpretation in the evolving landscape of music analysis and appreciation.
 

### 2023-12-01 outputs/paper/paper.qmd
rohan edits

@@
   - name: Michaela Drouillard
-    email: "michaela.drouillard@mail.utoronto.ca"
-thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop. We thank Rohan Alexander, Carl Wilson and Inessa De Angelis for valuable suggestions."
-date: "November 24 2023"
+thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop.
+We thank Rohan Alexander, Carl Wilson and Inessa De Angelis for valuable suggestions.
+For any comments or suggestions, please contact michaela.drouillard@mail.utoronto.ca"
+date: "today"
+date-format: "long"
 abstract: "This study investigates the feasibility of predicting whether a song is produced by Jack Antonoff using Spotify's audio feature API.
@@
 csl: modern-language-association.csl
+number-depth: 2
 ---
@@
 Examining the temporal evolution of artists' sounds, especially in correlation with their collaboration with Antonoff, provides insights into the broader trends in pop music.
-Each point in figures 3-8 represent one track.
+Each point in Figures 3-8 represent one track.
 
@@
 The model's accuracy is 0.757, and the ROC AUC score is 0.819.
-Based on the confusion matrix presented in Table 2, the precision score is 0.385, meaning that approximately 39% of songs predicted to be produced by Antonoff were actually produced by him.
+Based on the confusion matrix presented in @tbl-logres, the precision score is 0.385, meaning that approximately 39% of songs predicted to be produced by Antonoff were actually produced by him.
 The recall of 0.756 indicates that about 76% of songs produced by Antonoff were actually correctly identified in the model.
@@
 
-: Confusion Matrix for Logistic Regression Results
+: Confusion Matrix for Logistic Regression Results {#tbl-logres}
 
@@
 `Valence` was the second most important, with a positive score of 2.559.
-The third, `instrumentalness` correlates postively with Antonoff tracks with a score of 2.309.
+The third, `instrumentalness` correlates positively with Antonoff tracks with a score of 2.309.
 
@@
 
-: Overall Performance Statistics
+: Overall Performance Statistics {#tbl-perfstatsRF}
 
-As presented in Table 5, the accuracy of the model (0.6111) being significantly higher than the No Information Rate (0.4274) means that it is doing more than just blind guessing based on class prevalence.
+As presented in @tbl-perfstatsRF, the accuracy of the model (0.6111) being significantly higher than the No Information Rate (0.4274) means that it is doing more than just blind guessing based on class prevalence.
 A p-value below 0.05 suggests the model may be better than a naive approach.
@@
 
-The variables with the most variance in importance between each each producer were danceability, loudness, and energy (@tbl-rangecorr).
+The variables with the most variance in importance between each producer were danceability, loudness, and energy (@tbl-rangecorr).
 
@@
 The variance in variable importance scores are visualized in @fig-varimpbyprod.
-Similar to logistic regression, danceability stands out as a variable with a high degree of variance from producer to producer, meaning that it contributes to succesful classification.
+Similar to logistic regression, danceability stands out as a variable with a high degree of variance from producer to producer, meaning that it contributes to successful classification.
 
@@
 
-### Defining Style
+## Defining Style
 
@@
 
-### Our Results
+## Our Results
 
@@
 
-Antonoff's tracks have been described as anthemic (@citeRosen), with a distinct vocal treatment that makes it sound, as Caleb Gamman, a Youtuber who went viral for his critique of Jack Antonoffs sound, told us, "the way it would sound to the person who's singing them" (@citegammanint).
+Antonoff's tracks have been described as anthemic (@citeRosen), with a distinct vocal treatment that makes it sound, as Caleb Gamman, a Youtuber who went viral for his critique of Jack Antonoff's sound, told us, "the way it would sound to the person who's singing them" (@citegammanint).
 
@@
 You hear more of your own voice bouncing around in your head.
-And so it sort of creates this effect of -- **like he mixes his vocals the way it would sound to the person who's singing , them.
+And so it sort of creates this effect of -- **like he mixes his vocals the way it would sound to the person who's singing, them.
 Which is sort of strange.** He often has no vocals on that upper end, which is very unusual, and then often there are random little bits of noise happening up there which is the sort of thing he likes to do.
@@
 *I* *Wanna Get Better* by Bleachers typifies the Antonoff style as identified by our model.
-This song is more anthemic, resonating with a sort of neo-Springsteen aesthetic that Antonoff is often characterized as (or admitting to) being influenced by, than danceable (@citePareles, @citeHorn) .
+This song is more anthemic, resonating with a sort of neo-Springsteen aesthetic that Antonoff is often characterized as (or admitting to) being influenced by, than danceable (@citePareles, @citeHorn).
 It also does sound like it's trying to demand your attention, and the speechiness, loudness, and energy which are distinctly correlated with danceability in Antonoff's RF results, are all apparent.
@@
 
-### "Danceability" and "Valence" Features' Crowd-sourced Construction
+## "Danceability" and "Valence" Features' Crowd-sourced Construction
 
 We spoke with Glen McDonald, a Principal Engineer at Spotify who worked at The Echo Nest, a music intelligence start-up that was acquired by Spotify in 2014.
-While Macdonald did not build the features, he has worked with them, and provided information on their provenance.
-Macdonald revealed that `valence` and `danceability` were created by giving tracks to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
+While McDonald did not build the features, he has worked with them, and provided information on their provenance.
+McDonald revealed that `valence` and `danceability` were created by giving tracks to college interns and asking them to tag whether a song was positive or gloomy, or danceable or un-danceable.
 Variables like `energy` or `instrumentalness`, on the other hand, were determined primarily through machine listening techniques, with human subjectivity being applied to fine-tune the features[^2] (@citeglenint).:
@@
 
-Glen confirmed that this process didn't account for lyrics, which made valence a particularly difficult variable to build.
+McDonald confirmed that this process didn't account for lyrics, which made valence a particularly difficult variable to build.
 It could, in theory, pick up on aspects of vocal performance, but it didn't process anything about languages or words or the meanings of songs.
@@
 
-Mcdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence) (@citeglenint): "High energy and high valence could be generally happy, cheerful, upbeat.
+McDonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence) (@citeglenint): "High energy and high valence could be generally happy, cheerful, upbeat.
 Low energy, low valence could be sad or downbeat.
@@
 
-### Conclusion
+## Conclusion
 

### 2023-12-01 outputs/paper/paper.qmd
abstract lol

@@
 A significant finding is that 'danceability', a metric developed from human-tagged data, emerges as a crucial predictor in both models.
-This research contributes to the broader understanding of computational analysis in cultural contexts, illustrating how a blend of human judgment and algorithmic processing can approximate musical styles"
+This research contributes to the broader understanding of computational analysis in cultural contexts, illustrating how a blend of human judgment and algorithmic processing can approximate musical styles."
 format: pdf

### 2023-12-04 outputs/paper/paper.qmd
thanks
Minor tweaks throughout

@@
 thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop.
-We thank Rohan Alexander, Carl Wilson and Inessa De Angelis for valuable suggestions.
+We thank Rohan Alexander, Carl Wilson and Inessa De Angelis, Lindsay Katz, Michael Chong, and Ciara Zogheib for valuable suggestions.
 For any comments or suggestions, please contact michaela.drouillard@mail.utoronto.ca"
@@
 date-format: "long"
-abstract: "This study investigates the feasibility of predicting whether a song is produced by Jack Antonoff using Spotify's audio feature API.
-We developed two models: a logistic regression model and a random forest model.
+abstract: "We investigate the feasibility of predicting whether a song was produced by Jack Antonoff using Spotify's audio feature API.
+We develop two models: a logistic regression model and a random forest model.
 The logistic regression model, focusing on Antonoff's collaborators' discographies, demonstrates an accuracy of 76%, precision of 39%, and a recall of 76% when predicting whether a track has an Antonoff credit.
@@
 library(yardstick)
-
+library(kableExtra)
+library(dplyr)
 ```
@@
 Style, which is hard enough to define without building computational representations of it, is often constructed using lower-level perceptual features such as a pitch and tempo.
-We are interested in understanding how to create an approximation for production style based on a data set built using the Spotify audio features API, which includes both crowdsourced features and features derived from machine listening techniques.
-To do this, we explore different audio feature metrics and identify which were the most import predictors in identifying pop producers' style.
+We are interested in understanding how to create an approximation for production style based on a dataset built using the Spotify audio features API.
+This includes both crowdsourced features and features derived from machine listening techniques.
+To do this, we explore different audio feature metrics and identify which are the most import predictors in identifying pop producers' style.
 We focus on Jack Antonoff, as he is a prolific and widely used producer, who also has his own solo acts.
@@
 We also obtained tracks from similarly prolific pop producers: Joel Little, Ariel Rechstaid, Max Martin, Greg Kurstin, Paul Epworth and Rick Nowels.
-Our data set contains 3738 tracks in total, which were split into a training and test sets.
+Our dataset contains 3,738 tracks in total, which were split into a training and test sets.
 
@@
 The visualizations and analysis offer a closer look at Antonoff's oeuvre, along with a portrait of the trends in female singer-songwriter contemporary pop (according to Spotify's API).
-It draws on a dataset of artists who have collaborated with Antonoff (1954 tracks total), with a binary variable to tag whether the song was an Antonoff production.
-The training set included 1465 tracks, and the testing set included 489 tracks.
+It draws on a dataset of artists who have collaborated with Antonoff (1,954 tracks total), with a binary variable to tag whether the song was an Antonoff production.
+The training set included 1,465 tracks, and the testing set included 489 tracks.
 
-The second model, a Random Forest (RF), trains on a data set that includes the discographies of other prominent pop producers to classify producers.
-This expands upon and complements the results of the logistic regression -- where a logistic regression predicting an Antonoff-produced track in an artist's discography might actually be picking up on an artist releasing pop music, training a Random Forest on a bigger dataset with more producers captures both non-linear relationships in the data and accounts for intra-genre differences.
-In obtaining the strongest predictor variables in the Random Forest, we can better understand how it distinguishes between different pop producers' styles.
+The second model, a random forest, trains on a dataset that includes the discographies of other prominent pop producers to classify producers.
+This expands on and complements the results of the logistic regression -- where a logistic regression predicting an Antonoff-produced track in an artist's discography might be picking up on an artist releasing pop music, training a random forest on a bigger dataset with more producers captures both non-linear relationships in the data and accounts for intra-genre differences.
+In obtaining the strongest predictor variables in the random forest, we can better understand how it distinguishes between different pop producers' styles.
 
@@
 In an interview with Spotify principal engineer Glen McDonald, we learned that the danceability and valence features were built using crowdsourced methods.
-With the understanding that what we understand as "data-driven" in recommendation systems is often intertwined with human subjectivity -- engineers' understandings of taste and taste-making also shape the recommendation systems that they build (@citeSeaver) -- our results suggest that it's possible to create representations of style by combining features derived from both automated audio analysis and more crowdsourced, subjective contributions.
-This research adds to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.
+With the understanding that what we understand as "data-driven" in recommendation systems is often intertwined with human subjectivity -- engineers' understandings of taste and taste-making also shape the recommendation systems that they build (@citeSeaver) -- our results suggest that it is possible to create representations of style by combining features derived from both automated audio analysis and more crowdsourced, subjective contributions.
+This research adds to a more informed understanding of how style and tastes can be algorithmically rendered using Spotify's audio features.
 
-The remainder of this paper is structured as follows: @sec-data details the data curation and extraction processes, including documentation on features from Spotify's audio features data set.
+The remainder of this paper is structured as follows: @sec-data details the data curation and extraction processes, including documentation on features from Spotify's audio features dataset.
 @sec-model specifies the models used and @sec-results details their performance and most important predictors.
@@
 
-We collected data from the Spotify API, which we accessed using the `spotifyr` package in R [@citeSpotify, @citeR].
-For our first dataset, we acquired audio feature data on the complete Lorde, Taylor Swift, St.
+We collected data from the Spotify API, which we accessed using the `spotifyr` package in R [@citeSpotify; @citeR].
+We first acquired audio feature data on the complete Lorde, Taylor Swift, St.
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.
-HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Van Etten, and Mitski were included to represent other contemporary pop artists with overlapping fan bases who have never collaborated with Antonoff.
+HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Van Etten, and Mitski were included to represent other contemporary pop artists with similar fan bases who have never collaborated with Antonoff.
 Of the artists who did collaborate with Antonoff, we chose artists who have produced multiple albums, and at least one album with Antonoff.
-This dataset contains 1954 tracks in total.
+This dataset contains 1,954 tracks in total.
 
-For our second dataset, we acquired the discographies of comparable contemporary pop producers' by scraping their respective discography Wikipedia pages.
-Tracks were included if producers were either the solo producer, the primary producer, or a co-producer (a limitation of this study being that we can not know or truly parse who takes ownership for what during creative collaborations, especially when our audio features use one number to describe an entire song).
+For our second dataset, we acquired the discographies of comparable contemporary pop producers by scraping their respective discography Wikipedia pages.
+Tracks were included if producers were either the solo producer, the primary producer, or a co-producer (a limitation of this study being that we cannot know or truly parse who takes ownership for what during creative collaborations, especially when our audio features use one number to describe an entire song).
 The producers include Epworth, Rechstaid, Max Martin, Nowels, Joel Little, and Greg Kustin.
 These producers were chosen based both on their popularity, and because some have collaborated with the same artists as Antonoff.
-We combined these tracks with the first datasets tracks, for a total of 3738 tracks from 828 different artists[^1].
+We combined these tracks with the first datasets tracks, for a total of 3,738 tracks from 828 different artists[^1].
 
@@
 
-We created a binary variable, `is_antonoff`, which contains 1 if Jack Antonoff produced or co-produced the song, and 0 if it was produced by somebody else.
+We created a binary variable, `is_antonoff`, which is 1 if Jack Antonoff produced or co-produced the song, and 0 if it was produced by somebody else.
 The information underpinning this discography was drawn from the "Jack Antonoff Production Discography" Wikipedia page [@citewiki].
@@
 #| label: tbl-producercounts
-#| tbl-cap: Counts of Tracks Per Producer in Original Dataset
+#| tbl-cap: Number of tracks per producer in the original dataset
 #| echo: false
 #| warning: false
-lr_data<- read_csv(here::here("inputs/data/lr_data.csv"))
 
-library(kableExtra)
+lr_data <- read_csv(here::here("inputs/data/lr_data.csv"))
 
@@
 tbl |>
-  kable(col.names = c("Producer", "Number of Tracks"),
+  kable(
+    col.names = c("Producer", "Number of Tracks"),
     digits = 1,
     booktabs = TRUE,
-  linesep = "")
-
+    linesep = ""
+  )
 ```
@@
 
-Spotify provides variables for each song, some of which were developed by them, which enables comparisons across artists.
-The final audio features in our data set are: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `jack`.
-
-@fig-means depicts the range of scores for the tracks across all variables except for tempo, mode, and loudness, which are measured using different metrics.
+Spotify provides variables for each song, some of which they developed, which enables comparisons across artists.
+The final audio features in our dataset are: `artist_name`, `track_name`, `energy`, `danceability`, `key`, `loudness`, `mode`, `speechiness`, `acousticness`, `instrumentalness`, `liveness`, `valence`, `tempo`, `is_antonoff`.
 Spotify documents the `get_artist_audio_features` variables as follows (@citeSpotifyDocumentation):
@@
 >
-> **danceability**: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
+> **danceability**: Describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity.
 A value of 0.0 is least danceable and 1.0 is most danceable.
 >
-> **energy**: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
+> **energy**: A measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity.
 Typically, energetic tracks feel fast, loud, and noisy.
@@
 >
-> **speechiness:** Speechiness detects the presence of spoken words in a track.
+> **speechiness:** Detects the presence of spoken words in a track.
 The more exclusively speech-like the recording (e.g.
@@
 
+By analyzing the cumulative distribution functions of various audio features, particularly when categorizing songs based on whether they were produced by Antonoff (@fig-cdf), distinct patterns emerge.
+Notably, while attributes such as liveness, instrumentalness, and speechiness exhibit relative homogeneity, there are slight disparities in features like danceability, energy, loudness, acousticness, and valence.
+(@fig-means in @sec-datarange depicts the range of scores for the tracks across all variables except for tempo, mode, and loudness, which are measured using different metrics.)
+
 ```{r, fig.width= 8, fig.height=6}
-#| label: fig-means
-#| fig-cap: Comparing the Range and Mean Values of Each Variable From the Spotify Audio Features API for our Full Data Set
+#| label: fig-cdf
+#| fig-cap: Cumulative Distribution Function of each variable from the Spotify Audio Features API for our full dataset
 #| echo: false
@@
 
-lr_data <- read_csv(here::here("inputs/data/lr_data.csv"))
-
-lr_data %>%
-  select(-producer, -tempo, -loudness, -mode) %>%
-  pivot_longer(cols = -is_antonoff, names_to = "feature", values_to = "value") %>%
-  ggplot(aes(x = feature, y = value, fill = factor(is_antonoff))) +
-  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
-  scale_fill_manual(values = c("#B0C4DE", "#9ACD32"), labels = c("No", "Yes")) +
-  theme_minimal() +
-  labs(fill = "Antonoff", x = "Feature", y = "Value") + 
-  coord_flip()
-
-```
-
-By analyzing the cumulative distribution functions of various audio features, particularly when categorizing songs based on whether they were produced by Antonoff (@fig-cdf), distinct patterns emerge.
-Notably, while attributes such as liveness, instrumentalness, and speechiness exhibit relative homogeneity, there are slight disparities in features like danceability, energy, loudness, acousticness, and valence.
-
-```{r, fig.width= 8, fig.height=6}
-#| label: fig-cdf
-#| fig-cap: Cumulative Distribution Function of Each Variable From the Spotify Audio Features API for our Full Data Set
-#| echo: false
-#| warning: false
-av<- read_csv(here::here("inputs/data/antonoff_viz.csv"))
+av <- read_csv(here::here("inputs/data/antonoff_viz.csv"))
 
@@
 levels(av$is_antonoff) <- c("No", "Yes")
+
 av |>
@@
        color = "Antonoff")
-
 ```
@@
 #| label: fig-energy
-#| fig-cap: Energy Scores For Contemporary Pop Artists Over Time
+#| fig-cap: Energy scores for contemporary pop artists over time
 #| echo: false
@@
 av <- av |>
-  mutate(is_antonoff= factor(is_antonoff))
+  mutate(is_antonoff = factor(is_antonoff))
 
@@
 #| label: fig-danceability
-#| fig-cap: Danceability Scores For Contemporary Pop Artists Over Time
+#| fig-cap: Danceability scores for contemporary pop artists over time
 #| echo: false
@@
   labs(color = "Antonoff", x = "Album Release Date", y = "Danceability") +
-  facet_wrap( ~ artist_name, ncol = 3)
+  facet_wrap(~ artist_name, ncol = 3)
 ```
@@
 #| label: fig-valence
-#| fig-cap: Valence Scores For Contemporary Pop Artists Over Time
+#| fig-cap: Valence scores for contemporary pop artists over time
 #| echo: false
@@
   labs(color = "Antonoff", x = "Album Release Date", y = "Valence") +
-  facet_wrap( ~ artist_name, ncol = 3)
+  facet_wrap(~ artist_name, ncol = 3)
 ```
@@
 #| label: fig-acousticness
-#| fig-cap: Acousticness Scores For Contemporary Pop Artists Over Time
+#| fig-cap: Acousticness scores for contemporary pop artists over time
 #| echo: false
@@
   labs(color = "Antonoff", x = "Album Release Date", y = "Acousticness") +
-  facet_wrap( ~ artist_name, ncol = 3)
+  facet_wrap(~ artist_name, ncol = 3)
 ```
@@
 #| label: fig-instrumentalness
-#| fig-cap: Instrumentalness Scores For Contemporary Pop Artists Over Time
+#| fig-cap: Instrumentalness scores for contemporary pop artists over time
 #| echo: false
@@
   labs(color = "Antonoff", x = "Album Release Date", y = "Instrumentalness") +
-  facet_wrap( ~ artist_name, ncol = 3)
+  facet_wrap(~ artist_name, ncol = 3)
 ```
@@
 #| label: fig-tempo
-#| fig-cap: Tempo Scores For Contemporary Pop Artists Over Time
+#| fig-cap: Tempo scores for contemporary pop artists over time
 #| echo: false
@@
   labs(color = "Antonoff", x = "Album Release Date", y = "Tempo") +
-  facet_wrap( ~ artist_name, ncol = 3)
+  facet_wrap(~ artist_name, ncol = 3)
 ```
@@
 Arguments of "Antonoffication" risk removing agency and artistic credit from (female) artists, and attributing both their critical success and failures to Antonoff (@citeWilson).
-In an industry primarily led by male producers, and using features so limited in their ability to describe anything about the process of creation, it would not be appropriate the use these trends to suggest that Antonoff's involvement is the cause behind changes in sound.
+In an industry primarily led by male producers, and using features so limited in their ability to describe anything about the process of creation, it would not be appropriate to use these trends to suggest that Antonoff's involvement is the cause behind changes in sound.
 The purpose of these visualizations is to observe the distribution of variables across artists, over time, with a focus on Antonoff.
 
-This exploratory data analysis sets the stage for a) employing logistic regression to quantify the likelihood of Antonoff's involvement in a song based on these audio features, b) classifying tracks produced by a wider range of producers using a Random Forest, c) identifying the most influential variables in this classification to better understand how style can be inferred from the provided data set features.
+This exploratory data analysis sets the stage for:
+
+a) employing logistic regression to quantify the likelihood of Antonoff's involvement in a song based on these audio features, 
+b) classifying tracks produced by a wider range of producers using a random forest, 
+c) identifying the most influential variables in this classification to better understand how style can be inferred from the provided dataset features.
 
@@
 The aim of the logistic regression is to determine whether we can predict Antonoff's influence on artists' sound.
-We select a total of 1465 tracks from the dataset from 12 artists: Lorde, Taylor Swift, St.
+We select a total of 1,465 tracks from the dataset from 12 artists: Lorde, Taylor Swift, St.
 Vincent, Lana Del Rey, The Chicks, Florence and the Machine, HAIM, Marina and the Diamonds, Maggie Rogers, Sharon Van Etten, Mitski, and Bleachers.
@@
 \begin{aligned}
-Pr(y_i = 1) = & \text{logit}^{-1}(β_0 + β_1 \times \text{danceability} + β_2 \times \text{energy} \\
+\mbox{Pr}(y_i = 1) = & \text{logit}^{-1}(β_0 + β_1 \times \text{danceability} + β_2 \times \text{energy} \\
 & + β_3 \times \text{loudness} + β_4 \times \text{mode} + β_5 \times \text{speechiness} \\
@@
 
-where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1, \beta_2, \beta_3, \beta_4, \beta_5, \beta_6, \beta_7, \beta_8, \beta_9, \beta_{10}$ are the coefficients corresponding to the 10 predictor variables.
+where $\text{Pr}(y_i=1)$ is the probability of a song being produced by Antonoff, $\beta_0$ is the intercept, and $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$, $\beta_6$, $\beta_7$, $\beta_8$, $\beta_9$, and $\beta_{10}$ are the coefficients corresponding to the 10 predictor variables.
 
-After some minor pre-processing using the `tidyverse` package [@citetidyverse] in R [@citeR], we followed Julia Silge's downsampling methodology to build a model that accounts for imbalanced classes (@citeSilge).
+After pre-processing using the `tidyverse` package [@citetidyverse] in R [@citeR], we followed Julia Silge's downsampling methodology to build a model that accounts for imbalanced classes (@citeSilge).
 We downsampled the majority class using the **`step_downsample()`** function from the **`themis`** package.
@@
 
-## Random Forest
+## Random forest
 
-Random forests (RF) build an ensemble of decision trees during the training phase, and then use majority voting for classification tasks.
+Random forests build an ensemble of decision trees during the training phase, and then use majority voting for classification tasks.
 For each input of data, each decision tree in the forest makes a prediction, and the final classification is determined by the majority vote of these predictions.
-For example, if more trees in the ensemble predict a certain class for the input data than any other class, the RF model assigns that class to the input.
+For example, if more trees in the ensemble predict a certain class for the input data than any other class, the random forest model assigns that class to the input.
 Each tree in the ensemble is trained on a random sample of data taken with replacement, and only a random subset of features are considered when splitting at each node.
@@
 
-The training process for the RF model can be summarized as follows (@citehastie) :
+The training process for the random forest model can be summarized as follows (@citehastie):
 
@@
         iii. Split the node into two daughter nodes.
-2.  The ensemble of trees $( \{ T_b \}_{b=1}^B )$ constitutes the random forest model.
+2.  The ensemble of trees $\left( \{ T_b \}_{b=1}^B \right)$ constitutes the random forest model.
 
-For a new instance $( x )$, the RF model prediction $( \hat{C}^{RF}(x) )$ is determined by the majority vote from the ensemble of trees:
+For a new instance $( x )$, the random forest model prediction $( \hat{C}^{RF}(x) )$ is determined by the majority vote from the ensemble of trees:
 
@@
 
-Where $( \hat{C}_b(x) )$ is the class prediction of the $(b)$-th tree.
+Where $\left( \hat{C}_b(x) \right)$ is the class prediction of the $(b)$-th tree.
 
@@
 
-The RF was trained using the `ranger` package in R (@citeR, @citeRanger).
+The random forest was trained using the `ranger` package in R (@citeR, @citeRanger).
 We set the model to interpret variable importance using Gini impurity, which is a measure of misclassification.
@@
 
-@tbl-importance and @fig-importance show variables' respective importance in determining whether a track was produced by Antonoff.
+@tbl-importance and @fig-importance show the respective importance of the variables in determining whether a track was produced by Antonoff.
 
@@
                linesep = "")
-
 ```
@@
     "#A4D4AE"))
-
 ```
@@
   linesep = "")
-
 ```
@@
 
-## Random Forest
+## Random forest
 
@@
 
-library(dplyr)
-
 table_data <- read_csv(here::here("inputs/data/table_data.csv"))
@@
 range_data <- table_data %>%
-  gather(variable, value, -producer, -term) %>%
+  gather(variable, value,-producer,-term) %>%
   group_by(variable) %>%
@@
 range_data |>
-  kable(col.names = 
+  kable(
+    col.names =
       c("Variable", "Min Value", "Max Value", "Range"),
@@
     booktabs = TRUE,
-        linesep = "")
-
+    linesep = ""
+  )
 ```
@@
 
-df_importance<- read_csv(here::here("inputs/data/df_importance.csv"))
+df_importance <-
+  read_csv(here::here("inputs/data/df_importance.csv"))
 
 vibrant_colors <- c(
-  "#FAD4A6", # Peach Puff: A light, peachy tone
-  "#D3AE8C", # Tan: A muted, earthy tan color
-  "#B0C4DE", # Light Steel Blue: A soft, pastel blue
-  "#F1BAC3", # Light Pink: A pale pink shade
-  "#D1E0E0", # Powder Blue: A light, airy grey-blue
-  "#E7BE88", # Burlywood: A warm, beige color
-  "#C8A680", # Pale Goldenrod: A deeper, more saturated tan
-  "#A4D4AE", # Pale Green: A light, refreshing green
-  "#E2B0B0", # Rosy Brown: A soft coral tone
+  "#FAD4A6",
+  # Peach Puff: A light, peachy tone
+  "#D3AE8C",
+  # Tan: A muted, earthy tan color
+  "#B0C4DE",
+  # Light Steel Blue: A soft, pastel blue
+  "#F1BAC3",
+  # Light Pink: A pale pink shade
+  "#D1E0E0",
+  # Powder Blue: A light, airy grey-blue
+  "#E7BE88",
+  # Burlywood: A warm, beige color
+  "#C8A680",
+  # Pale Goldenrod: A deeper, more saturated tan
+  "#A4D4AE",
+  # Pale Green: A light, refreshing green
+  "#E2B0B0",
+  # Rosy Brown: A soft coral tone
   "#9ACD32"  # Yellow Green: A bright, springy green
@@
 df_importance_long <- df_importance %>%
-  gather(variable, importance, -producer)
+  gather(variable, importance,-producer)
 
-ggplot(df_importance_long, aes(x = reorder(variable, importance), y = importance, fill = variable)) +
+ggplot(df_importance_long,
+       aes(
+         x = reorder(variable, importance),
+         y = importance,
+         fill = variable
+       )) +
   geom_bar(stat = "identity") +
-  facet_wrap(~producer, scales = "fixed") + 
+  facet_wrap( ~ producer, scales = "fixed") +
   coord_flip() +
@@
   scale_fill_manual(values = vibrant_colors)
-
 ```
@@
 
-selected_data <- read_csv(here::here("inputs/data/selected_data.csv"))
+selected_data <-
+  read_csv(here::here("inputs/data/selected_data.csv"))
 
-ggplot(selected_data, aes(x=term, y=danceability, fill=producer)) +
-  geom_bar(stat="identity", position="dodge") +
-  labs(y="Danceability Score", x= "Feature") +
+ggplot(selected_data, aes(x = term, y = danceability, fill = producer)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(y = "Danceability Score", x = "Feature") +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_fill_manual(values = c(
-    "#9ACD32", #springy green
-    "#E7BE88", # Rich warm beige
-    "#A4D4AE", # Pale green
-    "#D3AE8C", # Vibrant taupe
-    "#B0C4DE", # Lively soft blue
-    "#F1BAC3", # Vivid blush pink
-    "#D1E0E0", # Bright light gray
-    "#B0C4DE", # Vivid sky blue
+  scale_fill_manual(
+    values = c(
+      "#9ACD32",
+      #springy green
+      "#E7BE88",
+      # Rich warm beige
+      "#A4D4AE",
+      # Pale green
+      "#D3AE8C",
+      # Vibrant taupe
+      "#B0C4DE",
+      # Lively soft blue
+      "#F1BAC3",
+      # Vivid blush pink
+      "#D1E0E0",
+      # Bright light gray
+      "#B0C4DE",
+      # Vivid sky blue
       "#A4D4AE" # Pale green
-  ))
+    )
+  )
 ```
@@
 Style is bizarre and complex.
-In order to define it, Roger Dannenberg differentiates between texture and style: ""Texture" usually refers to sound and the activity of making sound, while "style" is most often used to describe the general impression or intention provided by a texture." (@dannenberg).
+To define it Roger Dannenberg differentiates between texture and style: ""Texture" usually refers to sound and the activity of making sound, while "style" is most often used to describe the general impression or intention provided by a texture." (@dannenberg).
 According to Dannenberg, texture is a composite of aspects of music like that we can hear on longer or shorter time scales, such as harmonic rhythms, or differences in pitches, or loudness, among many others.
@@
 
-We can build on these lower level perceptual features to arrive at approximations of style.
+We can build on these lower-level perceptual features to arrive at approximations of style.
 For instance, David Cope's Experiments in Musical Intelligence measures composers' style by detecting recurring patterns (@cope).
 First, recurring patterns are detected within a piece, which indicate the significant stylistic elements specific to that piece.
-Then, many compositions are analyzed by the same producer in order to find patterns specific to that producer.
+Then, many compositions are analyzed by the same producer to find patterns specific to that producer.
 When the patterns specific to individual pieces are ruled out, a more latent "composer pattern" emerges.
@@
 
-Our study follows a similar line of inquiry -- the Spotify API mostly provides metrics related to lower level perceptual features (with danceability and valence being the obvious anomalies).
-However, a major limitation to our study is that it doesn't account for patterns within a track or the overall architecture of songs (or, for that matter, albums).
+Our study follows a similar line of inquiry -- the Spotify API mostly provides metrics related to lower-level perceptual features (with danceability and valence being the obvious anomalies).
+However, a major limitation to our study is that it does not account for patterns within a track or the overall architecture of songs (or, for that matter, albums).
 Our dataset provides metrics that stand in for an entire piece, making a major part of the song inaccessible.
@@
 
-For instance, our logistic regression model was not able to strongly predict whether a song was produced by Jack Antonoff using the Spotify API's audio features data.
+Our logistic regression model was not able to strongly predict whether a song was produced by Jack Antonoff using the Spotify API's audio features data.
 However, when classifying the tracks, `danceability` was the strongest predictor.
 
-Our Random Forest results, which were moderately strong at classifying tracks, indicate that Antonoff's music, characterized by its `danceability` score, exhibits a significant positive correlation with factors such as `speechiness`, `loudness`, and `energy`, while showing a negative correlation with `acousticness`, and a less pronounced positive correlation with `valence` compared to other music producers (@fig-dancevariance).
+Our random forest results, which were moderately strong at classifying tracks, indicate that Antonoff's music, characterized by its `danceability` score, exhibits a significant positive correlation with factors such as `speechiness`, `loudness`, and `energy`, while showing a negative correlation with `acousticness`, and a less pronounced positive correlation with `valence` compared to other music producers (@fig-dancevariance).
 This profile---marked by speechiness, loudness, energy, and a blend of positivity with a touch of nostalgia---aligns with both our interpretations and critical assessments of Antonoff's sound.
 
-Antonoff's tracks have been described as anthemic (@citeRosen), with a distinct vocal treatment that makes it sound, as Caleb Gamman, a Youtuber who went viral for his critique of Jack Antonoff's sound, told us, "the way it would sound to the person who's singing them" (@citegammanint).
+Antonoff's tracks have been described as anthemic (@citeRosen), with a distinct vocal treatment that makes it sound, as Caleb Gamman, a YouTuber who went viral for his critique of Jack Antonoff's sound, told us, "the way it would sound to the person who's singing them" (@citegammanint).
 
@@
 
-On that note, even Antonoff-produced songs with the highest `danceability` scores aren't exactly dance hits, compared to production styles like that of Max Martin.
+On that note, even Antonoff-produced songs with the highest `danceability` scores are not exactly dance hits, compared to production styles like that of Max Martin.
 
-For instance, the highest scoring song in danceability*, I Think He Knows* by Taylor Swift, can pass as danceable, yet the next highest-scoring songs -- Taylor Swift's *Vigilante Shit* or Florence and the Machine's *Heaven is Here* -- don't resonate as dance tracks.
-But they aren't understated, ambient tracks either -- they exhibit some sort of visceral quality.
-Like Gamman's description of Antonoff's vocal design sounding like how they would sound to the person who's singing them, these tracks seem like they're trying to get inside of your skin.
-But they don't have the rhythm of a dance track.
+For instance, the highest scoring song in danceability, *I Think He Knows* by Taylor Swift, can pass as danceable, yet the next highest-scoring songs -- Taylor Swift's *Vigilante Shit* or Florence and the Machine's *Heaven is Here* -- do not resonate as dance tracks.
+But they are not understated, ambient tracks either -- they exhibit some sort of visceral quality.
+Like Gamman's description of Antonoff's vocal design sounding like how they would sound to the person who is singing them, these tracks seem like they are trying to get inside of your skin.
+But they do not have the rhythm of a dance track.
 
-*I* *Wanna Get Better* by Bleachers typifies the Antonoff style as identified by our model.
+*I Wanna Get Better* by Bleachers typifies the Antonoff style as identified by our model.
 This song is more anthemic, resonating with a sort of neo-Springsteen aesthetic that Antonoff is often characterized as (or admitting to) being influenced by, than danceable (@citePareles, @citeHorn).
-It also does sound like it's trying to demand your attention, and the speechiness, loudness, and energy which are distinctly correlated with danceability in Antonoff's RF results, are all apparent.
-The redemptive, broken-but-getting-better tone can also be heard in the lyrics, but it still doesn't have the kind of rhythm for dancing -- which maps onto how Antonoff's "valence" score isn't as a high as other pop artists in @fig-dancevariance.
+It also does sound like it is trying to demand your attention, and the speechiness, loudness, and energy which are distinctly correlated with danceability in Antonoff's random forest results, are all apparent.
+The redemptive, broken-but-getting-better tone can also be heard in the lyrics, but it still does not have the kind of rhythm for dancing -- which maps onto how Antonoff's "valence" score isn't as a high as other pop artists in @fig-dancevariance.
 
@@
 
-> "You could imagine writing a formula for energy that combines loudness and tempo and degree of harmonic variation or something.
+> You could imagine writing a formula for energy that combines loudness and tempo and degree of harmonic variation or something.
 So that feature was \[machine learning\], but that one's more like a human helping a machine figure out a formula.
@@
 So there's a spectrum from, loudness as purely analytical, and then energy is a little like loudness, with a little more subjectivity.
-And then danceability and valence are purely subjective."
+And then danceability and valence are purely subjective.
 
-McDonald confirmed that this process didn't account for lyrics, which made valence a particularly difficult variable to build.
-It could, in theory, pick up on aspects of vocal performance, but it didn't process anything about languages or words or the meanings of songs.
-Take an upbeat, happy-sounding Elliot Smith song with devastating lyrics: the machine might register it as happy, where human listeners understand that it's sad.
+McDonald confirmed that this process did not account for lyrics, which made valence a particularly difficult variable to build.
+It could, in theory, pick up on aspects of vocal performance, but it did not process anything about languages or words or the meanings of songs.
+Take an upbeat, happy-sounding Elliot Smith song with devastating lyrics: the machine might register it as happy, where human listeners understand that it is sad.
 Furthermore, two humans might even disagree on the song's valence: "Plus, we have the confounding factor of like, a song that seems happy today could be sad tomorrow because the singer was killed in a plane crash.
-The song didn't change, but our world changed and our reactions changed" (@citeglenint).
+The song did not change, but our world changed and our reactions changed" (@citeglenint).
 
@@
 #| warning: false
+
 av <- read.csv(here::here("inputs/data/antonoff_viz.csv"))
@@
   theme_minimal() +
-  scale_color_manual(values = c("lightblue", "orange"),
+  scale_color_manual(
+    values = c("lightblue", "orange"),
     labels = c("No", "Yes"),
-                     name = "Antonoff")
+    name = "Antonoff"
+  )
 ```
@@
 
-This approach not only mirrors the multifaceted nature of auditory perception, but also somehow enriches our model's capacity to capture the essence of an artist's style, or at least distinguish between producers in a limited data set.
+This approach not only mirrors the multifaceted nature of auditory perception, but also somehow enriches our model's capacity to capture the essence of an artist's style, or at least distinguish between producers in a limited dataset.
 While crowdsourced metrics might initially seem rudimentary, this subjective aspect can play a crucial role in this process.
@@
 
+# Appendix
+
+## Data range and mean {#sec-datarange}
+
+```{r, fig.width= 8, fig.height=6}
+#| label: fig-means
+#| fig-cap: Comparing the range and mean of each variable from the Spotify Audio Features API for our full dataset
+#| echo: false
+#| warning: false
+
+lr_data <- read_csv(here::here("inputs/data/lr_data.csv"))
+
+lr_data %>%
+  select(-producer,-tempo,-loudness,-mode) %>%
+  pivot_longer(cols = -is_antonoff,
+               names_to = "feature",
+               values_to = "value") %>%
+  ggplot(aes(
+    x = feature,
+    y = value,
+    fill = factor(is_antonoff)
+  )) +
+  geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
+  scale_fill_manual(values = c("#B0C4DE", "#9ACD32"),
+                    labels = c("No", "Yes")) +
+  theme_minimal() +
+  labs(fill = "Antonoff", x = "Feature", y = "Value") +
+  coord_flip()  
+```)
+
+\newpage
+
 # References

### 2023-12-05 outputs/paper/paper.qmd
fixing appendix

@@
   select(-producer,-tempo,-loudness,-mode) %>% 
-  pivot_longer(cols = -is_antonoff,
-               names_to = "feature",
-               values_to = "value") %>%
-  ggplot(aes(
-    x = feature,
-    y = value,
-    fill = factor(is_antonoff)
-  )) +
+  pivot_longer(cols = -is_antonoff, names_to = "feature", values_to = "value") %>% 
+  ggplot(aes( x = feature, y = value, fill = factor(is_antonoff) )) +
   geom_boxplot(alpha = 0.6, position = position_dodge(width = 0.8)) +
-  scale_fill_manual(values = c("#B0C4DE", "#9ACD32"),
-                    labels = c("No", "Yes")) +
+  scale_fill_manual(values = c("#B0C4DE", "#9ACD32"), labels = c("No", "Yes")) + 
   theme_minimal() + 
@@
   coord_flip()
-```)
+```
 
